---
title: "Input Files"
editor_options:
  chunk_output_type: console
---

```{r set-knitr-opts, echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r packages}
library(data.tree)
library(tidyverse)
```

## Directory Structure

The GRAVI workflow requires a set directory structure. 
If using the template repository [as advised](quickstart.html) this will be mostly taken care of.
The required directory structure is

```{r}
data.frame(
    pathString = file.path(
        "project_home/", c("analysis", "config", "data", "docs", "output", "workflow")
    )
) %>% 
    as.Node()
```

- Rmarkdown scripts will be added and executed from the `analysis` directory
- Key configuration files are provided in the `config` directory
- *Your data* should be placed in the `data` directory as described below
- The `html` output summarising all results will be produced in the `docs` directory
- Additional output files will be placed in `output`
- The workflow itself is run by all code supplied in the `workflow` directory

## Alignments

The GRAVI workflow currently takes `bam` files as the primary input.
Multiple workflows exist for quality control, adapter removal and de-duplication and it is assumed that supplied reads will have been pre-processed with the above steps, then aligned to the genome of interest.

Files should be placed in the `data/aligned` directory as set in `config.yml`, although this can be changed if desired.
Each ChIP target should be placed in a separate directory using the example layout as given below.
As IgG input may be shared between ChIP targets, these can all be placed in the directory `data/Input` and this is effectively hard-wired into the workflow.

```{r, echo = FALSE, results='markup'}
file.path(
    "project_home", "data", "aligned", 
    c(
        "TF1/control_rep1.bam", "TF1/control_rep2.bam", "TF1/control_rep3.bam",
        "TF1/treat_rep1.bam", "TF1/treat_rep2.bam", "TF1/treat_rep3.bam",
        "TF2/control_rep1.bam", "TF2/control_rep2.bam", "TF2/control_rep3.bam",
        "TF2/treat_rep1.bam", "TF2/treat_rep2.bam", "TF2/treat_rep3.bam",
        "Input/pooled_input.bam"
    )
) %>% 
    data.frame() %>% 
    setNames("pathString") %>% 
    as.Node()
```

## Additional Files

Optional files can also be supplied and is it customary to place these in `data/external` with relative paths (to `project_home`) added to `config.yml`

```{r}
file.path(
    "project_home", "data",
    c(
        "aligned/TF1", "aligned/TF2",
        "external/rnaseq_topTable.tsv",
        "external/external_features.gtf",
        "external/hic_interactions.bedpe",
        "external/additional_coverage.bw"
    )
)%>% 
    data.frame() %>% 
    setNames("pathString") %>% 
    as.Node()
```

