## Comparison With Transcriptomic Data

```{r define-rna-objects}
library(fgsea)
library(metap)
library(ggside)
results_tbl <- combined_results %>% 
  splitAsList(.$status) %>% 
  endoapply(unique) %>% 
  unlist() %>%
  unname() %>% 
  as_tibble() %>% 
  unnest(gene_id) %>% 
  mutate(gene_name = id2gene[gene_id])
rna <- lapply(rna, addDiffStatus)
has_de <- map_lgl(rna, \(x) any(x$FDR < 0.05))
pw_gene_status <- results_tbl %>%
  dplyr::select(status, gene_id) %>% 
  unnest(gene_id) %>% 
  distinct(status, gene_id) %>% 
  mutate(
    cat = str_count(status, "(In|De)creased") + str_count(status, "Unch") / 2
  ) %>% 
  arrange(gene_id, desc(cat)) %>% 
  dplyr::filter(cat == max(cat), .by = gene_id) %>% 
  dplyr::select(status, gene_id)
enrich_cols$pathway <- enrich_cols$gs_name
enrich_cols$NES <- colDef(
  format = colFormat(digits = 2), maxWidth = 100,
  style = \(value) {
    text_colour <- ifelse(
      value > 0, 
      direction_colours[["Increased"]], 
      direction_colours[["Decreased"]]
    )
    list(color = text_colour)
  }
)
enrich_cols$meta_p <- enrich_cols$p
enrich_cols$n_cmn <- enrich_cols$observed_gene_hits
enrich_cols$n_cmn$name <- "Nbr Common"
enrich_cols$genes <- enrich_cols$genes_with_hits
enrich_cols$genes$name <- "Common Genes"
enrich_cols$cmn_ids <- enrich_cols$gene_id
```

In order to compare pairwise results across both `r pander(unname(unlist(short_names)))`, gene-to-range mappings were prepared by selecting the category showing the greatest change with either comparison.
Pairwise status was scored based on the presence of an Increased/Decreased status (score = 1) in either comparison, with a lower score assigned to an Unchanged status (score = 0.5).
The most higly ranked pairwise status was assigned to each gene, allowing for ties, giving `r comma(length(unique(pw_gene_status$gene_id)))` unique genes with `r comma(nrow(pw_gene_status))` mapped pairwise status levels.

These results were applied where summarised results were able to be presented, however, when comparing actual estimates for logFC etc, *all ranges mapped to a gene* were retained.


```{r summary-tables}
summary_tables <- names(rna) %>% 
  lapply(
    \(i) {
      x <- rna[[i]]
      id <- paste0("tbl-summary-", i) %>% 
        str_to_lower() %>% 
        str_replace_all("_", "-")
      tbl <- results_tbl %>%
        full_join(x, by = "gene_id") %>% 
        mutate(
          status.x = fct_na_value_to_level(status.x, "Undetected - Undetected"),
          status.y = fct_na_value_to_level(status.y, "Undetected")
        ) %>% 
        arrange(desc(status.x)) %>% 
        distinct(gene_id, .keep_all = TRUE) %>% 
        dplyr::select(gene_id, starts_with("status")) %>% 
        summarise(
          n = dplyr::n(), .by = starts_with("status")
        ) %>%
        pivot_wider(names_from = "status.y", values_from = "n", values_fill = 0) %>% 
        mutate(
          `% Changed` = rowSums(dplyr::select(., ends_with("creased"))) / rowSums(
            dplyr::select(., any_of(c("Increased", "Decreased", "Unchanged")))
          )
        ) %>% 
        dplyr::select(
          status.x, `% Changed`, any_of(c("Increased", "Decreased")), 
          everything()
        ) %>% 
        separate(status.x, unlist(short_names)) %>% 
        reactable(
          showPageSizeOptions = nrow(.) > 10, filterable = TRUE,
          columns = list2(
            "{short_names[[1]]}" := colDef(
              filterInput = \(values, name) drop_filter(values, name, id)
            ),
            "{short_names[[2]]}" := colDef(
              filterInput = \(values, name) drop_filter(values, name, id)
            ),
            "% Changed" = colDef(filterMethod = js_greater, format = percent_col),
            Increased = colDef(filterMethod = js_greater),
            Decreased = colDef(filterMethod = js_greater),
            Unchanged = colDef(filterMethod = js_greater),
            Undetected = colDef(filterMethod = js_greater),
          )[colnames(.)],
          defaultColDef = colDef(
            sortable = TRUE, format = comma_col,
            footerStyle = list(fontWeight = "bold")
          ),
          columnGroups = list(
            colGroup(
              "Gene Expression", str_subset(names(.), "ed$")
            )
          ),
          elementId = id
        )
      cp <- htmltools::em(
        glue(
          "Comparison between pairwise ",
          glue_collapse(short_names, last = " and "),
          " status with gene expression from {i}."
        )
      )
      htmltools::div(
        htmltools::div(
          id = id,
          class = "section level5",
          htmltools::h5("Summary Table"),
          htmltools::div(
            class = "table", tbl, htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(rna))
```



```{r volcano-plots, echo = any(has_de)}
volcano_plots <- list()
if (any(has_de)) {
  volcano_plots <- names(which(has_de)) %>% 
    lapply(
      \(i) {
        p <- rna[[i]] %>% 
          arrange(PValue) %>% 
          left_join(pw_gene_status, by = "gene_id") %>% 
          mutate(
            status.y = fct_na_value_to_level(status.y, "Undetected - Undetected"),
            gene_name = id2gene[gene_id]
          ) %>% 
          separate(status.y, unlist(short_names), sep = " - ") %>% 
          mutate(
            "{short_names[[1]]}" := paste(short_names[[1]], !!sym(short_names[[1]])),
            "{short_names[[2]]}" := paste(short_names[[2]], !!sym(short_names[[2]])),
          ) %>% 
          droplevels() %>% 
          ggplot(aes(logFC, -log10(PValue), colour = status.x)) +
          geom_point(alpha = 0.7) +
          geom_label_repel(
            aes(label = gene_name),
            data = . %>%
              dplyr::filter(grepl("creased", status.x)) %>%
              distinct(
                gene_name, status.x, !!!syms(short_names), .keep_all = TRUE
              ) %>%
              dplyr::slice(
                1:2,
                .by = all_of(c("status.x", unname(unlist(short_names))))
              ),
            show.legend = FALSE, alpha = 0.7, size = 3, fontface = "italic"
          ) +
          geom_text(
            aes(label = label),
            data = . %>%
              summarise(
                logFC = max(logFC), PValue = 1,
                n = dplyr::n(), .by = all_of(unname(unlist(short_names)))
              ) %>%
              mutate(
                logFC = max(logFC), label = paste("n =", comma(n))
              ),
            colour = "grey20", hjust = 1, vjust = 0, size = 3.5
          ) +
          facet_grid(
            rows = vars(!!sym(short_names[[2]])), 
            cols = vars(!!sym(short_names[[1]])),
          ) +
          labs(
            x = "logFC (RNA)", y = expr(paste(-log[10], "PValue")), 
            colour = "DE Status"
          ) +
          scale_colour_manual(values = direction_colours)
        
        ## Export the image
        img_out <- file.path(fig_path, glue("volcano_{i}.{fig_type}"))
        fig_fun(filename = img_out, width = fw, height = fh)
        print(p)
        dev.off()
        
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "Volcano plot for {i} showing genes separated by pairwise status. ",
            "The number of genes in each panel is given in the bottom right corner. ",
            "Where possible, the two most highly ranked DE genes are labelled.",
            "Given that multiple loci may be associated with a single gene, ",
            "genes may appear labelled in multiple panels"
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>% 
              basename() %>% 
              str_remove_all(glue(".{fig_type}$")) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level5",
            htmltools::h5("Volcano Plot"),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      }
    ) %>% 
    setNames(names(which(has_de)))
}
```
  
```{r pval-by-status}
pval_hist <- names(rna) %>% 
  lapply(
    \(i) {
      p <- rna[[i]] %>% 
        left_join(pw_gene_status, by = "gene_id") %>% 
        mutate(
          status.y = fct_na_value_to_level(status.y, "Undetected - Undetected") %>% 
            fct_relabel(
              \(x) x %>% 
                str_replace_all("Increased.+Increased", "Both Increased") %>% 
                str_replace_all("Decreased.+Decreased", "Both Decreased") %>% 
                str_replace_all(
                  "(.+) - (.+)",
                  paste(short_names[[1]], "\\1 -", short_names[[2]], "\\2")
                )
            ) %>% 
            fct_lump(5),
          gene_name = id2gene[gene_id]
        ) %>% 
        droplevels() %>% 
        ggplot(aes(PValue, after_stat(density))) +
        geom_histogram(
          fill = "grey70", colour = "black", binwidth = 0.01
        ) +
        facet_wrap(~status.y) +
        scale_y_continuous(expand = expansion(c(0, 0.05))) +
        scale_x_continuous(expand = expansion(0.025))
      y_max <- max(layer_scales(p)$y$range$range)
      p <- p +
        geom_text(
          aes(y = y, label = label),
          data = . %>% 
            summarise(
              n = dplyr::n(), .by = status.y
            ) %>% 
            mutate(
              PValue = 0.02, 
              y = y_max,
              label = paste("n =", comma(n))
            ),
          hjust = 0, vjust = 1, size = 3.5
        ) +
        labs(x = "PValue (RNA)", y = "Density")
      
      ## Export the image
      img_out <- file.path(fig_path, glue("pvals_{i}.{fig_type}"))
      fh <- ifelse(
        length(levels(p$data$status.y)) == 3, 2 * fh / 3, fh
      )
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "P-values for {i} separated by pairwise status. ",
          "Only the most common 5 groups are shown with all others grouped ",
          "into the status 'Other'"
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level5",
          htmltools::h5("P-Value Histogram"),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(rna))
```

```{r pw-deg-tables}
pw_deg_tables <- list()
if (any(has_de)) {
  pw_deg_tables <- has_de %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(i) {
        fs <- 12
        id <- paste0("tbl-", i, "-pairwise") %>% 
          str_to_lower() %>% 
          str_replace_all("_", "-")
        tbl <- rna[[i]] %>% 
          dplyr::filter(FDR < 0.05) %>% 
          left_join(as_tibble(gtf$gene)[c("range", "gene_id")]) %>% 
          inner_join(
            dplyr::filter(results_tbl, grepl("(In|De)creased", status)),
            by = "gene_id"
          ) %>% 
          mutate(d = distance(GRanges(range.x), GRanges(centre))) %>% 
          dplyr::select(-range.x) %>% 
          dplyr::rename(range = range.y) %>% 
          dplyr::select(
            starts_with("gene"), all_of(c("logCPM", "logFC", "FDR")),
            starts_with("status"), range, d, any_of(c("region", "feature"))
          ) %>% 
          dplyr::select(-starts_with("PValue")) %>% 
          dplyr::filter(str_detect(status.y, "(In|De)creased")) %>% 
          separate(status.y, unlist(short_names), sep = " - ") %>% 
          reactable(
            showPageSizeOptions = nrow(.) > 10, groupBy = "gene_name",
            filterable = TRUE, sortable = TRUE,
            columns = list2(
              gene_name = colDef(
                name = "Gene", maxWidth = fs*11,
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)",
                  fontSize = fs
                )
              ),
              gene_id = colDef(
                name = "ID", aggregate = "unique", cell = function(value) "",
                show = FALSE
              ),
              logCPM = colDef(
                maxWidth = fs*6, aggregate = range_js,
                format = colFormat(digits = 2),
                style = function(value) bar_style(
                  width = value / max(.$logCPM), fontSize = fs
                ),
                filterMethod = js_greater
              ),
              logFC = colDef(
                aggregate = "mean", 
                style = JS(
                  glue(
                    "function(rowInfo) {
                      var value = rowInfo.row['logFC']
                      if (value < 0) {
                        var color = '{{direction_colours[['Decreased']]}}'
                      } else if (value > 0) {
                        var color = '{{direction_colours[['Increased']]}}'
                      } else {
                        var color = '#000000'
                      }
                      return { color: color, fontSize: {{fs}} }
                    }",
                    .open = "{{", .close = "}}"
                  )
                ),
                format = colFormat(digits = 2), maxWidth = fs*6,
              ),
              FDR = colDef(
                aggregate = "min", maxWidth = fs*6,
                format = colFormat(digits = 4),
                cell = \(value) sprint_pval(value),
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)", 
                  fontSize = fs
                ),
                filterMethod = js_less
              ),
              status.x = colDef("Status", aggregate = "unique", show = FALSE),
              "{short_names[[1]]}" := colDef(
                filterInput = \(values, name) drop_filter(values, name, id),
                aggregate = "unique"
              ),
              "{short_names[[2]]}" := colDef(
                filterInput = \(values, name) drop_filter(values, name, id),
                aggregate = "unique"
              ),
              range = colDef(
                name = "Location", minWidth = 160,
                aggregate = granges_js
              ),
              region = colDef(
                name = "Genomic Region", aggregate = "unique",
                filterInput = \(values, name) drop_filter(values, name, id)
              ),
              feature = colDef(
                name = "Feature", aggregate = "unique",
                filterInput = \(values, name) drop_filter(unlist(values), name, id)
              ),
              d = colDef(
                "Distance To Locus", maxWidth = fs * 8, format = comma_col,
                aggregate = range_js, filterMethod = js_less
              )
            )[colnames(.)],
            columnGroups = list(
              colGroup("RNA", c("logFC", "logCPM", "FDR")),
              colGroup("Pairwise Status", unname(unlist(short_names))),
              colGroup(
                "Signal Location", 
                intersect(c("range", "region", "feature", "d"), colnames(.))
              )
              # colGroup(target, str_subset(names(.), "\\.y$"))
            ),
            theme = reactableTheme(style = list(fontSize = fs)),
            elementId = id
          )
        cp <- htmltools::em(
          glue(
            "Comparison of all DE genes with any pairwise status ",
            "considered as showing evidence of changed signal. "
          )
        )
        htmltools::div(
          id = id, class = "section level5", 
          htmltools::h5("DEGs with Differential Signal"), 
          tagList(
            tags$button(
              tagList(fontawesome::fa("download"), "Download as CSV"),
              onclick = glue(
                "Reactable.downloadDataCSV('{id}', '{i}_deg_{params$full_comp}_dsa.csv')"
              )
            ),
            tags$button(
              "Expand/Collapse All",
              onclick = glue("Reactable.toggleAllRowsExpanded('{id}')")
            ),
            tags$button(
              "Clear All Filters",
              onclick = glue("Reactable.setAllFilters('{id}', [])")
            ),
            tbl, tags$caption(cp)
          )
        )
      }
    ) %>% 
    setNames(names(which(has_de)))
}
```

### Direct Comparisons


```{r print-tables-and-figures, echo=FALSE}
tagList(
  names(rna) %>% 
    lapply(
      \(x) {
        tagList(
          list(
            htmltools::div(
              id = x, class = "section level4 tabset",
              htmltools::h4(x, class = "tabset"),
              summary_tables[[x]],
              volcano_plots[[x]],
              pval_hist[[x]],
              pw_deg_tables[[x]]
            )
          )
        )
      }
    )
)
```

### GSEA Using Pairwise Patterns

GSEA analysis [@SubramanianGsea2005] was first performed by taking the genes mapped to each set of pairwise loci, and checking for enrichment amongst the RNA-Seq results.
Two analyses were performed:

1. Ranking transcriptomic data by significance only, to test for any overall association with changed signal levels and DGE, then
2. Ranking transcriptomic data by both *direction of change* and *significance*, to test for any overall association between the direction of change in either -omics layer

The `R` package `fgsea` [@KorotkevichFgsea2019] was used for all downstream GSEA analysis.

#### Overall Significance {.tabset}

```{r run-gsea-sig}
ids_by_status <- results_tbl %>% 
  dplyr::select(range, status, gene_id, any_of(c("region", "feature"))) %>% 
  dplyr::filter(grepl("(In|De)creased", status)) %>% 
  unnest(any_of("feature")) %>% 
  distinct(!!!syms(colnames(.))) %>% 
  pivot_longer(cols = any_of(c("region", "feature"))) %>% 
  unite(group, all_of(c("status", "value"))) %>% 
  split(.$group) %>% 
  lapply(pull, "gene_id") %>% 
  lapply(unique)
rna_sig_scores <- lapply(rna, \(x) setNames(-log10(x$PValue), x$gene_id))
rna_gsea_sig <- rna_sig_scores %>%
  lapply(
    \(x) fgsea(
      ids_by_status, x, minSize = min(msigdb_params$size), scoreType = "pos"
    )
  ) %>% 
  lapply(as_tibble) %>% 
  lapply(
    separate, pathway, into = c("status", "region"),
    sep = "_", extra = "merge"
  ) %>%
  lapply(
    separate, status, into = unlist(short_names), sep = " - "
  ) %>% 
  lapply(arrange, pval) %>% 
  lapply(
    mutate,
    padj = p.adjust(pval, enrich_params$adj),
    region = str_remove_all(region, ".+: "), 
    leSize = map_int(leadingEdge, length),
    leadingEdge = map_chr(
      leadingEdge, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
    )
  ) %>% 
  lapply(left_join, region_sources, by = "region") %>% 
  lapply(
    dplyr::select, 
    all_of(unname(unlist(short_names))), source, region,
    padj, NES, ends_with("ize"), leadingEdge
  ) %>% 
  lapply(dplyr::filter, leSize >= enrich_params$min_sig)
any_gsea_sig <- map_lgl(rna_gsea_sig, \(x) any(x$padj < enrich_params$alpha))
```


```{r tbl-gsea-sig, echo=FALSE}
if (any(any_gsea_sig)) {
  tagList(
    any_gsea_sig %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(i) {
          tbl_id <- paste0("tbl-gsea-sig", i) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-")
          tbl <- rna_gsea_sig[[i]] %>% 
            dplyr::filter(padj < enrich_params$alpha) %>% 
            reactable(
              sortable = TRUE, filterable = TRUE, 
              showPageSizeOptions = nrow(.) > 10,
              columns = list2(
                "{short_names$comp1}" := colDef(
                  filterInput = \(values, name) drop_filter(values, name, tbl_id),
                  maxWidth = 120
                ),
                "{short_names$comp2}" := colDef(
                  filterInput = \(values, name) drop_filter(values, name, tbl_id),
                  maxWidth = 120
                ),
                source = colDef(
                  "Source", maxWidth = 150,
                  filterInput = \(values, name) drop_filter(values, name, tbl_id)
                ),
                region = colDef(
                  "Region Type", maxWidth = 120,
                  filterInput = \(values, name) drop_filter(values, name, tbl_id)
                ),
                padj = enrich_cols$adj_p,
                NES = enrich_cols$NES,
                size = colDef("Group Size"),
                leSize = colDef("LE Size"),
                leadingEdge = colDef(
                  "Leading Edge", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150, maxWidth = 500
                )
              ),
              defaultColDef = colDef(maxWidth = 75),
              theme = reactableTheme(style = list(fontSize = 12)),
              elementId = tbl_id
            )
          cp <- htmltools::em(
            glue(
              "GSEA results for {i}, testing an association any changed genes ", 
              "for all pairwise groups incorporating all available ", 
              "annotations and features. Only pairwise groupings where changed ", 
              "signal was observed in at least one comparison are shown. ", 
              "A Normalized Enrichment Score (NES) > 0 ", 
              "indicates an association with differentially expressed genes. ", 
              "The total number of genes mapped to each type of group is ", 
              "provided along with those found beyond the Leading Edge. ", 
              "Given that multiple peaks may map to a gene, ",
              "genes may be included in more than one grouping."
            )
          )
          tl <- tagList(
            tags$button(
              tagList(fontawesome::fa("download"), "Download as CSV"),
              onclick = glue(
                "Reactable.downloadDataCSV('{tbl_id}',",
                "'{params$full_comp}-{i}-gsea-sig.csv')"
              )
            ),
            tags$button(
              "Clear All Filters", 
              onclick = glue("Reactable.setAllFilters('{tbl_id}', [])")
            ),
            tbl, htmltools::tags$caption(cp)
          )
          htmltools::div(
            id = tbl_id,
            class = "section level5",
            htmltools::h5(i),
            htmltools::div(class = "table", tl)
          )
        }
      )
  )
}
```


#### Directional Association {.tabset}

```{r run-gsea-dir}
rna_dir_scores <- rna %>% 
  lapply(
    \(x) setNames(-sign(x$logFC) * log10(x$PValue), x$gene_id)
  )
rna_gsea_dir <- rna_dir_scores %>% 
  lapply(\(x) fgsea(ids_by_status, x, minSize = min(msigdb_params$size))) %>% 
  lapply(as_tibble) %>% 
  lapply(
    separate, pathway, into = c("status", "region"),
    sep = "_", extra = "merge"
  ) %>%
  lapply(
    separate, status, into = unlist(short_names), sep = " - "
  ) %>% 
  lapply(arrange, pval) %>% 
  mclapply(
    mutate,
    padj = p.adjust(pval, enrich_params$adj),
    region = str_remove_all(region, ".+: "), 
    leSize = map_int(leadingEdge, length),
    leadingEdge = map_chr(
      leadingEdge, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
    ),
    mc.cores = threads
  ) %>% 
  mclapply(left_join, region_sources, by = "region", mc.cores = threads) %>% 
  mclapply(
    dplyr::select, 
    all_of(unname(unlist(short_names))), source, region,
    padj, NES, ends_with("ize"), leadingEdge,
    mc.cores = threads
  ) %>% 
  mclapply(dplyr::filter, leSize >= enrich_params$min_sig, mc.cores = threads)
any_gsea_dir <- map_lgl(rna_gsea_dir, \(x) any(x$padj < enrich_params$alpha))
```

```{r tbl-gsea-dir, echo=FALSE}
if (any(any_gsea_dir)) {
  tagList(
    any_gsea_dir %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(i) {
          tbl_id <- paste0("tbl-gsea-dir", i) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-")
          tbl <- rna_gsea_dir[[i]] %>% 
            dplyr::filter(padj < enrich_params$alpha) %>% 
            reactable(
              sortable = TRUE, filterable = TRUE, 
              showPageSizeOptions = nrow(.) > 10,
              columns = list2(
                "{short_names$comp1}" := colDef(
                  filterInput = \(values, name) drop_filter(values, name, tbl_id),
                  maxWidth = 120
                ),
                "{short_names$comp2}" := colDef(
                  filterInput = \(values, name) drop_filter(values, name, tbl_id),
                  maxWidth = 120
                ),
                source = colDef(
                  "Source", maxWidth = 150,
                  filterInput = \(values, name) drop_filter(values, name, tbl_id)
                ),
                region = colDef(
                  "Region Type", maxWidth = 120,
                  filterInput = \(values, name) drop_filter(values, name, tbl_id)
                ),
                padj = enrich_cols$adj_p,
                NES = enrich_cols$NES,
                size = colDef("Group Size"),
                leSize = colDef("LE Size"),
                leadingEdge = colDef(
                  "Leading Edge", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150, maxWidth = 500
                )
              ),
              defaultColDef = colDef(maxWidth = 75),
              theme = reactableTheme(style = list(fontSize = 12)),
              elementId = tbl_id
            )
          cp <- htmltools::em(
            glue(
              "GSEA results for {i}, testing an association between either ", 
              "up- or down-regulated genes for all pairwise groups ", 
              "incorporating all available annotations and features. ", 
              "A Normalized Enrichment Score (NES) > 0 ", 
              "indicates an association with up-regulated genes, with a negative score ",
              "being associated with down-regulation. The total number of genes ", 
              "mapped to each grouping is provided along with those found ",
              "beyond the Leading Edge. Given that multiple peaks may map to a gene, ",
              "genes may be included in more than one grouping."
            )
          )
          tl <- tagList(
            tags$button(
              tagList(fontawesome::fa("download"), "Download as CSV"),
              onclick = glue(
                "Reactable.downloadDataCSV('{tbl_id}',",
                "'{params$full_comp}-{i}-gsea-dir.csv')"
              )
            ),
            tags$button(
              "Clear All Filters", 
              onclick = glue("Reactable.setAllFilters('{tbl_id}', [])")
            ),
            tbl, htmltools::tags$caption(cp)
          )
          htmltools::div(
            id = tbl_id,
            class = "section level5",
            htmltools::h5(i),
            htmltools::div(class = "table", tl)
          )
        }
      )
  )
}
```


### Combined Enrichment {.tabset}

```{r load-gsea-rds}
gsea_dir <- file.path(all_paths$annotation, "gsea_dir.rds") %>% read_rds()
gsea_sig <- file.path(all_paths$annotation, "gsea_sig.rds") %>% read_rds()
```


```{r changed-rna-enrichment}
changed_rna_enrich <- gsea_sig %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(
          bind_rows(enrich_by_status, .id = "comparison"),
          by = c("pathway" = "gs_name")
        ) %>% 
        dplyr::select(comparison, pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(comparison, pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  ) %>% 
  .[map_int(., nrow) > 0]
tbl_changed_rna_enrich <- names(changed_rna_enrich) %>% 
  lapply(
    \(x) {
      id <- x %>% str_to_lower() %>% str_replace_all("_", "-")
      id <- paste0(id, "-changed-enrich-tbl")
      df <- changed_rna_enrich[[x]]
      tbl <- reactable(
        df, 
        searchable = TRUE, filterable = TRUE, 
        showPageSizeOptions = nrow(df) > 10,
        columns = c(
          list(
            comparison = colDef(
              "Pairwise Grouping", minWidth = 150,
              filterInput = \(values, name) drop_filter(values, name, id)
            ),
            pathway = c(
              list(filterInput = \(values, name) drop_filter(values, name, id)),
              enrich_cols$pathway
            ) %>% set_class("colDef")
          ),
          enrich_cols[c("NES", "meta_p", "adj_p", "n_cmn", "genes", "cmn_ids")]
        ),
        elementId = id
      )
      cp <- tags$em(
        glue(
          "Results when combining directional GSEA results from {x} ",
          "with pairwise enrichment results across comparisons."
        )
      )
      htmltools::div(
        id = id, class = "section level4", h4(x),
        tbl, tags$caption(cp)
      )
    }
  ) %>% 
  setNames(names(changed_rna_enrich))
# network_changed_rna_enrich <- changed_rna_enrich %>% 
#   map_int(nrow) %>% 
#   is_weakly_greater_than(min(network_params$size)) %>% 
#   which() %>% 
#   names() %>%  
#   sapply(
#     \(x) {
#       p <- changed_rna_enrich[[x]] %>% 
#         mutate(
#           gene_set_size = msigdb %>% 
#             dplyr::filter(gs_name %in% pathway) %>% 
#             split(.$gs_name) %>% 
#             .[pathway] %>% 
#             map_int(nrow),
#           prop = n_cmn / gene_set_size,
#           cmn_ids = setNames(cmn_ids, pathway)
#         ) %>% 
#         make_tbl_graph(
#           gs = .$cmn_ids,
#           alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
#           hit_col = "n_cmn", size_col = "gene_set_size",
#           min_dist = network_params$max_distance,
#           max_gs = max(network_params$size)
#         ) %>% 
#         ggraph(layout = net_layout, weights = oc) +
#         geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
#         geom_node_point(
#           aes(fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop),
#           shape = 21
#         ) +
#         geom_node_text(
#           aes(label = label, alpha = prop),
#           colour = "black", size = 3, 
#           data = . %>%
#             mutate(
#               label = str_replace_all(label, "_", " ") %>% 
#                 str_trunc(60) %>% 
#                 str_wrap(width = 18)
#             ),
#           repel = TRUE, bg.color = "white", bg.r = 0.1, 
#         ) +
#         scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
#         scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
#         scale_fill_viridis_c(option = "inferno", begin = 0.2) +
#         scale_alpha(range = c(0.5, 1), trans = "sqrt") +
#         scale_size_continuous(range = c(1, 10)) +
#         scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
#         scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
#         guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
#         labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
#         theme_void() 
#       
#       ## Export the image
#       img_out <- file.path(
#         fig_path, glue("network_changed_rna_enrich_{x}.{fig_type}")
#       )
#       fig_fun(filename = img_out, width = fw, height = fh)
#       print(p)
#       dev.off()
#       
#       ## Create html tags
#       fig_link <- str_extract(img_out, "assets.+")
#       cp <- htmltools::tags$em(
#         glue(
#           "The relationship between gene-sets considered enriched when ",
#           "including GSEA results from {x} and changed sites from {target} ",
#           "DSA analysis."
#         )
#       )
#       htmltools::div(
#         id = img_out %>% 
#           basename() %>% 
#           str_remove_all(glue(".{fig_type}$")) %>% 
#           str_to_lower() %>% 
#           str_replace_all("_", "-"),
#         class = "section level5",
#         htmltools::h5("Enrichment Network: DSA Changed"),
#         htmltools::div(
#           class = "figure", style = "text-align: center",
#           htmltools::img(src = fig_link, width = "100%"),
#           htmltools::tags$caption(cp)
#         )
#       )
#       
#     }, simplify = FALSE
#   )  
```


Enrichment analysis was performed combined transcriptomic data with the enrichment analysis from pairwise groupings.
A simple strategy was used taking Wilkinson's maximum p-value method, which returns the probability of returning two p-values below the value of the largest p-value.
Transcriptomic enrichment was taken as the results from GSEA analysis, whilst pairwise enrichment results were those obtained above.
Genes within the Leading Edge for GSEA and which were also mapped to the same pathway within DSA enrichment are returned as the common genes.



```{r print-dsa-rna-enrichment, echo=FALSE}
tagList(
  tbl_changed_rna_enrich %>% 
    map_lgl(is, "shiny.tag") %>% 
    which() %>% 
    names() %>% 
    unique() %>% 
    lapply(
      \(x) {
        tagList(
          list(
            htmltools::div(tbl_changed_rna_enrich[[x]])
          )
        )
      }
    )
)
```



### Highly Ranked Genes {.tabset}


```{r rna-grl-to-plot}
grl_to_plot <- rna %>%
  lapply(inner_join, results_tbl, by = 'gene_id') %>% 
  bind_rows(.id = "dataset") %>% 
  arrange(PValue) %>% 
  dplyr::filter(grepl("(In|De)creased", status.y)) %>% 
  dplyr::slice(1:4, .by = status.y) %>% 
  dplyr::select(
    range, starts_with("gene"), starts_with("status"),
    any_of(c("region", "feature"))
  ) %>% 
  distinct(gene_id, .keep_all = TRUE) %>% 
  dplyr::slice(1, .by = status.y) %>% 
  colToRanges("range", seqinfo = sq) %>% 
  splitAsList(.$gene_id)
```

The most highly ranked gene for differential expression was chosen within each pairwise grouping, and in the case of a gene being allocated to multiple groups, the next-most significant was chosen.
This set of `r length(grl_to_plot)` loci is shown below, noting that for some genes, the peaks may be beyond the coding region of the gene.

`r if (length(rna) > 1) "When assigning genes to Increased, Decreased or Unchanged for the purposes of visualisation, the status associated with the lowest p-value was chosen across datasets. If transcriptomic datasets are strongly discordant, these classifications may be less informative than expected"`

```{r plot-all-top-rna-ranges}
.makeCaption <- function(.gr) {
  if (is.null(.gr)) return(NULL)
  id <- unique(.gr$gene_id)
  gn <- .gr$gene_name
  reg <- glue_collapse(.gr$region, sep = ", ", last = " & ")
  reg <- ifelse(
    grepl("^[AEIOUaeiou]", reg), paste("an", reg), paste("a", reg)
  )
  reg <- ifelse(grepl("on$", reg), paste0(reg, "ic"), reg)
  feat <- c()
  if (has_features) {
    feat <- glue_collapse(unique(unlist(.gr$feature)), sep = ", ", last = " & ")
    feat <- ifelse(
      grepl("^[AEIOUaeiou]", feat), paste("an", feat), paste("a", feat)
    )
    feat <- glue(
      "When comparing to external features, this locus was most associated with those designated as {feat}. "
    )
  }
  glue(
    "This {width(range(.gr))}bp region ({as.character(.gr)}) was considered ",
    "as {pw_relabel(.gr$status.y)}, overlapping {reg} region, and with the ", 
    "associated gene, {gn}, being considered as {.gr$status.x}. ",
    feat, 
    "Loci associated with differential signal in either dataset are annotated by colour."
  )
}
htmltools::tagList(
  lapply(
    seq_along(grl_to_plot),
    \(i) {
      ## Export the image
      nm <- str_to_lower(names(grl_to_plot)[[i]])
      id <- unique(grl_to_plot[[i]]$gene_id)
      img_out <- file.path(fig_path, glue("{nm}.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      plotHFGC(
        grl_to_plot[[i]],
        hic = hic,
        features = feat_gr, featcol = hfgc_feat_col, featname = featnames,
        featsize = max(1, 0.8 * length(feat_gr)), featstack = "dense",
        genes = hfgc_genes, genecol = gene_col, 
        genesize = rep_len(c(2, 1), length(hfgc_genes)),
        coverage = bwfl, linecol = line_col, ylim = y_lim,
        annotation = annot, annotcol = direction_colours,
        cytobands = bands_df,
        collapseTranscripts = collapseTrans, maxTrans = 15,
        zoom = 30, max = max(1e4, 1.1 * width(range(grl_to_plot[[i]]))), 
        col.title = "black", background.title = "white", 
        rotation.title = 90, showAxis = FALSE, fontsize = 10,
        highlight = rgb(0, 0, 1, 0.3)
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(.makeCaption(grl_to_plot[[i]]))
      htmltools::div(
        htmltools::div(
          id = nm,
          class = "section level4",
          htmltools::h4(id2gene[id]),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  )
)
```

