# Transcriptomic Datasets

```{r fig-params, results='hide'}
fig_params <- here::here("config", "rmarkdown.yml") %>% 
  read_yaml() %>% 
  pluck("knitr_opts")
fw <- fig_params$fig.width
fh <- fig_params$fig.height
fig_type <- fig_params$dev[[1]]
if (is.null(fig_type)) stop("Couldn't detect figure type")
fig_fun <- match.fun(fig_type)
if (fig_type %in% c("bmp", "jpeg", "png", "tiff")) {
  ## These figure types require dpi & resetting to be inches
  formals(fig_fun)$units <- "in"
  formals(fig_fun)$res <- 300
}
fig_path <- here::here("docs", "assets", names(rna))
fig_path %>% 
  lapply(
    \(x) {
      if (!dir.exists(x)) dir.create(x, recursive = TRUE)
    }
  )
```


```{r load-gsea, echo=FALSE}
library(fgsea)
gs_url <- setNames(msigdb$gs_url, msigdb$gs_name) %>% 
  .[!duplicated(names(.))]
ids_by_gs <- msigdb %>% 
  split(.$gs_name) %>% 
  lapply(pull, "ensembl_gene")
gsea_dir <- read_rds(all_out$gsea_dir)
gsea_sig <- read_rds(all_out$gsea_sig)
dir_ranks <- rna %>% 
  lapply(\(x) sort(setNames(-sign(x$logFC) * log10(x$PValue), x$gene_id)))
sig_ranks <- rna %>% 
  lapply(\(x) sort(setNames(-log10(x$PValue), x$gene_id)))
gsea_dir <- gsea_dir %>% 
  lapply(dplyr::filter, padj < enrich_params$alpha) %>% 
  lapply(arrange, pval) %>% 
  lapply(
    mutate, 
    leSize = map_int(leadingEdge, length),
    leadingEdge = map_chr(
      leadingEdge, 
      \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
    )
  ) %>% 
  lapply(dplyr::filter, leSize > enrich_params$min_sig) %>% 
  .[map_int(., nrow) > 0]
gsea_sig <- gsea_sig %>% 
  lapply(dplyr::filter, padj < enrich_params$alpha & NES > 0) %>% 
  lapply(arrange, pval) %>% 
  lapply(
    mutate, 
    leSize = map_int(leadingEdge, length),
    leadingEdge = map_chr(
      leadingEdge, 
      \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
    )
  ) %>% 
  lapply(dplyr::filter, leSize > enrich_params$min_sig) %>% 
  .[map_int(., nrow) > 0]
```

`r length(rna)` transcriptomic `r ifelse(length(rna) > 1, "datasets were", "dataset was")` provided, as is summarised in this section.
In the section below, a differentially expressed (DE) gene is simply considered to be one with an FDR-adjusted p-value < 0.05, noting that no differentially expressed genes are formally required for data integration.

In addition to checking the input file structures, Gene Set Enrichment Analysis (GSEA) [@SubramanianGsea2005] was performed using two approaches.
The first analysis takes a ranked list with the direction of change being included, testing for any association between pathways and up- or down-regulated genes.
The second analysis simply ranks genes by significance, performing a one-sided test for enrichment of pathways within the set of genes the most highly-ranked for statistical evidence of changed expression.
For both analyses, the `R` package `fgsea` was used [@KorotkevichFgsea2019].

## Dataset Summary

```{r show-rna}
rna %>% 
  bind_rows(.id = "Dataset") %>% 
  summarise(
    Detected = dplyr::n(), 
    `Up-Regulated` = sum(logFC > 0 & FDR < 0.05),
    `Down-Regulated` = sum(logFC < 0 & FDR < 0.05),
    `Total DE` = `Up-Regulated` + `Down-Regulated`,
    .by = Dataset
  ) %>% 
  pander(
    justify = "lrrrr",
    caption = "Summary of all provided transcriptomic datasets"
  )
```

## Directional GSEA 

### Results {.tabset}

```{r table-gsea-dir, eval = length(gsea_dir) > 0, echo = FALSE}
htmltools::tagList(
  names(gsea_dir) %>% 
    lapply(
      \(x) {
        id <- paste0('gsea-dir-', str_replace_all(str_to_lower(x), "_", "-"))
        tbl <- gsea_dir[[x]] %>%
          dplyr::select(pathway, padj, NES, contains("size"), leadingEdge) %>% 
          reactable(
            showPageSizeOptions = TRUE, filterable = TRUE,
            columns = list(
              pathway = colDef(
                "Pathway", minWidth = 150,
                cell = function(value) htmltools::tags$a(
                  href = gs_url[[value]],
                  target = "_blank",
                  str_replace_all(value, "_", " ")
                ),
                html = TRUE
              ),
              padj = colDef(
                "P<sub>adj</sub>", html = TRUE,
                cell = \(value) sprint_pval(value), maxWidth = 80,
                filterMethod = js_less
              ),
              NES = colDef(
                format = colFormat(digits = 2), maxWidth = 100,
                filterMethod = js_greater
              ),
              size = colDef("Pathway Size", maxWidth = 80, filterMethod = js_greater),
              leSize = colDef("Leading Edge Size", maxWidth = 100, filterMethod = js_greater),
              leadingEdge = colDef(
                "Leading Edge", minWidth = 150,
                cell = \(value) with_tooltip(value, width = 60)
              )
            ),
            elementId = id
          )
        cp = htmltools::tags$em(
          paste(
            "Taking genes ranked by significance and incorporating direction,",
            "the following gene sets were returned as significant using GSEA.",
            "A normalised enrichment score (NES) greater than zero indicates",
            "the pathway was enriched amongst the upregulated genes, whilst for",
            "a negative NES, enrichment will be amongst the down-regulated genes."
          )
        )
        htmltools::div(
          htmltools::div(
            id = id,
            class = "section level4",
            htmltools::h4(x),
            htmltools::div(
              class = "table", tbl, htmltools::tags$caption(cp)
            )
          )
        )
      }
    )
)
```

### Barcode Plots {.tabset}

```{r barcode-gsea-dir}
htmltools::tagList(
  names(gsea_dir) %>% 
    lapply(
      \(x) {
        p <- gsea_dir[[x]] %>% 
          dplyr::slice(1:9) %>% 
          pull("pathway") %>% 
          lapply(
            \(i) {
              plotEnrichment(ids_by_gs[[i]], dir_ranks[[x]]) +
                ggtitle(str_replace_all(i, "_", " ") %>% str_wrap(20))
            }
          ) %>% 
          wrap_plots(
            ncol = case_when(
              length(.) == 1 ~ 1,
              length(.) %in% c(2, 4) ~ 2,
              TRUE ~ 3
            )
          )
        img_out <- here::here(
          "docs", "assets", x, "barcode_gsea_dir.{fig_type}"
        ) %>% 
          glue()
        fig_fun(
          filename = img_out, 
          width = ifelse(length(p) > 1, fw, fw / 2),
          height = ifelse(length(p) > 3, fh, fh / 2)
        )
        print(p)
        dev.off()
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "Barcode plots for the most highly ranked, significantly enriched ",
            "pathways in the transcriptomic dataset {x} when ",
            "taking direction of estimated logFC into account"
          )
        )
        htmltools::div(
          htmltools::div(
            id = x %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-") %>% 
              paste("barcode-gsea-dir", sep = "-"),
            class = "section level4",
            htmltools::h4(x),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      }
    ) 
)
```


## Non-Directional GSEA 

### Results {.tabset}

```{r table-gsea-sig, eval = length(gsea_sig) > 0, echo = FALSE}
htmltools::tagList(
  names(gsea_sig) %>% 
    lapply(
      \(x) {
        id <- paste0('gsea-sig-', str_replace_all(str_to_lower(x), "_", "-"))
        tbl <- gsea_sig[[x]] %>%
          dplyr::select(pathway, padj, NES, contains("size"), leadingEdge) %>% 
          reactable(
            showPageSizeOptions = TRUE, filterable = TRUE,
            columns = list(
              pathway = colDef(
                "Pathway", minWidth = 150,
                cell = function(value) htmltools::tags$a(
                  href = gs_url[[value]],
                  target = "_blank",
                  str_replace_all(value, "_", " ")
                ),
                html = TRUE
              ),
              padj = colDef(
                "P<sub>adj</sub>", html = TRUE,
                cell = \(value) sprint_pval(value), maxWidth = 80,
                filterMethod = js_less
              ),
              NES = colDef(
                format = colFormat(digits = 2), maxWidth = 100,
                filterMethod = js_greater
              ),
              size = colDef("Pathway Size", maxWidth = 80, filterMethod = js_greater),
              leSize = colDef("Leading Edge Size", maxWidth = 100, filterMethod = js_greater),
              leadingEdge = colDef(
                "Leading Edge", minWidth = 150,
                cell = \(value) with_tooltip(value, width = 60)
              )
            ),
            elementId = id
          )
        cp = htmltools::tags$em(
          paste(
            "Taking genes ranked only by significance and without incorporating direction,",
            "the following gene sets were returned as significant using GSEA.",
            "This siply tests whether pathways are enriched within genes more",
            "likely to be differentially expressed, with no preference for up.",
            "or down-regulation."
          )
        )
        htmltools::div(
          htmltools::div(
            id = id,
            class = "section level4",
            htmltools::h4(x),
            htmltools::div(
              class = "table", tbl, htmltools::tags$caption(cp)
            )
          )
        )
      }
    )
)
```

### Barcode Plots {.tabset}


```{r barcode-gsea-sig}
htmltools::tagList(
  names(gsea_sig) %>% 
    lapply(
      \(x) {
        p <- gsea_sig[[x]] %>% 
          dplyr::slice(1:9) %>% 
          pull("pathway") %>% 
          lapply(
            \(i) {
              plotEnrichment(
                ids_by_gs[[i]], dir_ranks[[x]]
              ) +
                ggtitle(str_replace_all(i, "_", " ") %>% str_wrap(20))
            }
          ) %>% 
          wrap_plots(
            ncol = case_when(
              length(.) == 1 ~ 1,
              length(.) %in% c(2, 4) ~ 2,
              TRUE ~ 3
            )
          )
        img_out <- here::here(
          "docs", "assets", x, "barcode_gsea_sig.{fig_type}"
        ) %>% 
          glue()
        fig_fun(
          filename = img_out, 
          width = ifelse(length(p) > 2, fw, fw * length(p) / 3),
          height = ifelse(length(p) > 3, fh, fh / 3)
        )
        print(p)
        dev.off()
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "Barcode plots for the most highly ranked, significantly enriched ",
            "pathways in the transcriptomic dataset {x} when ",
            "only taking significance of estimated logFC into account"
          )
        )
        htmltools::div(
          htmltools::div(
            id = x %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-") %>% 
              paste("barcode-gsea-dir", sep = "-"),
            class = "section level4",
            htmltools::h4(x),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      }
    ) 
)
```

