
```{r ihw-all}
ihw_all <- external_features %>% 
  split(f = .$feature) %>% 
  c(
    list(
      no_feature = setdiff(GRanges(sq), external_features)
    )
  ) %>% 
  lapply(
    function(x) propOverlap(merged_results, x)
  ) %>% 
  as_tibble() %>% 
  mutate(
    range = as.character(merged_results), id = seq_along(range)
  ) %>% 
  pivot_longer(
    cols = c(all_of(unique(external_features$feature)), "no_feature"),
    names_to = "feature",
    values_to = "prop"
  ) %>% 
  group_by(id, range) %>% 
  dplyr::filter(prop == max(prop)) %>% 
  ungroup() %>% 
  distinct(id, range, .keep_all = TRUE) %>% 
  mutate(
    feature = factor(feature, levels = names(colours$features))
  ) %>% 
  arrange(id) %>% 
  .[["feature"]]
n_levels <- sum(table(ihw_all) > 1e3) - 1
merged_results$ihw_covariate <- fct_lump_n(ihw_all, n = n_levels)
ihw_proceed <- length(levels(merged_results$ihw_covariate)) > 1
if (ihw_proceed)  fdr_column <- "fdr_ihw"
ihw_cols <- colours$features
if (ihw_proceed) {
  ihw <- ihw(
    pvalues = merged_results$P.Value,
    covariates = merged_results$ihw_covariate,
    alpha <- fdr_alpha,
    covariate_type = "nominal"
  )
  merged_results <- mutate(merged_results, fdr_ihw = adj_pvalues(ihw))
} 
merged_results <- merged_results %>% 
  mutate(
    status = case_when(
      !!sym(fdr_column) < fdr_alpha ~ as.character(direction),
      !!sym(fdr_column) >= fdr_alpha ~ "Unchanged"
    ) %>% 
      factor(levels = c(levels(direction), "Unchanged")) %>% 
      droplevels()
  )
```

Independent Hypothesis Weighting (IHW) [@IgnatiadisIHW2016] was then used to partition the raw p-values for `r target` differential binding, by features provided in `r basename(config$external$features)`, and estimate weights to apply to each p-value as a divisor.
This allows recalculation of the FDR using *weighted p-values* instead of *raw p-values*.
In order for IHW to be a viable strategy, partitions should be greater than 1000.
The provided features were used as initial partitions, merging the smallest groups below this size until all partitions were of a suitable size.

Windows were classified as overlapping a feature if any section of the window overlapped the feature.
Where multiple overlapping features were found, the feature with the *largest* overlap was selected.

`r if (!ihw_proceed) "Only one partition > 1000 was able to be formed and IHW was not executed."`

`r if (ihw_proceed) "### Summary Table"`

```{r ihw-tab, eval = ihw_proceed, echo = ihw_proceed}
ihw %>%
  as.data.frame() %>% 
  mutate(
    fdr = p.adjust(pvalue, "fdr"),
    status = case_when(
      fdr < fdr_alpha & adj_pvalue > fdr_alpha ~ "Lost",
      fdr < fdr_alpha & adj_pvalue < fdr_alpha ~ "Retained",
      fdr > fdr_alpha & adj_pvalue < fdr_alpha ~ "Gained",
      fdr > fdr_alpha & adj_pvalue > fdr_alpha ~ "Never Significant"
    ) %>% 
      factor(
        levels = c("Never Significant", "Retained", "Gained", "Lost")
      )
  ) %>% 
  as_tibble() %>% 
  group_by(covariate, status) %>% 
  tally() %>% 
  ungroup() %>% 
  complete(status = status, covariate = covariate, fill = list(n = 0)) %>% 
  pivot_wider(
    names_from = status, values_from = n, values_fill = 0
  ) %>% 
  mutate(
    `Nett Change` = Gained - Lost
  ) %>% 
  bind_rows(
    summarise_if(., is.numeric, sum)
  ) %>% 
  mutate(
    covariate = covariate %>% 
      as.character() %>% 
      str_sep_to_title() %>% 
      str_replace_na("Total"),
    Significant = Retained + Gained
  ) %>% 
  dplyr::select(
    Covariate = covariate, ends_with("Significant"), Retained, everything()
  ) %>% 
  pander(
    justify = "lrrrrrr",
    caption = paste(
      "*Summary of changes introduced by IHW for windows considered as",
      paste0("differentially bound by ", target, "."),
      "This corresponds to a nett change of", 
        dplyr::slice(., nrow(.)) %>% 
        mutate(p = `Nett Change` / (Significant - `Nett Change`)) %>% 
        .[["p"]] %>% 
        percent(0.1),
      "from the initial list.*"
    ),
    emphasize.strong.rows = nrow(.)
  )
```



### Initial Group Sizes

```{r ihw-partitions-raw, fig.height=6, fig.cap = "*Breakdown of all windows which overlapped the features provided. Any window with no defined feature overlap was simply assigned as 'No Feature'. Any partitions with fewer than 1000 windows (indicated as the blue horizontal line) were combined into the next smallest partition consecutively, until all partitions contained > 1000 windows.*"}
tibble(ihw = ihw_all) %>% 
  mutate(
    ihw_label = fct_relabel(ihw, str_sep_to_title) %>% 
      fct_infreq()
  ) %>% 
  ggplot(
    aes(ihw_label, fill = ihw)
  ) +
  geom_bar() +
  geom_hline(yintercept = 1e3, colour = "blue") +
  geom_text(
    aes(y = n + 0.025*max(n), label = comma(n, 1)),
    data = . %>%
      group_by(ihw_label, ihw) %>%
      summarise(n = dplyr::n(), .groups = "drop") 
  ) +
  scale_y_continuous(expand = expansion(c(0, 0.05)), labels = comma) +
  scale_fill_manual(values = ihw_cols) +
  labs(
    x = str_to_title(ihw_method), y = "Total"
  ) +
  theme(
    legend.position = "none"
  )
```

### P-Value Distributions

```{r plot-ihw-pvals, fig.height= 3 + 2 * ceiling(length(levels(merged_results$ihw_covariate)) / 3), fig.cap = "*P-Value distributions within all final data partitions. The size of each partition is given within each panel.*"}
y <- ggplot_build(
  tibble(P.Value = merged_results$P.Value,) %>% 
    ggplot(aes(P.Value, stat(density))) + 
    geom_histogram(bins = 100) 
) %>% .[["data"]] %>% 
  .[[1]] %>% 
  .[["y"]] %>% 
  max()
merged_results %>% 
  select(P.Value, ihw_covariate) %>% 
  mcols() %>% 
  as_tibble() %>% 
  mutate(
    label = fct_relabel(ihw_covariate, str_sep_to_title)
  ) %>% 
  ggplot(aes(P.Value, stat(density))) +
  geom_histogram(fill = "grey", colour = "black", bins = 100) +
  geom_text(
    aes(0.8, y, label = lab),
    data = . %>% 
      group_by(label) %>% 
      summarise(n = dplyr::n(), .groups = "drop") %>% 
      mutate(
        lab = glue("N = {comma(n, 1)}"),
        y = y
      )
  ) +
  facet_wrap(~label, ncol = 3) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  labs(x = "P Value", y = "Density")
```

`r ifelse(ihw_proceed, "### IHW Weights", "")`

```{r plot-ihw-weights, eval = ihw_proceed, echo = ihw_proceed, fig.height = 5, fig.cap = "*Weights applied to p-values within each partition. 'Folds' represent random sub-partitions within each larger partition generated as part of the IHW process.*"}
plot(ihw) +
  geom_hline(yintercept = 1) +
  facet_wrap(~group, labeller = as_labeller(str_sep_to_title)) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  labs(x = "Fold", y = "Weight", fill = "Fold")
```

`r ifelse(ihw_proceed, "### FDR Comparison", "")`

```{r plot-fdr-comparison, dev = 'png', eval = ihw_proceed, echo = ihw_proceed, fig.height = 5, fig.cap = glue("*Comparison of raw and weighted p-values for each partition. Blue dashed lines indicate FDR = {fdr_alpha} for each set of p-values. Those in the lower-right quadrant would no longer be considered significant after IHW, whilst those in the upper-left quadrant would only be considered as significant after the IHW process. Those in the upper-right quadrant would be considered as significant regardless of the methodology.*")}
ihw %>%
  as.data.frame() %>% 
  ggplot(
    aes(-log10(pvalue), -log10(weighted_pvalue), colour = fold)
  ) +
  geom_point(size = 0.4) +
  geom_hline(
    aes(yintercept = -log10(weighted_pvalue)),
    data = . %>% 
      dplyr::select(-group) %>% 
      dplyr::filter(adj_pvalue < fdr_alpha) %>% 
      dplyr::filter(weighted_pvalue == max(weighted_pvalue)),
    linetype = 2, colour = "blue"
  ) +
  geom_vline(
    aes(xintercept = -log10(pvalue)),
    data =. %>% 
      dplyr::filter(p.adjust(pvalue, "fdr") < fdr_alpha) %>% 
      dplyr::filter(pvalue == max(pvalue)) %>% 
      dplyr::select(-group),
    linetype = 2, colour = "blue"
  ) +
  facet_wrap(~group, scales = "free", labeller = as_labeller(str_sep_to_title)) +
  coord_cartesian(xlim = c(0, 5), ylim = c(0, 5)) +
  labs(
    x = expression(-log[10](p)),
    y = expression(-log[10](p[IHW])),
    colour = "Fold"
  )
```

