```{r ihw-all}
other_targets <- here::here(config$samples$file) %>% 
  read_tsv() %>% 
  pull("target") %>% 
  unique() %>% 
  setdiff(target)
ihw_all <- other_targets %>% 
  sapply(
    function(x) {
      here::here(dirname(macs2_path), x, "consensus_peaks.bed") %>%
        import.bed(seqinfo = sq)
    },
    simplify = FALSE
  ) %>% 
  lapply(granges) %>% 
  lapply(
    function(x) overlapsAny(merged_results, x)
  ) %>% 
  as_tibble() %>% 
  mutate(
    range = as.character(merged_results), id = seq_along(range)
  ) %>% 
  pivot_longer(
    cols = all_of(other_targets),
    names_to = "targets", 
    values_to = "overlap"
  ) %>% 
  group_by(id, range) %>% 
  summarise(
    targets = ifelse(
      any(overlap), 
      paste(sort(targets[overlap]), collapse = "+ "),
      "No Other Targets"
    ),
    .groups = "drop"
  ) %>% 
  arrange(id) %>% 
  mutate(targets = fct_infreq(targets)) %>% 
  .[["targets"]]
n_levels <- sum(table(ihw_all) > 1e3) - 1
merged_results$ihw_covariate <- fct_lump_n(ihw_all, n = n_levels)
ihw_proceed <- length(levels(merged_results$ihw_covariate)) > 1
if (ihw_proceed)  fdr_column <- "fdr_ihw"
if (ihw_proceed) {
  ihw <- ihw(
    pvalues = merged_results$P.Value,
    covariates = merged_results$ihw_covariate,
    alpha <- fdr_alpha,
    covariate_type = "nominal"
  )
  merged_results <- mutate(merged_results, fdr_ihw = adj_pvalues(ihw))
} 
merged_results <- merged_results %>% 
  mutate(
    status = case_when(
      !!sym(fdr_column) < fdr_alpha ~ as.character(direction),
      !!sym(fdr_column) >= fdr_alpha ~ "Unchanged"
    ) %>% 
      factor(levels = c(levels(direction), "Unchanged")) %>% 
      droplevels()
  )
```


Independent Hypothesis Weighting (IHW) [@IgnatiadisIHW2016] was then used to partition the raw p-values for `r target` differential binding by the detection of the other ChIP targets under investigation in this workflow (`r glue_collapse(other_targets, sep = ", ", last = " and ")`).
The presence of `r glue_collapse(other_targets, sep = ", ", last = " and ")` was defined simply using the consensus peaks detected under any treatment by `macs2 callpeak`, as determined [previously](`r glue("{target}_macs2_summary.html")`).
This allows recalculation of the FDR using *weighted p-values* instead of *raw p-values*.
In order for IHW to be a viable strategy, partitions should be greater than 1000.
The provided consensus peaks were used as initial partitions in combination, merging the smallest groups below this size until all partitions were of a suitable size.

Windows were classified as overlapping a peak from a secondary ChIP target if *any* section of the window overlapped the secondary peak.
`r if (!ihw_proceed) "Only one partition > 1000 was able to be formed and IHW was not executed."`

`r if (ihw_proceed) "### Summary Table"`

```{r ihw-tab, eval = ihw_proceed, echo = ihw_proceed}
ihw %>%
  as.data.frame() %>% 
  mutate(
    fdr = p.adjust(pvalue, "fdr"),
    status = case_when(
      fdr < fdr_alpha & adj_pvalue > fdr_alpha ~ "Lost",
      fdr < fdr_alpha & adj_pvalue < fdr_alpha ~ "Retained",
      fdr > fdr_alpha & adj_pvalue < fdr_alpha ~ "Gained",
      fdr > fdr_alpha & adj_pvalue > fdr_alpha ~ "Never Significant"
    ) %>% 
      factor(
        levels = c("Never Significant", "Retained", "Gained", "Lost")
      )
  ) %>% 
  as_tibble() %>% 
  group_by(covariate, status) %>% 
  tally() %>% 
  ungroup() %>% 
  complete(status = status, covariate = covariate, fill = list(n = 0)) %>% 
  pivot_wider(
    names_from = status, values_from = n, values_fill = 0
  ) %>% 
  mutate(
    `Nett Change` = Gained - Lost
  ) %>% 
  bind_rows(
    summarise_if(., is.numeric, sum)
  ) %>% 
  mutate(
    covariate = str_replace_na(covariate, "Total"),
    Significant = Retained + Gained
  ) %>% 
  dplyr::select(
    Covariate = covariate, ends_with("Significant"), Retained, everything()
  ) %>% 
  pander(
    justify = "lrrrrrr",
    caption = paste(
      "*Summary of changes introduced by IHW for windows considered as being", 
      paste0("differentially bound by ", target, "."),
      "This corresponds to a __nett change of", 
        dplyr::slice(., nrow(.)) %>% 
        mutate(p = `Nett Change` / (Significant - `Nett Change`)) %>% 
        .[["p"]] %>% 
        percent(0.1),
      "from the initial list__.*"
    ),
    emphasize.strong.rows = nrow(.)
  )
```


### Initial Group Sizes

```{r ihw-partitions-raw, fig.height=6, fig.cap = "*Breakdown of all windows which overlapped peaks from additional ChIP targets. Any partitions with fewer than 1000 windows (indicated as the blue horizontal line) were combined into the next smallest partition consecutively, until all partitions contained > 1000 windows.*"}
tibble(ihw = ihw_all) %>% 
  ggplot(
    aes(ihw, fill = ihw)
  ) +
  geom_bar() +
  geom_hline(yintercept = 1e3, colour = "blue") +
  geom_text(
    aes(y = n + 0.025*max(n), label = comma(n, 1)),
    data = . %>%
      group_by(ihw) %>%
      summarise(n = dplyr::n(), .groups = "drop") 
  ) +
  scale_y_continuous(expand = expansion(c(0, 0.05)), labels = comma) +
  labs(x = "Peaks Overlapping Other ChIP Targets", y = "Total") +
  theme(legend.position = "none")
```

### P-Value Distributions

```{r plot-ihw-pvals, fig.height= 3 + 2 * ceiling(length(levels(merged_results$ihw_covariate)) / 3), fig.cap = "*P-Value distributions within all final data partitions. The size of each partition is given within each panel.*"}
y <- ggplot_build(
  tibble(
    P.Value = merged_results$P.Value,
    ihw_covariate = merged_results$ihw_covariate
    ) %>% 
    ggplot(aes(P.Value, stat(density))) + 
    geom_histogram(bins = 100) +
    facet_wrap(~ihw_covariate) 
) %>% .[["data"]] %>% 
  .[[1]] %>% 
  .[["y"]] %>% 
  max()
merged_results %>% 
  select(P.Value, ihw_covariate) %>% 
  mcols() %>% 
  as_tibble() %>% 
  ggplot(aes(P.Value, stat(density))) +
  geom_histogram(fill = "grey", colour = "black", bins = 100) +
  geom_text(
    aes(0.8, y, label = lab),
    data = . %>% 
      group_by(ihw_covariate) %>% 
      summarise(n = dplyr::n(), .groups = "drop") %>% 
      mutate(
        lab = glue("N = {comma(n, 1)}"),
        y = y
      )
  ) +
  facet_wrap(~ihw_covariate) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  labs(x = "P Value", y = "Density")
```

`r ifelse(ihw_proceed, "### IHW Weights", "")`

```{r plot-ihw-weights, eval = ihw_proceed, echo = ihw_proceed, fig.height = 5, fig.cap = "*Weights applied to p-values within each partition. 'Folds' represent random sub-partitions within each larger partition generated as part of the IHW process.*"}
plot(ihw) +
  geom_hline(yintercept = 1) +
  facet_wrap(~group) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  labs(x = "Fold", y = "Weight", fill = "Fold")
```

`r ifelse(ihw_proceed, "### FDR Comparison", "")`

```{r plot-fdr-comparison, dev = 'png', eval = ihw_proceed, echo = ihw_proceed, fig.height = 5, fig.cap = glue("*Comparison of raw and weighted p-values for each partition. Blue dashed lines indicate FDR = {fdr_alpha} for each set of p-values. Those in the lower-right quadrant would no longer be considered significant after IHW, whilst those in the upper-left quadrant would only be considered as significant after the IHW process. Those in the upper-right quadrant would be considered as significant regardless of the methodology.*")}
ihw %>%
  as.data.frame() %>% 
  ggplot(
    aes(-log10(pvalue), -log10(weighted_pvalue), colour = fold)
  ) +
  geom_point(size = 0.4) +
  geom_hline(
    aes(yintercept = -log10(weighted_pvalue)),
    data = . %>% 
      dplyr::select(-group) %>% 
      dplyr::filter(adj_pvalue < fdr_alpha) %>% 
      dplyr::filter(weighted_pvalue == max(weighted_pvalue)),
    linetype = 2, colour = "blue"
  ) +
  geom_vline(
    aes(xintercept = -log10(pvalue)),
    data =. %>% 
      dplyr::filter(p.adjust(pvalue, "fdr") < fdr_alpha) %>% 
      dplyr::filter(pvalue == max(pvalue)) %>% 
      dplyr::select(-group),
    linetype = 2, colour = "blue"
  ) +
  facet_wrap(~group, scales = "free") +
  coord_cartesian(xlim = c(0, 5), ylim = c(0, 5)) +
  labs(
    x = expression(-log[10](p)),
    y = expression(-log[10](p[IHW])),
    colour = "Fold"
  )
```


