## Data Inspection {.tabset}

No normalisation was performed beyond library size, with library sizes being determined from the *entire bam file*, not just areas where `r target` was considered detected

### Density Plot: All Sliding Windows

```{r plot-density-windows, fig.height = 6, fig.width = 8, fig.cap = "*Density plot for __all windows prior to the selection of windows__ more likely to contain true signal. Retained windows will be those at the upper end, whilst discarded windows will be at the lower end.*"}
plotAssayDensities(
  window_counts, trans = "log1p", n_max = n_max, colour = "treat",
  linetype = "target"
) +
  scale_colour_manual(values = colours$treat) + 
  labs(
    x = "log(counts + 1)", y = "Density", colour = "Treat", linetype = "Target"
  )
```

### logCPM Densities

```{r plot-logcpm-densities, fig.height = 6, fig.width = 8, fig.cap = "*logCPM distributions for all retained windows.*"}
filtered_counts %>% 
  plotAssayDensities("logCPM", colour = "treat", n_max = n_max) +
  scale_colour_manual(values = treat_colours) +
  labs(colour = "Treat")
```

### RLE 

```{r plot-rle, dev = 'png', fig.height=6, fig.cap = glue("*RLE plot showing logCPM values. RLE values were calculated within each treatment group to account for the potentially different binding dynamics of {target}.*")}
filtered_counts %>% 
  plotAssayRle(
  "logCPM", fill = "treat", rle_group = "treat", n_max = n_max
) +
  geom_hline(yintercept = 0) + 
  facet_grid(.~treat, scales = "free_x", space = "free_x") +
  scale_x_discrete(labels = sample_labeller) +
  scale_fill_manual(values = treat_colours) +
  labs(x = "Sample", fill = "Treat")
```


### PCA

```{r plot-pca, fig.height = 6, fig.width = 8, fig.cap = "*PCA plot based on logCPM values*"}
plotAssayPCA(filtered_counts, "logCPM", colour = "treat", label = "label") +
  scale_colour_manual(values = treat_colours)
```


# Differential Binding Analysis

## Primary Analysis {.tabset}

```{r fit-model}
fm <- ifelse(
  config$comparisons$paired & length(unique(vapply(split(samples, samples$treat), nrow, integer(1)))) == 1, 
  glue("~ {rep_col} + treat"), 
  glue("~ treat")
) %>% 
  as.formula()
X <- model.matrix(fm, data = colData(filtered_counts)) %>%
  set_colnames(str_remove(colnames(.), "treat"))
colData(filtered_counts)$design <- X
fc <- ifelse(is.null(config$comparisons$fc), 1, config$comparisons$fc)
dge <- DGEList(
  counts = assay(filtered_counts, "counts") %>% 
    set_rownames(as.character(rowRanges(filtered_counts))),
  lib.size = colData(filtered_counts)$totals,
  genes = as_tibble(rowRanges(filtered_counts)) %>% 
    dplyr::select(range, overlaps_ref)
) %>% 
  calcNormFactors(method = "none") %>% 
  estimateDisp(design = X)
fit <- glmQLFit(dge, design = X) %>% 
  glmTreat(coef = treat_levels[[2]], lfc = log2(fc))
fit_mu0 <- dge %>% 
  glmQLFit() %>% 
  glmQLFTest(coef = treat_levels[[2]])
res_cols <- c("logFC", "logCPM", "PValue", "FDR")
limma_cols <- setNames(
  c("logFC", "AveExpr", "P.Value", "fdr"),
  res_cols
) # This ensures compatibility with the structure of sq-lt results
rowData(filtered_counts) <- rowData(filtered_counts) %>%
  .[!colnames(.) %in% c(res_cols, limma_cols, "p_mu0")] %>% 
  as.list() %>% 
  c(
    fit %>% 
      topTags(sort.by = "none", n = Inf) %>% 
      .[["table"]] %>% 
      as.list %>%
      .[res_cols] %>% 
      setNames(limma_cols[names(.)]) %>% 
      c(
        list(
          p_mu0 = topTags(fit_mu0, sort.by = "none", n = Inf)$table$PValue
        )
      ) 
  ) 
```

Using simple library-size normalisation, the Quasi-Likelihood model was fitted across the set of sliding windows, taking `r treat_levels[[1]]` as the baseline and estimating the effects of `r treat_levels[[2]]` on `r target` binding within each sliding window, with all default settings.
`r if (str_detect(as.character(fm)[[2]], "\\+")) paste("A separate baseline expression level was estimated for each", rep_col, "with a common treatment response being fitted.")`
