```{r set-knitr-opts, echo=FALSE, child = here::here('analysis/setup_chunk.Rmd')}
```


```{r get-params}
target <- params$target
threads <- parallel::detectCores() - 1
# if (interactive()) threads - 4
```

```{r packages}
library(tidyverse)
library(yaml)
library(pander)
library(glue)
library(extraChIPs)
library(plyranges)
library(scales)
library(patchwork)
library(parallel)
library(reactable)
library(htmltools)
library(universalmotif)
library(motifTestR)
library(rlang)
library(GenomicInteractions)
source(here::here("workflow/scripts/custom_functions.R"))
theme_set(theme_bw())
config <- read_yaml(here::here("config", "config.yml"))
all_paths <- here::here("config", "json", "all_paths.json") %>% 
    jsonlite::fromJSON() %>% 
    lapply(here::here)
ucsc <- get_ucsc(config$genome$build)
bs_pkg <- glue("BSgenome.{ucsc$sp}.UCSC.{ucsc$build}")
library(bs_pkg, character.only = TRUE)
```

```{r load-samples}
samples <- all_paths$macs2 %>%
  file.path(target, glue("{target}_qc_samples.tsv")) %>%
  read_tsv()
treat_levels <- unique(samples$treat)
all_targets <- here::here(config$samples) %>% 
  read_tsv() %>% 
  pull("target") %>% 
  unique() 
all_samples <- all_paths$macs2 %>% 
  file.path("{all_targets}", "{all_targets}_qc_samples.tsv") %>% 
  glue() %>% 
  lapply(read_tsv) %>% 
  bind_rows() %>% 
  dplyr::filter(qc == "pass") %>% 
  dplyr::select(-all_of("qc"))
all_targets <- unique(all_samples$target)
n_targets <- all_targets
```

```{r config-params-paths}
all_params <-  here::here("config", "params.yml") %>% read_yaml() 
enrich_params <- all_params$enrichment
region_params <- all_params$gene_regions
regioner_params <- all_params$regioner
nfr_params <- all_params$nfr
motif_params <- here::here("config", "json", "motif_analysis_param.json") %>%
  jsonlite::fromJSON() %>%
  pluck("nfr")
knitr_params <- here::here("config", "rmarkdown.yml") %>% 
  read_yaml() %>% 
  pluck("knitr_opts")
source(here::here("workflow/scripts/table_funs.R"))
source(here::here("workflow/scripts/enrichment_funs.R"))
```

```{r figure-params}
fh <- knitr_params$fig.height
fw <- knitr_params$fig.width
fig_path <- here::here("docs", "assets", paste0(target, "_NFR"))
if (!dir.exists(fig_path)) dir.create(fig_path, recursive = TRUE)
fig_type <- knitr_params$dev[[1]]
fig_fun <- match.fun(fig_type)
if (fig_type %in% c("bmp", "jpeg", "png", "tiff")) {
  ## These figure types require dpi & resetting to be inches
  formals(fig_fun)$units <- "in"
  formals(fig_fun)$res <- 300
}
```



```{r load-annotations}
colours <- file.path(all_paths$annotation, "colours.rds") %>% read_rds()
names(colours$target)[names(colours$target) == target] <- paste0(target, "_NFR")
sq <- file.path(all_paths$annotation, "seqinfo.rds") %>% read_rds()
gene_regions <- file.path(all_paths$annotation, "gene_regions.rds") %>% 
  read_rds()
regions <- map_chr(gene_regions, \(x) x$region[1])
region_colours <- unlist(colours$regions) %>% 
  setNames(map_chr(gene_regions, \(x) x$region[[1]])[names(.)])
full_palette <- colours %>% 
  .[setdiff(names(.), "heatmaps")] %>% 
  unname() %>% 
  unlist()
## External features
external_features <- read_rds(file.path(all_paths$annotation, "features.rds"))
feat_names <- names(external_features) %>% 
  lapply(
    \(x) paste(x, external_features[[x]]$feature, sep = ": ")
  ) %>% 
  lapply(fct_infreq) %>% 
  lapply(levels) %>% 
  unlist()
## Combine features & regions for testing
all_regions <- c(unname(regions), feat_names)
hic <- file.path(all_paths$annotation, "hic.rds") %>% read_rds()
has_hic <- length(hic) > 0
rna <- file.path(all_paths$annotation, "rna.rds") %>% read_rds()
has_rna <- length(rna) > 0
```


## Introduction

Nucleosome Free Regions (NFRs) were identified for `r target` using the set of filtered peaks obtained for each treatment (`r pander(treat_levels)`).
Peaks within `r nfr_params$merge_peaks_within`bp of each other were merged into a single peak, with this distance approximating the distance occupied by 2 nucleosomes, and approximating broader peaks that may have been otherwise called by `macs2` [@Zhang18798982].
NFRs within merged peaks were called using Histrader [@Yan2020-nz] for each treatment group with the following settings.

- Any merged peaks <`r nfr_params$min_peak_width`bp were excluded as being too narrow to identify NFRs
- Any NFRs beyond `r max(nfr_params$nfr_width)/1e3`kb were discarded as not being viable NFRs
- Regions of the peak with signal below `r percent(nfr_params$p_max)` of the maximal signal were also not assessed for NFRs as these were more likely to be regions on the outer edges of wider peaks.
- Any NFRs within `r nfr_params$merge_nfrs_within`bp of the nearest NFR (~1/3 of a nucleosome) were also merged into a single NFR

Treatment-specific NFRs were then merged as consensus NFRs, taking only shared regions between treatment groups which were wider than `r min(nfr_params$nfr_width)`bp and which were identified in all groups.
Given the inclusive nature of HisTrader, this is a more stringent approach to ensure overly-inclusive NFR calls were minimised.

Whilst a motif enrichment analysis was performed, along with an assessment of the relationship between NFRs and other targets, no gene-set enrichment testing was performed.
The mapping from NFR to genes would very closely replicate that of the larger set of `r target` peaks, and as such, this was considered to be a redundant analysis.


```{r load-nfr}
nfr_by_treat <- all_paths$nfr %>% 
  file.path(target, "{target}_{treat_levels}.nfr.bed.gz") %>% 
  glue() %>% 
  importPeaks(type = "bed", seqinfo = sq) %>% 
  setNames(treat_levels)
nfr_consensus <- all_paths$nfr %>% 
  file.path(target, "{target}_consensus_nfr.rds") %>% 
  glue() %>% 
  read_rds()
consensus_peaks <- all_paths$peaks %>% 
  file.path("{all_targets}", "{all_targets}_consensus_peaks.rds") %>% 
  glue() %>% 
  lapply(read_rds) %>% 
  setNames(all_targets) %>% 
  GRangesList()
consensus_peaks[[paste0(target, "_NFR")]] <- nfr_consensus
has_features <- "feature" %in% colnames(mcols(consensus_peaks[[target]]))
```

## NFR Summary {.tabset}

### Shared NFRs

```{r plot-overlaps, fig.width=fh, fig.cap = "*Summary of NFRs identified in each treatment group separately, before identifying consensus NFRs*"}
plotOverlaps(nfr_by_treat, set_col = unlist(colours$treat[treat_levels]))
```

### Peak & NFR Width

```{r plot-width, fig.cap = glue("*A) Relationship between {target} peak width and NFR identification. {target} peaks within 300bp were merged prior to plotting, and B) NFR width and numbers by treatment group.*")}
A <- consensus_peaks[[target]] %>% 
  reduce(min.gapwidth = 300) %>% 
  mutate(has_nfr = overlapsAny(., nfr_consensus)) %>% 
  split(.$has_nfr) %>% 
  plotGrlCol(
    .id = 'has_nfr', fill = 'has_nfr', q = 0, geom = "violin",
    draw_quantiles = 0.5, trim = FALSE
  ) +
  scale_y_log10(labels = comma) +
  scale_fill_manual(values = c("indianred", "forestgreen")) +
  labs(
    x = "NFR Detected",
    y = glue("{target} Peak Width (bp)")
  )
B <- nfr_by_treat %>% 
  c(GRangesList(`Consensus NFR` = nfr_consensus)) %>% 
  plotGrlCol(
    df = tibble(treat = treat_levels), .id = "treat", fill = treat,
    geom = 'violin', q = 0, draw_quantiles = 0.5, trim = FALSE
  ) +
  scale_fill_manual(values = unlist(colours$treat)) +
  scale_y_log10(labels = comma) +
  labs(
    x = "NFR Group",
    y = "NFR Width (bp)"
  ) 
A + B + plot_annotation(tag_levels = "A") & theme(legend.position = 'none')
```

### NFRs Per Peak

```{r plot-nfr-by-peak-width, fig.cap = glue("*Number of NFRs found within each {target} peak, separated by genomic region. {target} peaks within {nfr_params$merge_peaks_within}bp were merged prior ro plotting*")}
consensus_peaks[[target]] %>% 
  reduce(min.gapwidth = 300) %>% 
  mutate(
    w = width, 
    region = regions[bestOverlap(., gene_regions)] %>% factor(levels = regions),
    nfr = countOverlaps(., nfr_consensus)
  ) %>% 
  as_tibble() %>% 
  ggplot(aes(w, nfr)) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, method = "lm") +
  geom_label(
    aes(w, nfr, label = lab) ,
    data = . %>% 
      summarise(
        n = sum(nfr), 
        w = min(w), nfr = max(nfr),
        .by = "region"
      ) %>% 
      mutate(
        lab = glue("n = {comma(n, 1)}"),
        w = min(w), nfr = max(nfr)
      ),
    hjust = 0, vjust = 1
  ) +
  facet_wrap(~ region) +
  scale_x_continuous(labels = comma_format(scale = 1e-3)) +
  labs(
    x = glue("{target} Merged Peak Width (kb)"),
    y = "NFR / Peak"
  )
```

### Distance Between NFRs

```{r plot-dist, fig.cap = glue("*Distances between each NFR and the nearest upstream NFR within each broader {target} peak*")}
nfr_consensus %>% 
  granges() %>% 
  join_overlap_inner(
    consensus_peaks[[target]] %>% 
      reduce(min.gapwidth = 300) %>% 
      mutate(id = seq_along(.))
  ) %>% 
  as_tibble(rangeAsChar = FALSE) %>% 
  split(.$id) %>% 
  .[map_int(., nrow) > 1] %>% 
  mclapply(
    \(x) x$start[-1] - x$end[-nrow(x)], 
    mc.cores = threads
  ) %>% 
  unlist() %>% 
  as_tibble_col("distance") %>% 
  ggplot(aes(distance, after_stat(count))) +
  geom_histogram(
    fill = "grey70", colour = 'black', bins = 100
  ) +
  scale_x_log10() +
  scale_y_continuous(expand = expansion(c(0, 0.05)), labels = comma) +
  labs(
    x = "Distance between NFRs", y = "Totals"
  )
```

### NFR By Region

```{r plot-pie, fig.width = fh, fig.cap = glue("*Distribution of NFRs by genomic region, noting that wider {target} peaks will accumulate more NFRs*")}
nfr_consensus %>% 
  plotPie(
    fill = region, total_size = 4,
    cat_glue = "{str_wrap(.data[[fill]], 15)}\n{comma(n, 1)}\n({percent(p, 0.1)})",
    cat_alpha = 0.7, cat_adj = 0.1
  ) +
  scale_fill_manual(values = region_colours) +
  theme(legend.position = "none") 
```

`r ifelse(has_features, "### NFR By Feature", "")`

```{r plot-nfr-by-feature, echo = has_features, fig.height = ifelse(length(feat_names) < 3, fh / 2, fh), fig.cap = glue("*{target} NFRs which overlap each provided feature, coloured by gene-centric regions defined previously. Features may overlap with each other, and as such, NFRs may overlap multiple feature types and be included within multiple categories. Total overlapping NFRs for each feature type are shown in bold, wth percentages being relative to the total number of consensus NFRs ({comma(length(nfr_consensus))}). Individual segments are labelled if representing >= 3% of the total number of NFRs Features with fewer than 1% of NFRs aligning to them are collapsed into the category 'Other'*")}
if (has_features) {
  nfr_consensus %>% 
    as_tibble() %>% 
    dplyr::select(range, region, feature) %>% 
    unnest(feature, keep_empty = TRUE) %>%
    mutate(
      feature = case_when(
        is.na(feature)~ "no_feature",
        is.null(names(feature)) ~ feature,
        TRUE ~ paste(names(feature), feature, sep = ": ")
      ) %>% 
        str_remove_all("^ *:")
    ) %>% 
    distinct(range, region, feature) %>% 
    mutate(feature = fct_infreq(feature) %>% fct_lump_prop(0.01)) %>% 
    dplyr::summarise(n = dplyr::n(), .by = c(region, feature)) %>% 
    arrange(desc(region)) %>% 
    mutate(label_x = cumsum(n) - 0.5*n, .by = c(feature)) %>% 
    ggplot(aes(n, y = fct_rev(feature))) +
    geom_col(aes(fill = region)) +
    geom_label(
      aes(x = N, label = label),
      data = . %>% 
        summarise(N = sum(n), .by = feature) %>% 
        mutate(
          p = N / length(nfr_consensus),
          label = glue("{comma(N)}\n{percent(p, 0.1)}"),
          N = N + 0.05 * max(N)
        ),
      hjust = 0.5, size = 3, fontface = 'bold'
    ) +
    geom_label(
      aes(label_x, fct_rev(feature), label = label),
      data = . %>% 
        dplyr::filter(n > 0.03 * length(nfr_consensus)) %>% 
        mutate(
          label = glue("{comma(n)}\n{percent(n / length(nfr_consensus), 0.1)}")
        ),
      alpha = 0.6, size = 3
    ) +
    scale_fill_manual(values = region_colours, labels = label_wrap(10)) +
    scale_x_continuous(expand = expansion(c(0, 0.1)), labels = comma) +
    labs(
      x = glue("Total Overlapping {target} NFRs"), 
      y = "Feature", fill = "Region"
    ) +
    theme(
      legend.key.height = unit(0.5 / length(regions), "npc")
    )
}
```


## Motif Analysis

```{r load-motif-res}
bs_genome <- get(bs_pkg)
fw_peaks <- resize(
  nfr_consensus, width = motif_params$peak_width, fix = "center"
)
genome(fw_peaks) <- ucsc$build
fw_seq <- getSeq(bs_genome, fw_peaks) %>%
  .[letterFrequency(., "N")[,1] == 0]
motif_adj <- match.arg(motif_params$adj, p.adjust.methods)
motif_list <- file.path(all_paths$annotation, "motif_list.rds") %>% 
  read_rds() %>% 
  to_df %>% 
  as_tibble()
motif_uri <- file.path(all_paths$annotation, "motif_uri.rds") %>% 
  read_rds()
motif_pos_res <- all_paths$nfr %>%
  file.path("{target}", "{target}_motif_position.tsv.gz") %>% 
  glue() %>% 
  read_tsv() %>% 
  dplyr::select(-fdr) %>% 
  mutate(adj_p = p.adjust(p, motif_adj))
n_pos <- sum(motif_pos_res$adj_p < motif_params$alpha)
motif_enrich_res <- all_paths$nfr %>% 
  file.path("{target}", "{target}_motif_enrichment.tsv.gz") %>%
  glue() %>% 
  read_tsv() %>% 
  dplyr::select(-fdr) %>% 
  mutate(adj_p = p.adjust(p, motif_adj))
n_enrich <- sum(
  motif_enrich_res$adj_p < motif_params$alpha & motif_enrich_res$Z > 0
)
```

Using the motifs defined during preparation, testing was performed for 1) enrichment above background and 2) any positional bias within the tested sequences.
For this step peaks were set to a fixed width of `r motif_params$peak_width`bp, with sequences taken from these regions.

### Motif Enrichment {.tabset}

A simple enrichment analysis was performed by selecting a set of `r comma(motif_enrich_res$sequences[[1]] * motif_enrich_res$n_iter[[1]])` *background sequences* matched by genomic regions to the `r comma(motif_enrich_res$sequences[[1]])` *test sequences*.
This gave a set of genomically matched control sequences to gauge enrichment by.
Analysis was performed using the `motifTestR` package and the `r motif_params$model` model.

Whilst statistical testing was two-sided, only enriched motifs (i.e. observed > expected) are shown in the results below, with `r n_enrich` of the `r comma(nrow(motif_list))` included motifs matching the criteria for being considered as enriched.

`r if (n_enrich > 0) "#### Enrichment Table"`

```{r tbl-res-enrich, eval = n_enrich > 0, echo = n_enrich > 0}
tbl <- motif_enrich_res %>%
  arrange(p) %>% 
  dplyr::filter(adj_p < motif_params$alpha, Z > 0) %>% 
  mutate(cluster = fct_inorder(as.character(cluster))) %>% 
  dplyr::select(
    cluster, ends_with("name"), Z, p, adj_p, matches, expected, enrichment
  ) %>% 
  mutate(
    PWM = lapply(altname, \(x) motif_uri[[x]])
  ) %>% 
  reactable(
    groupBy = "cluster",
    showPageSizeOptions = TRUE, filterable = TRUE,
    theme = reactableTheme(style = list(fontSize = 12)),
    defaultSorted = "adj_p", defaultSortOrder = "asc",
    columns = list(
      cluster = colDef("Cluster", maxWidth = 70),
      altname = colDef(
        "Motif", minWidth = 180, aggregate = "unique",
        style = list(borderLeft = "1px solid rgba(0, 0, 0, 0.1)")
      ),
      name = colDef(
        "Name", minWidth = 120, aggregate = "unique",
        style = list(borderRight = "1px solid rgba(0, 0, 0, 0.1)")
      ),
      matches = colDef(
        "Total", maxWidth = 60, format = comma_col, aggregate = "max",
        filterMethod = js_greater
      ),
      expected = colDef(
        "Expected", maxWidth = 80, format = comma_col, aggregate = "max",
        filterMethod = js_greater
      ),
      enrichment = colDef(
        "Enrichment", maxWidth = 90, format = colFormat(digits = 3),
        aggregate = "max", filterMethod = js_greater,
        style = list(borderRight = "1px solid rgba(0, 0, 0, 0.1)")
      ),
      Z = colDef(
        maxWidth = 60, format = colFormat(digits = 2), aggregate = "max",
        filterMethod = js_greater
      ),
      p = colDef(show = FALSE),
      adj_p = colDef(
        glue("p<sub>{motif_adj}</sub>"), html = TRUE, maxWidth = 70,
        cell = \(value) sprint_pval(value), aggregate = "min",
        format = colFormat(digits = 3), filterMethod = js_less
      ),
      PWM = colDef(
        name = "IC Matrix", minWidth = 180, filterable = FALSE,
        cell = function(value) tags$img(src = value, height = '80px')
      )
    ),
    elementId = "motif-enrich-table"
  )
cp <- htmltools::em(
  glue(
    "All {n_enrich} motifs considered as enriched relative to the background set
    of sequences, using the {motif_params$model} model and an adjusted 
    p-value ({motif_adj}) < {motif_params$alpha} as the criteria for 
    significance. Motifs are grouped by clusters formed earlier, with all motifs
    visible by clicking the arrow to the arbitrary cluster identifier. All 
    columns are searchable and filterable."
  )
)
tl <- tagList(
  tags$button(
    tagList(fontawesome::fa("download"), "Download as CSV"),
    onclick = glue(
      "Reactable.downloadDataCSV('motif-enrich-table', '{target}_motif_enrich_res.csv')"
    )
  ),
  tags$button(
    "Expand/Collapse All",
    onclick = "Reactable.toggleAllRowsExpanded('motif-enrich-table')"
  ),
  tags$button(
    "Clear All Filters",
    onclick = "Reactable.setAllFilters('motif-enrich-table', [])"
  ),
  tbl,
  cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```

`r if (n_enrich > 1) "#### Comparison of Motif Co-Occurrence"`

```{r enrich-corrplot, echo = n_enrich > 1, fig.cap = glue("*The top {length(top_enrich)} motifs ranked for enrichment were compared to each other using correlations derived from simple presence/absence calls. Highly correlated motifs may indicate similarity between TFBMs and the same sequence match being identified. Motifs detected in all sequences were excluded after the initial selction and as such, fewer than the selected {length(top_enrich)} motifs may be shown*")}
if (n_enrich > 1) {
  top_enrich <- motif_enrich_res %>%
    dplyr::filter(Z > 0, adj_p < motif_params$alpha) %>% 
    dplyr::slice(1:20) %>% 
    pull("altname")
  cor_mat <- motif_list %>% 
    dplyr::filter(altname %in% top_enrich) %>% 
    to_list() %>% 
    getPwmMatches(fw_seq, mc.cores = threads) %>% 
    lapply(\(x) tibble(seq = unique(x$seq))) %>% 
    bind_rows(.id = "altname") %>%
    distinct(altname, seq) %>% 
    left_join(
      dplyr::select(motif_enrich_res, altname, name), by = "altname"
    ) %>% 
    mutate(detected = TRUE) %>% 
    dplyr::select(seq, name, detected) %>% 
    pivot_wider(
      names_from = "name", values_from = "detected", values_fill = FALSE
    ) %>% 
    dplyr::select(-seq) %>% 
    dplyr::select(where(\(x) mean(x) < 1 & mean(x) > 0)) %>%
    as.matrix() %>%
    cor()
  hc <- hclust(dist(1 - cor_mat), method = "ward.D2")
  dend <- ggdendro::dendro_data(as.dendrogram(hc))
  levels <- dend$labels$label
  ym <- max(dend$segments$xend)
  dend_aes <- aes(yend, xend, xend = y, yend = x)
  dend_plot <- ggplot(dend$segments) +
    geom_segment(dend_aes) + 
    scale_x_reverse(expand = expansion(0)) + 
    scale_y_reverse(limits = c(ym, 0) + 0.5, expand = expansion(0)) +
    labs(x = "", y = c()) + 
    theme_minimal() + 
    theme(
      panel.grid = element_blank(), 
      axis.text = element_blank(), axis.ticks = element_blank(), 
      plot.margin = unit(c(5.5, 0, 5.5, 5.5), "points")
    )
  p <- cor_mat %>% 
    as_tibble(rownames = "m1") %>% 
    pivot_longer(cols = all_of(.$m1), names_to = "m2", values_to = "cor") %>% 
    mutate(across(starts_with("m"), \(x) factor(x, levels = levels))) %>% 
    dplyr::filter(m1 != m2) %>% 
    ggplot(aes(m1, fct_rev(m2))) +
    geom_point(aes(size = abs(cor), colour = cor, alpha = abs(cor))) +
    geom_text(
      aes(label = round(cor*100, 0), alpha = abs(cor)), show.legend = FALSE
    ) +
    scale_size(range = c(0, 10), guide = "none") +
    scale_alpha(range = c(0, 1), guide = "none") +
    scale_colour_gradient2(
      low = colours$direction[["decreased"]], 
      high = colours$direction[["increased"]],
      limits = c(-1, 1), labels = percent_format(suffix = "")
    ) +
    scale_x_discrete(position = "top") +
    labs(x = NULL, y = NULL, colour = expr(paste(rho))) +
    theme(
      panel.grid = element_blank(), 
      axis.text.x.top = element_text(angle = 90, hjust = 0, vjust = 0.5)
    )
  dend_plot + p + plot_layout(widths = c(0.1, 0.9), axes = "collect")
}
```



### Positional Bias {.tabset}

```{r get-top-pos, eval = n_pos > 0, echo = n_pos > 0}
top_pos <- motif_pos_res %>% 
  dplyr::filter(adj_p < motif_params$alpha) %>% 
  dplyr::filter(matches_in_region == max(matches_in_region), .by = cluster) %>% 
  dplyr::slice(1:9) %>% 
  dplyr::select(altname, name, cluster)
clust_names <- motif_pos_res %>% 
  dplyr::filter(
    adj_p < motif_params$alpha,
    cluster %in% top_pos$cluster
  ) %>% 
  arrange(desc(matches_in_region)) %>% 
  summarise(
    altname = altname[[1]], .by = cluster,
    name = paste(name, collapse = "/")
  ) %>% 
  mutate(name = setNames(name, altname)) %>% 
  pull(name)
```

Positional bias was tested using the package `motifTestR`, bin-sizes of `r motif_params$binwidth`bp and testing `r ifelse(motif_params$abs, "distance from sequence centres", "position either side of the sequence centres")`.
Any motifs which matched fewer than `r percent(motif_params$ignore_below)` of sequences were not tested for any positional bias.

Using an adjusted p-value (`r motif_adj`) less than `r motif_params$alpha` to be considered as significant, `r n_pos` motifs were considered as showing any type of positional bias.

It should also be noted that this type of analysis is agnostic to the number of matches found for a TFBM, and only assesses any bias within the distribution of matches.


`r if (n_pos > 0) "#### Results Table"`

```{r tbl-res-pos, eval = n_pos > 0, echo = n_pos > 0}
tbl <- motif_pos_res %>%
  dplyr::filter(adj_p < motif_params$alpha) %>% 
  mutate(cluster = fct_inorder(as.character(cluster))) %>% 
  dplyr::select(
    cluster, contains("name"), ends_with("p"), start, end, centre, width,
    contains("total"), matches_in_region, starts_with("e"), 
  ) %>% 
  mutate(
    PWM = lapply(altname, \(x) motif_uri[[x]])
  ) %>% 
  reactable(
    groupBy = "cluster", showPageSizeOptions = TRUE, filterable = TRUE,
    columns = list(
      cluster = colDef(maxWidth = 60),
      name = colDef("Name", aggregate = "unique", maxWidth = 80),
      altname = colDef("Motif", aggregate = "unique", minWidth = 180),
      start = colDef(
        "Start", aggregate = "min", maxWidth = 50,
        filterMethod = js_greater
      ),
      end = colDef(
        "End", aggregate = "max", maxWidth = 50, 
        filterMethod = js_greater
      ),
      centre = colDef(
        "Centre", aggregate = "median", maxWidth = 60, show = FALSE
      ),
      width = colDef(
        "Width", aggregate = "median", maxWidth = 60, 
        filterMethod = js_less
      ),
      total_matches = colDef(
        "Total", aggregate = "max", maxWidth = 60, format = comma_col,
        filterMethod = js_greater
      ),
      matches_in_region = colDef(
        "Region", aggregate = "max", maxWidth = 60, 
        filterMethod = js_greater
      ),
      expected = colDef(
        "Expected", aggregate = "max", maxWidth = 75, format = comma_col,
        filterMethod = js_greater
      ),
      enrichment = colDef(
        "Region Enrichment", aggregate = "max", format = colFormat(digits = 2),
        maxWidth = 90, filterMethod = js_greater
      ),
      prop_total = colDef(
        "% In Region", aggregate = "max", format = percent_col, maxWidth = 60,
        filterMethod = js_greater
      ),
      p = colDef(show = FALSE),
      adj_p = colDef(
        glue("p<sub>{motif_adj}</sub>"), html = TRUE, maxWidth = 70,
        cell = \(value) sprint_pval(value), format = colFormat(digits = 3),
        aggregate = "min", filterMethod = js_less
      ),
      PWM = colDef(
        name = "IC Matrix", minWidth = 180, filterable = FALSE,
        cell = function(value) tags$img(src = value, height = '80px')
      )
    ),
    columnGroups = list(
      colGroup(
        name = "Matches", 
        columns = c(
          "total_matches", "prop_total", "matches_in_region", "expected",
          "enrichment"
        )
      ),
      colGroup(
        name = "Positional Bias", 
        columns = c("start", "end", "centre", "width")
      )
    ),
    defaultSorted = "adj_p", defaultSortOrder = "asc",
    theme = reactableTheme(style = list(fontSize = 12)),
    elementId = "motif-pos-table"
  )
cp <- htmltools::em(
  glue(
    "
    All {sum(motif_pos_res$adj_p < motif_params$alpha)} TFBMs found to show
    positional bias using an adjusted p-value ({motif_adj}) below
    {motif_params$alpha} as the criteria for significance. Highly 
    similar motifs were clustered during preparation and are grouped accordingly.
    Clicking the arrow next to the arbitrary cluster id will reveal full details
    for all motifs within the cluster which were found to be significant, 
    including the sequence logo showing information content (ICM).
    "
  )
)
tl <- tagList(
  tags$button(
    tagList(fontawesome::fa("download"), "Download as CSV"),
    onclick = glue(
      "Reactable.downloadDataCSV('motif-pos-table', '{target}_position_res.csv')"
    )
  ),
  tags$button(
    "Expand/Collapse All",
    onclick = "Reactable.toggleAllRowsExpanded('motif-pos-table')"
  ),
  tags$button(
    "Clear All Filters",
    onclick = "Reactable.setAllFilters('motif-pos-table', [])"
  ),
  tbl,
  cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```


`r if (n_pos > 0) "#### Total Matches"`

```{r plot-total-pos-matches, eval = n_pos > 0, echo = n_pos > 0, fig.cap = glue("*All best matches by their distance from the sequence centre, without binning by position. Black lines represent the loess curve through all matches. The motif with the most matches within the enriched region is shown for each cluster, with the other motifs within the cluster indicated below in brackets.*")}
lb <- glue("{names(clust_names)}\n({clust_names})") %>% 
  setNames(names(clust_names))
top_matches <- motif_list %>% 
  dplyr::filter(altname %in% top_pos$altname) %>% 
  mutate(fct = factor(altname, levels = top_pos$altname)) %>% 
  arrange(fct) %>% 
  to_list() %>% 
  getPwmMatches(
    fw_seq, abs = motif_params$abs, best_only = TRUE, mc.cores = threads
  )
top_matches %>% 
  plotMatchPos(
    geom = "point", abs = motif_params$abs, use_totals = TRUE,
    binwidth = 1
  ) +
  geom_smooth(se = FALSE, colour = "black", linewidth = 1/2, method = 'loess') +
  facet_wrap(~name, labeller = as_labeller(lb)) +
  theme(legend.position = "none") +
  labs(x = "Distance From Sequence Centre", y = "Total Matches")
```

`r if (n_pos > 0) "#### Proportional Match Distributions"`

```{r plot-smoothed-pos, eval = n_pos > 0, echo = n_pos > 0, fig.height = 0.75 * fh, fig.cap = glue("*Best matches for the representative motif within each cluster, shown as (A) proportions and (B) as a cumulative distribution*")}
A <- top_matches %>% 
  plotMatchPos(
    se = FALSE, abs = motif_params$abs, linewidth = 1/2,
    binwidth = motif_params$binwidth / 2
  ) +
  labs(
    x = "Distance From Centre", y = "Proportion", colour = "Motifs In Cluster"
  ) +
  theme(legend.position = "none")
B <- top_matches %>% 
  plotMatchPos(
    type = "cdf", geom = "line", abs = motif_params$abs,
    linewidth = 1 / 2, binwidth = 1
  ) +
  labs(x = "Distance From Centre", y = NULL, colour = "Representative Motif")
A + B + plot_layout(guides = "collect") 
```

`r if (n_pos > 0) "#### Match Heatmap"`

```{r heatmap-pos, echo = n_pos > 0, fig.cap = glue("*The top-ranked motifs by positional bias showing the total number of matches within each bin. Motifs are clustered by similarity of totals within this figure when more than two are being displayed. A maximum of 25 motifs are shown.*")}
if (n_pos > 0) {
  top_pos <- motif_pos_res %>% 
    dplyr::filter(adj_p < motif_params$alpha) %>% 
    dplyr::slice(1:25) %>% 
    pull(altname)
  motif_list %>% 
    dplyr::filter(altname %in% top_pos) %>% 
    to_list() %>% 
    getPwmMatches(
      fw_seq, abs = motif_params$abs, best_only = TRUE, mc.cores = threads
    ) %>% 
    plotMatchPos(
      abs = motif_params$abs, type = "heatmap", cluster = n_pos > 2,
      use_totals = TRUE,
      # binwidth = floor(motif_params$binwidth / 2)
      binwidth = motif_params$binwidth 
    ) + 
    labs(x = "Bin", fill = "Total\n Matches")
}
```



## Relationship to Other Targets

```{r load-regioner}
localz_regions <- here::here(
  all_paths$peaks, "{all_targets}", "{all_targets}_regions_localz.rds"
) %>% 
  glue() %>%
  c(
    here::here(all_paths$nfr, target, "{target}_nfr_regions_localz.rds") %>% 
      glue()
  ) %>% 
  setNames(c(all_targets, glue("{target}_NFR"))) %>% 
  lapply(read_rds)
localz_df <- localz_regions %>%
  lapply(slot, "multiLocalZscores") %>% 
  lapply(pluck, "resumeTab") %>% 
  bind_rows(.id = "target") %>% 
  as_tibble() %>% 
  dplyr::rename(region = name) %>% 
  mutate(
    p_value = 1 - pchisq(norm_zscore^2, 1),
    adj.p_value = p.adjust(p_value, regioner_params$adj),
    region = case_when(
      region %in% names(regions) ~ regions[region],
      TRUE ~ region
    ) %>% 
      factor(all_regions),
    enrichment = log2(n_overlaps / mean_perm_test),
    target_peaks = map_int(consensus_peaks, length)[target],
    prop = n_overlaps / target_peaks
  ) %>% 
  dplyr::select(
    target, region, target_peaks, n_overlaps, prop, Z = norm_zscore, enrichment,
    ends_with("p_value"), ends_with("perm_test")
  ) 
shifted_z <- localz_regions %>% 
  lapply(slot, "multiLocalZscores") %>% 
  lapply(
    \(x) {
      tbl <- as_tibble(x$shifed_ZSs)
      tbl$shift <- x$shifts
      tbl
    }
  ) %>% 
  bind_rows(.id = "target") %>% 
  pivot_longer(
    cols = -all_of(c("target", "shift")), names_to = "region", values_to = "Z"
  ) %>% 
  mutate(
    n = map_int(consensus_peaks, length)[target],
    Z = Z / sqrt(n),
    region =  case_when(
      region %in% names(regions) ~ regions[region],
      TRUE ~ region
    ) %>% 
      factor(all_regions)
  ) 
## For regioneR analysis
heat_col <- c(
  rev(RColorBrewer::brewer.pal(9, "PuBuGn")),
  RColorBrewer::brewer.pal(9, "YlOrRd")
)
```

### Region Comparisons {.tabset}

The package `regionerReloaded` [@Malinverni2023-ee] was again used to compare the NFRs with defined regions, along with regions with detected signal from all targets.
The following analysis includes the NFRs as well as each target's set of peaks, as analysed in the individual summaries.
This provides an excellent frame of reference for comparison between targets, and the nature of the NFRs in this context.


#### Combined Heatmap

```{r zscore-heatmap, fig.cap = glue("*Normalised Z-scores showing the effect of shifting peaks for all targets and the NFRs, within a $\\pm$ {slot(localz_regions[[1]], 'parameters')$window / 1e3}kb window around the existing peaks.*")}
shifted_z %>% 
  arrange(abs(shift), desc(Z)) %>% 
  mutate(
    region = fct_inorder(region),
    source = ifelse(region %in% regions, "Region", "Feature")
  ) %>% 
  ggplot(aes(shift, region, fill = Z)) +
  geom_raster() +
  facet_grid(
    cols = vars(target), rows = vars(source), scales = "free", space = "free"
  ) +
  scale_x_continuous(
    expand = rep_len(0, 4), labels = \(x) x / 1e3, name = "Shift (kb)"
  ) +
  scale_y_discrete(expand = rep_len(0, 4), name = NULL) +
  scale_fill_gradientn(
    colours = heat_col,
    limits = c(-1.05, 1.05) * max(abs(shifted_z$Z))
  )
```

#### Results Table

```{r tbl-localz-regions}
tbl <- localz_df %>%
  arrange(p_value) %>% 
  dplyr::select(-sd_perm_test, -p_value) %>% 
  dplyr::select(target, region, target_peaks, mean_perm_test, everything()) %>% 
  mutate(
    source = str_extract(region, ".+: ") %>% 
      str_remove_all(": ") %>% 
      str_replace_na(""),
    region = str_remove_all(region, ".+: ")
  ) %>% 
  dplyr::select(target, source, region, everything()) %>% 
  dplyr::select(where(\(x) !all(x == ""))) %>% 
  dplyr::mutate(
    across(any_of("source"), \(x) str_replace(x, "$^", "GRAVI")),
    Sig = adj.p_value < 0.05
  ) %>% 
  reactable(
    filterable = TRUE, showPageSizeOptions = TRUE,
    columns = list(
      target = colDef(
        "Target",
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('localz-res-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px; font-size: 12px;"
          )
        }
      ),
      source = colDef(
        "Source",      
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('localz-res-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px; font-size: 12px;"
          )
        }
      ),
      region = colDef(
        "Region", minWidth = 120,
        style = function(value) {
          bg <- c(region_colours, unlist(colours$features))[[value]]
          list(backgroundColor = bg)
        },
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('localz-res-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px; font-size: 12px;"
          )
        }
      ),
      target_peaks = colDef(
        "N Peaks", format = comma_col, maxWidth = 80, filterMethod = js_greater
      ),
      mean_perm_test = colDef(
        "Expected", format = comma_col, filterMethod = js_greater
      ),
      n_overlaps = colDef(
        "Observed", format = comma_col, maxWidth = 85,
        filterMethod = js_greater
      ),
      prop = colDef(
        "% Peaks", format = percent_col, maxWidth = 80, 
        filterMethod = js_greater
      ),
      Z = colDef(
        format = colFormat(digits = 2), maxWidth = 80,
        style = \(value) list(color = ifelse(value > 0, "black", "red")),
        filterMethod = js_greater
      ),
      enrichment = colDef(
        "log<sub>2</sub> Enrichment", format = colFormat(digits = 2),
        html = TRUE, 
        style = \(value) list(color = ifelse(value > 0, "black", "red")),
        filterMethod = js_greater
      ),
      adj.p_value = colDef(
        glue("p<sub>{enrich_params$adj}</sub>"), html = TRUE,
        cell = \(value) sprint_pval(value), maxWidth = 80,
        style = \(value) list(
          color = ifelse(
            value < regioner_params$alpha,
            "black", colours$direction[["unchanged"]]
          )
        ),
        filterMethod = js_less
      ),
      Sig = colDef(
        show = FALSE,
        filterMethod = JS("function(rows, columnId, filterValue) {
          if (filterValue === true) {
            return rows.filter(function(row) {
              const OnlySignificant = row.values[columnId]
              return OnlySignificant
            })
          }
          return rows
        }")
      )
    )[colnames(.)],
    columnGroups = list(
      colGroup("Overlaps", columns = c("mean_perm_test", "n_overlaps", "prop"))
    ),
    theme = reactableTheme(style = list(fontSize = 12)),
    elementId = "localz-res-table"
  )
cp <- htmltools::em(
  glue(
    "
    Results for association between all peaks, including NFRs and the pre-defined 
    genomic regions{ifelse(has_features, ', including the supplied external features.', '.')}
    Expected overlaps are those obtained through random resampling of the genome
    to match the consensus peaks. Enrichment is given taking log2 of 
    the ratio between observed and expected overlapping peaks.
    "
  )
)
tl <- tagList(
  tags$button(
    tagList(fontawesome::fa("download"), "Download as CSV"),
    onclick = glue(
      "Reactable.downloadDataCSV('localz-res-table', 'shared_localz_res.csv')"
    )
  ),
  tags$button(
    "Clear All Filters",
    onclick = "Reactable.setAllFilters('localz-res-table', [])"
  ),
  tags$label(
    tags$input(
      type = "checkbox", 
      onclick = "Reactable.setFilter('localz-res-table', 'Sig', event.target.checked)"
    ),
    "Only show significant associations"
  ),
  tbl,
  cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```

### Comparison Between Targets {.tabset}

```{r load-localz-peaks}
mlz_list <- all_paths$nfr %>% 
  here::here(target, glue("{target}_nfr_targets_localz.rds")) %>% 
  read_rds() 
dist_df <- mlz_list %>% 
  lapply(
    function(x) {
      mlz <- slot(x, "multiLocalZscores")
      tbl <- as_tibble(mlz$shifed_ZSs)
      tbl$shift <- mlz$shifts
      tbl
    }
  ) %>% 
  bind_rows(.id = "Query") %>% 
  pivot_longer(
    cols = any_of(names(consensus_peaks)), names_to = "Target", values_to = "Z"
  ) %>% 
  dplyr::filter(!is.na(Z)) %>% 
  mutate(
    Target = Target %>% 
      str_replace_all(target, paste0(target, "_NFR")) %>% 
      factor(levels = names(consensus_peaks)) %>% 
      fct_rev(),
    Query = Query %>% 
      str_replace_all(target, paste0(target, "_NFR")) %>% 
      factor(levels = names(consensus_peaks)) %>% 
      fct_rev(),
    Z = Z / sqrt(map_int(consensus_peaks, length)[Query]),
  ) %>% 
  droplevels()
pairwise_df <- mlz_list %>% 
  lapply(slot, "multiLocalZscores") %>% 
  lapply(pluck, "resumeTab") %>% 
  bind_rows(.id = "Query") %>% 
  dplyr::rename(Target = name) %>% 
  mutate(
    Query = str_replace_all(Query, target, paste0(target, "_NFR")),
    Target = str_replace_all(Target, target, paste0(target, "_NFR")),
    n_regionA = map_int(consensus_peaks, length)[Query],
    n_regionB = map_int(consensus_peaks, length)[Target],
    prop = n_overlaps / map_int(consensus_peaks, length)[Query],
    enrichment = log2(n_overlaps / mean_perm_test),
    p_value = 1 - pchisq(norm_zscore^2, 1),
    adj.p_value = p.adjust(p_value, regioner_params$adj)
  ) 
```


The NFRs were compared to the set of peaks derived from each target to check for any relationship between their locations on the genome.
Shifted Z-Scores indicate proximity and fine-grained relative binding patterns, with additional Z-scores derived from any enrichment of direct overlaps above the expected values used to determine the statistical significance of these overlaps compared to random expectations.

#### Shifted Z-Scores

```{r plot-shiftedz, fig.cap = glue("*Distributions of shifted Z-scores for each target relative to NFRs. Narrow peaks indicate close proximity of overlaps, whilst dips around zero indicate very close by not overlapping signal, as may be observed with histone marks and associated transcription factors.*")}
dist_df %>% 
  arrange(desc(Z)) %>%
  mutate(
    Query = fct_inorder(Query),
    Target = fct_inorder(Target)
  ) %>%
  ggplot(aes(shift, Z, colour = Target)) +
  geom_line() +
  facet_wrap(~Query) +
  scale_x_continuous(
    labels = \(x) x * 1e-3, name = "Shift (kb)"
  ) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  scale_colour_manual(values = colours$target)
```


#### Overlap Heatmap

```{r plot-overlap-heatmap, fig.cap = glue("*Heatmap showing the strength of overlaps between each target and the alternative targets, here represented as the Query and Target respectively. Fill colours are derived from Z-scores and indicate the significance of overlaps, with asterisks indicating a formal result passing statistical significance as detailed in the reults table. The proportion of peaks from each query and alternative targets is also shown within each tile.*")}
pairwise_df %>% 
  ggplot(aes(Target, fct_rev(Query), fill = norm_zscore)) +
  geom_raster() +
  geom_label(
    aes(label = prop),
    data = . %>% 
      mutate(
        prop = percent(prop, 0.1),
        prop = case_when(
          adj.p_value < enrich_params$alpha ~ paste0(prop, "*"),
          TRUE ~ prop
        )
      ),
    fill = "white", alpha = 0.6
  ) +
  scale_x_discrete(expand = expansion(0, 0)) +
  scale_y_discrete(expand = expansion(0, 0)) +
  scale_fill_gradientn(
    colours = heat_col,
    limits = c(-1.05, 1.05) * max(abs(pairwise_df$norm_zscore)),
  ) +
  labs(y = "Query", fill = "Z") +
  theme(panel.grid = element_blank())
```


#### Results Table

```{r pairwise-tbl}
tbl <- pairwise_df %>%
  dplyr::select(
    Query, Target, n_regionA, n_regionB, prop, n_overlaps, mean_perm_test, 
    enrichment, norm_zscore, adj.p_value
  ) %>% 
  arrange(adj.p_value) %>% 
  mutate(Sig = adj.p_value < 0.05) %>% 
  reactable(
    showPageSizeOptions = TRUE, filterable = TRUE,
    columns = list(
      Query = colDef(
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('pairwise-res-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px; font-size: 12px;"
          )
        }
      ),
      Target = colDef(
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('pairwise-res-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px; font-size: 12px;"
          )
        }
      ),
      n_regionA = colDef("Query", format = comma_col),
      n_regionB = colDef(
        "Target", format = comma_col
      ),
      prop = colDef(
        "% Query", format = colFormat(percent = TRUE, digits = 1),
        maxWidth = 80
      ),
      n_overlaps = colDef("Observed", format = comma_col, maxWidth = 80),
      mean_perm_test = colDef("Expected", format = comma_col, maxWidth = 80),
      enrichment = colDef(
        "log<sub>2</sub> Enrichment", format = colFormat(digits = 2),
        html = TRUE,
        style = \(value) {
          text_colour = ifelse(value < 0, colours$direction$decreased, "black")
          list(color = text_colour)
        }
      ),
      norm_zscore = colDef(
        "Z", format = colFormat(digits = 2), maxWidth = 80,
        style = \(value) {
          text_colour = ifelse(value < 0, colours$direction$decreased, "black")
          list(color = text_colour)
        }
      ),
      adj.p_value = colDef(
        glue("p<sub>{enrich_params$adj}</sub>"), maxWidth = 80, html = TRUE,
        cell = \(value) sprint_pval(value),
        style = \(value) {
          text_colour = ifelse(
            value < regioner_params$alpha,
            "black", colours$direction[["unchanged"]]
          )
          list(color = text_colour)
        }
      ),
      Sig = colDef(
        show = FALSE,
        filterMethod = JS("function(rows, columnId, filterValue) {
          if (filterValue === true) {
            return rows.filter(function(row) {
              const OnlySignificant = row.values[columnId]
              return OnlySignificant
            })
          }
          return rows
        }")
      )
    ),
    columnGroups = list(
      colGroup(name = "Nbr. Regions", columns = str_subset(names(.), "n_reg")),
      colGroup(
        name = "Overlaps", columns = c("prop", "n_overlaps", "mean_perm_test")
      )
    ),
    theme = reactableTheme(style = list(fontSize = 12)),
    elementId = 'pairwise-res-table'
  )
cp <- htmltools::em(
  glue(
    "
    Results showing significance of the overlap between peaks from each target 
    and the other targets, including NFRs, defined here as the Query and Target respectively. 
    Expected values represent those estimated using permutations, with the log2 
    ratio between the two values expressed as enrichment
    "
  )
)
tl <- tagList(
  tags$button(
    "Clear All Filters",
    onclick = "Reactable.setAllFilters('pairwise-res-table', []);"
  ),
  tags$label(
    tags$input(
      type = "checkbox", 
      onclick = "Reactable.setFilter('pairwise-res-table', 'Sig', event.target.checked)"
    ),
    "Only show significant associations"
  ), tbl, cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```

## Example NFRs {.tabset}

```{r grl_to_plot}
grl_to_plot <- consensus_peaks[all_targets] %>% 
  endoapply(subsetByOverlaps, nfr_consensus) %>% 
  endoapply(filter, score == max(score)) %>% 
  .[!duplicated(unlist(.))]
```

```{r plot-objects}
cb <- config$genome$build %>% 
  str_to_lower() %>% 
  paste0(".cytobands") 
data(list = cb)
bands_df <- get(cb)
stopifnot(
  colnames(bands_df) == c("chrom", "chromStart", "chromEnd", "name", "gieStain")
)
## The coverage
bwfl <- all_samples %>% 
  distinct(target, treat) %>% 
  split(.$target) %>% 
  lapply(
    \(x) {
      file.path(
        all_paths$macs2, x$target, 
        paste(x$target, x$treat, "merged_treat_pileup.bw", sep = "_")
      ) %>% 
        BigWigFileList() %>% 
        setNames(x$treat)
    }
  )
line_col <- lapply(bwfl, function(x) unlist(colours$treat[names(x)]))
## The features track
feat_gr <- gene_regions %>% 
  lapply(granges) %>% 
  GRangesList()
feature_colours <- colours$regions
featnames <- "Regions"
if (has_features) {
  feat_gr <- list(Regions = feat_gr) %>% 
    c(
      external_features %>% 
        lapply(\(x) endoapply(splitAsList(x, x$feature), granges))
    )
  feature_colours <- feat_gr %>% lapply(\(x) full_palette[names(x)])
  featnames <- names(feat_gr) %>% str_sep_to_title(replacement = "\n")
}
## The genes track defaults to transcript models
hfgc_genes <- read_rds(
  here::here("output", "annotations", "trans_models.rds")
)
gene_col <- "grey"
collapseTrans <- "auto"
if (has_rna) {
  hfgc_genes <- hfgc_genes %>% 
    mutate(
      Detected = ifelse(
        gene %in% bind_rows(rna)$gene_id, "Detected", "Undetected"
      )
    ) %>% 
    split(.$Detected)
  gene_col <- c(
    Detected = colours$direction[["unchanged"]],
    Undetected = colours$direction[["undetected"]]
  )
  collapseTrans <- list(Detected = "auto", Undetected = "meta")
}
## Ad the NFRs as coverage annotations
annot <- lapply(
  all_targets, 
  \(x) {
    grl <- GRangesList()
    seqinfo(grl) <- sq
    grl
  }
)
names(annot) <- all_targets
annot[[target]] <- GRangesList(NFR = nfr_consensus)
## Define the y-limts for consistency across plots
y_lim <- bwfl %>%
  lapply(
    function(x) {
      gr <- unlist(grl_to_plot) %>% 
        resize(width = 30 * width(.), fix = 'center')
      x %>% 
        lapply(import.bw, which = gr) %>%
        lapply(function(rng) c(0, max(rng$score))) %>%
        unlist() %>%
        range()
    }
  )
```

For each target, the peak with the highest score, and which overlapped one or more NFRs was plotted for the purposes of visual inspection.
Gene-centric regions and any additional annotations were also included.
`r if (has_rna) "Genes were split into detected and undetected based on the supplied transcriptomic data."`

```{r plot-nfrs, echo=FALSE}
.makeCaption <- function(.gr, target) {
  if (is.null(.gr)) return(NULL)
  reg <- case_when(
    str_detect(.gr$region, "Inter") ~ paste("an", .gr$region, "region"),
    str_detect(.gr$region, "Upstream") ~ paste("an", .gr$region),
    str_detect(.gr$region, "(Ex|Intr)on") ~ paste("an", .gr$region),
    str_detect(.gr$region, "^Prom") ~ paste("a", .gr$region)
  )
  feat <- c()
  feat <- unlist(.gr$feature) %>% 
    sort() %>% 
    unique() %>% 
    glue_collapse(sep = ", ", last = " & ")
  add_feat <- has_features & !str_detect(feat, 'no_feature')
  if (add_feat) 
    feat <- case_when(
      str_detect(feat, "^[AEIOUaeiou]") ~ paste("an", feat),
      !str_detect(feat, "^[AEIOUaeiou]") ~ paste("a", feat)
    )
  gn <- unlist(.gr$gene_name)
  cp <- c(
    glue(
      "The {width(.gr)}bp region with the highest score for {target} signal, 
      and which overlaps one or more NFRs, highlighted as black bars in the 
      coverage tracks. The range mostly overlapped {reg}, 
      with all defined regions shown as a contiguous bar in the upper panel. ",
      ifelse(
        add_feat,
        glue(
          "Using the features supplied this mostly overlapped {feat}, shown as 
          a separate track beneath the gene-centric regions. "
        ),
        glue("")
      )
    ),
    ifelse(
      length(gn) > 0,
      paste0(
        "Using the above mapping strategy, this range is likely to regulate ",
        collapseGenes(gn, format = ""), "."
      ),
      "No genes were able to be assigned to this region."
    ),
    paste(
      "For each sample, the y-axis limits represent the values from the window", 
      "with the highest signal."
    )
  )
  paste(cp, collapse = " ")
}
htmltools::tagList(
  lapply(
    names(grl_to_plot),
    \(i) {
      ## Export the image
      img_out <- file.path(fig_path, glue("{i}_NFR.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      plotHFGC(
        grl_to_plot[[i]],
        hic = hic,
        features = feat_gr, featcol = feature_colours, featname = featnames,
        featsize = max(1, 0.8 * length(feat_gr)), featstack = "dense",
        genes = hfgc_genes, genecol = gene_col, 
        genesize = rep_len(c(2, 1), length(hfgc_genes)),
        coverage = bwfl, linecol = line_col, ylim = y_lim,
        annotation = annot, annotcol = list(NFR = "black"),
        cytobands = bands_df,
        collapseTranscripts = collapseTrans, maxTrans = 15,
        zoom = 10, max = max(
          10 * max(nfr_params$nfr_width), 
          width(range(grl_to_plot[[i]]))
        ),
        col.title = "black", background.title = "white", 
        rotation.title = 90, showAxis = FALSE, fontsize = 10,
        highlight = rgb(0, 0, 1, 0.3)
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(.makeCaption(grl_to_plot[[i]], i))
      tag <- glue("{i} Peak")
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level3",
          htmltools::h3(tag),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  )
)
```



## References

<div id="refs"></div>

<br>
<button type="button" class="btn btn-default btn-sessioninfo" data-toggle="collapse" data-target="#sessioninfo" style="display: block;">
<span class="glyphicon glyphicon-info-sign" aria-hidden="true"></span> Session information
</button>
</p>
<div id="sessioninfo" class="collapse">
```{r session-info, echo=FALSE}
pander::pander(sessionInfo())
```
</div>
