```{r set-knitr-opts, echo=FALSE, child = here::here('analysis/setup_chunk.Rmd')}
```


```{r get-params}
target <- params$target
threads <- parallel::detectCores() - 1
```

```{r packages}
library(tidyverse)
library(yaml)
library(pander)
library(glue)
library(extraChIPs)
library(plyranges)
library(scales)
library(patchwork)
library(parallel)
library(reactable)
library(htmltools)
library(universalmotif)
library(corrplot)
library(motifTestR)
library(rlang)
source(here::here("workflow/scripts/custom_functions.R"))
source(here::here("workflow/scripts/table_funs.R"))
theme_set(theme_bw())
config <- read_yaml(here::here("config", "config.yml"))
all_paths <- here::here("config", "json", "all_paths.json") %>% 
    jsonlite::fromJSON() %>% 
    lapply(here::here)
ucsc <- get_ucsc(config$genome$build)
bs_pkg <- glue("BSgenome.{ucsc$sp}.UCSC.{ucsc$build}")
library(bs_pkg, character.only = TRUE)
```

```{r load-samples}
samples <- all_paths$macs2 %>%
  file.path(target, glue("{target}_qc_samples.tsv")) %>%
  read_tsv()
treat_levels <- unique(samples$treat)
all_targets <- here::here(config$samples) %>% 
  read_tsv() %>% 
  pull("target") %>% 
  unique()
all_samples <- all_paths$macs2 %>% 
  file.path("{all_targets}", "{all_targets}_qc_samples.tsv") %>% 
  glue() %>% 
  lapply(read_tsv) %>% 
  bind_rows() %>% 
  dplyr::filter(qc == "pass") %>% 
  dplyr::select(-all_of("qc"))
all_targets <- unique(all_samples$target)
stopifnot(target %in% all_targets)
n_targets <- all_targets
```

```{r config-params-paths}
all_params <-  here::here("config", "params.yml") %>% read_yaml() 
region_params <- all_params$gene_regions
regioner_params <- all_params$regioner
motif_params <- here::here("config", "json", "motif_analysis_param.json") %>%
  jsonlite::fromJSON() %>%
  pluck(target)
knitr_params <- here::here("config", "rmarkdown.yml") %>% 
  read_yaml() %>% 
  pluck("knitr_opts")
```

```{r figure-params}
fh <- knitr_params$fig.height
fw <- knitr_params$fig.width
fig_path <- here::here("docs", "assets", target)
if (!dir.exists(fig_path)) dir.create(fig_path, recursive = TRUE)
fig_type <- knitr_params$dev[[1]]
fig_fun <- match.fun(fig_type)
if (fig_type %in% c("bmp", "jpeg", "png", "tiff")) {
  ## These figure types require dpi & resetting to be inches
  formals(fig_fun)$units <- "in"
  formals(fig_fun)$res <- 300
}
```



```{r load-annotations}
sq <- file.path(all_paths$annotation, "seqinfo.rds") %>% read_rds()
gene_regions <- file.path(all_paths$annotation, "gene_regions.rds") %>% 
  read_rds()
regions <- map_chr(gene_regions, \(x) x$region[1])
colours <- file.path(all_paths$annotation, "colours.rds") %>% read_rds()
region_colours <- unlist(colours$regions) %>% 
  setNames(map_chr(gene_regions, \(x) x$region[[1]])[names(.)])
```


## Introduction

Nucleosome Free Regions (NFRs) were identified for `r target` using the set of filtered peaks obtained for each treatment (`r pander(treat_levels)`).
Peaks within 300bp of each other were merged into a single peak, with this distance approximating the distance occupied by 2 nucleosomes, and approximating broader peaks that may have been otherwise called by `macs2` [@Zhang18798982].
NFRs within merged peaks were called using Histrader [@Yan2020-nz] for each treatment group with the following settings.

- Any merged peaks <500bp were excluded as being too short
- Any NFRs beyond 1kb were discarded as not being viable NFRs
- Regions of the peak with signal below 10% of the maximal signal were also not assessed for NFRs as these were more likely to be regions on the other edges of wider peaks.

Treatment-specific NFRs were merged as consensus NFRs, taking only shared regions between treatment groups which were wider than 75bp and which were identified in all groups.
Given the inclusive nature of HisTrader, this is a more stringent approach to ensure any overly-inclusive NFR calls were minimised.

Whilst a motif enrichment analysis was performed, along with an assessement of the relationship between NFRs and other targets, no gene-set enrichment testing was performed.
The mapping from NFR to genes would very closely replicate that of the larger set of `r target` peaks, and as such, this was considered to be a redundant analysis.


```{r load-nfr}
nfr_by_treat <- all_paths$nfr %>% 
  file.path(target, "{target}_{treat_levels}.nfr.bed.gz") %>% 
  glue() %>% 
  importPeaks(type = "bed", seqinfo = sq) %>% 
  setNames(treat_levels)
nfr_consensus <- all_paths$nfr %>% 
  file.path(target, "{target}_consensus_nfr.rds") %>% 
  glue() %>% 
  read_rds()
consensus_peaks <- all_paths$peaks %>% 
  file.path("{target}", "{target}_consensus_peaks.rds") %>% 
  glue() %>% 
  read_rds()
has_features <- "feature" %in% colnames(mcols(consensus_peaks))
```

## NFR Summary {.tabset}

### Shared NFRs

```{r plot-overlaps, fig.width=fh, fig.cap = "*Summary of NFRs identified in each treatment group separately, before identifying consensus NFRs*"}
plotOverlaps(nfr_by_treat, set_col = unlist(colours$treat[treat_levels]))
```

### Peak & NFR Width

```{r plot-width, fig.cap = glue("*A) Relationship between {target} peak width and NFR identification. {target} peaks within 300bp were merged prior to plotting, and B) NFR width and numbers by treatment group.*")}
A <- consensus_peaks %>% 
  reduce(min.gapwidth = 300) %>% 
  mutate(has_nfr = overlapsAny(., nfr_consensus)) %>% 
  split(.$has_nfr) %>% 
  plotGrlCol(
    .id = 'has_nfr', fill = 'has_nfr', q = 0, geom = "violin",
    draw_quantiles = 0.5, trim = FALSE
  ) +
  scale_y_log10(labels = comma) +
  scale_fill_manual(values = c("indianred", "forestgreen")) +
  labs(
    x = "NFR Detected",
    y = glue("{target} Peak Width (bp)")
  )
B <- nfr_by_treat %>% 
  c(GRangesList(`Consensus NFR` = nfr_consensus)) %>% 
  plotGrlCol(
    df = tibble(treat = treat_levels), .id = "treat", fill = treat,
    geom = 'violin', q = 0, draw_quantiles = 0.5, trim = FALSE
  ) +
  scale_fill_manual(values = unlist(colours$treat)) +
  scale_y_log10(labels = comma) +
  labs(
    x = "NFR Group",
    y = "NFR Width (bp)"
  ) 
A + B + plot_annotation(tag_levels = "A") & theme(legend.position = 'none')
```

### NFRs Per Peak

```{r plot-nfr-by-peak-width, fig.cap = glue("*Number of NFRs found within each {target} peak, separated by genomic region. {target} peaks within 300bp were merged prior ro plotting*")}
consensus_peaks %>% 
  reduce(min.gapwidth = 300) %>% 
  mutate(
    w = width, 
    region = regions[bestOverlap(., gene_regions)] %>% factor(levels = regions),
    nfr = countOverlaps(., nfr_consensus)
  ) %>% 
  as_tibble() %>% 
  ggplot(aes(w, nfr)) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, method = "lm") +
  geom_label(
    aes(w, nfr, label = lab) ,
    data = . %>% 
      summarise(
        n = sum(nfr), 
        w = min(w), nfr = max(nfr),
        .by = "region"
      ) %>% 
      mutate(
        lab = glue("n = {comma(n, 1)}"),
        w = min(w), nfr = max(nfr)
      ),
    hjust = 0, vjust = 1
  ) +
  facet_wrap(~ region) +
  scale_x_continuous(labels = comma_format(scale = 1e-3)) +
  labs(
    x = glue("{target} Merged Peak Width (kb)"),
    y = "NFR / Peak"
  )
```

### Distance Between NFRs

```{r plot-dist, fig.cap = glue("*Distances between each NFR and the nearest upstream NFR within each broader {target} peak*")}
nfr_consensus %>% 
  granges() %>% 
  join_overlap_inner(
    consensus_peaks %>% 
      reduce(min.gapwidth = 300) %>% 
      mutate(id = seq_along(.))
  ) %>% 
  as_tibble(rangeAsChar = FALSE) %>% 
  split(.$id) %>% 
  .[map_int(., nrow) > 1] %>% 
  mclapply(
    \(x) x$start[-1] - x$end[-nrow(x)], 
    mc.cores = threads
  ) %>% 
  unlist() %>% 
  as_tibble_col("distance") %>% 
  ggplot(aes(distance, after_stat(count))) +
  geom_histogram(
    fill = "grey70", colour = 'black', bins = 100
  ) +
  scale_x_log10() +
  scale_y_continuous(expand = expansion(c(0, 0.05)), labels = comma) +
  labs(
    x = "Distance between NFRs", y = "Totals"
  )
```

### NFRs By Region

```{r plot-pie, fig.width = fh, fig.cap = glue("*Distribution of NFRs by genomic region, noting that wider {target} peaks will accumulate more NFRs*")}
nfr_consensus %>% 
  plotPie(
    fill = region, hole_width = 0.2,
    total_size = 4,
    cat_glue = "{str_wrap(.data[[fill]], 15)}\n{comma(n, 1)}\n({percent(p, 0.1)})",
    cat_alpha = 0.7, cat_adj = 0.1
  ) +
  scale_fill_manual(values = region_colours) +
  theme(legend.position = "none")
```

`r if (has_features) "### NFRs By Region and Feature"`

```{r plot-donut, echo = has_features, fig.width = fh, fig.cap = 'Breakdown of NFRs by provided features and genomic region.'}
if (has_features) {
  nfr_consensus %>% 
    plotSplitDonut(
      inner = 'feature', outer = 'region',
      inner_palette = unlist(colours$features),
      outer_palette = region_colours,
      inner_glue = "{str_wrap(.data[[inner]], 15)}\n{comma(n)}\n{percent(p,0.1)}",
      outer_glue = "{str_wrap(.data[[outer]], 15)}\n{comma(n)}\n{percent(p,0.1)}",
      outer_min_p = 0.03
    )
}
```


## Motif Analysis

```{r load-motif-res}
bs_genome <- get(bs_pkg)
fw_peaks <- resize(
  nfr_consensus, width = motif_params$peak_width, fix = "center"
)
genome(fw_peaks) <- ucsc$build
fw_seq <- getSeq(bs_genome, fw_peaks)
motif_adj <- match.arg(motif_params$adj, p.adjust.methods)
motif_list <- file.path(all_paths$annotation, "motif_list.rds") %>% 
  read_rds() %>% 
  to_df %>% 
  as_tibble()
motif_uri <- file.path(all_paths$annotation, "motif_uri.rds") %>% 
  read_rds()
motif_pos_res <- all_paths$nfr %>%
  file.path("{target}", "{target}_motif_position.tsv.gz") %>% 
  glue() %>% 
  read_tsv() %>% 
  dplyr::select(-fdr) %>% 
  mutate(adj_p = p.adjust(p, motif_adj))
n_pos <- sum(motif_pos_res$adj_p < motif_params$alpha)
motif_enrich_res <- all_paths$nfr %>% 
  file.path("{target}", "{target}_motif_enrichment.tsv.gz") %>%
  glue() %>% 
  read_tsv() %>% 
  dplyr::select(-fdr) %>% 
  mutate(adj_p = p.adjust(p, motif_adj))
n_enrich <- sum(
  motif_enrich_res$adj_p < motif_params$alpha & motif_enrich_res$Z > 0
)
```

Using the motifs defined during preparation, testing was performed for 1) enrichment above background and 2) any positional bias within the tested sequences.
For this step peaks were set to a fixed width of `r motif_params$peak_width`bp, with sequences taken from these regions.

### Motif Enrichment {.tabset}

A simple enrichment analysis was performed by selecting a set of `r comma(motif_enrich_res$sequences[[1]] * motif_enrich_res$n_iter[[1]])` *background sequences* matched by genomic regions to the `r comma(motif_enrich_res$sequences[[1]])` *test sequences*.
This gave a set of genomically matched control sequences to gauge enrichment by.
Analysis was performed using the `motifTestR` package and the `r motif_params$model` model.

Whilst statistical testing was two-sided, only enriched motifs (i.e. observed > expected) are shown in the results below, with `r n_enrich` of the `r comma(nrow(motif_list))` included motifs matching the criteria for being considered as enriched.

`r if (n_enrich > 0) "#### Enrichment Table"`

```{r tbl-res-enrich, eval = n_enrich > 0, echo = n_enrich > 0}
tbl <- motif_enrich_res %>%
  arrange(p) %>% 
  dplyr::filter(adj_p < motif_params$alpha, Z > 0) %>% 
  mutate(cluster = fct_inorder(as.character(cluster))) %>% 
  dplyr::select(
    cluster, ends_with("name"), Z, p, adj_p, matches, expected, enrichment
  ) %>% 
  mutate(
    PWM = lapply(altname, \(x) motif_uri[[x]])
  ) %>% 
  reactable(
    groupBy = "cluster",
    showPageSizeOptions = TRUE, filterable = TRUE,
    theme = reactableTheme(style = list(fontSize = 12)),
    defaultSorted = "adj_p", defaultSortOrder = "asc",
    columns = list(
      cluster = colDef("Cluster", maxWidth = 70),
      altname = colDef(
        "Motif", minWidth = 180, aggregate = "unique",
        style = list(borderLeft = "1px solid rgba(0, 0, 0, 0.1)")
      ),
      name = colDef(
        "Name", minWidth = 120, aggregate = "unique",
        style = list(borderRight = "1px solid rgba(0, 0, 0, 0.1)")
      ),
      matches = colDef(
        "Total", maxWidth = 60, format = comma_col, aggregate = "max",
        filterMethod = js_greater
      ),
      expected = colDef(
        "Expected", maxWidth = 80, format = comma_col, aggregate = "max",
        filterMethod = js_greater
      ),
      enrichment = colDef(
        "Enrichment", maxWidth = 90, format = colFormat(digits = 3),
        aggregate = "max", filterMethod = js_greater,
        style = list(borderRight = "1px solid rgba(0, 0, 0, 0.1)")
      ),
      Z = colDef(
        maxWidth = 60, format = colFormat(digits = 2), aggregate = "max",
        filterMethod = js_greater
      ),
      p = colDef(show = FALSE),
      adj_p = colDef(
        glue("p<sub>{motif_adj}</sub>"), html = TRUE, maxWidth = 70,
        cell = \(value) sprint_pval(value), aggregate = "min",
        format = colFormat(digits = 3), filterMethod = js_less
      ),
      PWM = colDef(
        name = "IC Matrix", minWidth = 180, filterable = FALSE,
        cell = function(value) tags$img(src = value, height = '80px')
      )
    ),
    elementId = "motif-enrich-table"
  )
cp <- htmltools::em(
  glue(
    "All {n_enrich} motifs considered as enriched relative to the background set
    of sequences, using the {motif_params$model} model and an adjusted 
    p-value ({motif_adj}) < {motif_params$alpha} as the criteria for 
    significance. Motifs are grouped by clusters formed earlier, with all motifs
    visible by clicking the arrow to the arbitrary cluster identifier. All 
    columns are searchable and filterable."
  )
)
tl <- tagList(
  tags$button(
    tagList(fontawesome::fa("download"), "Download as CSV"),
    onclick = glue(
      "Reactable.downloadDataCSV('motif-enrich-table', '{target}_motif_enrich_res.csv')"
    )
  ),
  tags$button(
    "Expand/Collapse All",
    onclick = "Reactable.toggleAllRowsExpanded('motif-enrich-table')"
  ),
  tags$button(
    "Clear All Filters",
    onclick = "Reactable.setAllFilters('motif-enrich-table', [])"
  ),
  tbl,
  cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```

`r if (n_enrich > 1) "#### Comparison of Motif Co-Occurrence"`

```{r enrich-corrplot, eval = n_enrich > 1, echo = n_enrich > 1, fig.width = 0.8 * fw, fig.height = 0.8 * fw, fig.cap = glue("*The top {length(top_enrich)} motifs ranked for enrichment were compared to each other using correlations derived from simple presence/absence calls. Highly correlated motifs may indicate similarity between TFBMs and the same sequence match being identified. Motifs detected in all sequences were excluded after the initial selction and as such, fewer than the selected {length(top_enrich)} motifs may be shown*")}
top_enrich <- motif_enrich_res %>%
  dplyr::filter(Z > 0, adj_p < motif_params$alpha) %>% 
  dplyr::slice(1:20) %>% 
  pull("altname")
motif_list %>% 
  dplyr::filter(altname %in% top_enrich) %>% 
  to_list() %>% 
  getPwmMatches(fw_seq, mc.cores = threads) %>% 
  lapply(\(x) tibble(seq = unique(x$seq))) %>% 
  bind_rows(.id = "altname") %>%
  distinct(altname, seq) %>% 
  left_join(
    dplyr::select(motif_enrich_res, altname, name), by = "altname"
  ) %>% 
  mutate(detected = TRUE) %>% 
  dplyr::select(seq, name, detected) %>% 
  pivot_wider(
    names_from = "name", values_from = "detected", values_fill = FALSE
  ) %>% 
  dplyr::select(-seq) %>% 
  dplyr::select(where(\(x) mean(x) < 1 & mean(x) > 0)) %>%
  cor(method = "spearman") %>% 
  corrplot(
    addCoefasPercent = TRUE, addCoef.col = "grey20", 
    tl.cex = 0.75, number.cex = 0.8, order = "hclust"
  )
```



### Positional Bias {.tabset}

```{r get-top-pos, eval = n_pos > 0, echo = n_pos > 0}
top_pos <- motif_pos_res %>% 
  dplyr::filter(adj_p < motif_params$alpha) %>% 
  dplyr::filter(matches_in_region == max(matches_in_region), .by = cluster) %>% 
  dplyr::slice(1:9) %>% 
  dplyr::select(altname, name, cluster)
clust_names <- motif_pos_res %>% 
  dplyr::filter(
    adj_p < motif_params$alpha,
    cluster %in% top_pos$cluster
  ) %>% 
  arrange(desc(matches_in_region)) %>% 
  summarise(
    altname = altname[[1]], .by = cluster,
    name = paste(name, collapse = "/")
  ) %>% 
  mutate(name = setNames(name, altname)) %>% 
  pull(name)
```

Positional bias was tested using the package `motifTestR`, bin-sizes of `r motif_params$binwidth`bp and testing `r ifelse(motif_params$abs, "distance from sequence centres", "position either side of the sequence centres")`.
Any motifs which matched fewer than `r percent(motif_params$ignore_below)` of sequences were not tested for any positional bias.

Using an adjusted p-value (`r motif_adj`) less than `r motif_params$alpha` to be considered as significant, `r n_pos` motifs were considered as showing any type of positional bias.

It should also be noted that this type of analysis is agnostic to the number of matches found for a TFBM, and only assesses any bias within the distribution of matches.


`r if (n_pos > 0) "#### Results Table"`

```{r tbl-res-pos, eval = n_pos > 0, echo = n_pos > 0}
tbl <- motif_pos_res %>%
  dplyr::filter(adj_p < motif_params$alpha) %>% 
  mutate(cluster = fct_inorder(as.character(cluster))) %>% 
  dplyr::select(
    cluster, contains("name"), ends_with("p"), start, end, centre, width,
    contains("total"), matches_in_region, starts_with("e"), 
  ) %>% 
  mutate(
    PWM = lapply(altname, \(x) motif_uri[[x]])
  ) %>% 
  reactable(
    groupBy = "cluster", showPageSizeOptions = TRUE, filterable = TRUE,
    columns = list(
      cluster = colDef(maxWidth = 60),
      name = colDef("Name", aggregate = "unique", maxWidth = 80),
      altname = colDef("Motif", aggregate = "unique", minWidth = 180),
      start = colDef(
        "Start", aggregate = "min", maxWidth = 50,
        filterMethod = js_greater
      ),
      end = colDef(
        "End", aggregate = "max", maxWidth = 50, 
        filterMethod = js_greater
      ),
      centre = colDef(
        "Centre", aggregate = "median", maxWidth = 60, show = FALSE
      ),
      width = colDef(
        "Width", aggregate = "median", maxWidth = 60, 
        filterMethod = js_less
      ),
      total_matches = colDef(
        "Total", aggregate = "max", maxWidth = 60, format = comma_col,
        filterMethod = js_greater
      ),
      matches_in_region = colDef(
        "Region", aggregate = "max", maxWidth = 60, 
        filterMethod = js_greater
      ),
      expected = colDef(
        "Expected", aggregate = "max", maxWidth = 75, format = comma_col,
        filterMethod = js_greater
      ),
      enrichment = colDef(
        "Region Enrichment", aggregate = "max", format = colFormat(digits = 2),
        maxWidth = 90, filterMethod = js_greater
      ),
      prop_total = colDef(
        "% In Region", aggregate = "max", format = percent_col, maxWidth = 60,
        filterMethod = js_greater
      ),
      p = colDef(show = FALSE),
      adj_p = colDef(
        glue("p<sub>{motif_adj}</sub>"), html = TRUE, maxWidth = 70,
        cell = \(value) sprint_pval(value), format = colFormat(digits = 3),
        aggregate = "min", filterMethod = js_less
      ),
      PWM = colDef(
        name = "IC Matrix", minWidth = 180, filterable = FALSE,
        cell = function(value) tags$img(src = value, height = '80px')
      )
    ),
    columnGroups = list(
      colGroup(
        name = "Matches", 
        columns = c(
          "total_matches", "prop_total", "matches_in_region", "expected",
          "enrichment"
        )
      ),
      colGroup(
        name = "Positional Bias", 
        columns = c("start", "end", "centre", "width")
      )
    ),
    defaultSorted = "adj_p", defaultSortOrder = "asc",
    theme = reactableTheme(style = list(fontSize = 12)),
    elementId = "motif-pos-table"
  )
cp <- htmltools::em(
  glue(
    "
    All {sum(motif_pos_res$adj_p < motif_params$alpha)} TFBMs found to show
    positional bias using an adjusted p-value ({motif_adj}) below
    {motif_params$alpha} as the criteria for significance. Highly 
    similar motifs were clustered during preparation and are grouped accordingly.
    Clicking the arrow next to the arbitrary cluster id will reveal full details
    for all motifs within the cluster which were found to be significant, 
    including the sequence logo showing information content (ICM).
    "
  )
)
tl <- tagList(
  tags$button(
    tagList(fontawesome::fa("download"), "Download as CSV"),
    onclick = glue(
      "Reactable.downloadDataCSV('motif-pos-table', '{target}_position_res.csv')"
    )
  ),
  tags$button(
    "Expand/Collapse All",
    onclick = "Reactable.toggleAllRowsExpanded('motif-pos-table')"
  ),
  tags$button(
    "Clear All Filters",
    onclick = "Reactable.setAllFilters('motif-pos-table', [])"
  ),
  tbl,
  cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```


`r if (n_pos > 0) "#### Total Matches"`

```{r plot-total-pos-matches, eval = n_pos > 0, echo = n_pos > 0, fig.cap = glue("*All best matches by their distance from the sequence centre, without binning by position. Black lines represent the loess curve through all matches. The motif with the most matches within the enriched region is shown for each cluster, with the other motifs within the cluster indicated below in brackets.*")}
lb <- glue("{names(clust_names)}\n({clust_names})") %>% 
  setNames(names(clust_names))
top_matches <- motif_list %>% 
  dplyr::filter(altname %in% top_pos$altname) %>% 
  mutate(fct = factor(altname, levels = top_pos$altname)) %>% 
  arrange(fct) %>% 
  to_list() %>% 
  getPwmMatches(
    fw_seq, abs = motif_params$abs, best_only = TRUE, mc.cores = threads
  )
top_matches %>% 
  plotMatchPos(
    geom = "point", abs = motif_params$abs, use_totals = TRUE,
    binwidth = 1
  ) +
  geom_smooth(se = FALSE, colour = "black", linewidth = 1/2, method = 'loess') +
  facet_wrap(~name, labeller = as_labeller(lb)) +
  theme(legend.position = "none") +
  labs(x = "Distance From Sequence Centre", y = "Total Matches")
```

`r if (n_pos > 0) "#### Proportional Match Distributions"`

```{r plot-smoothed-pos, eval = n_pos > 0, echo = n_pos > 0, fig.height = 0.75 * fh, fig.cap = glue("*Best matches for the representative motif within each cluster, shown as (A) proportions and (B) as a cumulative distribution*")}
A <- top_matches %>% 
  plotMatchPos(
    se = FALSE, abs = motif_params$abs, linewidth = 1/2,
    binwidth = motif_params$binwidth / 2
  ) +
  labs(
    x = "Distance From Centre", y = "Proportion", colour = "Motifs In Cluster"
  ) +
  theme(legend.position = "none")
B <- top_matches %>% 
  plotMatchPos(
    type = "cdf", geom = "line", abs = motif_params$abs,
    linewidth = 1 / 2, binwidth = 1
  ) +
  labs(x = "Distance From Centre", y = "Proportion", colour = "Representative Motif")
A + B + plot_layout(guides = "collect") 
```

`r if (n_pos > 0) "#### Match Heatmap"`

```{r heatmap-pos, echo = n_pos > 0, fig.cap = glue("*The top-ranked motifs by positional bias showing the total number of matches within each bin. Motifs are clustered by similarity of totals within this figure when more than two are being displayed. A maximum of 25 motifs are shown.*")}
if (n_pos > 0) {
  top_pos <- motif_pos_res %>% 
    dplyr::filter(adj_p < motif_params$alpha) %>% 
    dplyr::slice(1:25) %>% 
    pull(altname)
  motif_list %>% 
    dplyr::filter(altname %in% top_pos) %>% 
    to_list() %>% 
    getPwmMatches(
      fw_seq, abs = motif_params$abs, best_only = TRUE, mc.cores = threads
    ) %>% 
    plotMatchPos(
      abs = motif_params$abs, type = "heatmap", cluster = n_pos > 2,
      use_totals = TRUE,
      # binwidth = floor(motif_params$binwidth / 2)
      binwidth = motif_params$binwidth 
    ) + 
    labs(x = "Bin", fill = "Total\n Matches")
}
```



## Relationship to Other Targets

```{r}
localz_regions <- here::here(
  all_paths$peaks, "{all_targets}", "{all_targets}_regions_localz.rds"
) %>% 
  glue() %>%
  c(
    here::here(all_paths$nfr, target, "{target}_nfr_regions_localz.rds") %>% 
      glue()
  ) %>% 
  setNames(c(all_targets, glue("{target}_NFR"))) %>% 
  lapply(read_rds)
```


## References

<div id="refs"></div>

<br>
<button type="button" class="btn btn-default btn-sessioninfo" data-toggle="collapse" data-target="#sessioninfo" style="display: block;">
<span class="glyphicon glyphicon-info-sign" aria-hidden="true"></span> Session information
</button>
</p>
<div id="sessioninfo" class="collapse">
```{r session-info, echo=FALSE}
pander::pander(sessionInfo())
```
</div>
