```{r set-knitr-opts, echo=FALSE, child = here::here('analysis/setup_chunk.Rmd')}
```


```{r get-params}
target <- params$target
threads <- parallel::detectCores() - 1
```

```{r plot-cars}
plot(cars)
```

```{r}
library(tidyverse)
library(yaml)
library(pander)
library(glue)
library(extraChIPs)
library(plyranges)
library(scales)
theme_set(theme_bw())
config <- read_yaml(here::here("config", "config.yml"))
all_paths <- here::here("config", "json", "all_paths.json") %>% 
    jsonlite::fromJSON() %>% 
    lapply(here::here)
```

```{r config-params-paths}
all_params <-  here::here("config", "params.yml") %>% read_yaml() 
region_params <- all_params$gene_regions
regioner_params <- all_params$regioner
motif_params <- here::here("config", "json", "motif_analysis_param.json") %>%
  jsonlite::fromJSON() %>%
  pluck(target)
knitr_params <- here::here("config", "rmarkdown.yml") %>% 
  read_yaml() %>% 
  pluck("knitr_opts")
```

```{r load-samples}
samples <- all_paths$macs2 %>%
  file.path(target, glue("{target}_qc_samples.tsv")) %>%
  read_tsv()
treat_levels <- unique(samples$treat)
```

```{r load-annotations}
sq <- file.path(all_paths$annotation, "seqinfo.rds") %>% read_rds()
gene_regions <- file.path(all_paths$annotation, "gene_regions.rds") %>% 
  read_rds()
colours <- file.path(all_paths$annotation, "colours.rds") %>% read_rds()
region_colours <- unlist(colours$regions) %>% 
  setNames(map_chr(gene_regions, \(x) x$region[[1]])[names(.)])
```


## Introduction

Nucleosome Free Regions (NFRs) were identified for `r target` using the set of filtered peaks obtained for each treatment (`r pander(treat_levels)`).
Peaks within 300bp of each other were merged into a single peak, with this distance approximating the distance occupied by 2 nucleosomes, and approximating broader peaks that may have been otherwise called by `macs2` [@Zhang18798982].
NFRs within merged peaks were called using Histrader [@Yan2020-nz] for each treatment group, and any overlapping NFRs merged as consensus NFRs.

- Any merged peaks <500bp were excluded as being too short
- Any NFRs beyond 1kb were discarded as not being viable NFRs
- Regions of the peak with signal below 10% of the maximal signal were also not assessed for NFRs as these were more likely to be regions on the other edges of wider peaks.



In this module the key points are to

- Summarise how many NFR per peak
    + Maybe include distances between NFRs?
- Present results from a motif analysis
- Compare to peaks from all other targets using localz
- No enrichment analysis is required, as that really is covered by the larger peak analysis

```{r}
nfr_by_treat <- all_paths$nfr %>% 
  file.path(target, "{target}_{treat_levels}.nfr.bed.gz") %>% 
  glue() %>% 
  importPeaks(type = "bed", seqinfo = sq) %>% 
  setNames(treat_levels)
nfr_consensus <- all_paths$nfr %>% 
  file.path(target, "{target}_consensus_nfr.rds") %>% 
  glue() %>% 
  read_rds()
consensus_peaks <- all_paths$peaks %>% 
  file.path(target, "{target}_consensus_peaks.rds") %>% 
  glue() %>% 
  read_rds() %>% 
  mutate(nfr = countOverlaps(., nfr_consensus))
```


```{r}
plotOverlaps(nfr_by_treat, set_col = unlist(colours$treat[treat_levels]))
```

```{r}
nfr_by_treat %>% 
  plotGrlCol(
    df = tibble(treat = treat_levels), .id = "treat", fill = treat
  ) +
  scale_fill_manual(values = unlist(colours$treat))
```

```{r}
consensus_peaks %>% 
  mutate(w = width) %>% 
  as_tibble() %>% 
  ggplot(aes(w, nfr)) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, method = "lm") +
  geom_label(
    aes(w, nfr, label = lab) ,
    data = . %>% 
      summarise(
        n = sum(nfr), 
        w = min(w), nfr = max(nfr),
        .by = "region"
      ) %>% 
      mutate(
        lab = glue("n = {comma(n, 1)}"),
        w = min(w), nfr = max(nfr)
      ),
    hjust = 0, vjust = 1
  ) +
  facet_wrap(~ region) +
  scale_x_continuous(labels = comma_format(scale = 1e-3)) +
  labs(
    x = glue("{target} Peak Width (kb)"),
    y = "NFR / Peak"
  )
```


```{r}
nfr_consensus %>% 
  sort() %>% 
  mutate(d = c(0, start[-1] - end[-length(.)])) %>% 
  subset(d > 0) %>% 
  as_tibble() %>% 
  ggplot(aes(d, after_stat(count))) +
  geom_histogram(fill = "grey70", colour = 'black', bins = 100) +
  scale_x_log10(labels = comma_format()) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  labs(x = "Distance to nearest NFR (bp)", y = "Count")
```


```{r}
nfr_consensus %>% 
  plotPie(fill = region) +
  scale_fill_manual(values = region_colours)
```

```{r}
nfr_consensus %>% 
  plotPie(fill = feature) +
  scale_fill_manual(values = unlist(colours$features))
```

```{r}
nfr_consensus %>% 
  plotSplitDonut(
    inner = 'feature', outer = 'region',
    inner_palette = unlist(colours$features),
    outer_palette = region_colours,
    inner_glue = "{str_wrap(.data[[inner]], 12)}\n{comma(n)}\n{percent(p,0.1)}",
    outer_glue = "{str_wrap(.data[[outer]], 12)}\n{comma(n)}\n{percent(p,0.1)}"
  )
```



## References

<div id="refs"></div>

<br>
<button type="button" class="btn btn-default btn-sessioninfo" data-toggle="collapse" data-target="#sessioninfo" style="display: block;">
<span class="glyphicon glyphicon-info-sign" aria-hidden="true"></span> Session information
</button>
</p>
<div id="sessioninfo" class="collapse">
```{r session-info, echo=FALSE}
pander::pander(sessionInfo())
```
</div>
