```{r set-knitr-opts, echo=FALSE, child = here::here('analysis/setup_chunk.Rmd')}
```


```{r get-params}
target <- params$target
threads <- parallel::detectCores() - 1
```

```{r packages}
library(tidyverse)
library(yaml)
library(pander)
library(glue)
library(extraChIPs)
library(plyranges)
library(scales)
library(patchwork)
library(parallel)
theme_set(theme_bw())
config <- read_yaml(here::here("config", "config.yml"))
all_paths <- here::here("config", "json", "all_paths.json") %>% 
    jsonlite::fromJSON() %>% 
    lapply(here::here)
```

```{r load-samples}
samples <- all_paths$macs2 %>%
  file.path(target, glue("{target}_qc_samples.tsv")) %>%
  read_tsv()
treat_levels <- unique(samples$treat)
```

```{r config-params-paths}
all_params <-  here::here("config", "params.yml") %>% read_yaml() 
region_params <- all_params$gene_regions
regioner_params <- all_params$regioner
motif_params <- here::here("config", "json", "motif_analysis_param.json") %>%
  jsonlite::fromJSON() %>%
  pluck(target)
knitr_params <- here::here("config", "rmarkdown.yml") %>% 
  read_yaml() %>% 
  pluck("knitr_opts")
```

```{r figure-params}
fh <- knitr_params$fig.height
fw <- knitr_params$fig.width
fig_path <- here::here("docs", "assets", target)
if (!dir.exists(fig_path)) dir.create(fig_path, recursive = TRUE)
fig_type <- knitr_params$dev[[1]]
fig_fun <- match.fun(fig_type)
if (fig_type %in% c("bmp", "jpeg", "png", "tiff")) {
  ## These figure types require dpi & resetting to be inches
  formals(fig_fun)$units <- "in"
  formals(fig_fun)$res <- 300
}
```



```{r load-annotations}
sq <- file.path(all_paths$annotation, "seqinfo.rds") %>% read_rds()
gene_regions <- file.path(all_paths$annotation, "gene_regions.rds") %>% 
  read_rds()
regions <- map_chr(gene_regions, \(x) x$region[1])
colours <- file.path(all_paths$annotation, "colours.rds") %>% read_rds()
region_colours <- unlist(colours$regions) %>% 
  setNames(map_chr(gene_regions, \(x) x$region[[1]])[names(.)])
```


## Introduction

Nucleosome Free Regions (NFRs) were identified for `r target` using the set of filtered peaks obtained for each treatment (`r pander(treat_levels)`).
Peaks within 300bp of each other were merged into a single peak, with this distance approximating the distance occupied by 2 nucleosomes, and approximating broader peaks that may have been otherwise called by `macs2` [@Zhang18798982].
NFRs within merged peaks were called using Histrader [@Yan2020-nz] for each treatment group with the following settings.

- Any merged peaks <500bp were excluded as being too short
- Any NFRs beyond 1kb were discarded as not being viable NFRs
- Regions of the peak with signal below 10% of the maximal signal were also not assessed for NFRs as these were more likely to be regions on the other edges of wider peaks.

Treatment-specific NFRs were merged as consensus NFRs, taking only shared regions between treatment groups which were wider than 75bp and which were identified in all groups.
Given the inclusive nature of HisTrader, this is a more stringent approach to ensure any overly-inclusive NFR calls were minimised.

Whilst a motif enrichment analysis was performed, along with an assessement of the relationship between NFRs and other targets, no gene-set enrichment testing was performed.
The mapping from NFR to genes would very closely replicate that of the larger set of `r target` peaks, and as such, this was considered to be a redundant analysis.


```{r load-nfr}
nfr_by_treat <- all_paths$nfr %>% 
  file.path(target, "{target}_{treat_levels}.nfr.bed.gz") %>% 
  glue() %>% 
  importPeaks(type = "bed", seqinfo = sq) %>% 
  setNames(treat_levels)
nfr_consensus <- all_paths$nfr %>% 
  file.path(target, "{target}_consensus_nfr.rds") %>% 
  glue() %>% 
  read_rds()
consensus_peaks <- all_paths$peaks %>% 
  file.path(target, "{target}_consensus_peaks.rds") %>% 
  glue() %>% 
  read_rds() %>% 
  mutate(nfr = countOverlaps(., nfr_consensus))
has_features <- "feature" %in% colnames(mcols(consensus_peaks))
```

## NFR Summary {.tabset}

### Shared NFRs

```{r plot-overlaps, fig.width=fh, fig.cap = "*Summary of NFRs identified in each treatment group separately*"}
plotOverlaps(nfr_by_treat, set_col = unlist(colours$treat[treat_levels]))
```

### Peak & NFR Width

```{r plot-width, fig.cap = glue("*A) Relationship between {target} peak width and NFR identification. {target} peaks within 300bp were merged prior to plotting, and B) NFR width and numbers by treatment group.*")}
A <- consensus_peaks %>% 
  reduce(min.gapwidth = 300) %>% 
  mutate(has_nfr = overlapsAny(., nfr_consensus)) %>% 
  split(.$has_nfr) %>% 
  plotGrlCol(
    .id = 'has_nfr', fill = 'has_nfr', q = 0, geom = "violin",
    draw_quantiles = 0.5, trim = FALSE
  ) +
  scale_y_log10(labels = comma) +
  scale_fill_manual(values = c("indianred", "forestgreen")) +
  labs(
    x = "NFR Detected",
    y = glue("{target} Peak Width (bp)")
  )
B <- nfr_by_treat %>% 
  c(GRangesList(`Consensus NFR` = nfr_consensus)) %>% 
  plotGrlCol(
    df = tibble(treat = treat_levels), .id = "treat", fill = treat,
    geom = 'violin', q = 0, draw_quantiles = 0.5, trim = FALSE
  ) +
  scale_fill_manual(values = unlist(colours$treat)) +
  scale_y_log10(labels = comma) +
  labs(
    x = "NFR Group",
    y = "NFR Width (bp)"
  ) 
A + B + plot_annotation(tag_levels = "A") & theme(legend.position = 'none')
```

### NFRs Per Peak

```{r plot-nfr-by-peak-width, fig.cap = glue("*Number of NFRs found within each {target} peak, separated by genomic region. {target} peaks within 300bp were merged prior ro plotting*")}
consensus_peaks %>% 
  reduce(min.gapwidth = 300) %>% 
  mutate(
    w = width, 
    region = regions[bestOverlap(., gene_regions)] %>% factor(levels = regions),
    nfr = countOverlaps(., nfr_consensus)
  ) %>% 
  as_tibble() %>% 
  ggplot(aes(w, nfr)) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, method = "lm") +
  geom_label(
    aes(w, nfr, label = lab) ,
    data = . %>% 
      summarise(
        n = sum(nfr), 
        w = min(w), nfr = max(nfr),
        .by = "region"
      ) %>% 
      mutate(
        lab = glue("n = {comma(n, 1)}"),
        w = min(w), nfr = max(nfr)
      ),
    hjust = 0, vjust = 1
  ) +
  facet_wrap(~ region) +
  scale_x_continuous(labels = comma_format(scale = 1e-3)) +
  labs(
    x = glue("{target} Merged Peak Width (kb)"),
    y = "NFR / Peak"
  )
```

### Distance Between NFRs

```{r plot-dist, fig.cap = glue("*Distances between each NFR and the nearest upstream NFR within each broader {target} peak*")}
nfr_consensus %>% 
  granges() %>% 
  join_overlap_inner(
    consensus_peaks %>% 
      reduce(min.gapwidth = 300) %>% 
      mutate(id = seq_along(.))
  ) %>% 
  as_tibble(rangeAsChar = FALSE) %>% 
  split(.$id) %>% 
  .[map_int(., nrow) > 1] %>% 
  mclapply(
    \(x) x$start[-1] - x$end[-nrow(x)], 
    mc.cores = threads
  ) %>% 
  unlist() %>% 
  as_tibble_col("distance") %>% 
  ggplot(aes(distance, after_stat(count))) +
  geom_histogram(
    fill = "grey70", colour = 'black', bins = 100
  ) +
  scale_x_log10() +
  scale_y_continuous(expand = expansion(c(0, 0.05)), labels = comma) +
  labs(
    x = "Distance between NFRs", y = "Totals"
  )
```

### NFRs By Region

```{r plot-pie, fig.width = fh, fig.cap = glue("*Distribution of NFRs by genomic region, noting that wider {target} peaks will accumulate more NFRs*")}
nfr_consensus %>% 
  plotPie(
    fill = region, hole_width = 0.2,
    total_size = 4,
    cat_glue = "{str_wrap(.data[[fill]], 15)}\n{comma(n, 1)}\n({percent(p, 0.1)})",
    cat_alpha = 0.7, cat_adj = 0.1
  ) +
  scale_fill_manual(values = region_colours) +
  theme(legend.position = "none")
```

`r if (has_features) "### NFRs By Region and Feature"`

```{r plot-donut, echo = has_features, fig.width = fh, fig.cap = 'Breakdown of NFRs by provided features and genomic region.'}
if (has_features) {
  nfr_consensus %>% 
    plotSplitDonut(
      inner = 'feature', outer = 'region',
      inner_palette = unlist(colours$features),
      outer_palette = region_colours,
      inner_glue = "{str_wrap(.data[[inner]], 15)}\n{comma(n)}\n{percent(p,0.1)}",
      outer_glue = "{str_wrap(.data[[outer]], 15)}\n{comma(n)}\n{percent(p,0.1)}",
      outer_min_p = 0.03
    )
}
```

## Motif Analysis

## Relationship to Other Targets

## References

<div id="refs"></div>

<br>
<button type="button" class="btn btn-default btn-sessioninfo" data-toggle="collapse" data-target="#sessioninfo" style="display: block;">
<span class="glyphicon glyphicon-info-sign" aria-hidden="true"></span> Session information
</button>
</p>
<div id="sessioninfo" class="collapse">
```{r session-info, echo=FALSE}
pander::pander(sessionInfo())
```
</div>
