```{r dummy-params, eval = interactive()}
## For the setup, refer to the comparison which will appear on the x-axis with
## the subscript _x, whilst for that onb the y-axis, use _y. This can be applied
## to every data object and element of params
## Setup for AR (x) & ER (y) using E2_E2DHT as the comparison
x <- "AR"
y <- "ER"
comp_x <- glue::glue("{x}_E2_E2DHT")
comp_y <- glue::glue("{y}_E2_E2DHT")
treat_levels <- list(comp_x, comp_y) |>
  lapply(\(x) strsplit(x, split = "_")[[1]][2:3]) |>
  setNames(c(x, y))
params <- list(
  x = x, y = y, comp_x = comp_x, comp_y = comp_y,
  filtered_peaks_x_ctrl = "output/peak_analysis/{x}/{x}_{treat_levels[[x]][[1]]}_filtered_peaks.narrowPeak",
  filtered_peaks_x_treat = "output/peak_analysis/{x}/{x}_{treat_levels[[x]][[2]]}_filtered_peaks.narrowPeak",
  filtered_peaks_y_ctrl = "output/peak_analysis/{y}/{y}_{treat_levels[[y]][[1]]}_filtered_peaks.narrowPeak",
  filtered_peaks_y_treat = "output/peak_analysis/{y}/{y}_{treat_levels[[y]][[2]]}_filtered_peaks.narrowPeak",
  dsa_x = "output/differential_signal/{x}/{comp_x}-differential-signal.rds",
  dsa_y = "output/differential_signal/{y}/{comp_y}-differential-signal.rds"
) |>
  lapply(glue::glue)
rm(list = c("x", "y", "comp_x", "comp_y", "treat_levels"))
```



```{r packages}
library(tidyverse)
library(glue)
library(pander)
library(yaml)
library(extraChIPs)
library(scales)
library(ggrepel)
library(reactable)
library(htmltools)
library(rlang)
library(plyranges)
library(parallel)
# library(rtracklayer)
# library(VennDiagram)
# library(magrittr)
# library(rlang)
# library(ggside)
# library(msigdbr)
# library(goseq)
# library(GenomicInteractions)
# library(ggraph)
# library(tidygraph)
source(here::here("workflow", "scripts", "custom_functions.R"))
source(here::here("workflow", "scripts", "table_funs.R"))
threads <- detectCores() - 1
if (interactive()) threads <- 4
# register(MulticoreParam(workers = threads))
panderOptions("big.mark", ",")
panderOptions("missing", "")
panderOptions("table.split.table", Inf)
theme_set(
  theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5)
    )
)
```



```{r get_params, echo = FALSE}
config <- read_yaml(here::here("config", "config.yml"))
all_samples <- read_tsv(here::here(config$samples))
targets <- c(params$x, params$y)
n_targets = length(unique(targets))

target_regex <- paste(targets, collapse = "|")
treat_levels <- params[paste0("comp_", c("x", "y"))] %>% 
  lapply(str_remove_all, pattern = paste0("^(", target_regex, ")_")) %>% 
  lapply(\(x) strsplit(x, split = "_")[[1]])

full_comp <- params[c("comp_x", "comp_y")] %>% 
  unlist() %>% 
  paste(collapse = "_")
comp_names <- list(
  x = glue("{params$x}: {treat_levels$comp_x[[2]]} Vs {treat_levels$comp_x[[1]]}"),
  y = glue("{params$y}: {treat_levels$comp_y[[2]]} Vs {treat_levels$comp_y[[1]]}")
)
if (n_targets > 1) {
  short_names <- as.list(targets)
} else {
  short_names <- treat_levels %>% 
    lapply(\(x) paste(x[[2]], "Vs", x[[1]]))
}
names(short_names) <- c("x", "y")

all_paths <- here::here("config", "json", "all_paths.json") %>% 
  jsonlite::fromJSON()
diff_sig_param <- here::here(
  "config", "json", "differential_signal_param.json"
) %>% 
  jsonlite::fromJSON() %>% 
  .[targets]

## Move to json as these are defined
pairwise_params <- config$pairwise$default
pairwise_params[names(config$pairwise[[full_comp]])] <- config$pairwise[[full_comp]]
```



```{r load-annotations}
gtf_gene <- file.path(all_paths$annotation, "gtf_gene.rds") %>% read_rds()
id2gene <- structure(gtf_gene$gene_name, names = gtf_gene$gene_id) %>% 
  .[!duplicated(names(.))]
sq <- seqinfo(gtf_gene)

colours <- read_rds(file.path(all_paths$annotation, "colours.rds")) %>% 
  lapply(unlist)
comp_cols <- colours$target[targets]

pw_cols <- colours$pairwise
names(pw_cols) <- names(pw_cols) %>% 
  str_replace_all("^(.+)_(.+)$", "\\1 \\2") %>% 
  str_to_title() %>% 
  str_replace_all(
    "^(.+) (.+)$", paste(short_names$x, "\\1 -", short_names$y, "\\2")
  )
pw_cols <- c(
  pw_cols,
  "Both Increased" = colours$pairwise[["increased_increased"]],
  "Both Decreased" = colours$pairwise[["decreased_decreased"]],
  "Both Unchanged" = colours$pairwise[["unchanged_unchanged"]]
)

gene_regions <- file.path(all_paths$annotation, "gene_regions.rds") %>% 
  read_rds()
regions <- map_chr(gene_regions, \(x) unique(x$region))
region_colours <- colours$regions %>% 
  setNames(regions[names(colours$regions)]) %>% 
  .[unname(regions)]

feature_colours <- unlist(colours$features)
features <- here::here(all_paths$annotation, "features.rds") %>%
  read_rds()
has_features <- length(features) > 0
feat_names <- names(features) %>% 
  lapply(
    \(x) paste(x, features[[x]]$feature, sep = ": ")
  ) %>% 
  lapply(fct_infreq) %>% 
  lapply(levels) %>% 
  unlist()
## Combine features & regions for testing
all_regions <- c(unname(regions), feat_names)

hic <- file.path(all_paths$annotation, "hic.rds") %>% read_rds()
has_hic <- length(hic) > 0

rna <- file.path(all_paths$annotation, "rna.rds") %>% read_rds()
has_rna <- length(rna) > 0
detected_ids <- unique(gtf_gene$gene_id)
if (has_rna) {
  detected_ids <- detected_ids %>% intersect(bind_rows(rna)$gene_id)
}
include_genes <- id2gene[detected_ids]

direction_colours <- colours$direction %>% 
  setNames(str_to_title(names(.)))
treat_colours <- colours$treat[unique(unlist(treat_levels))]
full_palette <- c(
  feature_colours, region_colours, direction_colours, treat_colours,
  colours$target[targets],
  pw_cols,
  c(
    "None" = colours$direction[["undetected"]],
    "Other" = colours$direction[["unchanged"]])
)

fig_params <- here::here("config", "rmarkdown.yml") %>% 
  read_yaml() %>% 
  pluck("knitr_opts")
fw <- fig_params$fig.width
fh <- fig_params$fig.height
fig_type <- fig_params$dev[[1]]
if (is.null(fig_type)) stop("Couldn't detect figure type")
fig_fun <- match.fun(fig_type)
if (fig_type %in% c("bmp", "jpeg", "png", "tiff")) {
  ## These figure types require dpi & resetting to be inches
  formals(fig_fun)$units <- "in"
  formals(fig_fun)$res <- 300
}
fig_path <- here::here("docs", "assets", full_comp)
if (!dir.exists(fig_path)) dir.create(fig_path, recursive = TRUE)
```


```{r load-data}
filtered_peaks <- c(
  params$filtered_peaks_x_ctrl, params$filtered_peaks_x_treat,
  params$filtered_peaks_y_ctrl, params$filtered_peaks_y_treat
) %>% 
  unique() %>% 
  here::here() %>% 
  importPeaks(seqinfo = sq) %>% 
  setNames(str_remove_all(names(.), "_filtered.+"))
dsa_results <- list2(
  "{params$comp_x}" := params$dsa_x,
  "{params$comp_y}" := params$dsa_y
) %>% 
  lapply(here::here) %>% 
  lapply(read_rds) %>% 
  mclapply(
    \(x) {
      ## Set FDR as the common column, now IHW is sorted out
      fdr_col <- ifelse(metadata(x)$ihw == "none", "FDR", "fdr_ihw")
      ## Also switch any p_mu0 column to be PValue for declaring categories
      p_col <- ifelse(metadata(x)$fc == 1, "PValue", "p_mu0")
      x %>% 
        mutate(
          FDR = !!sym(fdr_col),
          PValue = !!sym(p_col)
        ) %>% 
        select(
          any_of(c("region", "feature")), starts_with("gene"),
          starts_with("log"), "PValue", "FDR", "status"
        )
    },
    mc.cores = 2
  ) %>% 
  GRangesList()
```


```{r define-outputs}
# out_path <- here::here(
#   "output", "pairwise_comparisons", paste(targets, collapse = "_")
# )
# if (!dir.exists(out_path)) dir.create(out_path, recursive = TRUE)
# all_out <- list(
#   csv = file.path(
#     out_path, glue(comp_name, "-pairwise_comparison.csv.gz"
#     )
#   ),
#   de_genes = file.path(
#     out_path, 
#     glue("{comp_name}-de_genes.csv")
#   ),
#   goseq = file.path(
#     out_path,
#     glue("{comp_name}-enrichment.csv")
#   ),
#   results = file.path(
#     out_path, glue(comp_name, "-all_windows.rds")
#   ),
#   rna_enrichment = file.path(
#     out_path,
#     glue("{comp_name}-rnaseq_enrichment.csv")
#   ),
#   renv = here::here(
#     "output", "envs",
#     glue(comp_name, "-pairwise_comparison.RData")
#   )
# )
# ## Empty objects for when RNA-Seq is absent
# de_genes_both_comps <- tibble()
# cmn_res <- list(tibble(gs_name = character(), leadingEdge = list()))
```


# Outline

This stage of the GRAVI workflow takes the results from two individual differential signal analyses and compares the results, by finding regions which directly overlap and considering the joint behaviour.
In this specific analyses the differential signal response for `r targets[[1]]` when comparing `r comp_names$x`, will be integrated with the differential signal response for `r targets[[2]]` comparing `r comp_names$y`.
Data will first be described from the perspective of peaks, regardless of treatment before moving onto the differential signal analysis (DSA) results.

## Integration Of Differential Signal Results

The initial steps of DSA detect regions with clear changes in signal, however the integration of datasets instead focusses on the *most correct* classification of a given range/window, where unchanged is not *simply lack of a statistically significant result*, but *a true state needing to be identified*.
Differential signal analyses generally focus on controlling the Type I errors (i.e. false positive), which comes with a commensurate increase in Type II errors (false negatives).
When seeking to correctly classify a window across two datasets, the detection as changed can be taken as a robust finding, whilst the consideration as unchanged may be less robust.
When comparing a window across two datasets, a detection as changed within one dataset is taken as primary evidence of 'something interesting' occurring, with the secondary consideration being the correct description of the combined signal patterns.
For this reason, and to minimise issues with hard thresholding, once a window is considered to be of interest, the p-value used in the second comparison is that obtained (and FDR-adjusted) using the test for $H_0: \mu = 0$ instead of the initial $H_0: 0 < |\mu| < \lambda$.
The insistence on the more stringent initial inclusion criteria still protects against false discoveries from the perspective of window identification, but this approach instead allows *improved classification* across both comparisons.

For *classification steps within each window*, each target is given the status 1) Increased, 2) Decreased, 3) Unchanged, or 4) Undetected.
Across two comparisons, this gives 15 possible classifications for each region found with signal from at least one comparison (e.g. Increased-Increased, Increased-Decreased, Increased-Unchanged, Increased-Undetected etc.), given that any potential Undetected-Undetected windows will not be included for obvious reasons.

## Enrichment Testing

After direct comparison of signal patterns, enrichment analysis is performed in a similar manner to for individual analyses.
If RNA-Seq data is provided, genes will be restricted to those present within the RNA-Seq dataset, as representative of detected genes.
Enrichment testing is performed on:

1. All genes mapped to any region with detected signal against those not mapped to detected signal
2. All genes mapped to both targets against genes not mapped 

## Comparison To RNA-Seq

If RNA-Seq data is supplied, the sets of genes associated with all signal patterns (e.g. Up-Up, Up-Down etc) were treated as gene-sets and Gene Set Enrichment Analysis (GSEA) [@SubramanianGsea2005] was used to determine enrichment and leading-edge genes associated with altered expression.
Genes will be ranked to incorporate direction of change (i.e. Directional GSEA) or significance only (i.e Non-Directional GSEA).
The results from enrichment analysis across ChIP target and RNA-Seq data will also be combined to reveal pathways under the regulatory influence in either comparison, for which we also have evidence of changed gene expression.

# Peak Comparison {.tabset}

Data was loaded for the initially-identified peaks to show the general level of similarity between signal regions from either dataset.


`r if (n_targets > 1) "## Shared Signal Regions"`

```{r plot-common, results = 'hide', eval = n_targets > 1, eval = n_targets > 1, fig.cap = glue("*Relationship between regions contained in {targets[[1]]} and {targets[[2]]}, and which of the two targets each peak is found in. Peaks are based on the union of both sets of peaks, which are themselves independent of treatment group.*"), fig.height = 8, fig.width = 8}
dsa_results %>% 
  setNames(
    str_replace_all(unlist(comp_names), "^(.+): (.+)$", "\\1\n(\\2)")
  ) %>% 
  plotOverlaps(
    set_col = comp_cols, alpha = 0.3, cex = 1.3, cat.cex = 1.3
  )
```


## Treatment Specific Peaks

```{r treatment-peaks-upset, fig.height = 6, fig.cap = glue("*Peaks detected for ", glue_collapse(unique(targets), sep = " and " ), " in the treatment groups ", glue_collapse(unique(unlist(treat_levels)), sep = ", ", last = " & "), ".*")}
filtered_peaks %>% 
  .[rev(names(.))] %>% 
  setNames(str_replace_all(names(.), "(.+)_(.+)", "\\1 (\\2)")) %>% 
  plotOverlaps(
    type = "upset", .sort_sets = FALSE, n_intersections = 20,
    set_col = treat_colours[str_replace_all(names(.), ".+\\((.+)\\)", "\\1")]
  )
```

# Differential Signal Windows {.tabset}

```{r all-windows}
grl_cols <- c("centre", "logCPM", "logFC", "PValue", "FDR", "status")
merge_cols <- c("feature", "gene_id", "gene_name")
combined_results <- dsa_results %>% 
  endoapply(mutate, centre = start + 0.5 * width) %>% 
  mapGrlCols(var = grl_cols, collapse = merge_cols) %>% 
  mutate(
    region = bestOverlap(., unlist(gene_regions), var = "region") %>% 
      factor(levels = names(region_colours))    
  ) %>% 
  as_tibble() %>% 
  mutate(
    across(ends_with("status"), \(x) fct_na_value_to_level(x, "Undetected")),
    d = abs(
      !!sym(paste0(params$comp_x, "_centre")) - !!sym(paste0(params$comp_y, "_centre"))
    ),
    is_sig = (
      !!sym(glue(params$comp_x, "_status")) %in% c("Increased", "Decreased") |
        !!sym(glue(params$comp_y, "_status")) %in% c("Increased", "Decreased")
    ),
    
    ## Reclassify those significant in y but which failed by lambda in x
    "{params$comp_x}_status" := case_when(
      ## Don't change if not sig in one comparison
      !is_sig ~ !!sym(glue("{params$comp_x}_status")),
      ## Don't change if not significant using the conventional H0
      p.adjust(
        !!sym(glue("{params$comp_x}_PValue")), pairwise_params$adj_mu0
      ) >= pairwise_params$alpha ~ !!sym(glue("{params$comp_x}_status")),
      # ## The remaining ones are candidates
      (
        !!sym(glue("{params$comp_x}_logFC")) > log2(diff_sig_param[[params$x]]$fc)
      ) ~ "Increased",
      (
        !!sym(glue("{params$comp_x}_logFC")) < -log2(diff_sig_param[[params$x]]$fc)
      ) ~ "Decreased",
      TRUE ~ as.character(!!sym(glue("{params$comp_x}_status")))
    ) %>%
      factor(levels = levels(dsa_results[[1]]$status)),
    
    ## Repeat for those significant in x but which failed by lambda in y
    "{params$comp_y}_status" := case_when(
      ## Don't change if not sig in one comparison
      !is_sig ~ as.character(!!sym(glue("{params$comp_y}_status"))),
      ## Don't change if not significant using the conventional H0
      p.adjust(
        !!sym(glue("{params$comp_y}_PValue")), pairwise_params$adj_mu0
      ) >= pairwise_params$alpha ~ as.character(!!sym(glue("{params$comp_y}_status"))),
      ## The remaining ones are candidates
      (
        !!sym(glue("{params$comp_y}_logFC")) > log2(diff_sig_param[[params$y]]$fc)
      ) ~ "Increased",
      (
        !!sym(glue("{params$comp_y}_logFC")) < -log2(diff_sig_param[[params$y]]$fc)
      ) ~ "Decreased",
      TRUE ~ as.character(!!sym(glue("{params$comp_y}_status")))
    ) %>%
      factor(levels = levels(dsa_results[[2]]$status)),
    
    ## Add a combined status
    status = fct_cross(
      !!sym(glue("{params$comp_x}_status")), 
      !!sym(glue("{params$comp_y}_status")), sep = " - "
    ) %>% 
      fct_relabel(
        \(f) str_replace_all(
          f,
          "^(.+) - (.+)$",
          paste(short_names$x, "\\1 -", short_names$y, "\\2")
        )
      )
    
  ) %>% 
  dplyr::select(-ends_with("centre"), -is_sig) %>% 
  colToRanges("range", seqinfo = sq) %>% 
  mutate(
     gene_id = gene_id %>% 
       mclapply(\(x) unique(strsplit(x, "; ")[[1]]), mc.cores = threads) %>% 
       as("CharacterList"),
     gene_name = endoapply(gene_id, \(x) sort(id2gene[x]))
  )
```


A set of `r comma(length(all_windows))` regions with signal was formed by obtaining the union of windows produced during the analysis of `r glue_collapse(unlist(comp_names), last = " and ")` individually.
These common windows were formed using any overlap between analysis-specific windows.
All mappings to genes and gneomic features were maintined form the original data

In the original datasets, `r comma(length(dsa_results[[1]]))` ranges were included for `r comp_names$x`, whilst `r comma(length(dsa_results[[2]]))` ranges were included for `r comp_names$y`.
The distance between each pair of windows was found by taking the centre of the original sliding window used for statistical analysis, and finding the difference. 
As windows were detected in an un-stranded manner, all distance values were set to be positive.
Taking the complete set of `r comma(length(all_windows))` combined windows, `r comma(sum(!is.na(all_windows$d)))` (`r percent(mean(!is.na(all_windows$d)), 0.1)`) were included in *both* datasets.
The median distance between overlapping windows found in both comparisons was `r  median(all_windows$d, na.rm = TRUE)`bp.

Merged windows were compared across both targets and in order to obtain the best classification for each window, the initial values of an FDR-adjusted p-value < `r fdr_alpha` from either comparison were used to consider a window as showing changed signal in at least one comparison.


A secondary step was then introduced for each window such that if one target was significant and the other target had 1) received an FDR-adjusted p-value < `r fdr_alpha` *using a point-based* $H_0$, and 2) had an estimated logFC beyond the range specified for the range-based null hypothesis ($H_0: 0 < |\mu| < \lambda$ for $\lambda = `r round(lambda, 2)`$), these windows were then considered significant in the second comparison.
This was to help *reduce the number of windows incorrectly classified as unchanged* for windows the alternative comparison was defined as changed.


```{r tbl-summary}
cp <- htmltools::tags$em(
  glue(
    "Summary of combined windows for ",
    glue_collapse(comp_names, sep = " and "),
    ". Ranges were classified as described above. ",
    "Where signal was deteted in both comparisons, the median distance ",
    "between peak centres is given.",
    "The number showing each combined pattern are given, along ",
    "with the number of unique genes mapped to each set of rangess, ",
    "noting that some genes will be mapped to multiple sets of ranges. ",
    "The % of all genes with a window mapped from either target is given in ",
    "the final column."
  )
)
tbl <- combined_results %>%
  as_tibble() %>%
  mutate(mapped = map_int(gene_id, length)) %>% 
  summarise(
    total = dplyr::n(),
    d = median(d, na.rm = TRUE),
    mapped_genes = gene_id %>% unlist() %>% unique() %>% length(),
    # .groups = "drop"
    .by = status
  ) %>% 
  mutate(
    `% genes` = mapped_genes / length(unique(unlist(combined_results$gene_id)))
  ) %>% 
  arrange(desc(total)) %>% 
  separate(status, into = unlist(short_names), sep = " - ") %>% 
  reactable(
    filterable = TRUE, showPageSizeOptions = nrow(.) > 10,
    columns = list2(
      !!sym(short_names$x) := colDef(
        cell = function(value) {
          value <- str_extract(value, "Increased|Decreased|Unchanged|Undetected")
          if (str_detect(value, "Increased")) value <- paste(value, "\u21E7")
          if (str_detect(value, "Decreased")) value <- paste(value, "\u21E9")
          paste(targets[[1]], value)
        },
        style = function(value) {
          colour <- case_when(
            str_detect(value, "Increased") ~ colours$direction[["increased"]],
            str_detect(value, "Decreased") ~ colours$direction[["decreased"]],
            str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
            str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
          )
          list(color = colour)
        },
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('summary-result-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        },
        footer = htmltools::tags$b("Total")
      ),
      !!sym(short_names$y) := colDef(
        cell = function(value) {
          value <- str_extract(value, "Increased|Decreased|Unchanged|Undetected")
          if (str_detect(value, "Increased")) value <- paste(value, "\u21E7")
          if (str_detect(value, "Decreased")) value <- paste(value, "\u21E9")
          paste(targets[[2]], value)
        },
        style = function(value) {
          colour <- case_when(
            str_detect(value, "Increased") ~ colours$direction[["increased"]],
            str_detect(value, "Decreased") ~ colours$direction[["decreased"]],
            str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
            str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
          )
          list(color = colour)
        },
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('summary-result-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        }
      ),
      total = colDef(
        name = "Number of Ranges", format = comma_col, 
        filterMethod = js_greater, maxWidth = 150,
        style = function(value) {
          bar_style(width = 0.9*value / length(combined_results), fill = "#B3B3B3")
        },
        footer = htmltools::tags$b(comma(length(combined_results), 1))
      ),
      d = colDef(
        "Mean Distance", format = comma_col, filterMethod = js_greater,
        maxWidth = 120
      ),
      mapped_genes = colDef(
        name = "Mapped Genes", format = comma_col, filterMethod = js_greater,
        maxWidth = 120,
        style = function(value) {
          bar_style(width = 0.9*value / length(unique(unlist(combined_results$gene_id))), fill = "#B3B3B3")
        },
        footer = htmltools::tags$b(comma(length(unique(unlist(combined_results$gene_id))), 1))
      ),
      `% genes` = colDef(
        name = "% Of All Mapped Genes", format = percent_col, 
        filterMethod = js_greater,
        style = function(value) {
          bar_style(width = 0.9*value, fill = "#B3B3B3")
        }
      )
    ),
    elementId = 'summary-result-table'
  )
tl <- tagList(tbl, cp)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```

## Overall Status

```{r plot-both-donut, fig.cap = "Summary of signal across both comparisons incorporating all sites where signal was detected in either copmarison. Ranges where signal is changed in either comparisons are exploded."}
combined_results %>% 
  as_tibble() %>% 
  dplyr::select(ends_with("_status")) %>% 
  setNames(unlist(short_names)) %>% 
  plotSplitDonut(
    inner = !!sym(short_names$x), outer = !!sym(short_names$y), 
    inner_palette = full_palette, outer_palette = full_palette,
    inner_glue = "{inner}\n{.data[[inner]]}\n{comma(n, 1)}\n({percent(p, 0.1)})",
    inner_label_alpha = 0.7,
    outer_glue = "{outer} {.data[[outer]]}\n{comma(n, 1)} ({percent(p, 0.1)})",
    outer_min_p = 0.02, outer_label = "text", outer_rotate = TRUE,
    outer_label_colour = "palette",
    explode_inner = "(In|De)creased", explode_outer = "(In|De)creased",
    explode_r = 0.4, explode_query = "AND", nudge_r = 0.5
  ) %>% 
    .[[1]]
```


## Pairwise Differential Signal


```{r plot-dbwin, fig.cap = glue("*Comparative changes in both ", glue_collapse(comp_names, sep = " and "), ". The range with most extreme combined change in signal is shown with any genes labelled. ", ifelse(has_rna, "Only genes detected within transcriptomic data are labelled. " , ""),  "The range around zero used for range-based hypothesis testing is indicated with grey dashed lines.*")}
df <- combined_results %>% 
  as_tibble() %>% 
  select(
    range,
    ends_with("logFC"), ends_with("status"), any_of(c("region", "feature")), 
    starts_with("gene")
  ) %>% 
  dplyr::rename(
    "{short_names$x} logFC" := !!sym(paste0(params$comp_x, "_logFC")),
    "{short_names$y} logFC" := !!sym(paste0(params$comp_y, "_logFC")),
    "{short_names$x} status" := !!sym(paste0(params$comp_x, "_status")),
    "{short_names$y} status" := !!sym(paste0(params$comp_y, "_status"))
  )
x_lim <- range(df[[glue("{short_names$x} logFC")]], na.rm = TRUE)
y_lim <- range(df[[glue("{short_names$y} logFC")]], na.rm = TRUE)
df %>% 
  arrange(status) %>% 
  ggplot(
    aes(
      x = !!sym(glue("{short_names$x} logFC")), 
      y = !!sym(glue("{short_names$y} logFC")),
      colour = status
    )
  ) +
  geom_point(
    data = . %>% dplyr::filter(!str_detect(status, "Undetected"))
  ) +
  geom_label_repel(
    aes(label = label),
    data = . %>% 
      mutate(
        across(ends_with("logFC"), list(sq = \(x) x^2)),
        diam = (
          !!sym(glue("{short_names$x} logFC_sq")) + 
            !!sym(glue("{short_names$y} logFC_sq"))
        )
      ) %>% 
      dplyr::filter(diam > 1) %>% 
      arrange(desc(diam)) %>% 
      dplyr::slice(1:10, .by = status) %>% 
      mutate(
        label = map_chr(
          gene_name, 
          \(x) paste(intersect(x, include_genes), collapse = "; ")
        ) %>% 
          str_wrap(15) %>% 
          str_trunc(30) 
      ) %>% 
      dplyr::filter(label != "") %>% 
      dplyr::slice(1:2, .by = status),
    size = 3, show.legend = FALSE, fill = rgb(1, 1, 1, 0.8),
    fontface = "italic"
  ) +
  geom_hline(
    yintercept = c(1, -1) * log2(diff_sig_param[[params$y]]$fc),
    linetype = 2, colour = "grey"
  ) +
  geom_vline(
    xintercept = c(1, -1) * log2(diff_sig_param[[params$x]]$fc),
    linetype = 2, colour = "grey"
  ) +
  annotate(
    "text",
    x = min(x_lim), y = max(y_lim), vjust = 1, hjust = 0,
    label = paste(
      "rho == ",
      cor(
        dplyr::select(df, ends_with("logFC")), use = "pairwise.complete.obs"
      )[1, 2] %>%  round(3)
    ),
    parse = TRUE, size = 3.4
  ) +
  geom_xsideboxplot(
    aes(
      x = !!sym(glue("{short_names$x} logFC")), 
      y = !!sym(glue("{short_names$y} status")),
      fill = !!sym(glue("{short_names$y} status"))
    ), 
    data = . %>% 
      dplyr::filter(!is.na(!!sym(glue("{short_names$x} logFC")))),
    orientation = "y", colour = "grey20", show.legend = FALSE
  ) +
  geom_ysideboxplot(
    aes(
      x = !!sym(glue("{short_names$x} status")),
      y = !!sym(glue("{short_names$y} logFC")),
      fill = !!sym(glue("{short_names$x} status"))
    ),
    data = . %>% 
      dplyr::filter(!is.na(!!sym(glue("{short_names$y} logFC")))),
    colour = "grey20", orientation = "x", show.legend = FALSE
  ) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_ysidex_discrete(
    position = "top",     
    labels = \(x) paste(short_names$x, x) %>% str_wrap(15)
  ) +
  scale_xsidey_discrete(
    labels = \(x) paste(short_names$y, x) %>% str_wrap(15)
  ) +
  scale_colour_manual(values = full_palette, labels = label_wrap(15)) +
  scale_fill_manual(values = full_palette) +
  labs(colour = "Combined Status") +
  theme(
    ggside.panel.scale.x = 0.1 + 0.05 * length(levels(df[[glue("{short_names$y} status")]])),
    ggside.panel.scale.y = 0.1 + 0.025 * length(levels(df[[glue("{short_names$x} status")]])),
    ggside.axis.text.x.top = element_text(angle = 90, hjust = 0, vjust = 0.5),
    legend.position = "bottom",
    axis.title.y.left = element_text(vjust = -10, hjust = 0.4),
    axis.title.x.bottom = element_text(hjust = 0.45)
  )
```


## logFC By Region

```{r plot-dbwin-by-region, fig.cap = glue("*Comparative changes in both ", glue_collapse(comp_names, sep = " and "), ", separated by defined gene-centric regions. The window with most extreme combined change in signal for each region is shown with any genes labelled as previously. The range around zero used for range-based hypothesis testing is indicated with grey dashed lines.*")}
df %>% 
  arrange(status) %>% 
  ggplot(
    aes(
      x = !!sym(glue("{short_names$x} logFC")), 
      y = !!sym(glue("{short_names$y} logFC")),
      colour = status
    )
  ) +
  geom_point(
    data = . %>% dplyr::filter(!str_detect(status, "Undetected"))
  ) +
  geom_label_repel(
    aes(label = label),
    data = . %>% 
      mutate(
        across(ends_with("logFC"), list(sq = \(x) x^2)),
        diam = (
          !!sym(glue("{short_names$x} logFC_sq")) + 
            !!sym(glue("{short_names$y} logFC_sq"))
        )
      ) %>% 
      dplyr::filter(diam > 1) %>% 
      arrange(desc(diam)) %>% 
      dplyr::slice(1:10, .by = c(status, region)) %>% 
      mutate(
        label = map_chr(
          gene_name, 
          \(x) paste(intersect(x, include_genes), collapse = "; ")
        ) %>% 
          str_wrap(15) %>% 
          str_trunc(30) 
      ) %>% 
      dplyr::filter(label != "") %>% 
      dplyr::slice(1, .by = c(status, region)),
    size = 3, show.legend = FALSE, fill = rgb(1, 1, 1, 0.8),
    fontface = "italic"
  ) +
  geom_hline(
    yintercept = c(1, -1) * log2(diff_sig_param[[params$y]]$fc),
    linetype = 2, colour = "grey"
  ) +
  geom_vline(
    xintercept = c(1, -1) * log2(diff_sig_param[[params$x]]$fc),
    linetype = 2, colour = "grey"
  ) +
  geom_smooth(
    data = . %>% 
      dplyr::filter(!grepl("Undetected", status)),
    se = FALSE, colour = "grey40", method = MASS::rlm
  ) +
  geom_text(
    aes(label = label),
    data = . %>%
      dplyr::filter(!grepl("Undetected", status)) %>% 
      dplyr::select(region, ends_with("logFC")) %>% 
      split(.$region) %>% 
      mclapply(
        \(x) {
          tibble(
            cor = cor(dplyr::select(x, ends_with("logFC")))[1, 2],
            "{short_names$x} logFC" := min(x[[glue("{short_names$x} logFC")]]),
            "{short_names$y} logFC" := max(x[[glue("{short_names$y} logFC")]])
          )
        }, mc.cores = threads
      ) %>% 
      bind_rows(.id = "region") %>% 
      mutate(
        label = paste("rho ==", round(cor, 3)),
        "{short_names$x} logFC" := min(!!sym(glue("{short_names$x} logFC"))),
        "{short_names$y} logFC" := max(!!sym(glue("{short_names$y} logFC")))
      ),
    colour = "grey20", parse = TRUE, vjust = 1, hjust = 0, size = 3.5
  ) +
  facet_wrap(~region) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_colour_manual(values = full_palette, labels = label_wrap(15)) +
  labs(colour = "Combined Status") +
  theme(legend.position = "bottom")
```



`r if (has_features) "## logFC By Feature"`

```{r plot-dbwin-by-feature, echo = has_features, fig.cap = glue("*Comparative changes in both ", glue_collapse(comp_names, sep = " and "), ", separated by feature. The range with most extreme combined change in signal for each separate feature is shown with any genes labelled as previously, noting that combined ranges may map to multiple features. The range around zero used for range-based hypothesis testing is indicated with grey dashed lines.*")}
if (has_features) {
  df %>% 
    mutate(
      feature = feature %>% 
        mclapply(strsplit, split = "; ", mc.cores = threads) %>% 
        mclapply(pluck, 1, mc.cores = threads)
    ) %>% 
    unnest(feature, keep_empty = TRUE) %>% 
    mutate(
      feature = str_replace_na(feature, "no_feature") %>% 
        fct_infreq() %>% 
        fct_lump_n(8)
    ) %>% 
    arrange(status) %>% 
    ggplot(
      aes(
        x = !!sym(glue("{short_names$x} logFC")), 
        y = !!sym(glue("{short_names$y} logFC")),
        colour = status
      )
    ) +
    geom_point(
      data = . %>% dplyr::filter(!str_detect(status, "Undetected"))
    ) +
    geom_label_repel(
      aes(label = label),
      data = . %>% 
        mutate(
          across(ends_with("logFC"), list(sq = \(x) x^2)),
          diam = (
            !!sym(glue("{short_names$x} logFC_sq")) + 
              !!sym(glue("{short_names$y} logFC_sq"))
          )
        ) %>% 
        dplyr::filter(diam > 1) %>% 
        arrange(desc(diam)) %>% 
        dplyr::slice(1:10, .by = c(status, feature)) %>% 
        mutate(
          label = map_chr(
            gene_name, 
            \(x) paste(intersect(x, include_genes), collapse = "; ")
          ) %>% 
            str_wrap(15) %>% 
            str_trunc(30) 
        ) %>% 
        dplyr::filter(label != "") %>% 
        dplyr::slice(1, .by = c(status, feature)),
      size = 3, show.legend = FALSE, fill = rgb(1, 1, 1, 0.8),
      fontface = "italic"
    ) +
    geom_hline(
      yintercept = c(1, -1) * log2(diff_sig_param[[params$y]]$fc),
      linetype = 2, colour = "grey"
    ) +
    geom_vline(
      xintercept = c(1, -1) * log2(diff_sig_param[[params$x]]$fc),
      linetype = 2, colour = "grey"
    ) +
    geom_smooth(
      data = . %>% 
        dplyr::filter(!grepl("Undetected", status)),
      se = FALSE, colour = "grey40", method = MASS::rlm
    ) +
    geom_text(
      aes(label = label),
      data = . %>%
        dplyr::filter(!grepl("Undetected", status)) %>% 
        dplyr::select(feature, ends_with("logFC")) %>% 
        split(.$feature) %>% 
        mclapply(
          \(x) {
            tibble(
              cor = cor(dplyr::select(x, ends_with("logFC")))[1, 2],
              "{short_names$x} logFC" := min(x[[glue("{short_names$x} logFC")]]),
              "{short_names$y} logFC" := max(x[[glue("{short_names$y} logFC")]])
            )
          }, mc.cores = threads
        ) %>% 
        bind_rows(.id = "feature") %>% 
        mutate(
          label = paste("rho ==", round(cor, 3)),
          "{short_names$x} logFC" := min(!!sym(glue("{short_names$x} logFC"))),
          "{short_names$y} logFC" := max(!!sym(glue("{short_names$y} logFC")))
        ),
      colour = "grey20", parse = TRUE, vjust = 1, hjust = 0, size = 3.5
    ) +
    facet_wrap(~feature) +
    scale_x_continuous() +
    scale_y_continuous() +
    scale_colour_manual(values = full_palette, labels = label_wrap(15)) +
    labs(colour = "Combined Status") +
    theme(legend.position = "bottom")
}
```


# HERE

## Pairwise Signal

```{r plot-pairwise-signal, fig.cap = glue("Pairwise signal levels with boxplots indicating the baseline signal whether signal was detected in both  or individual comparisons.")}
db_results %>% 
  plotPairwise(
    var = cpm_column, alpha = 0.6, side_panel_width = 0.2,
  ) +
  scale_fill_manual(
    values = setNames(
      comp_cols, paste(names(comp_cols), c("Only", "Only", "Detected"))
    )
  )
```


## Distribution By Region

```{r status-by-region, fig.cap = glue("*Distribution of signal patterns by genomic regions as defined in earlier sections. Windows are only shown if change was detected in one of the comparisons. Segments where changes were detected in both comparisons are exploded.*")}
all_windows %>% 
  filter(grepl("Up|Down", status)) %>% 
  mutate(status = fct_lump_prop(status, 0.01)) %>% 
  plotSplitDonut(
    outer = "region", 
    inner = "status",
    inner_palette = group_colours,
    outer_palette = region_colours,
    inner_glue = "{str_replace_all(.data[[inner]], ' - ', '\n')}\nn = {comma(n, 1)}\n{percent(p,0.1)}",
    outer_glue = "{str_wrap(.data[[outer]], 15)}\nn = {comma(n, 1)}\n{percent(p,0.1)}",
    outer_min_p = 0.025, inner_label_alpha = 0.8,
    explode_inner = "(Up|Down).+(Up|Down)", explode_r = 1/3
  )
```

`r if (has_features) "## Distribution By Feature"`

```{r status-by-feature, eval = has_features, echo = has_features, fig.cap = glue("*Distribution of signal patterns by external features as defined in earlier sections. Windows are only shown if change was detected in one of the comparisons.*")}
all_windows %>% 
  filter(grepl("Up|Down", status)) %>% 
  mutate(status = fct_lump_prop(status, 0.01)) %>% 
  plotSplitDonut(
    outer = "feature", 
    inner = "status",
    inner_palette = group_colours,
    outer_palette = feature_colours,
    inner_glue = "{str_replace_all(.data[[inner]], ' - ', '\n')}\nn = {comma(n, 1)}\n{percent(p,0.1)}",
    outer_glue = "{str_wrap(.data[[outer]], 15)}\nn = {comma(n, 1)}\n{percent(p,0.1)}",
    outer_min_p = 0.025, inner_label_alpha = 0.8,
    explode_inner = "(Up|Down).+(Up|Down)", explode_r = 1/3
  )
```


`r if (n_targets > 1) "## Distance By Status"`

```{r distances-bystatus, echo = n_targets > 1, eval = n_targets > 1, fig.height = 6, fig.cap = "*Distances between windows where maximal signal was detected for each target. Windows are only shown if change was detected in one or more comparisons, and if 10 or more windows were found in each group. The x-axis is truncated at 1kb*"}
all_windows %>% 
  as_tibble() %>% 
  dplyr::filter(!is.na(distance)) %>% 
  arrange(status, distance) %>% 
  group_by(status) %>% 
  mutate(
    n = dplyr::n(),
    q = seq_along(status) / n
  ) %>% 
  dplyr::filter(distance < 1e3, str_detect(status, "Up|Down"), n > 10) %>% 
  ggplot(aes(distance, q, colour = status)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(breaks = seq(0, 10, by = 2) * 100) +
  scale_y_continuous(
    labels = percent, expand = expansion(c(0, 0.05)), 
    breaks = seq(0, 1, by = 0.2)
  ) +
  scale_colour_manual(
    values = group_colours
  ) +
  labs(
    x = "Distance Between Window Centres (bp)",
    y = "Percentile",
    colour = "Combined Status"
  )
```


`r if (n_targets > 1) "## Distance by Region"`

```{r distance-by-region, echo = n_targets > 1, eval = n_targets > 1, fig.height = 6, fig.width = 8, fig.cap = "*Distribution of distances between peaks separated by genomic region. All sites are included regardless of changed signal. The x-axis is truncated 1kb.*"}
all_windows %>% 
  as_tibble() %>% 
  dplyr::filter(!is.na(distance)) %>% 
  arrange(region, distance) %>% 
  group_by(region) %>% 
  mutate(
    n = dplyr::n(),
    q = seq_along(region) / n
  ) %>% 
  dplyr::filter(distance < 1e3) %>% 
  ggplot(aes(distance, q, colour = region)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(breaks = seq(0, 10, by = 2) * 100) +
  scale_y_continuous(
    labels = percent, expand = expansion(c(0, 0.05)), 
    breaks = seq(0, 1, by = 0.2)
  ) +
  scale_colour_manual(values = region_colours) +
  labs(
    x = "Distance Between Window Centres (bp)",
    y = "Percentile",
    colour = "Genomic Region"
  )
```

`r if (has_features & n_targets > 1) "## Distance By Feature"`

```{r distance-by-feature, echo = has_features & n_targets > 1, eval = has_features & n_targets > 1, fig.height = 6, fig.width = 8, fig.cap = "*Distribution of distances between peaks separated by external feature. All sites are included regardless of changed signal. The x-axis is truncated 1kb.*"}
all_windows %>% 
  as_tibble() %>% 
  dplyr::filter(!is.na(distance)) %>% 
  mutate(feature = factor(feature, levels = names(feature_colours))) %>% 
  arrange(feature, distance) %>% 
  group_by(feature) %>% 
  mutate(
    n = dplyr::n(),
    q = seq_along(feature) / n
  ) %>% 
  dplyr::filter(distance < 1e3) %>% 
  droplevels() %>% 
  ggplot(aes(distance, q, colour = feature)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(breaks = seq(0, 10, by = 2) * 100) +
  scale_y_continuous(
    labels = percent, expand = expansion(c(0, 0.05)), 
    breaks = seq(0, 1, by = 0.2)
  ) +
  scale_colour_manual(values = unlist(feature_colours)) +
  labs(
    x = "Distance Between Window Centres (bp)",
    y = "Percentile",
    colour = "External Feature"
  )
```


# Combined Signal

## Heatmaps {.tabset}

A series of profile heatmaps were generated for any set of pairwise comparisons with more than 25 ranges, and where signal was detected as changed in at least one of the comparisons.
If comparing two targets, peaks were centred at the mid-point between the ranges of individual maximal signal.
If only one target is detected within a range, peaks are centred at the range with maximal signal.
Scales for fill and average signal (both logCPM) are held constant across all plots for easier comparison between groups.

```{r form-profile-heatmaps}
## Set the BWFL as a single object incorporating target & treatment names
fl <- seq_along(pairs) %>% 
  lapply(
    function(x) {
      file.path(
        macs2_path[[x]],
         # glue("{names(pairs)[x]}_{pairs[[x]]}_merged_FE.bw")
         glue("{names(pairs)[x]}_{pairs[[x]]}_merged_treat_pileup.bw")
      ) %>% 
        setNames(paste(names(pairs)[[x]], pairs[[x]]))
    } 
  ) %>% 
  unlist() %>% 
  .[!duplicated(.)]
bwfl_fe <- setNames(BigWigFileList(fl), names(fl))
n_bins <- 100
n_max <- 2000
grl_to_plot <- all_windows %>% 
  split(.$status) %>% 
  .[str_detect(names(.),"Up|Down")] %>% 
  .[vapply(., length, integer(1)) > 25] %>% # Only show groups with > 25 ranges
  lapply(
    mutate, 
    centre = case_when(
      is.na(!!sym(glue("{c1}_centre"))) ~ !!sym(glue("{c2}_centre")),
      !is.na(!!sym(glue("{c1}_centre"))) ~ !!sym(glue("{c1}_centre"))
    ),
    rng = paste0(seqnames, ":", as.integer(centre))
  ) %>% 
  lapply(function(x) GRanges(x$rng, seqinfo = sq)) %>% 
  GRangesList() %>% 
  endoapply(distinctMC) 
profile_width <- 5e3
x_lab <- c(
  seq(0, floor(0.5 * profile_width / 1e3), length.out = 3),
  seq(-floor(0.5 * profile_width / 1e3), 0, length.out = 3)
) %>% 
  sort() %>% 
  unique()
sig_profiles <- grl_to_plot %>% 
  endoapply(
    function(x) {
      if (length(x) <= n_max) return(x)
      id <- sample(length(x), n_max)
      sort(x[id])
    }
  ) %>% 
  mclapply(
    function(x) getProfileData(
      bwfl_fe, x, upstream = profile_width / 2, bins = n_bins, log = TRUE
    ),
    mc.cores = min(length(grl_to_plot), threads)
  )
profile_heatmaps <- sig_profiles %>% 
  mclapply(
    plotProfileHeatmap,
    profileCol = "profile_data", 
    xLab = "Distance from Centre (bp)",
    fillLab = "log2 CPM",
    colour = "name",
    mc.cores = min(length(.), threads)
  ) 
fill_range <- profile_heatmaps %>% 
  lapply(function(x) x$data[,"score"]) %>% 
  unlist() %>%
  c(0) %>% 
  range()
sidey_range <- profile_heatmaps %>% 
  lapply(function(x) x$layers[[3]]$data$y) %>% 
  unlist() %>% 
  range()
profile_heatmaps <- profile_heatmaps %>% 
  lapply(
    function(x) {
      x + 
        scale_x_continuous(
          breaks = x_lab * 1e3, labels = x_lab, expand = expansion(0)
        ) +
        scale_xsidey_continuous(
          limits = sidey_range, expand = expansion(c(0, 0.1))
        ) +
        scale_fill_gradientn(colours = colours$heatmaps, limits = fill_range) +
        scale_colour_manual(
          values = lapply(
            seq_along(pairs), 
            function(i) setNames(
              treat_colours[pairs[[i]]], 
              paste(names(pairs)[[i]], pairs[[i]])
            )
          ) %>%
            unlist() %>% 
            .[!duplicated(names(.))]
        ) +
        labs(
          x = "Distance from Centre (kb)",
          fill = "log2 CPM", linetype = "Union\nPeak\nOverlap",
          colour = "Sample"
        ) 
    }
  )
```

```{r plot-all-profile-heatmaps}
htmltools::tagList(
  mclapply(
    seq_along(profile_heatmaps),
    function(x) {
      ## Export the image
      pw <- names(profile_heatmaps)[[x]] 
      img_out <- file.path(
        fig_path,
        pw %>% 
          str_remove_all("(:|Vs. |- )") %>% 
          str_replace_all(" ", "_") %>% 
          str_replace_all("_-_", "-") %>% 
          paste0("_profile_heatmap.", fig_type)
      )
      n_ranges <- length(grl_to_plot[[x]])
      n_samples <- length(unique(profile_heatmaps[[x]]$data$name))
      fig_fun(
        filename = img_out,
        width = knitr::opts_current$get("fig.width") * n_samples / 4, 
        height = min(
          3.5 + knitr::opts_current$get("fig.height") * n_ranges / 1.5e3, 
          10
        )
      )
      print(profile_heatmaps[[x]])
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "
        {n_ranges} ranges were defined as showing the pattern {pw}. Ranges were 
        centered at the midpoint between ranges identified as the maximal 
        signal range, based on {c1} only. If more than {n_max} ranges were 
        defined, plots were restricted to {n_max} ranges. Values showb are log2 
        CPM taking the SPMR-based bigwig files produced by macs callpeak.
        "
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level3",
          htmltools::h3(pw),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    },
    mc.cores = threads
  )
)
```


## Windows With The Strongest Signal {.tabset}

For the initial set of visualisations, the sets of windows with the strongest signal were selected from within each group.
Groups were restricted to those where differential signal was seen in at least comparison.

```{r grl-to-plot-avexpr}
grl_to_plot <- all_windows %>% 
  filter(
    !str_detect(status, "Un.+Un"),
    vapply(detected, length, integer(1)) > 0
  ) %>% 
  as_tibble() %>% 
  rename_all(
    str_replace_all, pattern = "\\.Vs\\.\\.", replacement = " Vs. "
  ) %>% 
  rename_all(
    str_replace_all, pattern = "\\.\\.", replacement = ": "
  ) %>% 
  nest(
    AveExpr = ends_with("AveExpr")
  ) %>% 
  mutate(
    max_signal = vapply(AveExpr, function(x) max(unlist(x)), numeric(1))
  ) %>% 
  group_by(status) %>% 
  filter(max_signal == max(max_signal)) %>% 
  dplyr::select(
    range, status, ends_with("logFC")
  ) %>% 
  droplevels() %>% 
  split(.$status) %>% 
  lapply(pull, "range") %>% 
  lapply(function(x) x[1]) %>% 
  lapply(GRanges, seqinfo = sq) %>% 
  GRangesList()
```

```{r prepare-plot-objects}
## The coverage
bwfl <- seq_along(pairs) %>% 
  lapply(
    function(x) {
      file.path(
        macs2_path[[x]], 
        glue("{names(pairs)[x]}_{pairs[[x]]}_merged_treat_pileup.bw")
      ) %>% 
        BigWigFileList() %>% 
        setNames(pairs[[x]])
    } 
  ) %>% 
  setNames(comps)
line_col <- lapply(bwfl, function(x) treat_colours[names(x)])

## Coverage annotations
annot <- comps %>% 
  lapply(
    function(x) {
      col <- glue("{x}_status")
      select(all_windows, all_of(col)) %>% 
        mutate(
          status = str_extract(!!sym(col), "Down|Up|Unchanged|Undetected")
        ) %>% 
        splitAsList(.$status) %>% 
        lapply(granges) %>% 
        lapply(unique) %>% 
        GRangesList()
    }
  ) %>% 
  setNames(comps)
annot_col <- unlist(colours$direction) %>% 
  setNames(str_to_title(names(.)))

## Coverage y-limits
gr <- unlist(grl_to_plot)
y_lim <- lapply(
  bwfl,
  function(x) {
    cov <- lapply(x, import.bw, which = gr)
    unlist(lapply(cov, function(y) max(y$score))) %>% 
      c(0) %>% 
      range()
  }
)

## The features track
feat_gr <- gene_regions %>% 
  lapply(granges) %>% 
  GRangesList()
feat_col_hfgc <- colours$regions
if (has_features) {
  feat_gr <- list(Regions = feat_gr)
  feat_gr$Features <- splitAsList(external_features, external_features$feature)
  feat_col_hfgc <- list(
    Regions = unlist(colours$regions),
    Features = unlist(colours$features)
  )
}

## The genes track
hfgc_genes <- trans_models
gene_col <- "grey"
if (has_rnaseq) {
  rna_lfc_col <- colnames(rnaseq)[str_detect(str_to_lower(colnames(rnaseq)), "logfc")][1]
  rna_fdr_col <- colnames(rnaseq)[str_detect(str_to_lower(colnames(rnaseq)), "fdr|adjp")][1]
  if (!is.na(rna_lfc_col) & !is.na(rna_fdr_col)) {
    hfgc_genes <- trans_models %>% 
      mutate(
        status = case_when(
          !gene %in% rnaseq$gene_id ~ "Undetected",
          gene %in% dplyr::filter(
            rnaseq, !!sym(rna_lfc_col) > 0, !!sym(rna_fdr_col) < fdr_alpha
          )$gene_id ~ "Up",
          gene %in% dplyr::filter(
            rnaseq, !!sym(rna_lfc_col) < 0, !!sym(rna_fdr_col) < fdr_alpha
          )$gene_id ~ "Down",
          gene %in% dplyr::filter(
            rnaseq, !!sym(rna_fdr_col) >= fdr_alpha
          )$gene_id ~ "Unchanged",
        )
      ) %>% 
      splitAsList(.$status) %>% 
      lapply(select, -status) %>% 
      GRangesList()
    gene_col <- colours$direction %>% 
      setNames(str_to_title(names(.)))
  }
}

## External Coverage (Optional)
if (!is.null(config$external$coverage)) {
  ext_cov_path <- config$external$coverage %>% 
    lapply(unlist) %>% 
    lapply(function(x) setNames(here::here(x), names(x)))
  bwfl <- c(
    bwfl[comps],
    lapply(ext_cov_path, function(x) BigWigFileList(x) %>% setNames(names(x)))
  )
  line_col <- c(
    line_col[comps],
    ext_cov_path %>% 
      lapply(
        function(x) {
          missing <- setdiff(names(x), names(treat_colours))
          cmn <- intersect(names(x), names(treat_colours))
          col <- setNames(character(length(names(x))), names(x))
          if (length(cmn) > 0) col[cmn] <- treat_colours[cmn]
          if (length(missing) > 0) 
            col[missing] <- hcl.colors(
              max(5, length(missing)), "Zissou 1")[seq_along(missing)]
          col
        }
      )
  )
  
  y_ranges <- grl_to_plot %>% 
    unlist() %>% 
    granges() %>% 
    resize(w = 10 * width(.), fix = 'center')
  
  y_lim <- c(
    y_lim[comps],
    bwfl[names(config$external$coverage)] %>% 
      lapply(
        function(x) {
          GRangesList(lapply(x, import.bw, which = y_ranges)) %>% 
            unlist() %>% 
            filter(score == max(score)) %>% 
            mcols() %>% 
            .[["score"]] %>% 
            c(0) %>% 
            range()
        }
      )
  )
  
}
```



```{r plot-top-ranges-aveexpr}
htmltools::tagList(
  mclapply(
    seq_along(grl_to_plot),
    function(x) {
      ## Export the image
      img_out <- file.path(
        fig_path,
        names(grl_to_plot)[[x]] %>% 
          str_remove_all("(:|Vs. |- )") %>% 
          str_replace_all(" ", "_") %>% 
          paste0("_AveExpr.", fig_type)
      )
      fig_fun(
        filename = img_out, width = 10, height = 8
      )
      ct <- FALSE
      if (length(subsetByOverlaps(gtf_transcript, grl_to_plot[[x]])) > 10)
        ct <- "meta"
      plotHFGC(
        grl_to_plot[[x]],
        features = feat_gr, featcol = feat_col_hfgc, 
        featsize = 1 + has_features,
        genes = hfgc_genes, genecol = gene_col,
        coverage = bwfl, linecol = line_col,
        annotation = annot, annotcol = annot_col,
        cytobands = bands_df,
        rotation.title = 90,
        zoom = 10,
        ylim = y_lim,
        collapseTranscripts = ct,
        col.title = "black", background.title = "white", showAxis = FALSE
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      gn <- unlist(subsetByOverlaps(all_windows, grl_to_plot[[x]])$detected)
      cp <- htmltools::tags$em(
        glue(
          "
        Highlighted region corresponds to the highest signal within the group 
        {names(grl_to_plot)[[x]]}. The most likely target 
        {ifelse(length(gn) == 1, 'gene is ', 'genes are ')}
        {collapseGenes(gn, format = '')}
        "
        )
      )
      htmltools::div(
        htmltools::div(
          id = names(grl_to_plot)[[x]] %>% 
            str_remove_all("(:|Vs. |- )") %>% 
            str_replace_all(" ", "-") %>% 
            str_to_lower() %>% 
            paste0("-aveexpr"),
          class = "section level3",
          htmltools::h3(names(grl_to_plot)[[x]]),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = 960),
            htmltools::p(
              class = "caption", htmltools::tags$em(cp)
            )
          )
        )
      )
    },
    mc.cores = threads
  )
)  
```

## Windows With The Largest Change {.tabset}

For the initial set of visualisations, the sets of windows with the strongest signal were selected from within each group.
Groups were restricted to those where differential signal was seen in at least comparison.

```{r grl-to-plot-logfc}
grl_to_plot <- all_windows %>% 
  filter(
    !str_detect(status, "Un.+Un"),
    vapply(detected, length, integer(1)) > 0
  ) %>% 
  as_tibble() %>% 
  rename_all(
    str_replace_all, pattern = "\\.Vs\\.\\.", replacement = " Vs. "
  ) %>% 
  rename_all(
    str_replace_all, pattern = "\\.\\.", replacement = ": "
  ) %>% 
  nest(
    logFC = ends_with("logFC")
  ) %>% 
  mutate(
    max_logFC = vapply(logFC, function(x) max(abs(unlist(x))), numeric(1))
  ) %>% 
  group_by(status) %>% 
  filter(max_logFC == max(max_logFC)) %>% 
  dplyr::select(
    range, status, ends_with("logFC")
  ) %>% 
  droplevels() %>% 
  split(.$status) %>% 
  lapply(pull, "range") %>% 
  lapply(GRanges, seqinfo = sq) %>% 
  GRangesList()
```

```{r update-plot-objects}
## Coverage y-limits
gr <- unlist(grl_to_plot)
y_lim <- lapply(
  bwfl,
  function(x) {
    cov <- lapply(x, import.bw, which = gr)
    unlist(lapply(cov, function(y) max(y$score))) %>% 
      c(0) %>% 
      range()
  }
)

## External Coverage (Optional)
if (!is.null(config$external$coverage)) {
  y_ranges <- grl_to_plot %>% 
    unlist() %>% 
    granges() %>% 
    resize(w = 10 * width(.), fix = 'center')
  
  y_lim <- c(
    y_lim[comps],
    bwfl[names(config$external$coverage)] %>% 
      lapply(
        function(x) {
          GRangesList(lapply(x, import.bw, which = y_ranges)) %>% 
            unlist() %>% 
            filter(score == max(score)) %>% 
            mcols() %>% 
            .[["score"]] %>% 
            c(0) %>% 
            range()
        }
      )
  )
  
}
```

```{r plot-top-ranges-logfc}
htmltools::tagList(
  mclapply(
    seq_along(grl_to_plot),
    function(x) {
      ## Export the image
      img_out <- file.path(
        fig_path,
        names(grl_to_plot)[[x]] %>% 
          str_remove_all("(:|Vs. |- )") %>% 
          str_replace_all(" ", "_") %>% 
          paste0("_logFC.", fig_type)
      )
      fig_fun(
        filename = img_out, width = 10, height = 8
      )
      ct <- FALSE
      if (length(subsetByOverlaps(gtf_transcript, grl_to_plot[[x]])) > 10)
        ct <- "meta"
      plotHFGC(
        grl_to_plot[[x]],
        features = feat_gr, featcol = feat_col_hfgc,
        featsize = 1 + has_features,
        genes = hfgc_genes, genecol = gene_col,
        coverage = bwfl, linecol = line_col,
        annotation = annot, annotcol = annot_col,
        cytobands = bands_df,
        rotation.title = 90,
        zoom = 10,
        ylim = y_lim,
        collapseTranscripts = ct,
        col.title = "black", background.title = "white", showAxis = FALSE
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      gn <- unlist(subsetByOverlaps(all_windows, grl_to_plot[[x]])$detected)
      cp <- htmltools::tags$em(
        glue(
          "
        Highlighted region corresponds to the largest change within the group 
        {names(grl_to_plot)[[x]]}. The most likely target 
        {ifelse(length(gn) == 1, 'gene is ', 'genes are ')}
        {collapseGenes(gn, format = '')}
        "
        )
      )
      htmltools::div(
        htmltools::div(
          id = names(grl_to_plot)[[x]] %>% 
            str_remove_all("(:|Vs. |- )") %>% 
            str_replace_all(" ", "-") %>% 
            str_to_lower() %>% 
            paste0("-logfc"),
          class = "section level3",
          htmltools::h3(names(grl_to_plot)[[x]]),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = 960),
            htmltools::p(
              class = "caption", htmltools::tags$em(cp)
            )
          )
        )
      )
    },
    mc.cores = threads
  )
)  
```

# Enrichment Analysis 


```{r msigdb}
min_gs_size <- extra_params$enrichment$min_size
if (is.null(min_gs_size) | is.na(min_gs_size)) min_gs_size <- 0
max_gs_size <- extra_params$enrichment$max_size
if (is.null(max_gs_size) | is.na(max_gs_size)) max_gs_size <- Inf
min_sig <- extra_params$enrichment$min_sig
if (is.null(min_sig) | is.na(min_sig)) min_sig <- 1

all_ids <- gtf_gene %>%
  mutate(w = width) %>% 
  as_tibble() %>% 
  dplyr::select(gene_id, w) %>% 
  arrange(desc(w)) %>% 
  distinct(gene_id, w) %>% 
  arrange(gene_id) %>% 
  left_join(
    all_windows$gene_id %>%
      unlist() %>%
      table() %>%
      enframe(name = "gene_id", value = "n_windows"),
    by = "gene_id"
  ) %>% 
  dplyr::filter(gene_id %in% gtf_gene$gene_id) %>% 
  mutate(
    n_windows = as.integer(ifelse(is.na(n_windows), 0, n_windows))
  )
msigdb <- msigdbr(species = extra_params$enrichment$species) %>% 
  dplyr::filter(
    gs_cat %in% unlist(extra_params$enrichment$msigdb$gs_cat) |
      gs_subcat %in% unlist(extra_params$enrichment$msigdb$gs_subcat),
    str_detect(ensembl_gene, "^E")
  ) %>% 
  dplyr::rename(gene_id = ensembl_gene, gene_name = gene_symbol) %>% # For easier integration 
  dplyr::select(-starts_with("human"), -contains("entrez")) %>% 
  dplyr::filter(gene_id %in% all_ids$gene_id) %>% 
  group_by(gs_name) %>% 
  mutate(n = dplyr::n()) %>% 
  ungroup() %>% 
  dplyr::filter(n >= min_gs_size, n <= max_gs_size) 
gs_by_gsid <- msigdb %>% 
  split(.$gs_name) %>% 
  mclapply(pull, "gene_id", mc.cores = threads) %>% 
  mclapply(unique, mc.cores = threads)
gs_by_geneid <- msigdb %>% 
  split(.$gene_id) %>% 
  mclapply(pull, "gs_name", mc.cores = threads) %>% 
  mclapply(unique, mc.cores = threads)
gs_url <- msigdb %>% 
  distinct(gs_name, gs_url) %>%
  mutate(
    gs_url = ifelse(
      gs_url == "", "
      http://www.gsea-msigdb.org/gsea/msigdb/collections.jsp",
      gs_url
    )
  ) %>% 
  with(setNames(gs_url, gs_name))

min_network_size <- extra_params$networks$min_size
if (is.null(min_network_size) | is.na(min_network_size))
  min_network_size <- 2

max_network_size <- extra_params$networks$max_size
if (is.null(max_network_size) | is.na(max_network_size))
  max_network_size <- Inf

max_network_dist <- extra_params$networks$max_distance
if (is.null(max_network_dist) | is.na(max_network_dist))
  max_network_dist <- 1

net_layout <- extra_params$networks$layout
enrich_alpha <- extra_params$enrichment$alpha
adj_method <- match.arg(extra_params$enrichment$adj, p.adjust.methods)
adj_desc <- case_when(
  p.adjust.methods %in% c("fdr", "BH") ~ "the Benjamini-Hochberg FDR",
  p.adjust.methods %in% c("BY") ~ "the Benjamini-Yekutieli FDR",
  p.adjust.methods %in% c("bonferroni") ~ "the Bonferroni",
  p.adjust.methods %in% c("holm") ~ "Holm's",
  p.adjust.methods %in% c("hommel") ~ "Hommel's",
  p.adjust.methods %in% c("hochberg") ~ "Hochberg's",
  p.adjust.methods %in% c("none") ~ "no"
) %>% 
  setNames(p.adjust.methods)
comp_ids <- comps %>% 
  lapply(
    function(x) {
      subset(all_windows, !str_detect(status, paste(x, "Undetected")))$gene_id
    }
  ) %>% 
  lapply(unlist) %>% 
  lapply(unique) %>% 
  lapply(intersect, gtf_gene$gene_id) %>% 
  setNames(comps)
mapped_ids <- comp_ids %>% 
  unlist() %>% 
  unique()
gene_lengths <- structure(width(gtf_gene), names = gtf_gene$gene_id)[mapped_ids]
group_ids <- all_windows %>% 
  split(f = .$status) %>% 
  mclapply(
    function(x) unique(unlist(x$gene_id)),
    mc.cores = threads
  ) 
```

## Comparison of Genes Mapped To Targets

The first level of enrichment testing performed was to simply check the genes mapped to a site in either comparison, with no consideration being paid to the dynamics of signal in either comparison.
This will capture and describe the overall biological activity being jointly regulated.

The genes in each section and intersection in the following Venn Diagram will then be analysed against the complete set of `r comma(length(unique(unlist(comp_ids))))` genes mapped to either `r glue_collapse(comps, last = ' or ')`.
Firstly, all `r comma(length(mapped_ids))` genes mapped to a site were tested for enrichment in comparison to the set of `r comma(length(unique(gtf_gene$gene_id)))` `r ifelse(has_rnaseq, "detected", "annotated")` genes.
For this analysis, gene width was used as a potential source of sampling bias and Wallenius' Non-Central Hypergeometric was used as implemented in `goseq` [@YoungGoseq2010].

The sets of genes mapped to each section of the Venn Diagram were then tested for enrichment in comparison to the larger set of `r comma(length(mapped_ids))` mapped genes.
The standard hypergeometric distribution was used for these analyses.

If `r min_network_size` or more gene-sets were considered to be enriched, a network plot was produced using the same methodology as for results from differential signal for individual comparisons.

```{r venn-shared-genes, results = 'hide', fig.show = 'asis', fig.cap = glue("*Venn Diagram showing __genes__ mapped to a signal region associated with {glue_collapse(comps, sep = ' or ')}. Of the {comma(nrow(all_ids))} genes under consideration, a total of {comma(length(mapped_ids))} genes were mapped to a signal region across both comparisons, leaving {comma(nrow(all_ids) - length(mapped_ids))} genes unmapped to a signal region across either comparison.*")}
plotOverlaps(
  comp_ids, set_col = comp_cols[comps], cex = 1.2, cat.cex = 1.2
)
```

### Result Tables {.tabset}

#### Mapped Genes To Either `r glue_collapse(comps, last = " or ")`

```{r goseq-mapped-res, fig.show = 'hide'}
pwf_mapped <- gtf_gene %>%
  mutate(w = width) %>%
  as_tibble() %>% 
  arrange(gene_id, desc(w)) %>%
  distinct(gene_id, w) %>% 
  mutate(mapped = gene_id %in% mapped_ids) %>% 
  with(nullp(setNames(mapped, gene_id), bias.data = log10(w)))
goseq_mapped_res <- pwf_mapped %>% 
  goseq(gene2cat = gs_by_geneid) %>% 
  dplyr::filter(numDEInCat > 0) %>% 
  as_tibble() %>% 
  dplyr::select(
    gs_name = category, pval = over_represented_pvalue, starts_with("num")
  ) %>% 
  mutate(
    adj_p = p.adjust(pval, adj_method),
    enriched = adj_p < enrich_alpha
  ) %>% 
  dplyr::filter(numDEInCat >= min_sig) %>% 
  left_join(
    dplyr::filter(msigdb, gene_id %in% mapped_ids)[c("gs_name", "gene_name", "gs_url", "gs_description")],
    by = "gs_name"
  ) %>% 
  chop(gene_name) %>% 
  mutate(
    gene_name = vapply(gene_name, paste, character(1), collapse = "; ")
  )
any_goseq_mapped <- sum(goseq_mapped_res$adj_p < enrich_alpha) > 0
tg_mapped <- make_tbl_graph(
  goseq_mapped_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, rownames(subset(pwf_mapped, DEgenes))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_mapped <- length(tg_mapped) >= min_network_size
```


```{r tbl-goseq-mapped-res, eval = any_goseq_mapped, echo = any_goseq_mapped}
tbl <- goseq_mapped_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = "Mapped Genes",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes in Gene-Set",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
cp <- glue(
  "
  All {comma(length(mapped_ids))} genes with a mapped window from 
  either comparison were compared to the complete set of 
  {comma(length(unique(gtf_gene$gene_id)))} annotated genes, in order to 
  ascertain which key gene-sets and pathways were likely to be targeted in either analysis. 
  ",
  ifelse(
    has_rnaseq,
    "Genes were restricted to those in the set of detected genes from the provided RNA-Seq dataset. ", 
    ""
  ),
  "
  Gene-length was included as the source of any potential sampling bias for a 
  gene being mapped to any target-signal range. This analysis provides an overall 
  viewpoint as to which pathways are regulated across either dataset.
  "
)
div(class = "table",
    div(class = "table-header", div(class = "caption", cp)),
    tbl
)
```

`r if(!any_goseq_mapped) "No pathway enrichment was detected amongst genes mapped in either comparison."`


`r if (n_targets == 2) glue("#### Genes Mapped to Both ", glue_collapse(comps, last = " and " ))`

```{r empty-goseq-t1t2-res, echo = FALSE}
goseq_t1t2_res <- tibble(
  gs_name = character(), mapped = numeric(), enriched = logical(),
  adj_p = numeric()
)
vec_t1t2 <- structure(
  mapped_ids %in% intersect(comp_ids[[2]], comp_ids[[1]]),
  names = mapped_ids
)
any_goseq_t1t2 <- plot_network_t1t2 <- FALSE
para <- glue(
  "
  All {comma(length(intersect(comp_ids[[2]], comp_ids[[1]])))} genes with a 
  mapped window to both {glue_collapse(comps, last = ' and ')} were compared to 
  the {comma(length(mapped_ids))} genes mapped to either, 
  in order to ascertain whether any key gene-sets were likely to be targeted by 
  both factors in combination. No sampling bias offset was incorporated, instead 
  using a simple Hypergeometric model for enrichment testing.
  "
)
```


```{r goseq_t1t2_res, eval = n_targets == 2, fig.show = 'hide'}
pwf_t1t2 <- nullp(
  vec_t1t2,
  bias.data = log10(setNames(all_ids$w, all_ids$gene_id)[mapped_ids]), 
  plot.fit = TRUE
) 
if (sum(pwf_t1t2$DEgenes) > 0) {
  goseq_t1t2_res <- pwf_t1t2 %>% 
    goseq(gene2cat = gs_by_geneid, method = gs_method) %>% 
    dplyr::filter(numDEInCat > 0) %>% 
    mutate(
      adj_p = p.adjust(pval, adj_method),
      enriched = adj_p < enrich_alpha,
      DEgenes = lapply(DEgenes, function(x) id2gene[x])
    ) %>% 
    dplyr::filter(numDEInCat >= min_sig) %>% 
    left_join(
      distinct(msigdb, gs_name, gs_url, gs_description),
      by = "gs_name"
    ) %>% 
    dplyr::rename(gene_name = DEgenes) %>% 
    mutate(gene_name = vapply(gene_name, paste, character(1), collapse = "; "))
}
any_goseq_t1t2 <- sum(goseq_t1t2_res$adj_p < enrich_alpha) > 0
tg_t1t2 <- make_tbl_graph(
  goseq_t1t2_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, names(which(vec_t1t2))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_t1t2 <- length(tg_t1t2) >= min_network_size
if (!any_goseq_t1t2) para <- glue(para, " No enrichment was found.")
```

`r if (!any_goseq_t1t2) para`

```{r tbl-goseq-t1t2-res, eval = any_goseq_t1t2 & n_targets == 2, echo = any_goseq_t1t2}
tbl <- goseq_t1t2_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = "Genes Mapped to Both",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes Mapped to Either",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
div(class = "table",
    div(class = "table-header", div(class = "caption", para)),
    tbl
)
```


`r if (n_targets == 2) glue("#### Genes Mapped to {comps[[1]]} Only")`


```{r empty-goseq-t1only}
goseq_t1only_res <- tibble(
  gs_name = character(), mapped = numeric(), enriched = logical(),
  adj_p = numeric()
)
vec_t1only <- structure(
  mapped_ids %in% setdiff(comp_ids[[1]], comp_ids[[2]]),
  names = mapped_ids
)
gs_method <- "Hypergeometric"
any_goseq_t1only <- plot_network_t1only <- FALSE
para <- glue(
  "
  All {comma(length(setdiff(comp_ids[[1]], comp_ids[[2]])))} genes with a 
  mapped window to only {comps[[1]]} were compared to 
  the {comma(length(mapped_ids))} genes mapped to either
  {glue_collapse(comps, last = ' or ')}, in order to ascertain whether any gene-sets 
  were likely to be targeted exclusively. No sampling bias offset was 
  incorporated, instead using a simple Hypergeometric model for enrichment testing.
  "
)
```

```{r goseq-t1only-res, fig.show = 'hide'}
pwf_t1only <- nullp(
  vec_t1only,
  bias.data = log10(setNames(all_ids$w, all_ids$gene_id)[mapped_ids]), 
  plot.fit = TRUE
) 
if (sum(pwf_t1only$DEgenes) > 0) {
  goseq_t1only_res <- pwf_t1only %>% 
    goseq(gene2cat = gs_by_geneid, method = gs_method) %>% 
    dplyr::filter(numDEInCat > 0) %>% 
    mutate(
      adj_p = p.adjust(pval, adj_method),
      enriched = adj_p < enrich_alpha,
      DEgenes = lapply(DEgenes, function(x) id2gene[x])
    ) %>% 
    dplyr::filter(numDEInCat >= min_sig) %>% 
    left_join(
      distinct(msigdb, gs_name, gs_url, gs_description),
      by = "gs_name"
    ) %>% 
    dplyr::rename(gene_name = DEgenes) %>% 
    mutate(gene_name = vapply(gene_name, paste, character(1), collapse = "; "))
}
any_goseq_t1only <- sum(goseq_t1only_res$adj_p < enrich_alpha) > 0
tg_t1only <- make_tbl_graph(
  goseq_t1only_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, names(which(vec_t1only))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_t1only <- length(tg_t1only) >= min_network_size
if (!any_goseq_t1only) para <- glue(para, " No enrichment was found.")
```

`r if (!any_goseq_t1only) para`

```{r tbl-goseq-t1only-res, eval = n_targets == 2 & any_goseq_t1only, echo = any_goseq_t1only}
tbl <- goseq_t1only_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = glue("Genes Mapped to {comps[[1]]} Only"),
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes Mapped to Both",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
div(class = "table",
    div(class = "table-header", div(class = "caption", para)),
    tbl
)
```


`r if (n_targets == 2) glue("#### Genes Mapped to {comps[[2]]} Only")`


```{r empty-goseq-t2only}
goseq_t2only_res <- tibble(
  gs_name = character(), mapped = numeric(), enriched = logical(),
  adj_p = numeric()
)
vec_t2only <- structure(
  mapped_ids %in% setdiff(comp_ids[[2]], comp_ids[[1]]),
  names = mapped_ids
)
gs_method <- "Hypergeometric"
para <- glue(
  "
  All {comma(length(setdiff(comp_ids[[2]], comp_ids[[1]])))} genes with a 
  mapped window to only {comps[[2]]} were compared to 
  the {comma(length(mapped_ids))} genes mapped to either 
  {glue_collapse(comps, last = ' or ')}, in order to ascertain whether any gene-sets 
  were likely to be targeted exclusively. No sampling bias offset was 
  incorporated, instead using a simple Hypergeometric model for enrichment testing.
  "
)
any_goseq_t2only <- plot_network_t2only <- FALSE
```

```{r goseq-t2only-res, fig.show = 'hide'}
pwf_t2only <- nullp(
  vec_t2only,
  bias.data = log10(setNames(all_ids$w, all_ids$gene_id)[mapped_ids]), 
  plot.fit = TRUE
) 
if (sum(pwf_t2only$DEgenes) > 0) {
  goseq_t2only_res <- pwf_t2only %>% 
    goseq(gene2cat = gs_by_geneid, method = gs_method) %>% 
    dplyr::filter(numDEInCat > 0) %>% 
    mutate(
      adj_p = p.adjust(pval, adj_method),
      enriched = adj_p < enrich_alpha,
      DEgenes = lapply(DEgenes, function(x) id2gene[x])
    ) %>% 
    dplyr::filter(numDEInCat >= min_sig) %>% 
    left_join(
      distinct(msigdb, gs_name, gs_url, gs_description),
      by = "gs_name"
    ) %>% 
    dplyr::rename(gene_name = DEgenes) %>% 
    mutate(gene_name = vapply(gene_name, paste, character(1), collapse = "; "))
}
any_goseq_t2only <- sum(goseq_t2only_res$adj_p < enrich_alpha) > 0
tg_t2only <- make_tbl_graph(
  goseq_t2only_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, names(which(vec_t2only))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_t2only <- length(tg_t2only) >= min_network_size
if (!any_goseq_t2only) para <- glue(para, " No enrichment was found.")
```

`r if (!any_goseq_t2only) para`

```{r tbl-goseq-t2only-res, eval = any_goseq_t2only, echo = any_goseq_t2only}
tbl <- goseq_t2only_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = glue("Genes Mapped to {comps[[2]]} Only"),
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes Mapped to Both",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
div(class = "table",
    div(class = "table-header", div(class = "caption", para)),
    tbl
)
```


### Network Plots {.tabset}

`r if (plot_network_mapped) "#### All Mapped Genes"`

```{r plot-network-mapped, eval = plot_network_mapped, echo = plot_network_mapped, fig.cap = glue("*Network plot showing  gene-sets enriched amongst the overall set of sites with signal detected in __either__ comparison.*")}
tg_mapped %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_mapped) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```

`r if (plot_network_t1t2) paste("#### Genes Mapped to Both", glue_collapse(comps, last = ' and '))`

```{r plot-network-t1t2, eval = plot_network_t1t2, echo = plot_network_t1t2, fig.cap = glue("*Network plot showing gene-sets enriched amongst the overall set of sites with detected signal in __both__ {comps[[1]]} and {comps[[2]]}.*")}
tg_t1t2 %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_t1t2) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```

`r if (plot_network_t1only) glue("#### Genes Mapped to {comps[[1]]} Only")`

```{r plot-network-t1only, eval = plot_network_t1only, echo = plot_network_t1only, fig.cap = glue("*Network plot showing all gene-sets enriched amongst the overall set of sites with detected signal in __{comps[[1]]} only__.*")}
tg_t1only %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_t1only) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```

`r if (plot_network_t2only) glue("#### Genes Mapped to {comps[[2]]} Only")`

```{r plot-network-t2only, eval = plot_network_t2only, echo = plot_network_t2only, fig.cap = glue("*Network plot showing all gene-sets enriched amongst the overall set of sites with detected signal in __{comps[[2]]} only__.*")}
tg_t2only %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_t2only) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```


## Comparison of Differential Signal Regions

```{r goseq-groups, fig.show = 'hide'}
n_peaks <- structure(all_ids$n_windows, names = all_ids$gene_id)[mapped_ids]
goseq_group_res <- group_ids %>% 
  .[str_detect(names(.), "Up|Down")] %>% 
  .[vapply(., length, integer(1)) > 0] %>% 
  mclapply(
    function(x) {
      vec <- structure(mapped_ids %in% x, names = mapped_ids)
      res <- tibble(adj_p = numeric())
      if (sum(vec) > 0) {
        res <- data.frame(DEgenes = vec) %>% 
          goseq_hyper(gene2cat = gs_by_geneid, cores = 1) %>% 
          dplyr::filter(numDEInCat > 0) %>% 
          mutate(adj_p = p.adjust(pval, adj_method)) %>% 
          dplyr::filter(numDEInCat >= min_sig) %>% 
          dplyr::select(gs_name, pval, adj_p, ends_with("InCat"), DEgenes) %>% 
          left_join(
            dplyr::distinct(msigdb, gs_name, gs_url, gs_description),
            by = "gs_name"
          ) %>% 
          mutate(
            gene_name = vapply(
              DEgenes, function(x) paste(id2gene[x], collapse = "; "), character(1)
            )
          )
      }
      res
    },
    mc.cores = threads
  ) %>% 
  ## This step at the end basically removes any groups with no results
  lapply(list) %>% 
  as_tibble() %>% 
  pivot_longer(everything(), names_to = "group") %>% 
  unnest(everything()) %>% 
  split(.$group)
goseq_group_sig <- goseq_group_res %>% 
  lapply(dplyr::filter, adj_p < enrich_alpha) %>% 
  .[vapply(., nrow, integer(1)) > 0]
tg_group_res <- names(goseq_group_sig) %>% 
  lapply(
    function(x) {
      make_tbl_graph(
        goseq_group_sig[[x]],
        gs_by_gsid[goseq_group_sig[[x]]$gs_name] %>%
          lapply(intersect, group_ids[[x]])
      )
    }
  ) %>% 
  setNames(names(goseq_group_sig)) %>% 
  .[vapply(., length, integer(1)) >= min_network_size]
```

The same MSigDB gene sets were used for associating any pathways with the combined changes in both `r collapseGenes(comps)`.
The 'universe' of genes was defined as all `r comma(length(mapped_ids))` gene ids with a window mapped from either target.
No sampling bias was included, with test results using a simple Hypergeometric Distribution.
`r if(length(goseq_group_sig) == 0) "No enrichment was found amongst any group of windows"`


`r if(length(goseq_group_sig) > 0) "### Results {.tabset}"`


```{r tbl-goseq-groups, eval = length(goseq_group_sig) > 0, echo = length(goseq_group_sig) > 0}
htmltools::tagList(
  names(goseq_group_sig) %>% 
    mclapply(
      function(x) {
        cp <- glue(
          "
          All {comma(length(group_ids[[x]]), 1)} genes mapped to windows
           with the pattern {x} were compared to the complete set of 
          {comma(length(mapped_ids), 1)} genes mapped to at least one
          window. {nrow(goseq_group_sig[[x]])} gene-sets were considered as 
          enriched using a threshold of {enrich_alpha} for adjusted-p-values.
          "
        )
        htmltools::div(
          htmltools::div(
            id = x %>% 
              str_to_lower %>% 
              str_replace_all(" ", "-") %>% 
              str_replace_all("-+", "-"),
            class = "section level4",
            htmltools::h4(class = "tabset", x),
            htmltools::tags$em(cp),
            goseq_group_sig[[x]] %>% 
              separate(group, into = comps, sep = " - ") %>% 
              dplyr::select(
                all_of(comps), gs_name, gs_description, pval, adj_p, everything()
              ) %>% 
              reactable(
                filterable = TRUE,
                showPageSizeOptions = TRUE,
                columns = list2(
                  "{comps[[1]]}" := colDef(
                    cell = function(value) {
                      value <- str_extract(value, "Up|Down|Unchanged|Undetected")
                      if (str_detect(value, "Up")) value <- paste(value, "\u21E7")
                      if (str_detect(value, "Down")) value <- paste(value, "\u21E9")
                      paste(targets[[1]], value)
                    },
                    style = function(value) {
                      colour <- case_when(
                        str_detect(value, "Up") ~ colours$direction[["up"]],
                        str_detect(value, "Down") ~ colours$direction[["down"]],
                        str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
                        str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
                      )
                      list(color = colour)
                    },
                    maxWidth = 120
                  ),
                  "{comps[[2]]}" := colDef(
                    cell = function(value) {
                      value <- str_extract(value, "Up|Down|Unchanged|Undetected")
                      if (str_detect(value, "Up")) value <- paste(value, "\u21E7")
                      if (str_detect(value, "Down")) value <- paste(value, "\u21E9")
                      paste(targets[[2]], value)
                    },
                    style = function(value) {
                      colour <- case_when(
                        str_detect(value, "Up") ~ colours$direction[["up"]],
                        str_detect(value, "Down") ~ colours$direction[["down"]],
                        str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
                        str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
                      )
                      list(color = colour)
                    },
                    maxWidth = 120
                  ),
                  gs_name = colDef(
                    name = "Gene Set",
                    cell = function(value) {
                      htmltools::tags$a(
                        href = gs_url[[value]], target = "_blank",
                        str_replace_all(value, "_", " ")
                      )
                    },
                    html = TRUE,
                    maxWidth = 150
                  ),
                  gs_description = colDef(
                    name = "Description",
                    cell = function(value) str_trunc(value, width = 150)
                  ),
                  gs_url = colDef(show = FALSE),
                  pval = colDef(
                    name = "P-Value",
                    cell = function(value) ifelse(
                      value < 0.001,
                      sprintf("%.2e", value),
                      sprintf("%.3f", value)
                    ),
                    maxWidth = 80
                  ),
                  adj_p = colDef(
                    name = "p<sub>adj</sub>", html = TRUE,
                    cell = function(value) ifelse(
                      value < 0.001,
                      sprintf("%.2e", value),
                      sprintf("%.3f", value)
                    ),
                    maxWidth = 80
                  ),
                  numDEInCat = colDef(
                    name = "Mapped Genes",
                    cell = function(value) comma(value, 1),
                    maxWidth = 80
                  ),
                  numInCat = colDef(
                    name = "Genes in Gene-Set",
                    cell = function(value) comma(value, 1),
                    maxWidth = 80
                  ),
                  gene_name = colDef(
                    name = "Gene Names",
                    cell = function(value) with_tooltip(value, width = 150)
                  ),
                  DEgenes = colDef(show = FALSE)
                )
              )
          )
        )
      },
      mc.cores = threads
    )
)
```

`r if (length(goseq_group_sig)) "#### All Groups"`

```{r plot-all-enriched, eval = length(goseq_group_sig) > 0, echo = length(goseq_group_sig) > 0, fig.width = min(5 + length(goseq_group_sig), knitr::opts_chunk$get("fig.width")), fig.cap = glue("*Combined enrichment across all groups, showing only significant results for enrichment. The top panel shows how many genes were mapped to sites for each group, whilst the right panel shows gene set size. Point sizes indicate the proportion of mapped genes which are from each pathway.*")}
goseq_group_sig %>%  
  bind_rows() %>% 
  arrange(pval) %>% 
  left_join(n_mapped, by = c("group" = "status")) %>% 
  mutate(
    gs_name = fct_inorder(gs_name) %>% 
      fct_relabel(str_replace_all, "_", " ") %>% 
      fct_relabel(str_trunc, width = 60),
    prop = numDEInCat / mapped,
    group = factor(group, levels = n_mapped$status)
  ) %>%
  droplevels() %>% 
  ggplot(aes(group, fct_rev(gs_name))) + 
  geom_point(aes(size = prop, fill = -log10(pval)), shape = 21) +
  geom_ysidecol(
    aes(x = numInCat), data = . %>% distinct(gs_name, numInCat),
  ) +
  geom_xsidecol(
    aes(y = mapped),
    data = n_mapped %>% 
      dplyr::filter(status %in% names(goseq_group_sig)) %>% 
      dplyr::rename(group = status),
    width = 0.5
  ) +
  scale_x_discrete(labels = label_wrap(10)) +
  scale_ysidex_continuous(expand = expansion(c(0, 0.15))) +
  scale_xsidey_continuous(
    expand = expansion(c(0, 0.15))
  ) +
  scale_fill_viridis_c(option = "inferno", begin = 0.2) +
  scale_size(range = c(0, 5), labels = percent) + 
  labs(
    x = c(), y = c(), size = "% Mapped\nGenes", fill = expr(paste(-log[10], "p"))
  ) +
  theme(
    panel.grid = element_blank(), 
    axis.text = element_text(size = 8),
    legend.position = "right",
    ggside.panel.scale.x = 0.2,
    ggside.panel.scale.y = 0.3,
    ggside.axis.text.x.bottom = element_text(angle = 270, hjust = 0, vjust = 0.5)
  ) 
```



`r if(length(tg_group_res) > 0) "### Network Plots {.tabset}"`


```{r network-plots-goseq-groups, eval = length(tg_group_res) > 0, echo = length(tg_group_res) > 0}
htmltools::tagList(
  names(tg_group_res) %>% 
    mclapply(
      function(x) {
        ## Export the image
        img_out <- file.path(
          fig_path,
          x %>% 
            str_replace_all(" ", "_") %>% 
            str_replace_all("_-_", "-") %>% 
            paste0("_network.", fig_type)
        )
        fig_fun(
          filename = img_out,
          width = knitr::opts_current$get("fig.width"),
          height = knitr::opts_current$get("fig.height")
        )
        p <- tg_group_res[[x]] %>%
          ggraph(layout = net_layout, weights = oc^2) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(fill = -log10(pval), size = numDEInCat),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
              ),
            repel = TRUE,
            max.overlaps = max(10, round(length(tg_group_res[[x]]) / 4, 0)),
            bg.color = "white", bg.r = 0.1
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.25) +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
          guides(edge_alpha = "none") +
          labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
          theme_void()
        print(p)
        dev.off()
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "
            Network plot showing enriched pathways mapped to genes 
            associated with {x}. 
            "
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>%
              basename() %>%
              str_remove_all(glue(".{fig_type}$")) %>%
              str_to_lower() %>%
              str_replace_all("_", "-"),
            class = "section level4",
            htmltools::h4(x),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      },
      mc.cores = threads
    )
)
```




```{r add-rna, eval = has_rnaseq, child = here::here('workflow/modules/rnaseq_pairwise.Rmd')}
```

# Data Export

```{r all-out}
write_rds(all_windows, all_out$results, compress = "gz")
save.image(all_out$renv)
```

```{r export-csv}
all_windows %>% 
  as_tibble() %>% 
  dplyr::select(-detected) %>% 
  unnest(all_of("gene_id")) %>% # This ensures correct id2gene mappings
  mutate(gene_name = id2gene[gene_id]) %>% 
  dplyr::select(
    starts_with("gene"), range, any_of(c("region", "feature", "hic")),
  ) %>% 
  write_csv(
    gzfile(all_out$csv)
  )
write_csv(de_genes_both_comps, all_out$de_genes)
goseq_group_sig %>% 
  bind_rows() %>% 
  write_csv(all_out$goseq)
cmn_res %>% 
  bind_rows() %>% 
  mutate(
    leadingEdge = vapply(leadingEdge, paste, character(1), collapse = "; "),
    gs_url = gs_url[gs_name]
  ) %>% 
  write_csv(all_out$rna_enrichment)
```


During this workflow, the following files were exported:

`r pander(lapply(all_out, str_extract, "output.+"))`

## References

<div id="refs"></div>
<br>

<button type="button" class="btn btn-default btn-sessioninfo" data-toggle="collapse" data-target="#sessioninfo" style="display: block;">
<span class="glyphicon glyphicon-info-sign" aria-hidden="true"></span> Session information
</button>
</p>
<div id="sessioninfo" class="collapse">

```{r session-info, echo=FALSE}
pander::pander(sessionInfo())
```

</div>
