```{r dummy-params, eval = interactive()}
params <- list(
  tgt1 = "AR",
  tgt2 = "ER",
  comp1 = "E2_E2DHT",
  comp2 = "E2_E2DHT",
  full_comp = "AR_E2_E2DHT-ER_E2_E2DHT",
  results = "output/pairwise_comparisons/AR_E2_E2DHT-ER_E2_E2DHT/AR_E2_E2DHT-ER_E2_E2DHT-pairwise_results.rds",
  motif_enrichment = "output/pairwise_comparisons/AR_E2_E2DHT-ER_E2_E2DHT/AR_E2_E2DHT-ER_E2_E2DHT-motif_enrichment.tsv.gz",
  motif_position = "output/pairwise_comparisons/AR_E2_E2DHT-ER_E2_E2DHT/AR_E2_E2DHT-ER_E2_E2DHT-motif_position.tsv.gz"
)
```



```{r packages}
library(tidyverse)
library(glue)
library(pander)
library(yaml)
library(extraChIPs)
library(scales)
library(ggrepel)
library(reactable)
library(htmltools)
library(rlang)
library(plyranges)
library(parallel)
library(rtracklayer)
library(VennDiagram)
library(patchwork)
# library(magrittr)
# library(rlang)
# library(ggside)
# library(msigdbr)
# library(goseq)
# library(GenomicInteractions)
# library(ggraph)
# library(tidygraph)
source(here::here("workflow", "scripts", "custom_functions.R"))
source(here::here("workflow", "scripts", "table_funs.R"))
threads <- detectCores() - 1
if (interactive()) threads <- 4
# register(MulticoreParam(workers = threads))
panderOptions("big.mark", ",")
panderOptions("missing", "")
panderOptions("table.split.table", Inf)
theme_set(
  theme_bw() +
    theme(
      text = element_text(size = 10), # TODO: Set in params
      plot.title = element_text(hjust = 0.5)
    )
)
```



```{r get_params, echo = FALSE}
config <- read_yaml(here::here("config", "config.yml"))
targets <- params[c("tgt1", "tgt2")] %>% unlist()
n_targets = length(unique(targets))
treat_levels <- params[paste0("comp", 1:2)] %>% 
  lapply(\(x) strsplit(x, split = "_")[[1]])
comp_names <- list(
  comp1 = glue("{params$tgt1}_{params$comp1}"),
  comp2 = glue("{params$tgt2}_{params$comp2}")
)
comp_labels <- list(
  comp1 = glue("{params$tgt1}: {treat_levels$comp1[[2]]} Vs {treat_levels$comp1[[1]]}"),
  comp2 = glue("{params$tgt2}: {treat_levels$comp2[[2]]} Vs {treat_levels$comp2[[1]]}")
)
if (n_targets > 1) {
  short_names <- as.list(targets)
} else {
  short_names <- treat_levels %>% 
    lapply(\(x) paste(x[[2]], "Vs", x[[1]]))
}
names(short_names) <- paste0("comp", 1:2)
all_paths <- here::here("config", "json", "all_paths.json") %>% 
  jsonlite::fromJSON()
diff_sig_param <- here::here(
  "config", "json", "differential_signal_param.json"
) %>% 
  jsonlite::fromJSON() %>% 
  .[targets]
pw_param <- here::here("config", "json", "pairwise_param.json") %>% 
  jsonlite::fromJSON() %>% 
  pluck(params$full_comp)
motif_param <- here::here("config", "json", "motif_analysis_param.json") %>% 
  jsonlite::fromJSON() %>% 
  pluck("pairwise")
heat_params <- here::here("config", "json", "profile_heatmap_param.json") %>% 
  jsonlite::fromJSON() %>% 
  .[targets]
```


```{r load-annotations}
## Regions & Features
gene_regions <- file.path(all_paths$annotation, "gene_regions.rds") %>% 
  read_rds()
regions <- map_chr(gene_regions, \(x) unique(x$region))
features <- here::here(all_paths$annotation, "features.rds") %>%
  read_rds()
has_features <- length(features) > 0
feat_names <- names(features) %>% 
  lapply(
    \(x) paste(x, features[[x]]$feature, sep = ": ")
  ) %>% 
  lapply(fct_infreq) %>% 
  lapply(levels) %>% 
  unlist()
all_regions <- c(unname(regions), feat_names)
sq <- seqinfo(gene_regions)
## Gene-level
gtf_gene <- file.path(all_paths$annotation, "gtf_gene.rds") %>% read_rds()
id2gene <- structure(gtf_gene$gene_name, names = gtf_gene$gene_id) %>% 
  .[!duplicated(names(.))]
rna <- file.path(all_paths$annotation, "rna.rds") %>% read_rds()
detected_ids <- unique(gtf_gene$gene_id)
if (length(rna) ) {
  detected_ids <- detected_ids %>% intersect(bind_rows(rna)$gene_id)
}
include_genes <- id2gene[detected_ids]
## HiC
hic <- file.path(all_paths$annotation, "hic.rds") %>% read_rds()
has_hic <- length(hic) > 0
```

```{r colours}
colours <- read_rds(file.path(all_paths$annotation, "colours.rds")) %>% 
  lapply(unlist)
comp_cols <- colours$target[targets]
## Pairwise palette
pw_cols <- colours$pairwise
names(pw_cols) <- names(pw_cols) %>% 
  str_replace_all("^(.+)_(.+)$", "\\1 - \\2") %>% 
  str_to_title() 
pw_cols <- c(
  pw_cols,
  "Both Increased" = colours$pairwise[["increased_increased"]],
  "Both Decreased" = colours$pairwise[["decreased_decreased"]],
  "Both Unchanged" = colours$pairwise[["unchanged_unchanged"]]
)
region_colours <- colours$regions %>% 
  setNames(regions[names(colours$regions)]) %>% 
  .[unname(regions)]
feature_colours <- unlist(colours$features)
direction_colours <- colours$direction %>% 
  setNames(str_to_title(names(.)))
treat_colours <- colours$treat[unique(unlist(treat_levels))]
full_palette <- c(
  feature_colours, region_colours, direction_colours, treat_colours,
  colours$target[targets],
  pw_cols,
  c(
    "None" = colours$direction[["undetected"]],
    "Other" = colours$direction[["unchanged"]])
)
scale_colour_pw <- function(width = 15, name = "Combined Status", ...) {
  scale_colour_manual(
    values = pw_cols, 
    labels = \(x) x %>% 
      str_replace_all("Increased.+Increased", "Both Increased") %>% 
      str_replace_all("Decreased.+Decreased", "Both Decreased") %>% 
      str_replace_all("Unchanged.+Unchanged", "Both Unchanged") %>% 
      str_replace_all(
        "(.+) - (.+)", 
        paste(short_names$comp1, "\\1 -", short_names$comp2, "\\2")
      ) %>% 
      str_wrap(width),
    name = name,
    ...
  ) 
}
scale_fill_pw <- function(width = 15, name = "Combined Status", ...) {
  scale_fill_manual(
    values = pw_cols, 
    labels = \(x) x %>% 
      str_replace_all("Increased.+Increased", "Both Increased") %>% 
      str_replace_all("Decreased.+Decreased", "Both Decreased") %>% 
      str_replace_all("Unchanged.+Unchanged", "Both Unchanged") %>% 
      str_replace_all(
        "(.+) - (.+)", 
        paste(short_names$comp1, "\\1 -", short_names$comp2, "\\2")
      ) %>% 
      str_wrap(width),
    name = name,
    ...
  ) 
}
```

```{r figure-params}
fig_params <- here::here("config", "rmarkdown.yml") %>% 
  read_yaml() %>% 
  pluck("knitr_opts")
fw <- fig_params$fig.width
fh <- fig_params$fig.height
fig_type <- fig_params$dev[[1]]
if (is.null(fig_type)) stop("Couldn't detect figure type")
fig_fun <- match.fun(fig_type)
if (fig_type %in% c("bmp", "jpeg", "png", "tiff")) {
  ## These figure types require dpi & resetting to be inches
  formals(fig_fun)$units <- "in"
  formals(fig_fun)$res <- 300
}
fig_path <- here::here("docs", "assets", params$full_comp)
if (!dir.exists(fig_path)) dir.create(fig_path, recursive = TRUE)
```


```{r load-data}
filtered_peaks <- list(
  file.path(all_paths$peaks, params$tgt1, "{params$tgt1}_{treat_levels$comp1}"),
  file.path(all_paths$peaks, params$tgt2, "{params$tgt2}_{treat_levels$comp2}")
) %>% 
  lapply(glue) %>% 
  unlist() %>% 
  paste0("_filtered_peaks.narrowPeak") %>% 
  unique() %>% 
  here::here() %>% 
  importPeaks(seqinfo = sq) %>% 
  setNames(str_remove_all(names(.), "_filtered.+"))
combined_results <- params$results %>% read_rds()
```

```{r reclass-df}
reclass_by_region <- combined_results %>% 
  plyranges::select(region, ends_with("logFC"), ends_with("FDR")) %>% 
  addDiffStatus(
    fc_col = paste0(comp_names$comp1, "_logFC"),
    sig_col = paste0(comp_names$comp1, "_FDR"),
    alpha = diff_sig_param[[params$tgt1]]$alpha,
    new_col = paste0(comp_names$comp1, "_orig_status")
  ) %>% 
  addDiffStatus(
    fc_col = paste0(comp_names$comp2, "_logFC"),
    sig_col = paste0(comp_names$comp2, "_FDR"),
    alpha = diff_sig_param[[params$tgt2]]$alpha,
    new_col = paste0(comp_names$comp2, "_orig_status")
  ) %>% 
  mcols() %>% 
  as_tibble() %>% 
  mutate(
    status = fct_cross(
      !!sym(paste0(comp_names$comp1, "_orig_status")),
      !!sym(paste0(comp_names$comp2, "_orig_status")),
      sep = " - "
    )
  ) %>% 
  summarise(
    orig_n = dplyr::n(), .by = c(region, status)
  )
if (length(features)) {
  reclass_by_feature <- combined_results %>% 
    plyranges::select(feature, ends_with("logFC"), ends_with("FDR")) %>% 
    addDiffStatus(
      fc_col = paste0(comp_names$comp1, "_logFC"),
      sig_col = paste0(comp_names$comp1, "_FDR"),
      alpha = diff_sig_param[[params$tgt1]]$alpha,
      new_col = paste0(comp_names$comp1, "_orig_status")
    ) %>% 
    addDiffStatus(
      fc_col = paste0(comp_names$comp2, "_logFC"),
      sig_col = paste0(comp_names$comp2, "_FDR"),
      alpha = diff_sig_param[[params$tgt2]]$alpha,
      new_col = paste0(comp_names$comp2, "_orig_status")
    ) %>% 
    mcols() %>% 
    as_tibble() %>% 
    unnest(feature, keep_empty = TRUE) %>% 
    mutate(
      feature = str_replace_na(feature, "no_feature"),
      status = fct_cross(
        !!sym(paste0(comp_names$comp1, "_orig_status")),
        !!sym(paste0(comp_names$comp2, "_orig_status")),
        sep = " - "
      )
    ) %>% 
    summarise(
      orig_n = dplyr::n(), .by = c(feature, status)
    )
}  
```



```{r define-outputs}
# out_path <- here::here(
#   "output", "pairwise_comparisons", paste(targets, collapse = "_")
# )
# if (!dir.exists(out_path)) dir.create(out_path, recursive = TRUE)
# all_out <- list(
#   csv = file.path(
#     out_path, glue(comp_name, "-pairwise_comparison.csv.gz"
#     )
#   ),
#   de_genes = file.path(
#     out_path, 
#     glue("{comp_name}-de_genes.csv")
#   ),
#   goseq = file.path(
#     out_path,
#     glue("{comp_name}-enrichment.csv")
#   ),
#   results = file.path(
#     out_path, glue(comp_name, "-all_windows.rds")
#   ),
#   rna_enrichment = file.path(
#     out_path,
#     glue("{comp_name}-rnaseq_enrichment.csv")
#   ),
#   renv = here::here(
#     "output", "envs",
#     glue(comp_name, "-pairwise_comparison.RData")
#   )
# )
# ## Empty objects for when RNA-Seq is absent
# de_genes_both_comps <- tibble()
# cmn_res <- list(tibble(gs_name = character(), leadingEdge = list()))
```


# Outline

This stage of the GRAVI workflow present combined results from two individual differential signal analyses and compares the results, by finding regions which directly overlap and considering the joint behaviour.
In this specific analyses the differential signal response for `r targets[[1]]` when comparing `r comp_labels[[1]]`, will be integrated with the differential signal response for `r targets[[2]]` comparing `r comp_labels[[2]]`.
Data will first be described from the perspective of peaks, regardless of treatment before moving onto the differential signal analysis (DSA) results.

## Integration Of Differential Signal Results

The initial steps of DSA detect regions with clear changes in signal, however the integration of datasets instead focusses on the *most correct* classification of a given range/window, where unchanged is not *simply lack of a statistically significant result*, but *a true state needing to be identified*.
Differential signal analyses generally focus on controlling the Type I errors (i.e. false positive), which comes with a commensurate increase in Type II errors (false negatives).
When seeking to correctly classify a window across two datasets, the detection as changed can be taken as a robust finding, whilst the consideration as unchanged may be less robust.
When comparing a window across two datasets, a detection as changed within one dataset is taken as primary evidence of 'something interesting' occurring, with the secondary consideration being the correct description of the combined signal patterns.
For this reason, and to minimise issues with hard thresholding, once a window is considered to be of interest (i.e. significant somewhere), the p-value used in the second comparison is from the test $H_0: \mu = 0$  instead of the initial $H_0: 0 < |\mu| < \lambda$.
Unchanged windows are then reclassified as significant if meeting the criteria:

1. Significant in the alternate comparison
2. The adjusted p-value for $H_0: \mu = 0$ (method = `r pw_param$adj_mu0`) is $<$ `r pw_param$alpha`, and
3. The estimated logFC is beyond the range spcified by $\lamda$ in the original analysis.


The insistence on the more stringent initial inclusion criteria still protects against false discoveries from the perspective of window identification, but this approach instead allows *improved classification* across both comparisons, taking the alternate comparison into account.

For *classification steps within each window*, each target is given the status 1) Increased, 2) Decreased, 3) Unchanged, or 4) Undetected.
Across two comparisons, this gives 15 possible classifications for each region found with signal from at least one comparison (e.g. Increased-Increased, Increased-Decreased, Increased-Unchanged, Increased-Undetected etc.), given that any potential Undetected-Undetected windows will not be included for obvious reasons.

<!-- ## Enrichment Testing -->

<!-- After direct comparison of signal patterns, enrichment analysis is performed in a similar manner to for individual analyses. -->
<!-- If RNA-Seq data is provided, genes will be restricted to those present within the RNA-Seq dataset, as representative of detected genes. -->
<!-- Enrichment testing is performed on: -->

<!-- 1. All genes mapped to any region with detected signal against those not mapped to detected signal -->
<!-- 2. All genes mapped to both targets against genes not mapped  -->

<!-- ## Comparison To RNA-Seq -->

<!-- If RNA-Seq data is supplied, the sets of genes associated with all signal patterns (e.g. Up-Up, Up-Down etc) were treated as gene-sets and Gene Set Enrichment Analysis (GSEA) [@SubramanianGsea2005] was used to determine enrichment and leading-edge genes associated with altered expression. -->
<!-- Genes will be ranked to incorporate direction of change (i.e. Directional GSEA) or significance only (i.e Non-Directional GSEA). -->
<!-- The results from enrichment analysis across ChIP target and RNA-Seq data will also be combined to reveal pathways under the regulatory influence in either comparison, for which we also have evidence of changed gene expression. -->

<!-- # Peak Comparison {.tabset} -->

<!-- Data was loaded for the initially-identified peaks to show the general level of similarity between signal regions from either dataset. -->


`r if (n_targets > 1) "## Shared Signal Regions"`

```{r plot-common, echo = n_targets > 1, fig.cap = glue("*Relationship between regions contained in {targets[[1]]} and {targets[[2]]}, and which of the two targets each peak is found in. Peaks are based on the union of both sets of peaks, which are themselves independent of treatment group.*"), fig.height = 8, fig.width = 8}
if (n_targets > 1) {
  library(VennDiagram)
  venn_args <- combined_results %>% 
    as_tibble() %>% 
    dplyr::select(range, ends_with("logCPM")) %>% 
    mutate(
      across(ends_with("logCPM"), \(x) !is.na(x))
    ) %>% 
    setNames(c("range", unlist(short_names))) %>% 
    summarise(
      area1 = sum(!!sym(short_names[[1]])),
      area2 = sum(!!sym(short_names[[2]])),
      cross.area = sum(!!sym(short_names[[1]]) & !!sym(short_names[[2]]))
    ) %>% 
    as.list()
  venn_args$fill <- unlist(colours$target)[targets]
  venn_args$category <- unname(unlist(short_names))
  grid::grid.newpage()
  do.call(draw.pairwise.venn, venn_args)
}
```


## Treatment Specific Peaks

```{r treatment-peaks-upset, fig.height = 6, fig.cap = glue("*Peaks detected for ", glue_collapse(unique(targets), sep = " and " ), " in the treatment groups ", glue_collapse(unique(unlist(treat_levels)), sep = ", ", last = " & "), ".*")}
filtered_peaks %>% 
  .[rev(names(.))] %>% 
  setNames(str_replace_all(names(.), "(.+)_(.+)", "\\1 (\\2)")) %>% 
  plotOverlaps(
    type = "upset", .sort_sets = FALSE, n_intersections = 20,
    set_col = treat_colours[str_replace_all(names(.), ".+\\((.+)\\)", "\\1")]
  )
```

# Combined Differential Signal Results 

A set of `r comma(length(combined_results))` loci with signal was formed by obtaining the union of windows produced during the analysis of `r glue_collapse(unlist(short_names), last = " and ")` individually.
These common windows were formed using any overlap between analysis-specific windows.
After merging across datasets, the merged ranges were re-mapped to regions, features and genes, using the same methods are previously.


The distance between each pair of windows was found by taking the centre of the original sliding window used for statistical analysis, and finding the difference. 
As windows were detected in an un-stranded manner, all distance values were set to be positive.
Taking the complete set of `r comma(length(combined_results))` combined windows, `r comma(sum(!is.na(combined_results$d)))` (`r percent(mean(!is.na(combined_results$d)), 0.1)`) were included in *both* datasets.
The median distance between overlapping windows found in both comparisons was `r  median(combined_results$d, na.rm = TRUE)`bp.

## Results Tables {.tabset}

### Summary Table

```{r tbl-summary, echo = FALSE}
cp <- htmltools::tags$caption(
  htmltools::tags$em(
    glue(
      "Summary of combined results for ",
      glue_collapse(short_names, sep = " and "),
      ". Ranges were classified as described above. ",
      "Where signal was deteted in both comparisons, the median distance ",
      "between peak centres is given.",
      "The number of loci showing each combined pattern are given summarised ",
      "across all overlapping gene-centric regions, with these able to be ",
      "included in filtering interactively, and expanded using the arrows in the ",
      "leftmost column. The number of loci re-classified when combining results ",
      "is also provided in the last column, with sites generally expected to be ",
      "reclassified from an unchanged category to either 'increased' or 'decreased'."
    )
  )
)
tbl <- combined_results %>%
  as_tibble() %>%
  summarise(
    total = dplyr::n(),
    d = median(d, na.rm = TRUE),
    .by = c(status, region)
  ) %>% 
  arrange(desc(total)) %>% 
  separate(status, into = unlist(short_names), sep = " - ", remove = FALSE) %>% 
  mutate(
    across(
      all_of(unname(unlist(short_names))), 
      \(x) factor(x, levels = names(direction_colours))
    )
  ) %>% 
  droplevels() %>% 
  arrange(!!!syms(unname(short_names)), region) %>% 
  left_join(
    reclass_by_region, by = c("region", "status")
  ) %>% 
  mutate(
    orig_n = ifelse(is.na(orig_n), 0, orig_n),
    reclass = total - orig_n
  ) %>% 
  dplyr::select(-orig_n) %>% 
  reactable(
    filterable = TRUE, showPageSizeOptions = nrow(.) > 10,
    groupBy = "status",
    columns = list2(
      status = colDef(
        "Click To<br>Expand", html = TRUE, maxWidth = 80,
        grouped = JS("function(cellInfo, state) {
            val = ''
            return val
          }"),
        footer = htmltools::tags$b("Total")
      ),
      !!sym(short_names$comp1) := colDef(
        aggregate = "unique",
        cell = function(value) {
          value <- str_extract(value, "Increased|Decreased|Unchanged|Undetected")
          if (str_detect(value, "Increased")) value <- paste(value, "\u21E7")
          if (str_detect(value, "Decreased")) value <- paste(value, "\u21E9")
          paste(targets[[1]], value)
        },
        style = function(value) {
          colour <- case_when(
            str_detect(value, "Increased") ~ colours$direction[["increased"]],
            str_detect(value, "Decreased") ~ colours$direction[["decreased"]],
            str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
            str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
          )
          list(color = colour)
        },
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('summary-result-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        }
      ),
      !!sym(short_names$comp2) := colDef(
        aggregate = "unique",        
        cell = function(value) {
          value <- str_extract(value, "Increased|Decreased|Unchanged|Undetected")
          if (str_detect(value, "Increased")) value <- paste(value, "\u21E7")
          if (str_detect(value, "Decreased")) value <- paste(value, "\u21E9")
          paste(targets[[2]], value)
        },
        style = function(value) {
          colour <- case_when(
            str_detect(value, "Increased") ~ colours$direction[["increased"]],
            str_detect(value, "Decreased") ~ colours$direction[["decreased"]],
            str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
            str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
          )
          list(color = colour)
        },
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('summary-result-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        }
      ),
      region = colDef(
        "Region",
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('summary-result-table', '%s', event.target.value || undefined)",
              name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        }
      ),
      total = colDef(
        name = "Number of Ranges", format = comma_col, aggregate = "sum",
        filterMethod = js_greater, maxWidth = 150,
        footer = JS("function(column, state) {
            let total = 0
            state.sortedData.forEach(function(row) {
              total += row[column.id]
            })
            return total.toLocaleString('en-US')
          }")
      ),
      d = colDef(
        "Mean Distance", format = comma_col, filterMethod = js_greater,
        maxWidth = 120, aggregate = "median",
        aggregated = JS("function(cellInfo, state) {
           let total = 0;
           let n = 0;
           state.sortedData.forEach(function(row) {
               total += parseFloat(cellInfo.value);
               n += 1;
           })
           if (!isNaN(total)){
             return total / n;
           } else {
             return '';
           }
          }
          "),
        footer = JS("function(column, state) {
            let total = 0
            let n = 0
            state.sortedData.forEach(function(row) {
              let value = parseFloat(row[column.id]);
              if (!isNaN(value)) {
                total += value;
                n += 1;
              }
            })
            mn = total / n
            return mn.toFixed(1)
          }")
      ),
      reclass = colDef(
        "Re-Classified", aggregate = "sum", filterMethod = js_greater
      )
    ),
    defaultColDef = colDef(footerStyle = list(fontWeight = "bold")),
    elementId = 'summary-result-table'
  )
tl <- tagList(
  tags$button(
    tagList(fontawesome::fa("download"), "Download as CSV"),
    onclick = sprintf(
      "Reactable.downloadDataCSV('summary-result-table', '%s_summary.csv')",
      params$full_comp
    )
  ),
  tags$button(
    "Expand/Collapse All",
    onclick = "Reactable.toggleAllRowsExpanded('summary-result-table')"
  ),
  tags$button(
    "Clear All Filters",
    onclick = "Reactable.setAllFilters('summary-result-table', [])"
  ),
  tbl, cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```

`r if (has_features) "### Summary Table By Feature"`

```{r tbl-summary-by-feature, eval = has_features, echo = FALSE}
id <- 'summary-table-by-feature'
cp <- htmltools::tags$caption(
  htmltools::tags$em(
    glue(
      "Summary of combined results for ",
      glue_collapse(short_names, sep = " and "),
      "grouping results by feature. Ranges were classified as described above. ",
      "Where signal was deteted in both comparisons, the median distance ",
      "between peak centres is given.",
      "The number of loci showing each combined pattern are given summarised ",
      "across all overlapping gene-centric regions, with these able to be ",
      "included in filtering interactively, and expanded using the arrows in the ",
      "leftmost column. The number of loci re-classified when combining results ",
      "is also provided in the last column, with sites generally expected to be ",
      "reclassified from an unchanged category to either 'increased' or 'decreased'.",
      "Given that loci may map to multiple features and contribute to ", 
      "multiple summary values, totals are not provided."
    )
  )
)
tbl <- combined_results %>%
  as_tibble() %>%
  unnest(feature, keep_empty = TRUE) %>% 
  mutate(feature = str_replace_na(feature, "no_feature")) %>% 
  summarise(
    total = dplyr::n(),
    d = median(d, na.rm = TRUE),
    .by = c(status, feature)
  ) %>% 
  arrange(desc(total)) %>% 
  separate(status, into = unlist(short_names), sep = " - ", remove = FALSE) %>% 
  mutate(
    across(
      all_of(unname(unlist(short_names))), 
      \(x) factor(x, levels = names(direction_colours))
    )
  ) %>% 
  droplevels() %>% 
  arrange(!!!syms(unname(short_names)), feature) %>% 
  left_join(
    reclass_by_feature, by = c("feature", "status")
  ) %>% 
  mutate(
    orig_n = ifelse(is.na(orig_n), 0, orig_n),
    reclass = total - orig_n
  ) %>% 
  dplyr::select(
    feature, all_of(unname(unlist(short_names))), total, d, reclass
  ) %>% 
  reactable(
    groupBy = "feature", showPageSizeOptions = nrow(.) > 10,
    filterable = TRUE,
    columns = list2(
      feature = colDef(
        "Feature",
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
              id, name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        }
      ),
      !!sym(short_names$comp1) := colDef(
        aggregate = "unique",
        cell = function(value) {
          value <- str_extract(value, "Increased|Decreased|Unchanged|Undetected")
          if (str_detect(value, "Increased")) value <- paste(value, "\u21E7")
          if (str_detect(value, "Decreased")) value <- paste(value, "\u21E9")
          paste(targets[[1]], value)
        },
        style = function(value) {
          colour <- case_when(
            str_detect(value, "Increased") ~ colours$direction[["increased"]],
            str_detect(value, "Decreased") ~ colours$direction[["decreased"]],
            str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
            str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
          )
          list(color = colour)
        },
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
              id, name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        }
      ),
      !!sym(short_names$comp2) := colDef(
        aggregate = "unique",        
        cell = function(value) {
          value <- str_extract(value, "Increased|Decreased|Unchanged|Undetected")
          if (str_detect(value, "Increased")) value <- paste(value, "\u21E7")
          if (str_detect(value, "Decreased")) value <- paste(value, "\u21E9")
          paste(targets[[2]], value)
        },
        style = function(value) {
          colour <- case_when(
            str_detect(value, "Increased") ~ colours$direction[["increased"]],
            str_detect(value, "Decreased") ~ colours$direction[["decreased"]],
            str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
            str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
          )
          list(color = colour)
        },
        filterInput = function(values, name) {
          tags$select(
            # Set to undefined to clear the filter
            onchange = sprintf(
              "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
              id, name
            ),
            # "All" has an empty value to clear the filter, and is the default option
            tags$option(value = "", "All"),
            lapply(unique(values), tags$option),
            "aria-label" = sprintf("Filter %s", name),
            style = "width: 100%; height: 28px;"
          )
        }
      ),
      total = colDef(
        name = "Number of Ranges", format = comma_col, aggregate = "sum",
        filterMethod = js_greater, maxWidth = 150
      ),
      d = colDef(
        "Mean Distance", format = comma_col, filterMethod = js_greater,
        maxWidth = 120, aggregate = "median",
        aggregated = JS("function(cellInfo, state) {
           let total = 0;
           let n = 0;
           state.sortedData.forEach(function(row) {
               total += parseFloat(cellInfo.value);
               n += 1;
           })
           if (!isNaN(total)){
             return total / n;
           } else {
             return '';
           }
          }
          ")
      ),
      reclass = colDef(
        "Re-Classified", aggregate = "sum", filterMethod = js_greater
      )
    ),
    elementId = id
  )
tl <- tagList(
  tags$button(
    tagList(fontawesome::fa("download"), "Download as CSV"),
    onclick = sprintf(
      "Reactable.downloadDataCSV('%s', '%s_summary_by_feature.csv')",
      id, params$full_comp
    )
  ),
  tags$button(
    "Expand/Collapse All",
    onclick = sprintf("Reactable.toggleAllRowsExpanded('%s')", id)
  ),
  tags$button(
    "Clear All Filters",
    onclick = sprintf("Reactable.setAllFilters('%s', [])", id)
  ),
  tbl, cp
)
if (interactive()) {
  htmltools::browsable(tl)
} else {
  tl
}
```

## Summary Figures {.tabset}

### Overall Status

```{r plot-both-donut, fig.cap = "Summary of signal across both comparisons incorporating all sites where signal was detected in either copmarison. Ranges where signal is changed in either comparisons are exploded."}
combined_results %>% 
  as_tibble() %>% 
  dplyr::select(ends_with("_status")) %>% 
  setNames(unlist(short_names)) %>% 
  plotSplitDonut(
    inner = !!sym(short_names$comp1), outer = !!sym(short_names$comp2), 
    inner_palette = full_palette, outer_palette = full_palette,
    inner_glue = "{inner}\n{.data[[inner]]}\n{comma(n, 1)}\n({percent(p, 0.1)})",
    inner_label_alpha = 0.7,
    outer_glue = "{outer} {.data[[outer]]}\n{comma(n, 1)} ({percent(p, 0.1)})",
    outer_min_p = 0.02, outer_label = "text", outer_rotate = TRUE,
    outer_label_colour = "palette",
    explode_inner = "(In|De)creased", explode_outer = "(In|De)creased",
    explode_r = 0.4, explode_query = "AND", nudge_r = 0.5
  ) %>% 
    .[[1]]
```


### Pairwise Differential Signal


```{r plot-dbwin, fig.cap = glue("*Comparative changes in both ", glue_collapse(comp_labels, sep = " and "), ". The range with most extreme combined change in signal is shown with any genes labelled. ", ifelse(has_rna, "Only genes detected within transcriptomic data are labelled. " , ""),  "The range around zero used for range-based hypothesis testing is indicated with grey dashed lines.*")}
df <- combined_results %>% 
  as_tibble() %>% 
  select(
    range,
    ends_with("logFC"), ends_with("status"), any_of(c("region", "feature")), 
    starts_with("gene")
  ) %>% 
  dplyr::rename(
    "{short_names$comp1} logFC" := !!sym(
      paste(params$tgt1, params$comp1, "logFC", sep = "_")
    ),
    "{short_names$comp2} logFC" := !!sym(
      paste(params$tgt2, params$comp2, "logFC", sep = "_")
    ),
    "{short_names$comp1} status" := !!sym(
      paste(params$tgt1, params$comp1, "status", sep = "_")
    ),
    "{short_names$comp2} status" := !!sym(
      paste(params$tgt2, params$comp2, "status", sep = "_")
    ),
  )
x_lim <- range(df[[glue("{short_names$comp1} logFC")]], na.rm = TRUE)
y_lim <- range(df[[glue("{short_names$comp2} logFC")]], na.rm = TRUE)
df %>% 
  arrange(status) %>% 
  ggplot(
    aes(
      x = !!sym(glue("{short_names$comp1} logFC")), 
      y = !!sym(glue("{short_names$comp2} logFC")),
      colour = status
    )
  ) +
  geom_point(
    data = . %>% dplyr::filter(!str_detect(status, "Undetected"))
  ) +
  geom_label_repel(
    aes(label = label),
    data = . %>% 
      mutate(
        across(ends_with("logFC"), list(sq = \(x) x^2)),
        diam = (
          !!sym(glue("{short_names$comp1} logFC_sq")) + 
            !!sym(glue("{short_names$comp2} logFC_sq"))
        )
      ) %>% 
      dplyr::filter(diam > 1) %>% 
      arrange(desc(diam)) %>% 
      dplyr::slice(1:10, .by = status) %>% 
      mutate(
        label = map_chr(
          gene_name, 
          \(x) paste(intersect(x, include_genes), collapse = "; ")
        ) %>% 
          str_wrap(15) %>% 
          str_trunc(30) 
      ) %>% 
      dplyr::filter(label != "") %>% 
      dplyr::slice(1:2, .by = status),
    size = 3, show.legend = FALSE, fill = rgb(1, 1, 1, 0.8),
    fontface = "italic"
  ) +
  geom_hline(
    yintercept = c(1, -1) * log2(diff_sig_param[[params$tgt2]]$fc),
    linetype = 2, colour = "grey"
  ) +
  geom_vline(
    xintercept = c(1, -1) * log2(diff_sig_param[[params$tgt1]]$fc),
    linetype = 2, colour = "grey"
  ) +
  annotate(
    "text",
    x = min(x_lim), y = max(y_lim), vjust = 1, hjust = 0,
    label = paste(
      "rho == ",
      cor(
        dplyr::select(df, ends_with("logFC")), use = "pairwise.complete.obs"
      )[1, 2] %>%  round(3)
    ),
    parse = TRUE, size = 3.4
  ) +
  geom_xsideboxplot(
    aes(
      x = !!sym(glue("{short_names$comp1} logFC")), 
      y = !!sym(glue("{short_names$comp2} status")),
      fill = !!sym(glue("{short_names$comp2} status"))
    ), 
    data = . %>% 
      dplyr::filter(!is.na(!!sym(glue("{short_names$comp1} logFC")))),
    orientation = "y", colour = "grey20", show.legend = FALSE
  ) +
  geom_ysideboxplot(
    aes(
      x = !!sym(glue("{short_names$comp1} status")),
      y = !!sym(glue("{short_names$comp2} logFC")),
      fill = !!sym(glue("{short_names$comp1} status"))
    ),
    data = . %>% 
      dplyr::filter(!is.na(!!sym(glue("{short_names$comp2} logFC")))),
    colour = "grey20", orientation = "x", show.legend = FALSE
  ) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_ysidex_discrete(
    position = "top",     
    labels = \(x) paste(short_names$comp1, x) %>% str_wrap(15)
  ) +
  scale_xsidey_discrete(
    labels = \(x) paste(short_names$comp2, x) %>% str_wrap(15)
  ) +
  scale_colour_pw() +
  scale_fill_manual(values = full_palette) +
  theme(
    ggside.panel.scale.x = 0.1 + 0.05 * length(levels(df[[glue("{short_names$comp2} status")]])),
    ggside.panel.scale.y = 0.1 + 0.025 * length(levels(df[[glue("{short_names$comp1} status")]])),
    ggside.axis.text.x.top = element_text(angle = 90, hjust = 0, vjust = 0.5),
    legend.position = "bottom",
    axis.title.y.left = element_text(vjust = -10, hjust = 0.4),
    axis.title.x.bottom = element_text(hjust = 0.45)
  )
```


### logFC By Region

```{r plot-dbwin-by-region, fig.cap = glue("*Comparative changes in both ", glue_collapse(comp_labels, sep = " and "), ", separated by defined gene-centric regions. The window with most extreme combined change in signal for each region is shown with any genes labelled as previously. The range around zero used for range-based hypothesis testing is indicated with grey dashed lines.*")}
df %>% 
  arrange(status) %>% 
  ggplot(
    aes(
      x = !!sym(glue("{short_names$comp1} logFC")), 
      y = !!sym(glue("{short_names$comp2} logFC")),
      colour = status
    )
  ) +
  geom_point(
    data = . %>% dplyr::filter(!str_detect(status, "Undetected"))
  ) +
  geom_label_repel(
    aes(label = label),
    data = . %>% 
      mutate(
        across(ends_with("logFC"), list(sq = \(x) x^2)),
        diam = (
          !!sym(glue("{short_names$comp1} logFC_sq")) + 
            !!sym(glue("{short_names$comp2} logFC_sq"))
        )
      ) %>% 
      dplyr::filter(diam > 1) %>% 
      arrange(desc(diam)) %>% 
      dplyr::slice(1:10, .by = c(status, region)) %>% 
      mutate(
        label = map_chr(
          gene_name, 
          \(x) paste(intersect(x, include_genes), collapse = "; ")
        ) %>% 
          str_wrap(15) %>% 
          str_trunc(30) 
      ) %>% 
      dplyr::filter(label != "") %>% 
      dplyr::slice(1, .by = c(status, region)),
    size = 3, show.legend = FALSE, fill = rgb(1, 1, 1, 0.8),
    fontface = "italic"
  ) +
  geom_hline(
    yintercept = c(1, -1) * log2(diff_sig_param[[params$tgt2]]$fc),
    linetype = 2, colour = "grey"
  ) +
  geom_vline(
    xintercept = c(1, -1) * log2(diff_sig_param[[params$tgt1]]$fc),
    linetype = 2, colour = "grey"
  ) +
  geom_smooth(
    data = . %>% 
      dplyr::filter(!grepl("Undetected", status)),
    se = FALSE, colour = "grey40", method = MASS::rlm
  ) +
  geom_text(
    aes(label = label),
    data = . %>%
      dplyr::filter(!grepl("Undetected", status)) %>% 
      dplyr::select(region, ends_with("logFC")) %>% 
      split(.$region) %>% 
      mclapply(
        \(x) {
          tibble(
            cor = cor(dplyr::select(x, ends_with("logFC")))[1, 2],
            "{short_names$comp1} logFC" := min(x[[glue("{short_names$comp1} logFC")]]),
            "{short_names$comp2} logFC" := max(x[[glue("{short_names$comp2} logFC")]])
          )
        }, mc.cores = threads
      ) %>% 
      bind_rows(.id = "region") %>% 
      mutate(
        label = paste("rho ==", round(cor, 3)),
        "{short_names$comp1} logFC" := min(!!sym(glue("{short_names$comp1} logFC"))),
        "{short_names$comp2} logFC" := max(!!sym(glue("{short_names$comp2} logFC")))
      ),
    colour = "grey20", parse = TRUE, vjust = 1, hjust = 0, size = 3.5
  ) +
  facet_wrap(~region) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_colour_pw() +
  theme(legend.position = "bottom")
```



`r if (has_features) "### logFC By Feature"`

```{r plot-dbwin-by-feature, echo = has_features, fig.cap = glue("*Comparative changes in both ", glue_collapse(comp_labels, sep = " and "), ", separated by feature. The range with most extreme combined change in signal for each separate feature is shown with any genes labelled as previously, noting that combined ranges may map to multiple features. The range around zero used for range-based hypothesis testing is indicated with grey dashed lines.*")}
if (has_features) {
  df %>% 
    unnest(feature, keep_empty = TRUE) %>% 
    mutate(
      feature = str_replace_na(feature, "no_feature") %>% 
        fct_infreq() %>% 
        fct_lump_n(8)
    ) %>% 
    arrange(status) %>% 
    ggplot(
      aes(
        x = !!sym(glue("{short_names$comp1} logFC")), 
        y = !!sym(glue("{short_names$comp2} logFC")),
        colour = status
      )
    ) +
    geom_point(
      data = . %>% dplyr::filter(!str_detect(status, "Undetected"))
    ) +
    geom_label_repel(
      aes(label = label),
      data = . %>% 
        mutate(
          across(ends_with("logFC"), list(sq = \(x) x^2)),
          diam = (
            !!sym(glue("{short_names$comp1} logFC_sq")) + 
              !!sym(glue("{short_names$comp2} logFC_sq"))
          )
        ) %>% 
        dplyr::filter(diam > 1) %>% 
        arrange(desc(diam)) %>% 
        dplyr::slice(1:10, .by = c(status, feature)) %>% 
        mutate(
          label = map_chr(
            gene_name, 
            \(x) paste(intersect(x, include_genes), collapse = "; ")
          ) %>% 
            str_wrap(15) %>% 
            str_trunc(30) 
        ) %>% 
        dplyr::filter(label != "") %>% 
        dplyr::slice(1, .by = c(status, feature)),
      size = 3, show.legend = FALSE, fill = rgb(1, 1, 1, 0.8),
      fontface = "italic"
    ) +
    geom_hline(
      yintercept = c(1, -1) * log2(diff_sig_param[[params$tgt2]]$fc),
      linetype = 2, colour = "grey"
    ) +
    geom_vline(
      xintercept = c(1, -1) * log2(diff_sig_param[[params$tgt1]]$fc),
      linetype = 2, colour = "grey"
    ) +
    geom_smooth(
      data = . %>% 
        dplyr::filter(!grepl("Undetected", status)),
      se = FALSE, colour = "grey40", method = MASS::rlm
    ) +
    geom_text(
      aes(label = label),
      data = . %>%
        dplyr::filter(!grepl("Undetected", status)) %>% 
        dplyr::select(feature, ends_with("logFC")) %>% 
        split(.$feature) %>% 
        mclapply(
          \(x) {
            tibble(
              cor = cor(dplyr::select(x, ends_with("logFC")))[1, 2],
              "{short_names$comp1} logFC" := min(x[[glue("{short_names$comp1} logFC")]]),
              "{short_names$comp2} logFC" := max(x[[glue("{short_names$comp2} logFC")]])
            )
          }, mc.cores = threads
        ) %>% 
        bind_rows(.id = "feature") %>% 
        mutate(
          label = paste("rho ==", round(cor, 3)),
          "{short_names$comp1} logFC" := min(!!sym(glue("{short_names$comp1} logFC"))),
          "{short_names$comp2} logFC" := max(!!sym(glue("{short_names$comp2} logFC")))
        ),
      colour = "grey20", parse = TRUE, vjust = 1, hjust = 0, size = 3.5
    ) +
    facet_wrap(~feature) +
    scale_x_continuous() +
    scale_y_continuous() +
    scale_colour_pw() +
    theme(legend.position = "bottom")
}
```


### Pairwise Signal

```{r plot-pairwise-signal, fig.cap = glue("Pairwise signal levels with boxplots indicating the baseline signal based on whether signal was detected in both or only the individual dataset. The blue, smoothed line represents a GAM line of best-fit.")}
combined_results %>% 
  as_tibble() %>% 
  ggplot(
    aes(
      !!sym(glue("{params$tgt1}_{params$comp1}_logCPM")),
      !!sym(glue("{params$tgt2}_{params$comp2}_logCPM"))
    )
  ) +
  geom_point(data = . %>% dplyr::filter(!grepl("Undetected", status))) +
  geom_smooth(
    method = 'gam', formula = y ~ s(x, bs = "cs"),
    se = FALSE, data = . %>% dplyr::filter(!grepl("Undetected", status))
  ) +
  geom_text(
    aes(label = label),
    data = . %>% 
      dplyr::filter(!grepl("Undetected", status)) %>% 
      summarise(
        cor = cor(
          !!sym(glue("{params$tgt1}_{params$comp1}_logCPM")),
          !!sym(glue("{params$tgt2}_{params$comp2}_logCPM"))
        ),
        "{params$tgt1}_{params$comp1}_logCPM" := min(
          !!sym(glue("{params$tgt1}_{params$comp1}_logCPM"))
        ),
        "{params$tgt2}_{params$comp2}_logCPM" := max(
          !!sym(glue("{params$tgt2}_{params$comp2}_logCPM"))
        )
      ) %>% 
      mutate(label = glue("rho == {round(cor, 3)}")),
    parse = TRUE, vjust = 1, hjust = 0
  ) + 
  geom_xsideboxplot(
    aes(y = detected, x = !!sym(glue("{params$tgt1}_{params$comp1}_logCPM"))),
    data = . %>% 
      mutate(
        detected = ifelse(
          is.na(!!sym(glue("{params$tgt2}_{params$comp2}_logCPM"))),
          paste(short_names$comp1, "Only"), "Both Detected"
        ) 
      ) %>% 
          dplyr::filter(!is.na(!!sym(glue("{params$tgt1}_{params$comp1}_logCPM")))
      ),
    orientation = "y"
  ) +
  geom_ysideboxplot(
    aes(x = detected, y = !!sym(glue("{params$tgt2}_{params$comp2}_logCPM"))),
    data = . %>% 
      mutate(
        detected = ifelse(
          is.na(!!sym(glue("{params$tgt1}_{params$comp1}_logCPM"))),
          paste(short_names$comp2, "Only"), "Both Detected"
        ) 
      ) %>% 
          dplyr::filter(!is.na(!!sym(glue("{params$tgt2}_{params$comp2}_logCPM")))
      ),
    orientation = "x"
  ) +
  labs(
    x = glue("logCPM ({short_names$comp1})"),
    y = glue("logCPM ({short_names$comp2})"),
  ) +
  theme(
    ggside.panel.scale.x = 0.25,
    ggside.panel.scale.y = 0.25,
    ggside.axis.text.x.top = element_text(angle = 90, hjust = 0, vjust = 0.5),
    legend.position = "bottom",
    axis.title.y.left = element_text(vjust = -10, hjust = 0.4),
    axis.title.x.bottom = element_text(hjust = 0.45)
  )
```

### Distribution By Region

```{r status-by-region, fig.cap = glue("*Distribution of signal patterns by genomic regions as defined in earlier sections. Windows are only shown if signal was detected in both datasets. Segments where changes were detected in both comparisons are exploded.*")}
combined_results %>% 
  filter(!grepl("Undetected", status)) %>%
  mutate(
    status = status %>% 
      str_replace_all("Increased.+Increased", "Both Increased") %>% 
      str_replace_all("Decreased.+Decreased", "Both Decreased") %>% 
      str_replace_all("Unchanged.+Unchanged", "Both Unchanged") 
  ) %>% 
  mutate(status = fct_lump_prop(status, 0.01)) %>%
  plotSplitDonut(
    outer = "status", inner = "region",
    inner_palette = region_colours, outer_palette = pw_cols,
    inner_glue = "{str_wrap(.data[[inner]], 12)}\nn = {comma(n, 1)}\n({percent(p,0.1)})",
    outer_glue = "{str_replace_all(.data[[outer]], '(.+) - (.+)', paste(short_names$comp1, '\\\\1\n', short_names$comp2, '\\\\2'))}\nn = {comma(n, 1)}\n({percent(p,0.1)})",
    min_p = 0.03, label_alpha = 0.8,
    explode_outer = "Both (In|De)creased", explode_r = 0.15
  ) %>% 
  .[[1]]
```

`r if (has_features) "### Distribution By Feature"`

```{r status-by-feature, echo = has_features, fig.cap = glue("*Distribution of signal patterns by external features as defined in earlier sections, noting again that features may be overlapping and loci may be mapped to multiple features. Features are grouped to capture 99% of mappings.*")}
if (has_features) {
  combined_results %>% 
    plyranges::select(feature, status) %>% 
    plyranges::filter(!grepl("Undetected", status)) %>% 
    as_tibble() %>% 
    unnest(feature, keep_empty = TRUE) %>% 
    mutate(
      feature = str_replace_na(feature, "no_feature") %>% 
        fct_infreq() %>% 
        fct_lump_prop(0.01),
      status  = status %>% 
        fct_relabel(
          \(x) x %>% 
            str_replace_all("Increased.+Increased", "Both Increased") %>% 
            str_replace_all("Decreased.+Decreased", "Both Decreased") %>% 
            str_replace_all("Unchanged.+Unchanged", "Both Unchanged") 
        )
    ) %>% 
    summarise(n = dplyr::n(), .by = c(status, feature)) %>% 
    ggplot(aes(n, fct_rev(feature), fill = fct_rev(status))) +
    geom_col() +
    geom_label(
      aes(x = x, label = comma(n)),
      data = . %>% 
        arrange(feature, status) %>% 
        mutate(x = cumsum(n) - 0.5 * n, .by = feature) %>% 
        dplyr::filter(n > sum(n) * 0.01),
      show.legend = FALSE, fill = "white", alpha = 0.7, size = 3
    ) +
    scale_x_continuous(expand = expansion(c(0, 0.05)), labels = comma) +
    scale_fill_pw(width = 80) +
    labs(y = "Feature", x = "Loci Mapped to Feature")
}
```

`r if (n_targets > 1) "## Distance Between Individual Loci {.tabset}"`

`r if (n_targets > 1) "### Distance By Status"`

```{r distances-by-status, echo = n_targets > 1, fig.height = 6, fig.cap = "*Distances between loci where signal was detected for each target. Windows are only shown if 10 or more windows were found in each group. The x-axis is truncated at a maximum of 1kb*"}
if (n_targets > 1) {
  combined_results %>% 
    plyranges::filter(!is.na(d)) %>% 
    as_tibble() %>% 
    arrange(status, d) %>% 
    group_by(status) %>% 
    mutate(
      n = dplyr::n(),
      q = seq_along(status) / n
    ) %>% 
    dplyr::filter(n > 10, d < 1e3) %>%
    ggplot(aes(d, q, colour = status)) +
    geom_line(linewidth = 1) +
    scale_y_continuous(
      labels = percent, expand = expansion(c(0, 0.05)), 
      breaks = seq(0, 1, by = 0.2)
    ) +
    scale_colour_pw(width = 80) +
    labs(
      x = "Distance Between Window Centres (bp)",
      y = "Percentile"
    )
}
```


`r if (n_targets > 1) "### Distance by Region"`

```{r distance-by-region, echo = n_targets > 1, fig.height = 6, fig.width = 8, fig.cap = "*Distribution of distances between peaks separated by genomic region. All sites are included regardless of changed signal. The x-axis is truncated 1kb.*"}
if (n_targets > 1) {
  combined_results %>% 
    plyranges::filter(!is.na(d)) %>% 
    as_tibble() %>% 
    arrange(region, d) %>% 
    group_by(region) %>% 
    mutate(
      n = dplyr::n(),
      q = seq_along(region) / n
    ) %>% 
    dplyr::filter(d < 1e3) %>% 
    ggplot(aes(d, q, colour = region)) +
    geom_line(linewidth = 1) +
    scale_y_continuous(
      labels = percent, expand = expansion(c(0, 0.05)), 
      breaks = seq(0, 1, by = 0.2)
    ) +
    scale_colour_manual(values = region_colours) +
    labs(
      x = "Distance Between Window Centres (bp)",
      y = "Percentile",
      colour = "Genomic Region"
    )
}
```

`r if (has_features & n_targets > 1) "### Distance By Feature"`

```{r distance-by-feature, echo = has_features & n_targets > 1, fig.height = 6, fig.width = 8, fig.cap = "*Distribution of distances between peaks separated by external feature. All sites are included regardless of changed signal. The x-axis is truncated 1kb.*"}
if (has_features & n_targets > 1) {
  combined_results %>% 
    plyranges::filter(!is.na(d)) %>% 
    as_tibble() %>% 
    unnest(feature, keep_empty = TRUE) %>% 
    mutate(
      feature = feature %>% 
        str_replace_na("no_feature") %>% 
        fct_infreq() %>% 
        fct_lump_prop(0.01)
    ) %>% 
    arrange(feature, d) %>% 
    mutate(
      n = dplyr::n(),
      q = seq_along(feature) / n,
      .by = feature
    ) %>% 
    dplyr::filter(d < 1e3) %>% 
    droplevels() %>% 
    ggplot(aes(d, q, colour = feature)) +
    geom_line(linewidth = 1) +
    scale_y_continuous(
      labels = percent, expand = expansion(c(0, 0.05)), 
      breaks = seq(0, 1, by = 0.2)
    ) +
    scale_colour_manual(values = unlist(feature_colours)) +
    labs(
      x = "Distance Between Window Centres (bp)",
      y = "Percentile",
      colour = "External Feature"
    )
}
```

# Combined Signal Visualisation

## Heatmaps {.tabset}

A series of profile heatmaps were generated for any set of pairwise comparisons with more than 25 ranges, and where signal was detected as changed in at least one of the comparisons.
If comparing two targets, peaks were centred at the mid-point between the ranges of individual maximal signal.
If only one target is detected within a range, peaks are centred at the range with maximal signal.
Scales for fill and average signal (both logCPM) are held constant across all plots for easier comparison between groups.

```{r form-profile-heatmaps}
min_ranges <- min(map_dbl(heat_params, \(x) x$min_ranges))
max_ranges <- max(map_dbl(heat_params, \(x) x$max_ranges))
sig_ranges <- combined_results %>% 
  subset(grepl("(In|De)creased", status)) %>% 
  subset(!grepl("Undetected", status)) %>% 
  splitAsList(.$status) %>% 
  endoapply(granges) %>% 
  .[map_int(., length) >= min_ranges] %>% 
  endoapply(
    \(x) {
      if (length(x) > max_ranges) {
        i <- sample.int(length(x), max_ranges)
        x <- sort(x[i])
      }
      x
    }
  )

## Hardwire to the first bw_type. Can't be different...
## Does this need to go into a separate spot for heat or pairwise params?
bw_type <- match.arg(heat_params[[1]]$bw_type, c("FE", "coverage"))
bw_tag <- c(coverage = "treat_pileup", FE = "FE")[[bw_type]]
fill_type <- c(coverage = "logCPM", FE = "log Enrichment")[[bw_type]]
bwfl <- seq_len(2) %>% 
  lapply(
    \(i) here::here(
      all_paths$macs2, targets[[i]], 
      "{targets[[i]]}_{treat_levels[[i]]}_merged_{bw_tag}.bw"
    )
  ) %>% 
  lapply(glue) %>% 
  unlist() %>% 
  unique() %>% 
  setNames(str_remove_all(basename(.), "_merged.+")) %>% 
  BigWigFileList()
prof_widths <- seq_len(2) %>% 
  lapply(
    \(i) {
      heat_params[[i]]$width %>% 
        rep_len(2) %>% 
        setNames(
          paste0(targets[[i]], "_", treat_levels[[i]])
        )
    }
  ) %>% 
  unlist() %>% 
  .[names(bwfl)]
n_bins <- max(map_dbl(heat_params, pluck, "n_bins"))
## Get the profile data
sig_profiles <- sig_ranges %>% 
  mclapply(
    \(x) {
      ## step through each bwfl separately to ensure variable widths
      gr <- names(bwfl) %>% 
        lapply(
          \(i) {
            getProfileData(
              bwfl[[i]], x, upstream = prof_widths[[i]] / 2, bins = n_bins,
              log = TRUE
            )
          }
        )
      names(gr) <- names(bwfl)
      GRangesList(gr)
    },
    mc.cores = min(length(.), threads)
  ) 
## Run as a list to enable using different gradients
p_list <- sig_profiles %>% 
  lapply(unlist) %>% 
  lapply(names_to_column, "name") %>% 
  lapply(
    mutate, 
    target = str_extract(name, paste(targets, collapse = "|")),
    treat = str_remove_all(
      name, paste0("^(", paste(targets, collapse = "|"), ")_")
    ),
    name = str_replace_all(
      name, paste0("^(", paste(targets, collapse = "|"), ")_"), "\\1: "
    )
  ) %>% 
  mclapply(plotProfileHeatmap, mc.cores = min(length(sig_profiles), threads))
x_axis <- list(
  ggplot(data.frame(x = c(0, 1)), aes(x = x)) + geom_blank() +
    theme_void() + 
    theme(axis.title.x = element_text()) + 
    labs(x = "Distance From Centre (kb)")
)
## Wrap using patchwork
profile_heatmaps <- p_list %>% 
  lapply(
    function(p){
      df <- p$data %>% 
        mutate(treat = factor(treat, levels = unique(unlist(treat_levels))))
      rng_levels <- df %>% 
        dplyr::filter(abs(bp) < 150, target == params$tgt1) %>% 
        dplyr::filter(score == max(score), .by = c(range, name)) %>% 
        mutate(rank = rank(-score), .by = name) %>% 
        summarise(rank = mean(rank), .by = range) %>% 
        arrange(rank) %>% 
        pull(range) %>% 
        fct_rev() %>% 
        levels()
      levels(df$range) <- rng_levels
      sidey_lim <- df %>% 
        summarise(mn = mean(score), .by = c(position, name)) %>%
        pull("mn") %>% 
        range()
      ## Plot each target separately then merge using patchwork
      df %>% 
        split(.$target) %>% 
        lapply(
          function(x) {
            tgt <- unique(x$target)
            p <- x %>% 
              ggplot(aes(bp, range)) +
              geom_tile(aes(fill = score)) +
              geom_xsideline(
                aes(y = y, colour = treat),
                data = . %>% 
                  summarise(
                    y = mean(score), .by = c(position, bp, treat, name)
                  )
              ) +
              facet_grid(cols = vars(name)) +
              scale_x_continuous(
                expand = rep_len(0, 4), labels = comma_format(scale = 1e-3),
                name = NULL
              ) +
              scale_y_discrete(expand = rep_len(0, 4)) +
              scale_xsidey_continuous(limits = sidey_lim) +
              scale_colour_manual(values = treat_colours, name = "Treat") +
              scale_fill_gradientn(
                colours = heat_params[[tgt]]$gradient,
                name = glue("{tgt}\n{fill_type}")
              ) +
              theme(
                ggside.panel.scale.x = 0.3,
                panel.grid = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.title.y = element_blank()
              )
            p
          }
        )
    }
  ) %>% 
  lapply(
    \(p) {
      p[[1]] <- p[[1]] +
        geom_text(
          aes(label = label),
          data = p[[1]]$data %>% 
            summarise(
              label = glue("n = {comma(length(levels(range)))}"),
              bp = min(bp), 
              range = levels(range)[[1]],
              name = name[[1]]
            ),
          hjust = -0.1, vjust = -0.2#, fontface = "bold"
        )
      a1 <- area(1, 1, 40, 1)
      a2 <- area(1, 2, 40, 2)
      a3 <- area(41, 1, 41, length(p))
      design <- c(a1, a2, a3)
      if (length(p) == 1) design <- c(a1, a3)
      wrap_plots(c(p, x_axis), design = design) + plot_layout(guides = "collect")
    }
  )
```


```{r plot-all-profile-heatmaps}
htmltools::tagList(
  lapply(
    seq_along(profile_heatmaps),
    function(i) {
      ## Export the image
      pw <- names(profile_heatmaps)[[i]] 
      pw_name <- pw %>% 
        str_replace_all(
          "(.+) - (.+)", 
          paste(short_names$comp1, "\\1 -", short_names$comp2, "\\2")
        )
      img_out <- file.path(
        fig_path,
        pw %>% 
          str_to_lower() %>% 
          str_replace_all(
            "(.+) - (.+)", 
            paste0(
              comp_names$comp1, "-\\1-", 
              comp_names$comp2, "-\\2-profile_heatmap.", fig_type
            )
          )
      )
      n_ranges <- length(sig_ranges[[i]])
      fig_fun(
        filename = img_out, width = fw,
        height = min(3.5 + fh * n_ranges / 1.5e3, fh)
      )
      print(profile_heatmaps[[i]])
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "
        {comma(n_ranges)} overlapping ranges were defined as showing the 
        pattern {pw_name}. Ranges were centered at the midpoint of the 
        overlapping range, and ordered by taking the average signal rank across 
        all merged {params$tgt1} samples within 150bp of the peak centre. In the case of
        large sets of results, plots were restricted to {max_ranges} ranges. Values 
        shown are {fill_type} taking the {bw_type} bigwig files produced by 
        macs callpeak.
        "
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level3",
          htmltools::h3(pw_name),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  )
)
```


## Windows With The Strongest Signal {.tabset}


```{r prepare-plot-objects}
cb <- config$genome$build %>% 
  str_to_lower() %>% 
  paste0(".cytobands") 
data(list = cb)
bands_df <- get(cb)
stopifnot(
  colnames(bands_df) == c("chrom", "chromStart", "chromEnd", "name", "gieStain")
)
## Setup the coverage 
bwfl <- seq_len(2) %>% 
  lapply(
    \(i) tibble(
      target = targets[[i]],
      short_name = short_names[[i]],
      treat = unlist(treat_levels[[i]])
    )
  ) %>% 
  bind_rows() %>% 
  distinct(!!!syms(colnames(.))) %>% 
  mutate(
    bw = here::here(
      all_paths$macs2, target, 
      glue("{target}_{treat}_merged_treat_pileup.bw")
    ) %>% 
      setNames(treat)
  ) %>% 
  split(.$short_name) %>% 
  lapply(pull, "bw") %>% 
  lapply(BigWigFileList)
line_col <- bwfl %>% lapply(\(x) treat_colours[names(x)])
## Coverage annotations. 
annot <- comp_names %>% 
  paste0("_status") %>% 
  lapply(
    \(x) {
      combined_results %>% 
        splitAsList(mcols(combined_results)[[x]]) %>% 
        .[!grepl("Undetected", names(.))] %>% 
        endoapply(granges)
    }
  ) %>% 
  setNames(short_names)
annot_col <- direction_colours
## Coverage y-limits
max_ranges <- comp_names %>% 
  lapply(
    \(x) {
      cols <- paste0(x, "_log", c("CPM", "FC"))
      combined_results %>% 
        plyranges::select(all_of(cols)) %>% 
        mutate(
          max = !!sym(cols[[1]]) + !!sym(cols[[2]])
        ) %>% 
        arrange(-1 * max) %>% 
        .[1:10] %>% 
        granges()
    }
  ) %>% 
  GRangesList() %>% 
  unlist()
y_lim <- bwfl %>% 
  lapply(
    \(x) {
      lapply(x, import.bw, which = max_ranges) %>% 
        GRangesList() %>% 
        unlist()
    }
  ) %>% 
  lapply(\(x) c(0, max(x$score)))
## The features track
feat_gr <- list(Regions = endoapply(gene_regions, granges))
if (has_features) {
  feat_gr <- c(
    feat_gr,
    lapply(features, \(x) endoapply(splitAsList(x, x$feature), granges))
  )
}
hfgc_feat_col <- feat_gr %>% 
  lapply(
    \(x) c(colours$regions, feature_colours)[names(x)]
  )
featnames <- names(feat_gr) %>% str_sep_to_title(replacement = "\n")
## The genes track defaults to transcript models
hfgc_genes <- read_rds(
  here::here("output", "annotations", "trans_models.rds")
)
gene_col <- "grey"
collapseTrans <- "auto"
if (length(rna)) {
  hfgc_genes <- hfgc_genes %>% 
    mutate(
      Detected = ifelse(
        gene %in% bind_rows(rna)$gene_id, "Detected", "Undetected"
      )
    ) %>% 
    split(.$Detected)
  gene_col <- c(
    Detected = colours$direction[["unchanged"]],
    Undetected = colours$direction[["undetected"]]
  )
  collapseTrans <- list(Detected = "auto", Undetected = "meta")
}
```



For the initial set of visualisations, the sets of windows with the strongest signal were selected from within each group.
Groups were restricted to those where differential signal was seen in at least comparison.

```{r grl-to-plot-avexpr}
grl_to_plot <- combined_results %>% 
  subset(str_detect(status, "(In|De)creased")) %>% 
  arrange(
    (-1) * case_when(
      !!sym(paste0(comp_names$comp1, "_status")) == "Undetected" ~ !!sym(paste0(comp_names$comp2, "_logCPM")),
            !!sym(paste0(comp_names$comp2, "_status")) == "Undetected" ~ !!sym(paste0(comp_names$comp2, "_logCPM")),
      TRUE ~ (
        !!sym(paste0(comp_names$comp1, "_logCPM")) +
          !!sym(paste0(comp_names$comp2, "_logCPM"))
      )
    )
  ) %>% 
  splitAsList(.$status) %>% 
  .[map_int(., length) > 0] %>% 
  endoapply(\(x) x[1])
```


```{r plot-top-ranges-aveexpr, echo=FALSE}
.makeCaption <- function(gr, feat = "") {
  status <- gr$status %>% 
    str_replace_all(
      "(.+) - (.+)", paste(short_names$comp1, "\\1 -", short_names$comp2, "\\2")
    )
  gn <- glue_collapse(unlist(gr$gene_name), sep = ", ", last = " & ")
  if (length(gn)) {
    gn <- glue("The most probable regulatory targets include ", gn, ". ")
  } else {
    gn <- ""
  }
  reg <- glue_collapse(gr$region, sep = ", ", last = " & ")
  reg <- ifelse(
    grepl("^[AEIOUaeiou]", reg), paste("an", reg), paste("a", reg)
  )
  if (has_features) {
    feat <- glue_collapse(unlist(gr$feature), sep = ", ", last = " & ")
    feat <- ifelse(
      grepl("^[AEIOUaeiou]", feat), paste("an", feat), paste("a", feat)
    )
    feat <- glue(
      "When comparing to external features, this locus was most associated with {feat}."
    )
  }
  w <- width(gr)
  rng <- as.character(gr)
  glue(
    "This {comma(w)}bp region ({rng}) was classified as {status}, and overlaps {reg}. ",
    gn, feat
  )
}
htmltools::tagList(
  lapply(
    seq_along(grl_to_plot),
    function(i) {
      ## Export the image
      pw <- names(grl_to_plot)[[i]]
      pw_name <- pw %>% 
        str_replace_all(
          "(.+) - (.+)", 
          paste(short_names$comp1, "\\1 -", short_names$comp2, "\\2")
        )
      img_out <- file.path(
        fig_path,
        pw %>% 
          str_to_lower() %>% 
          str_replace_all(
            "(.+) - (.+)", 
            paste0(
              comp_names$comp1, "-\\1-", 
              comp_names$comp2, "-\\2-by-aveexpr.", fig_type
            )
          )
      )
      fig_fun(filename = img_out, width = fw, height = fh)
      plotHFGC(
        grl_to_plot[[i]],
        hic = hic,
        features = feat_gr, featcol = hfgc_feat_col, featname = featnames,
        featsize = max(1, 0.8 * length(feat_gr)), featstack = "dense",
        genes = hfgc_genes, genecol = gene_col, 
        genesize = rep_len(c(2, 1), length(hfgc_genes)),
        coverage = bwfl, linecol = line_col,  ylim = y_lim,
        annotation = annot, annotcol = direction_colours,
        cytobands = bands_df,
        collapseTranscripts = collapseTrans, maxTrans = 15,
        zoom = 20, max = max(1e4, width(range(grl_to_plot[[i]]))),
        col.title = "black", background.title = "white", 
        rotation.title = 90, showAxis = FALSE, fontsize = 10,
        highlight = rgb(0, 0, 1, 0.3)
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$caption(
        htmltools::tags$em(
          .makeCaption(grl_to_plot[[i]])
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level3",
          htmltools::h3(pw_name),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = 960),
            cp
          )
        )
      )
    }
  )
)  
```

# HERE

## Windows With The Largest Change {.tabset}

For the initial set of visualisations, the sets of windows with the strongest signal were selected from within each group.
Groups were restricted to those where differential signal was seen in at least comparison.

```{r grl-to-plot-logfc}
grl_to_plot <- all_windows %>% 
  filter(
    !str_detect(status, "Un.+Un"),
    vapply(detected, length, integer(1)) > 0
  ) %>% 
  as_tibble() %>% 
  rename_all(
    str_replace_all, pattern = "\\.Vs\\.\\.", replacement = " Vs. "
  ) %>% 
  rename_all(
    str_replace_all, pattern = "\\.\\.", replacement = ": "
  ) %>% 
  nest(
    logFC = ends_with("logFC")
  ) %>% 
  mutate(
    max_logFC = vapply(logFC, function(x) max(abs(unlist(x))), numeric(1))
  ) %>% 
  group_by(status) %>% 
  filter(max_logFC == max(max_logFC)) %>% 
  dplyr::select(
    range, status, ends_with("logFC")
  ) %>% 
  droplevels() %>% 
  split(.$status) %>% 
  lapply(pull, "range") %>% 
  lapply(GRanges, seqinfo = sq) %>% 
  GRangesList()
```

```{r update-plot-objects}
## Coverage y-limits
gr <- unlist(grl_to_plot)
y_lim <- lapply(
  bwfl,
  function(x) {
    cov <- lapply(x, import.bw, which = gr)
    unlist(lapply(cov, function(y) max(y$score))) %>% 
      c(0) %>% 
      range()
  }
)

## External Coverage (Optional)
if (!is.null(config$external$coverage)) {
  y_ranges <- grl_to_plot %>% 
    unlist() %>% 
    granges() %>% 
    resize(w = 10 * width(.), fix = 'center')
  
  y_lim <- c(
    y_lim[comps],
    bwfl[names(config$external$coverage)] %>% 
      lapply(
        function(x) {
          GRangesList(lapply(x, import.bw, which = y_ranges)) %>% 
            unlist() %>% 
            filter(score == max(score)) %>% 
            mcols() %>% 
            .[["score"]] %>% 
            c(0) %>% 
            range()
        }
      )
  )
  
}
```

```{r plot-top-ranges-logfc}
htmltools::tagList(
  mclapply(
    seq_along(grl_to_plot),
    function(x) {
      ## Export the image
      img_out <- file.path(
        fig_path,
        names(grl_to_plot)[[x]] %>% 
          str_remove_all("(:|Vs. |- )") %>% 
          str_replace_all(" ", "_") %>% 
          paste0("_logFC.", fig_type)
      )
      fig_fun(
        filename = img_out, width = 10, height = 8
      )
      ct <- FALSE
      if (length(subsetByOverlaps(gtf_transcript, grl_to_plot[[x]])) > 10)
        ct <- "meta"
      plotHFGC(
        grl_to_plot[[x]],
        features = feat_gr, featcol = feat_col_hfgc,
        featsize = 1 + has_features,
        genes = hfgc_genes, genecol = gene_col,
        coverage = bwfl, linecol = line_col,
        annotation = annot, annotcol = annot_col,
        cytobands = bands_df,
        rotation.title = 90,
        zoom = 10,
        ylim = y_lim,
        collapseTranscripts = ct,
        col.title = "black", background.title = "white", showAxis = FALSE
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      gn <- unlist(subsetByOverlaps(all_windows, grl_to_plot[[x]])$detected)
      cp <- htmltools::tags$em(
        glue(
          "
        Highlighted region corresponds to the largest change within the group 
        {names(grl_to_plot)[[x]]}. The most likely target 
        {ifelse(length(gn) == 1, 'gene is ', 'genes are ')}
        {collapseGenes(gn, format = '')}
        "
        )
      )
      htmltools::div(
        htmltools::div(
          id = names(grl_to_plot)[[x]] %>% 
            str_remove_all("(:|Vs. |- )") %>% 
            str_replace_all(" ", "-") %>% 
            str_to_lower() %>% 
            paste0("-logfc"),
          class = "section level3",
          htmltools::h3(names(grl_to_plot)[[x]]),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = 960),
            htmltools::p(
              class = "caption", htmltools::tags$em(cp)
            )
          )
        )
      )
    },
    mc.cores = threads
  )
)  
```

# Enrichment Analysis 


```{r msigdb}
min_gs_size <- extra_params$enrichment$min_size
if (is.null(min_gs_size) | is.na(min_gs_size)) min_gs_size <- 0
max_gs_size <- extra_params$enrichment$max_size
if (is.null(max_gs_size) | is.na(max_gs_size)) max_gs_size <- Inf
min_sig <- extra_params$enrichment$min_sig
if (is.null(min_sig) | is.na(min_sig)) min_sig <- 1

all_ids <- gtf_gene %>%
  mutate(w = width) %>% 
  as_tibble() %>% 
  dplyr::select(gene_id, w) %>% 
  arrange(desc(w)) %>% 
  distinct(gene_id, w) %>% 
  arrange(gene_id) %>% 
  left_join(
    all_windows$gene_id %>%
      unlist() %>%
      table() %>%
      enframe(name = "gene_id", value = "n_windows"),
    by = "gene_id"
  ) %>% 
  dplyr::filter(gene_id %in% gtf_gene$gene_id) %>% 
  mutate(
    n_windows = as.integer(ifelse(is.na(n_windows), 0, n_windows))
  )
msigdb <- msigdbr(species = extra_params$enrichment$species) %>% 
  dplyr::filter(
    gs_cat %in% unlist(extra_params$enrichment$msigdb$gs_cat) |
      gs_subcat %in% unlist(extra_params$enrichment$msigdb$gs_subcat),
    str_detect(ensembl_gene, "^E")
  ) %>% 
  dplyr::rename(gene_id = ensembl_gene, gene_name = gene_symbol) %>% # For easier integration 
  dplyr::select(-starts_with("human"), -contains("entrez")) %>% 
  dplyr::filter(gene_id %in% all_ids$gene_id) %>% 
  group_by(gs_name) %>% 
  mutate(n = dplyr::n()) %>% 
  ungroup() %>% 
  dplyr::filter(n >= min_gs_size, n <= max_gs_size) 
gs_by_gsid <- msigdb %>% 
  split(.$gs_name) %>% 
  mclapply(pull, "gene_id", mc.cores = threads) %>% 
  mclapply(unique, mc.cores = threads)
gs_by_geneid <- msigdb %>% 
  split(.$gene_id) %>% 
  mclapply(pull, "gs_name", mc.cores = threads) %>% 
  mclapply(unique, mc.cores = threads)
gs_url <- msigdb %>% 
  distinct(gs_name, gs_url) %>%
  mutate(
    gs_url = ifelse(
      gs_url == "", "
      http://www.gsea-msigdb.org/gsea/msigdb/collections.jsp",
      gs_url
    )
  ) %>% 
  with(setNames(gs_url, gs_name))

min_network_size <- extra_params$networks$min_size
if (is.null(min_network_size) | is.na(min_network_size))
  min_network_size <- 2

max_network_size <- extra_params$networks$max_size
if (is.null(max_network_size) | is.na(max_network_size))
  max_network_size <- Inf

max_network_dist <- extra_params$networks$max_distance
if (is.null(max_network_dist) | is.na(max_network_dist))
  max_network_dist <- 1

net_layout <- extra_params$networks$layout
enrich_alpha <- extra_params$enrichment$alpha
adj_method <- match.arg(extra_params$enrichment$adj, p.adjust.methods)
adj_desc <- case_when(
  p.adjust.methods %in% c("fdr", "BH") ~ "the Benjamini-Hochberg FDR",
  p.adjust.methods %in% c("BY") ~ "the Benjamini-Yekutieli FDR",
  p.adjust.methods %in% c("bonferroni") ~ "the Bonferroni",
  p.adjust.methods %in% c("holm") ~ "Holm's",
  p.adjust.methods %in% c("hommel") ~ "Hommel's",
  p.adjust.methods %in% c("hochberg") ~ "Hochberg's",
  p.adjust.methods %in% c("none") ~ "no"
) %>% 
  setNames(p.adjust.methods)
comp_ids <- comps %>% 
  lapply(
    function(x) {
      subset(all_windows, !str_detect(status, paste(x, "Undetected")))$gene_id
    }
  ) %>% 
  lapply(unlist) %>% 
  lapply(unique) %>% 
  lapply(intersect, gtf_gene$gene_id) %>% 
  setNames(comps)
mapped_ids <- comp_ids %>% 
  unlist() %>% 
  unique()
gene_lengths <- structure(width(gtf_gene), names = gtf_gene$gene_id)[mapped_ids]
group_ids <- all_windows %>% 
  split(f = .$status) %>% 
  mclapply(
    function(x) unique(unlist(x$gene_id)),
    mc.cores = threads
  ) 
```

## Comparison of Genes Mapped To Targets

The first level of enrichment testing performed was to simply check the genes mapped to a site in either comparison, with no consideration being paid to the dynamics of signal in either comparison.
This will capture and describe the overall biological activity being jointly regulated.

The genes in each section and intersection in the following Venn Diagram will then be analysed against the complete set of `r comma(length(unique(unlist(comp_ids))))` genes mapped to either `r glue_collapse(comps, last = ' or ')`.
Firstly, all `r comma(length(mapped_ids))` genes mapped to a site were tested for enrichment in comparison to the set of `r comma(length(unique(gtf_gene$gene_id)))` `r ifelse(has_rnaseq, "detected", "annotated")` genes.
For this analysis, gene width was used as a potential source of sampling bias and Wallenius' Non-Central Hypergeometric was used as implemented in `goseq` [@YoungGoseq2010].

The sets of genes mapped to each section of the Venn Diagram were then tested for enrichment in comparison to the larger set of `r comma(length(mapped_ids))` mapped genes.
The standard hypergeometric distribution was used for these analyses.

If `r min_network_size` or more gene-sets were considered to be enriched, a network plot was produced using the same methodology as for results from differential signal for individual comparisons.

```{r venn-shared-genes, results = 'hide', fig.show = 'asis', fig.cap = glue("*Venn Diagram showing __genes__ mapped to a signal region associated with {glue_collapse(comps, sep = ' or ')}. Of the {comma(nrow(all_ids))} genes under consideration, a total of {comma(length(mapped_ids))} genes were mapped to a signal region across both comparisons, leaving {comma(nrow(all_ids) - length(mapped_ids))} genes unmapped to a signal region across either comparison.*")}
plotOverlaps(
  comp_ids, set_col = comp_cols[comps], cex = 1.2, cat.cex = 1.2
)
```

### Result Tables {.tabset}

#### Mapped Genes To Either `r glue_collapse(comps, last = " or ")`

```{r goseq-mapped-res, fig.show = 'hide'}
pwf_mapped <- gtf_gene %>%
  mutate(w = width) %>%
  as_tibble() %>% 
  arrange(gene_id, desc(w)) %>%
  distinct(gene_id, w) %>% 
  mutate(mapped = gene_id %in% mapped_ids) %>% 
  with(nullp(setNames(mapped, gene_id), bias.data = log10(w)))
goseq_mapped_res <- pwf_mapped %>% 
  goseq(gene2cat = gs_by_geneid) %>% 
  dplyr::filter(numDEInCat > 0) %>% 
  as_tibble() %>% 
  dplyr::select(
    gs_name = category, pval = over_represented_pvalue, starts_with("num")
  ) %>% 
  mutate(
    adj_p = p.adjust(pval, adj_method),
    enriched = adj_p < enrich_alpha
  ) %>% 
  dplyr::filter(numDEInCat >= min_sig) %>% 
  left_join(
    dplyr::filter(msigdb, gene_id %in% mapped_ids)[c("gs_name", "gene_name", "gs_url", "gs_description")],
    by = "gs_name"
  ) %>% 
  chop(gene_name) %>% 
  mutate(
    gene_name = vapply(gene_name, paste, character(1), collapse = "; ")
  )
any_goseq_mapped <- sum(goseq_mapped_res$adj_p < enrich_alpha) > 0
tg_mapped <- make_tbl_graph(
  goseq_mapped_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, rownames(subset(pwf_mapped, DEgenes))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_mapped <- length(tg_mapped) >= min_network_size
```


```{r tbl-goseq-mapped-res, eval = any_goseq_mapped, echo = any_goseq_mapped}
tbl <- goseq_mapped_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = "Mapped Genes",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes in Gene-Set",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
cp <- glue(
  "
  All {comma(length(mapped_ids))} genes with a mapped window from 
  either comparison were compared to the complete set of 
  {comma(length(unique(gtf_gene$gene_id)))} annotated genes, in order to 
  ascertain which key gene-sets and pathways were likely to be targeted in either analysis. 
  ",
  ifelse(
    has_rnaseq,
    "Genes were restricted to those in the set of detected genes from the provided RNA-Seq dataset. ", 
    ""
  ),
  "
  Gene-length was included as the source of any potential sampling bias for a 
  gene being mapped to any target-signal range. This analysis provides an overall 
  viewpoint as to which pathways are regulated across either dataset.
  "
)
div(class = "table",
    div(class = "table-header", div(class = "caption", cp)),
    tbl
)
```

`r if(!any_goseq_mapped) "No pathway enrichment was detected amongst genes mapped in either comparison."`


`r if (n_targets == 2) glue("#### Genes Mapped to Both ", glue_collapse(comps, last = " and " ))`

```{r empty-goseq-t1t2-res, echo = FALSE}
goseq_t1t2_res <- tibble(
  gs_name = character(), mapped = numeric(), enriched = logical(),
  adj_p = numeric()
)
vec_t1t2 <- structure(
  mapped_ids %in% intersect(comp_ids[[2]], comp_ids[[1]]),
  names = mapped_ids
)
any_goseq_t1t2 <- plot_network_t1t2 <- FALSE
para <- glue(
  "
  All {comma(length(intersect(comp_ids[[2]], comp_ids[[1]])))} genes with a 
  mapped window to both {glue_collapse(comps, last = ' and ')} were compared to 
  the {comma(length(mapped_ids))} genes mapped to either, 
  in order to ascertain whether any key gene-sets were likely to be targeted by 
  both factors in combination. No sampling bias offset was incorporated, instead 
  using a simple Hypergeometric model for enrichment testing.
  "
)
```


```{r goseq_t1t2_res, eval = n_targets == 2, fig.show = 'hide'}
pwf_t1t2 <- nullp(
  vec_t1t2,
  bias.data = log10(setNames(all_ids$w, all_ids$gene_id)[mapped_ids]), 
  plot.fit = TRUE
) 
if (sum(pwf_t1t2$DEgenes) > 0) {
  goseq_t1t2_res <- pwf_t1t2 %>% 
    goseq(gene2cat = gs_by_geneid, method = gs_method) %>% 
    dplyr::filter(numDEInCat > 0) %>% 
    mutate(
      adj_p = p.adjust(pval, adj_method),
      enriched = adj_p < enrich_alpha,
      DEgenes = lapply(DEgenes, function(x) id2gene[x])
    ) %>% 
    dplyr::filter(numDEInCat >= min_sig) %>% 
    left_join(
      distinct(msigdb, gs_name, gs_url, gs_description),
      by = "gs_name"
    ) %>% 
    dplyr::rename(gene_name = DEgenes) %>% 
    mutate(gene_name = vapply(gene_name, paste, character(1), collapse = "; "))
}
any_goseq_t1t2 <- sum(goseq_t1t2_res$adj_p < enrich_alpha) > 0
tg_t1t2 <- make_tbl_graph(
  goseq_t1t2_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, names(which(vec_t1t2))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_t1t2 <- length(tg_t1t2) >= min_network_size
if (!any_goseq_t1t2) para <- glue(para, " No enrichment was found.")
```

`r if (!any_goseq_t1t2) para`

```{r tbl-goseq-t1t2-res, eval = any_goseq_t1t2 & n_targets == 2, echo = any_goseq_t1t2}
tbl <- goseq_t1t2_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = "Genes Mapped to Both",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes Mapped to Either",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
div(class = "table",
    div(class = "table-header", div(class = "caption", para)),
    tbl
)
```


`r if (n_targets == 2) glue("#### Genes Mapped to {comps[[1]]} Only")`


```{r empty-goseq-t1only}
goseq_t1only_res <- tibble(
  gs_name = character(), mapped = numeric(), enriched = logical(),
  adj_p = numeric()
)
vec_t1only <- structure(
  mapped_ids %in% setdiff(comp_ids[[1]], comp_ids[[2]]),
  names = mapped_ids
)
gs_method <- "Hypergeometric"
any_goseq_t1only <- plot_network_t1only <- FALSE
para <- glue(
  "
  All {comma(length(setdiff(comp_ids[[1]], comp_ids[[2]])))} genes with a 
  mapped window to only {comps[[1]]} were compared to 
  the {comma(length(mapped_ids))} genes mapped to either
  {glue_collapse(comps, last = ' or ')}, in order to ascertain whether any gene-sets 
  were likely to be targeted exclusively. No sampling bias offset was 
  incorporated, instead using a simple Hypergeometric model for enrichment testing.
  "
)
```

```{r goseq-t1only-res, fig.show = 'hide'}
pwf_t1only <- nullp(
  vec_t1only,
  bias.data = log10(setNames(all_ids$w, all_ids$gene_id)[mapped_ids]), 
  plot.fit = TRUE
) 
if (sum(pwf_t1only$DEgenes) > 0) {
  goseq_t1only_res <- pwf_t1only %>% 
    goseq(gene2cat = gs_by_geneid, method = gs_method) %>% 
    dplyr::filter(numDEInCat > 0) %>% 
    mutate(
      adj_p = p.adjust(pval, adj_method),
      enriched = adj_p < enrich_alpha,
      DEgenes = lapply(DEgenes, function(x) id2gene[x])
    ) %>% 
    dplyr::filter(numDEInCat >= min_sig) %>% 
    left_join(
      distinct(msigdb, gs_name, gs_url, gs_description),
      by = "gs_name"
    ) %>% 
    dplyr::rename(gene_name = DEgenes) %>% 
    mutate(gene_name = vapply(gene_name, paste, character(1), collapse = "; "))
}
any_goseq_t1only <- sum(goseq_t1only_res$adj_p < enrich_alpha) > 0
tg_t1only <- make_tbl_graph(
  goseq_t1only_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, names(which(vec_t1only))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_t1only <- length(tg_t1only) >= min_network_size
if (!any_goseq_t1only) para <- glue(para, " No enrichment was found.")
```

`r if (!any_goseq_t1only) para`

```{r tbl-goseq-t1only-res, eval = n_targets == 2 & any_goseq_t1only, echo = any_goseq_t1only}
tbl <- goseq_t1only_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = glue("Genes Mapped to {comps[[1]]} Only"),
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes Mapped to Both",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
div(class = "table",
    div(class = "table-header", div(class = "caption", para)),
    tbl
)
```


`r if (n_targets == 2) glue("#### Genes Mapped to {comps[[2]]} Only")`


```{r empty-goseq-t2only}
goseq_t2only_res <- tibble(
  gs_name = character(), mapped = numeric(), enriched = logical(),
  adj_p = numeric()
)
vec_t2only <- structure(
  mapped_ids %in% setdiff(comp_ids[[2]], comp_ids[[1]]),
  names = mapped_ids
)
gs_method <- "Hypergeometric"
para <- glue(
  "
  All {comma(length(setdiff(comp_ids[[2]], comp_ids[[1]])))} genes with a 
  mapped window to only {comps[[2]]} were compared to 
  the {comma(length(mapped_ids))} genes mapped to either 
  {glue_collapse(comps, last = ' or ')}, in order to ascertain whether any gene-sets 
  were likely to be targeted exclusively. No sampling bias offset was 
  incorporated, instead using a simple Hypergeometric model for enrichment testing.
  "
)
any_goseq_t2only <- plot_network_t2only <- FALSE
```

```{r goseq-t2only-res, fig.show = 'hide'}
pwf_t2only <- nullp(
  vec_t2only,
  bias.data = log10(setNames(all_ids$w, all_ids$gene_id)[mapped_ids]), 
  plot.fit = TRUE
) 
if (sum(pwf_t2only$DEgenes) > 0) {
  goseq_t2only_res <- pwf_t2only %>% 
    goseq(gene2cat = gs_by_geneid, method = gs_method) %>% 
    dplyr::filter(numDEInCat > 0) %>% 
    mutate(
      adj_p = p.adjust(pval, adj_method),
      enriched = adj_p < enrich_alpha,
      DEgenes = lapply(DEgenes, function(x) id2gene[x])
    ) %>% 
    dplyr::filter(numDEInCat >= min_sig) %>% 
    left_join(
      distinct(msigdb, gs_name, gs_url, gs_description),
      by = "gs_name"
    ) %>% 
    dplyr::rename(gene_name = DEgenes) %>% 
    mutate(gene_name = vapply(gene_name, paste, character(1), collapse = "; "))
}
any_goseq_t2only <- sum(goseq_t2only_res$adj_p < enrich_alpha) > 0
tg_t2only <- make_tbl_graph(
  goseq_t2only_res, 
  gs = gs_by_gsid %>% 
    lapply(intersect, names(which(vec_t2only))) %>% 
    .[vapply(., length, integer(1)) > 0]
)
plot_network_t2only <- length(tg_t2only) >= min_network_size
if (!any_goseq_t2only) para <- glue(para, " No enrichment was found.")
```

`r if (!any_goseq_t2only) para`

```{r tbl-goseq-t2only-res, eval = any_goseq_t2only, echo = any_goseq_t2only}
tbl <- goseq_t2only_res %>%
  dplyr::filter(enriched) %>% 
  dplyr::select(
    gs_name, gs_description, pval, adj_p, starts_with("num"), gene_name
  ) %>% 
  reactable(
    filterable = TRUE,
    showPageSizeOptions = TRUE,
    columns = list2(
      gs_name = colDef(
        name = "Gene Set",
        cell = function(value) htmltools::tags$a(
          href = gs_url[[value]], 
          target = "_blank", 
          str_replace_all(value, "_", " ")
        ),
        html = TRUE,
        maxWidth = 150
      ),
      gs_description = colDef(
        name = "Description",
        cell = function(value) str_trunc(value, width = 150)
      ),
      pval = colDef(
        name = "P-Value",
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      adj_p = colDef(
        name = "p<sub>adj</sub>", html = TRUE,
        cell = function(value) ifelse(
          value < 0.001,
          sprintf("%.2e", value),
          sprintf("%.3f", value)
        ),
        maxWidth = 80
      ),
      numDEInCat = colDef(
        name = glue("Genes Mapped to {comps[[2]]} Only"),
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      numInCat = colDef(
        name = "Genes Mapped to Both",
        cell = function(value) comma(value, 1),
        maxWidth = 80
      ),
      gene_name = colDef(
        name = "Gene Names",
        cell = function(value) with_tooltip(value, width = 150)
      )
    )
  )
div(class = "table",
    div(class = "table-header", div(class = "caption", para)),
    tbl
)
```


### Network Plots {.tabset}

`r if (plot_network_mapped) "#### All Mapped Genes"`

```{r plot-network-mapped, eval = plot_network_mapped, echo = plot_network_mapped, fig.cap = glue("*Network plot showing  gene-sets enriched amongst the overall set of sites with signal detected in __either__ comparison.*")}
tg_mapped %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_mapped) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```

`r if (plot_network_t1t2) paste("#### Genes Mapped to Both", glue_collapse(comps, last = ' and '))`

```{r plot-network-t1t2, eval = plot_network_t1t2, echo = plot_network_t1t2, fig.cap = glue("*Network plot showing gene-sets enriched amongst the overall set of sites with detected signal in __both__ {comps[[1]]} and {comps[[2]]}.*")}
tg_t1t2 %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_t1t2) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```

`r if (plot_network_t1only) glue("#### Genes Mapped to {comps[[1]]} Only")`

```{r plot-network-t1only, eval = plot_network_t1only, echo = plot_network_t1only, fig.cap = glue("*Network plot showing all gene-sets enriched amongst the overall set of sites with detected signal in __{comps[[1]]} only__.*")}
tg_t1only %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_t1only) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```

`r if (plot_network_t2only) glue("#### Genes Mapped to {comps[[2]]} Only")`

```{r plot-network-t2only, eval = plot_network_t2only, echo = plot_network_t2only, fig.cap = glue("*Network plot showing all gene-sets enriched amongst the overall set of sites with detected signal in __{comps[[2]]} only__.*")}
tg_t2only %>% 
  ggraph(layout = net_layout, weights = oc^2) +
  geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
  geom_node_point(
    aes(fill = -log10(pval), size = numDEInCat),
    shape = 21
  ) +
  geom_node_text(
    aes(label = label),
    colour = "black", size = 3, 
    data = . %>%
      mutate(
        label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
      ),
    repel = TRUE, max.overlaps = max(10, round(length(tg_t2only) / 4, 0)),
    bg.color = "white", bg.r = 0.1, 
  ) +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_viridis_c(option = "inferno", begin = 0.25) +
  scale_size_continuous(range = c(1, 10)) +
  scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
  scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
  guides(edge_alpha = "none") +
  labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
  theme_void() 
```


## Comparison of Differential Signal Regions

```{r goseq-groups, fig.show = 'hide'}
n_peaks <- structure(all_ids$n_windows, names = all_ids$gene_id)[mapped_ids]
goseq_group_res <- group_ids %>% 
  .[str_detect(names(.), "Up|Down")] %>% 
  .[vapply(., length, integer(1)) > 0] %>% 
  mclapply(
    function(x) {
      vec <- structure(mapped_ids %in% x, names = mapped_ids)
      res <- tibble(adj_p = numeric())
      if (sum(vec) > 0) {
        res <- data.frame(DEgenes = vec) %>% 
          goseq_hyper(gene2cat = gs_by_geneid, cores = 1) %>% 
          dplyr::filter(numDEInCat > 0) %>% 
          mutate(adj_p = p.adjust(pval, adj_method)) %>% 
          dplyr::filter(numDEInCat >= min_sig) %>% 
          dplyr::select(gs_name, pval, adj_p, ends_with("InCat"), DEgenes) %>% 
          left_join(
            dplyr::distinct(msigdb, gs_name, gs_url, gs_description),
            by = "gs_name"
          ) %>% 
          mutate(
            gene_name = vapply(
              DEgenes, function(x) paste(id2gene[x], collapse = "; "), character(1)
            )
          )
      }
      res
    },
    mc.cores = threads
  ) %>% 
  ## This step at the end basically removes any groups with no results
  lapply(list) %>% 
  as_tibble() %>% 
  pivot_longer(everything(), names_to = "group") %>% 
  unnest(everything()) %>% 
  split(.$group)
goseq_group_sig <- goseq_group_res %>% 
  lapply(dplyr::filter, adj_p < enrich_alpha) %>% 
  .[vapply(., nrow, integer(1)) > 0]
tg_group_res <- names(goseq_group_sig) %>% 
  lapply(
    function(x) {
      make_tbl_graph(
        goseq_group_sig[[x]],
        gs_by_gsid[goseq_group_sig[[x]]$gs_name] %>%
          lapply(intersect, group_ids[[x]])
      )
    }
  ) %>% 
  setNames(names(goseq_group_sig)) %>% 
  .[vapply(., length, integer(1)) >= min_network_size]
```

The same MSigDB gene sets were used for associating any pathways with the combined changes in both `r collapseGenes(comps)`.
The 'universe' of genes was defined as all `r comma(length(mapped_ids))` gene ids with a window mapped from either target.
No sampling bias was included, with test results using a simple Hypergeometric Distribution.
`r if(length(goseq_group_sig) == 0) "No enrichment was found amongst any group of windows"`


`r if(length(goseq_group_sig) > 0) "### Results {.tabset}"`


```{r tbl-goseq-groups, eval = length(goseq_group_sig) > 0, echo = length(goseq_group_sig) > 0}
htmltools::tagList(
  names(goseq_group_sig) %>% 
    mclapply(
      function(x) {
        cp <- glue(
          "
          All {comma(length(group_ids[[x]]), 1)} genes mapped to windows
           with the pattern {x} were compared to the complete set of 
          {comma(length(mapped_ids), 1)} genes mapped to at least one
          window. {nrow(goseq_group_sig[[x]])} gene-sets were considered as 
          enriched using a threshold of {enrich_alpha} for adjusted-p-values.
          "
        )
        htmltools::div(
          htmltools::div(
            id = x %>% 
              str_to_lower %>% 
              str_replace_all(" ", "-") %>% 
              str_replace_all("-+", "-"),
            class = "section level4",
            htmltools::h4(class = "tabset", x),
            htmltools::tags$em(cp),
            goseq_group_sig[[x]] %>% 
              separate(group, into = comps, sep = " - ") %>% 
              dplyr::select(
                all_of(comps), gs_name, gs_description, pval, adj_p, everything()
              ) %>% 
              reactable(
                filterable = TRUE,
                showPageSizeOptions = TRUE,
                columns = list2(
                  "{comps[[1]]}" := colDef(
                    cell = function(value) {
                      value <- str_extract(value, "Up|Down|Unchanged|Undetected")
                      if (str_detect(value, "Up")) value <- paste(value, "\u21E7")
                      if (str_detect(value, "Down")) value <- paste(value, "\u21E9")
                      paste(targets[[1]], value)
                    },
                    style = function(value) {
                      colour <- case_when(
                        str_detect(value, "Up") ~ colours$direction[["up"]],
                        str_detect(value, "Down") ~ colours$direction[["down"]],
                        str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
                        str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
                      )
                      list(color = colour)
                    },
                    maxWidth = 120
                  ),
                  "{comps[[2]]}" := colDef(
                    cell = function(value) {
                      value <- str_extract(value, "Up|Down|Unchanged|Undetected")
                      if (str_detect(value, "Up")) value <- paste(value, "\u21E7")
                      if (str_detect(value, "Down")) value <- paste(value, "\u21E9")
                      paste(targets[[2]], value)
                    },
                    style = function(value) {
                      colour <- case_when(
                        str_detect(value, "Up") ~ colours$direction[["up"]],
                        str_detect(value, "Down") ~ colours$direction[["down"]],
                        str_detect(value, "Unchanged") ~ colours$direction[["unchanged"]],
                        str_detect(value, "Undetected") ~ colours$direction[["undetected"]]
                      )
                      list(color = colour)
                    },
                    maxWidth = 120
                  ),
                  gs_name = colDef(
                    name = "Gene Set",
                    cell = function(value) {
                      htmltools::tags$a(
                        href = gs_url[[value]], target = "_blank",
                        str_replace_all(value, "_", " ")
                      )
                    },
                    html = TRUE,
                    maxWidth = 150
                  ),
                  gs_description = colDef(
                    name = "Description",
                    cell = function(value) str_trunc(value, width = 150)
                  ),
                  gs_url = colDef(show = FALSE),
                  pval = colDef(
                    name = "P-Value",
                    cell = function(value) ifelse(
                      value < 0.001,
                      sprintf("%.2e", value),
                      sprintf("%.3f", value)
                    ),
                    maxWidth = 80
                  ),
                  adj_p = colDef(
                    name = "p<sub>adj</sub>", html = TRUE,
                    cell = function(value) ifelse(
                      value < 0.001,
                      sprintf("%.2e", value),
                      sprintf("%.3f", value)
                    ),
                    maxWidth = 80
                  ),
                  numDEInCat = colDef(
                    name = "Mapped Genes",
                    cell = function(value) comma(value, 1),
                    maxWidth = 80
                  ),
                  numInCat = colDef(
                    name = "Genes in Gene-Set",
                    cell = function(value) comma(value, 1),
                    maxWidth = 80
                  ),
                  gene_name = colDef(
                    name = "Gene Names",
                    cell = function(value) with_tooltip(value, width = 150)
                  ),
                  DEgenes = colDef(show = FALSE)
                )
              )
          )
        )
      },
      mc.cores = threads
    )
)
```

`r if (length(goseq_group_sig)) "#### All Groups"`

```{r plot-all-enriched, eval = length(goseq_group_sig) > 0, echo = length(goseq_group_sig) > 0, fig.width = min(5 + length(goseq_group_sig), knitr::opts_chunk$get("fig.width")), fig.cap = glue("*Combined enrichment across all groups, showing only significant results for enrichment. The top panel shows how many genes were mapped to sites for each group, whilst the right panel shows gene set size. Point sizes indicate the proportion of mapped genes which are from each pathway.*")}
goseq_group_sig %>%  
  bind_rows() %>% 
  arrange(pval) %>% 
  left_join(n_mapped, by = c("group" = "status")) %>% 
  mutate(
    gs_name = fct_inorder(gs_name) %>% 
      fct_relabel(str_replace_all, "_", " ") %>% 
      fct_relabel(str_trunc, width = 60),
    prop = numDEInCat / mapped,
    group = factor(group, levels = n_mapped$status)
  ) %>%
  droplevels() %>% 
  ggplot(aes(group, fct_rev(gs_name))) + 
  geom_point(aes(size = prop, fill = -log10(pval)), shape = 21) +
  geom_ysidecol(
    aes(x = numInCat), data = . %>% distinct(gs_name, numInCat),
  ) +
  geom_xsidecol(
    aes(y = mapped),
    data = n_mapped %>% 
      dplyr::filter(status %in% names(goseq_group_sig)) %>% 
      dplyr::rename(group = status),
    width = 0.5
  ) +
  scale_x_discrete(labels = label_wrap(10)) +
  scale_ysidex_continuous(expand = expansion(c(0, 0.15))) +
  scale_xsidey_continuous(
    expand = expansion(c(0, 0.15))
  ) +
  scale_fill_viridis_c(option = "inferno", begin = 0.2) +
  scale_size(range = c(0, 5), labels = percent) + 
  labs(
    x = c(), y = c(), size = "% Mapped\nGenes", fill = expr(paste(-log[10], "p"))
  ) +
  theme(
    panel.grid = element_blank(), 
    axis.text = element_text(size = 8),
    legend.position = "right",
    ggside.panel.scale.x = 0.2,
    ggside.panel.scale.y = 0.3,
    ggside.axis.text.x.bottom = element_text(angle = 270, hjust = 0, vjust = 0.5)
  ) 
```



`r if(length(tg_group_res) > 0) "### Network Plots {.tabset}"`


```{r network-plots-goseq-groups, eval = length(tg_group_res) > 0, echo = length(tg_group_res) > 0}
htmltools::tagList(
  names(tg_group_res) %>% 
    mclapply(
      function(x) {
        ## Export the image
        img_out <- file.path(
          fig_path,
          x %>% 
            str_replace_all(" ", "_") %>% 
            str_replace_all("_-_", "-") %>% 
            paste0("_network.", fig_type)
        )
        fig_fun(
          filename = img_out,
          width = knitr::opts_current$get("fig.width"),
          height = knitr::opts_current$get("fig.height")
        )
        p <- tg_group_res[[x]] %>%
          ggraph(layout = net_layout, weights = oc^2) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(fill = -log10(pval), size = numDEInCat),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% str_trunc(60) %>% str_wrap(width = 18)
              ),
            repel = TRUE,
            max.overlaps = max(10, round(length(tg_group_res[[x]]) / 4, 0)),
            bg.color = "white", bg.r = 0.1
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.25) +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1)) +
          guides(edge_alpha = "none") +
          labs(size = "Mapped Genes", edge_width = expr(paste(OC^2))) +
          theme_void()
        print(p)
        dev.off()
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "
            Network plot showing enriched pathways mapped to genes 
            associated with {x}. 
            "
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>%
              basename() %>%
              str_remove_all(glue(".{fig_type}$")) %>%
              str_to_lower() %>%
              str_replace_all("_", "-"),
            class = "section level4",
            htmltools::h4(x),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      },
      mc.cores = threads
    )
)
```




```{r add-rna, eval = has_rnaseq, child = here::here('workflow/modules/rnaseq_pairwise.Rmd')}
```

# Data Export

```{r all-out}
write_rds(all_windows, all_out$results, compress = "gz")
save.image(all_out$renv)
```

```{r export-csv}
all_windows %>% 
  as_tibble() %>% 
  dplyr::select(-detected) %>% 
  unnest(all_of("gene_id")) %>% # This ensures correct id2gene mappings
  mutate(gene_name = id2gene[gene_id]) %>% 
  dplyr::select(
    starts_with("gene"), range, any_of(c("region", "feature", "hic")),
  ) %>% 
  write_csv(
    gzfile(all_out$csv)
  )
write_csv(de_genes_both_comps, all_out$de_genes)
goseq_group_sig %>% 
  bind_rows() %>% 
  write_csv(all_out$goseq)
cmn_res %>% 
  bind_rows() %>% 
  mutate(
    leadingEdge = vapply(leadingEdge, paste, character(1), collapse = "; "),
    gs_url = gs_url[gs_name]
  ) %>% 
  write_csv(all_out$rna_enrichment)
```


During this workflow, the following files were exported:

`r pander(lapply(all_out, str_extract, "output.+"))`

## References

<div id="refs"></div>
<br>

<button type="button" class="btn btn-default btn-sessioninfo" data-toggle="collapse" data-target="#sessioninfo" style="display: block;">
<span class="glyphicon glyphicon-info-sign" aria-hidden="true"></span> Session information
</button>
</p>
<div id="sessioninfo" class="collapse">

```{r session-info, echo=FALSE}
pander::pander(sessionInfo())
```

</div>
