## Comparison With Transcriptomic Data

```{r define-rna-objects}
library(fgsea)
library(metap)
library(ggside)
results_tbl <- results %>% 
  as_tibble() %>% 
  dplyr::select(
    range, logCPM, logFC, PValue, FDR = !!sym(fdr_column), status,
    gene_id, any_of(c("region", "feature"))
  ) %>% 
  unnest(gene_id) %>% 
  mutate(gene_name = id2gene[gene_id])
rna <- lapply(rna, addDiffStatus)
has_de <- map_lgl(rna, \(x) any(x$FDR < 0.05))
dsa_status <- results_tbl %>% 
  dplyr::select(status, gene_id) %>% 
  unnest(gene_id) %>% 
  summarise(
    status = case_when(
      any(status == "Increased") & any(status == "Decreased") ~ "Ambiguous",
      any(status == "Increased") ~ "Increased",
      any(status == "Decreased") ~ "Decreased",
      any(status == "Ambiguous") ~ "Ambiguous",
      TRUE ~ "Unchanged"
    ) %>% 
      factor(levels = names(direction_colours)),
    .by = gene_id
  ) %>% 
  droplevels()
```

In order to compare DSA results for `r target`, gene-to-range mappings were summarised such that if a gene was mapped to any DSA results showing increased or decreased binding, they were given this status.
Genes with conflicting mappings were considered to be ambiguous, whilst those without any changed regions were assigned as being consistently unchanged.
These results were applied where summarised results were able to be presented, however, when comparing actual estimates for logFC etc, all DSA ranges were retained.


```{r summary-tables}
summary_tables <- names(rna) %>% 
  lapply(
    \(i) {
      x <- rna[[i]]
      id <- paste0("tbl-summary-", i) %>% 
        str_to_lower() %>% 
        str_replace_all("_", "-")
      tbl <- results_tbl %>% 
        full_join(x, by = "gene_id") %>% 
        mutate(
          status.x = fct_na_value_to_level(status.x, "Undetected"),
          status.y = fct_na_value_to_level(status.y, "Undetected")
        ) %>% 
        arrange(PValue.x) %>% 
        distinct(gene_id, .keep_all = TRUE) %>% 
        dplyr::select(gene_id, starts_with("status")) %>% 
        summarise(
          n = dplyr::n(), .by = starts_with("status")
        ) %>%
        pivot_wider(names_from = "status.y", values_from = "n", values_fill = 0) %>% 
        dplyr::select(
          starts_with("status"), any_of(c("Increased", "Decreased")), 
          everything()
        ) %>% 
        reactable(
          showPageSizeOptions = FALSE,
          columns = list(
            status.x = colDef(glue("{target} Status"), footer = "")
          ),
          defaultColDef = colDef(
            sortable = TRUE, format = comma_col,
            footer = function(values) comma(sum(values)),
            footerStyle = list(fontWeight = "bold")
          ),
          columnGroups = list(
            colGroup(
              "Gene Expression", str_subset(names(.), "ed$")
            )
          )
        )
      cp <- htmltools::em(
        glue(
          "Comparison between {target} status and gene expression in {i}."
        )
      )
      htmltools::div(
        htmltools::div(
          id = id,
          class = "section level4",
          htmltools::h4("Summary Table"),
          htmltools::div(
            class = "table", tbl, htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(rna))
```

```{r donut-plots}
donuts <- rna %>% 
  lapply(left_join, dsa_status, by = "gene_id") %>% 
  lapply(
    mutate, 
    status.x = fct_na_value_to_level(status.x, "Undetected"),
    status.y = fct_na_value_to_level(status.y, "Undetected"),
  ) %>% 
  lapply(
    dplyr::rename, RNA = status.x, "{target}" := status.y
  ) %>% 
  mclapply(
    plotSplitDonut,
    inner = !!sym(target), outer = "RNA",
    inner_palette = direction_colours, outer_palette = direction_colours,
    inner_glue = "{inner}\n{.data[[inner]]}\n{comma(n, 1)}\n{percent(p,0.1)}",
    inner_label_alpha = 0.8, inner_min_p = 0.04,
    outer_glue = "{.data[[outer]]} {comma(n, 1)}",
    outer_min_p = 0.001, outer_label_colour = "palette", 
    outer_rotate = TRUE,
    outer_label = "text",
    mc.cores = threads
  ) 
donut_plots <- names(donuts) %>% 
  lapply(
    \(x) {
      ## Export the image
      img_out <- file.path(fig_path, glue("donut_{x}.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      print(donuts[[x]])
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "All detected genes within the {x} dataset were compared to {target} ",
          "DSA results for {treat_levels[[2]]} Vs {treat_levels[[1]]}. Regions ",
          "with signal contained within the DSA results were previously mapped ",
          "to genes, and these mappings were used. Results for each gene were ",
          "summarised to account for multiple regions with different DSA results."
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level4",
          htmltools::h4("Donut Plot"),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(donuts))
```

```{r ma-plots}
ma_plots <- names(rna) %>% 
  lapply(
    \(i) {
      p <- rna[[i]] %>% 
        left_join(dsa_status, by = "gene_id") %>% 
        mutate(
          status.y = fct_na_value_to_level(status.y, "Undetected") %>% 
            fct_relabel(\(x) paste(target, x)),
          gene_name = id2gene[gene_id]
        ) %>% 
        droplevels() %>% 
        arrange(1 / PValue) %>% 
        ggplot(aes(logCPM, logFC)) +
        geom_point(aes(colour = status.x), alpha = 0.7) +
        geom_xsideboxplot(
          aes(logCPM, fct_rev(status.y), fill = status.y), orientation = "y",
          show.legend = FALSE
        ) +
        geom_ysideboxplot(
          aes(fct_rev(status.y), logFC, fill = status.y), orientation = "x"
        ) +
        geom_label_repel(
          aes(label = gene_name, colour = status.x),
          data = . %>% 
            dplyr::filter(grepl("(In|De)creased", status.x)) %>% 
            arrange(desc(abs(logFC))) %>% 
            dplyr::slice(1:5, .by = status.x),
          size = 3, show.legend = FALSE
        ) +
        annotate(
          "text", label = "logFC", angle = 90,
          x = min(rna[[i]]$logCPM) - 0.1 * diff(range(rna[[i]]$logCPM)),
          y = mean(range(rna[[i]]$logFC))
        ) +
        scale_xsidey_discrete(position = "right") +
        scale_y_continuous(position = "left") +
        scale_ysidex_discrete(labels = NULL) +
        scale_colour_manual(values = direction_colours) +
        scale_fill_manual(
          values = setNames(
            direction_colours, paste(target, names(direction_colours))
          )
        ) +
        coord_cartesian(xlim = range(rna[[i]]$logCPM), clip = 'off') +
        labs(colour = "DE Status", fill = paste(target, "Status"), y = NULL) +
        theme(
          axis.title.x = element_text(hjust = 0.4, vjust = 1),
          axis.title.y = element_text(hjust = 0.6, vjust = 1),
          ggside.panel.scale = 0.2, 
          ggside.axis.ticks.x = element_blank(),
          plot.margin = unit(c(1, 1, 1, 2), "lines")
          # legend.position = "inside",
          # legend.position.inside = c(0.8, 0.8),
          # legend.justification = c(1, 1)
        ) 
      
        ## Export the image
        img_out <- file.path(fig_path, glue("{i}_MA.{fig_type}"))
        fig_fun(filename = img_out, width = fw, height = fh)
        print(p)
        dev.off()
        
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "MA Plot for {i} with outer boxplots showing the distribution of ",
            "values for overall expression (logCPM) and logFC by {target} status"
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>% 
              basename() %>% 
              str_remove_all(glue(".{fig_type}$")) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level4",
            htmltools::h4("MA Plot"),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
    }
  ) %>% 
  setNames(names(rna))
```


```{r lfc-by-region}
pw_pl <- c(
  direction_colours[["Unchanged"]], RColorBrewer::brewer.pal(8, "Paired")
)
names(pw_pl) <- paste(
  target, 
  c(
    "Unchanged\nRNA Unchanged",
    "Unchanged\nRNA Decreased", "Decreased\nRNA Decreased", 
    "Decreased\nRNA Unchanged", "Increased\nRNA Unchanged", 
    "Unchanged\nRNA Increased", "Increased\nRNA Increased", 
    "Increased\nRNA Decreased", "Decreased\nRNA Increased"
  ) 
)
lfc_by_region <- names(rna) %>% 
  lapply(
    \(i) {
      x <- rna[[i]]
      p <- results_tbl %>% 
        inner_join(x, by = "gene_id") %>% 
        mutate(
          status.x = fct_relabel(status.x, \(x) paste(target, x)), 
          status.y = fct_relabel(status.y, \(x) paste("RNA", x)), 
          status = fct_cross(status.x, status.y, sep = "\n") %>% fct_rev()
        ) %>% 
        rename_with(
          \(x) str_replace_all(x, "\\.x$", paste0(" ", target))
        ) %>% 
        rename_with(
          \(x) str_replace_all(x, "\\.y$", " RNA")
        ) %>% 
        arrange(desc(`logFC RNA`), !!sym(glue("PValue {target}"))) %>% 
        distinct(gene_id, .keep_all = TRUE) %>%
        ggplot(aes(!!sym(glue("logFC {target}")), `logFC RNA`)) +
        geom_hline(yintercept = 0) +
        geom_vline(xintercept = 0) +
        geom_point(
          data = . %>% dplyr::filter(grepl("Unchanged.+Unchanged", status)),
          colour = direction_colours[["Unchanged"]], alpha = 0.5
        ) +
        geom_point(
          aes(colour = status),
          data = . %>% dplyr::filter(!grepl("Unchanged.+Unchanged", status)),
          alpha = 0.7
        ) +
        geom_smooth(se = FALSE, method = MASS::rlm, colour = "grey20") +
        geom_text(
          aes(label = lab),
          data = . %>% 
            dplyr::summarise(
              cor = cor(!!sym(glue("logFC {target}")), `logFC RNA`),
              "logFC {target}" := max(!!sym(glue("logFC {target}"))),
              `logFC RNA` = min(`logFC RNA`),
              lab = glue("rho == {round(cor, 3)}"),
              .by = region
            ) %>% 
            mutate(
              "logFC {target}" := max(!!sym(glue("logFC {target}"))),
              `logFC RNA` = min(`logFC RNA`)
            ),
          hjust = 1, vjust = 0, parse = TRUE
        ) +
        geom_label_repel(
          aes(label = gene_name, colour = status),
          data = . %>% 
            dplyr::filter(!grepl("Unchanged.+Unchanged", status)) %>% 
            dplyr::filter(
              (!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2 > 1
            ) %>% 
            dplyr::arrange(
              desc((!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2)
            ) %>% 
            dplyr::slice(1:2, .by = all_of(c("region", "status"))),
          show.legend = FALSE, size = 3, fontface = "italic"
        ) +
        facet_wrap(~region) +
        scale_colour_manual(values = pw_pl) +
        labs(
          x = glue("logFC ({target})", y = "logFC (RNA)"), 
          colour = "Combined\nStatus"
        ) +
        theme(legend.position = "bottom")
      
      ## Export the image
      img_out <- file.path(fig_path, glue("lfc_by_region_{i}.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "All detected genes within the {i} dataset were compared to {target}",
          "DSA results for {treat_levels[[2]]} Vs {treat_levels[[1]]}. ",
          "Where multiple signal regions wre mapped to a given gene, the most ",
          "highly-ranked by p-value was selected for visualisation. ",
          "Points are shown grouped by which genomic region the DSA result ", 
          "was mapped to. ",
          "Genes which were unchanged in {i} and with no changed DSA regions ",
          "are shown in grey. Robust linear regression lines are shown in ", 
          "black along with correlations in the bottom right corner of each panel."
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level4",
          htmltools::h4("Combined logFC"),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
      
    }
  ) %>% 
  setNames(names(rna))
```

```{r lfc-by-feature, echo = has_features}
lfc_by_feature <- list()
if (has_features) {
  lfc_by_feature <- names(rna) %>% 
    lapply(
      \(i) {
        x <- rna[[i]]
        p <- results_tbl %>% 
          inner_join(x, by = "gene_id") %>% 
          mutate(
            status.x = fct_relabel(status.x, \(x) paste(target, x)), 
            status.y = fct_relabel(status.y, \(x) paste("RNA", x)), 
            status = fct_cross(status.x, status.y, sep = "\n") %>% fct_rev(),
            gene_name = id2gene[gene_id]
          ) %>% 
          rename_with(
            \(x) str_replace_all(x, "\\.x$", paste0(" ", target))
          ) %>% 
          rename_with(
            \(x) str_replace_all(x, "\\.y$", " RNA")
          ) %>% 
          arrange(desc(`logFC RNA`), !!sym(glue("PValue {target}"))) %>% 
          distinct(gene_id, .keep_all = TRUE) %>% 
          ggplot(aes(!!sym(glue("logFC {target}")), `logFC RNA`)) +
          geom_hline(yintercept = 0) +
          geom_vline(xintercept = 0) +
          geom_point(
            data = . %>% dplyr::filter(grepl("Unchanged.+Unchanged", status)),
            colour = direction_colours[["Unchanged"]], alpha = 0.5
          ) +
          geom_point(
            aes(colour = status),
            data = . %>% dplyr::filter(!grepl("Unchanged.+Unchanged", status)),
            alpha = 0.7
          ) +
          geom_smooth(se = FALSE, method = MASS::rlm, colour = "grey20") +
          geom_text(
            aes(label = lab),
            data = . %>% 
              dplyr::summarise(
                cor = cor(!!sym(glue("logFC {target}")), `logFC RNA`),
                "logFC {target}" := max(!!sym(glue("logFC {target}"))),
                `logFC RNA` = min(`logFC RNA`),
                lab = glue("rho == {round(cor, 3)}"),
                .by = feature
              ) %>% 
              mutate(
                "logFC {target}" := max(!!sym(glue("logFC {target}"))),
                `logFC RNA` = min(`logFC RNA`)
              ),
            hjust = 1, vjust = 0, parse = TRUE
          ) +
          geom_label_repel(
            aes(label = gene_name, colour = status),
            data = . %>% 
              dplyr::filter(!grepl("Unchanged.+Unchanged", status)) %>% 
              dplyr::filter(
                (!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2 > 1
              ) %>% 
              dplyr::arrange(
                desc((!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2)
              ) %>% 
              dplyr::slice(1:2, .by = all_of(c("feature", "status"))),
            show.legend = FALSE, size = 3, fontface = "italic"
          ) +
          facet_wrap(~feature) +
          scale_colour_manual(values = pw_pl) +
          labs(
            x = glue("logFC ({target})", y = "logFC (RNA)"),
            colour = "Combined\nStatus"
          ) +
        theme(legend.position = "bottom")
        
        ## Export the image
        img_out <- file.path(fig_path, glue("lfc_by_feature_{i}.{fig_type}"))
        fig_fun(filename = img_out, width = fw, height = fh)
        print(p)
        dev.off()
        
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "All detected genes within the {i} dataset were compared to {target}",
            "DSA results for {treat_levels[[2]]} Vs {treat_levels[[1]]}. ",
            "Where multiple signal regions wre mapped to a given gene, the most ",
            "highly-ranked by p-value was selected for visualisation. ",
            "Points are shown by which feature (if any) the DSA result was ", 
            "mapped to. ",
            "Genes which were unchanged in {i} and with no changed DSA regions ",
            "are shown in grey. Robust linear regression lines are shown in ", 
            "black along with correlations in the bottom right corner of each panel."
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>% 
              basename() %>% 
              str_remove_all(glue(".{fig_type}$")) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level4",
            htmltools::h4("logFC By Feature"),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      }
    ) %>% 
    setNames(names(rna))
}
```

```{r volcano-plots, echo = any(has_de)}
volcano_plots <- list()
if (any(has_de)) {
  volcano_plots <- names(which(has_de)) %>% 
    lapply(
      \(i) {
        p <- rna[[i]] %>% 
          arrange(PValue) %>% 
          left_join(dsa_status, by = "gene_id") %>% 
          mutate(
            status.y = fct_na_value_to_level(status.y, "Undetected") %>% 
              fct_relabel(\(x) paste(target, x)),
            gene_name = id2gene[gene_id]
          ) %>% 
          droplevels() %>% 
          ggplot(aes(logFC, -log10(PValue), colour = status.x)) +
          geom_point(alpha = 0.7) +
          geom_label_repel(
            aes(label = gene_name),
            data = . %>% 
              dplyr::filter(grepl("creased", status.x)) %>% 
              distinct(gene_name, status.x, status.y, .keep_all = TRUE) %>% 
              dplyr::slice(1:2, .by = starts_with("status")),
            show.legend = FALSE, alpha = 0.7, size = 3, fontface = "italic"
          ) +
          geom_text(
            aes(label = label),
            data = . %>% 
              summarise(
                logFC = max(logFC), PValue = 1,
                n = dplyr::n(), .by = status.y
              ) %>% 
              mutate(
                logFC = max(logFC), label = paste("n =", comma(n))
              ),
            colour = "grey20", hjust = 1, vjust = 0, size = 3.5
          ) +
          facet_wrap(~status.y) +
          labs(
            x = "logFC (RNA)", y = expr(paste(-log[10], "PValue")), 
            colour = "DE Status"
          ) +
          scale_colour_manual(values = direction_colours)
        
        ## Export the image
        img_out <- file.path(fig_path, glue("volcano_{i}.{fig_type}"))
        fh <- ifelse(
          length(levels(p$data$status.y)) == 3, 2 * fh / 3, fh
        )
        fig_fun(filename = img_out, width = fw, height = fh)
        print(p)
        dev.off()
        
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "Volcano plot for {i} showing genes separated by whether any DSA ",
            "results corresponded to changed {target} signal. The number of ",
            "genes in each panel is given in the bottom right corner. ",
            "Where possible, the two most highly ranked DE genes are labelled."
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>% 
              basename() %>% 
              str_remove_all(glue(".{fig_type}$")) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level4",
            htmltools::h4("Volcano Plot"),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      }
    ) %>% 
    setNames(names(which(has_de)))
}
```
  
```{r pval-by-dsa}
pval_hist <- names(rna) %>% 
  lapply(
    \(i) {
      p <- rna[[i]] %>% 
        left_join(dsa_status, by = "gene_id") %>% 
        mutate(
          status.y = fct_na_value_to_level(status.y, "Undetected") %>% 
            fct_relabel(\(x) paste(target, x)),
        ) %>% 
        droplevels() %>% 
        ggplot(aes(PValue, after_stat(density))) +
        geom_histogram(
          fill = "grey70", colour = "black", binwidth = 0.01
        ) +
        facet_wrap(~status.y) +
        scale_y_continuous(expand = expansion(c(0, 0.05))) +
        scale_x_continuous(expand = expansion(0.025))
      y_max <- max(layer_scales(p)$y$range$range)
      p <- p +
        geom_text(
          aes(y = y, label = label),
          data = . %>% 
            summarise(
              n = dplyr::n(), .by = status.y
            ) %>% 
            mutate(
              PValue = 0.02, 
              y = y_max,
              label = paste("n =", comma(n))
            ),
          hjust = 0, vjust = 1, size = 3.5
        ) +
        labs(x = "PValue (RNA)", y = "Density")
      
      ## Export the image
      img_out <- file.path(fig_path, glue("pvals_{i}.{fig_type}"))
      fh <- ifelse(
        length(levels(p$data$status.y)) == 3, 2 * fh / 3, fh
      )
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "P-values for {i} separated by {target} status."
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level4",
          htmltools::h4("P-Value Histogram"),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(rna))
```

```{r dsa-deg-tables}
dsa_deg_tables <- list()
if (any(has_de) & has_ds) {
  dsa_deg_tables <- has_de %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(i) {
        fs <- 12
        id <- paste0("tbl-", i, "-dsa") %>% 
          str_to_lower() %>% 
          str_replace_all("_", "-")
        tbl <- rna[[i]] %>% 
          dplyr::filter(FDR < 0.05) %>% 
          left_join(as_tibble(gtf_gene)[c("range", "gene_id")]) %>% 
          inner_join(
            dplyr::filter(results_tbl, grepl("(In|De)creased", status)),
            by = "gene_id"
          ) %>% 
          mutate(d = distance(GRanges(range.x), GRanges(range.y))) %>% 
          dplyr::select(-range.x) %>% 
          dplyr::rename(range = range.y) %>% 
          dplyr::select(
            starts_with("gene"), ends_with(".x"), ends_with(".y"), 
            range, d, any_of(c("region", "feature"))
          ) %>% 
          dplyr::select(-starts_with("PValue")) %>% 
          reactable(
            showPageSizeOptions = TRUE, groupBy = "gene_name",
            filterable = TRUE, sortable = TRUE,
            columns = list(
              gene_name = colDef(
                name = "Gene", maxWidth = fs*11,
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)",
                  fontSize = fs
                )
              ),
              gene_id = colDef(
                name = "ID", aggregate = "unique", cell = function(value) "",
                show = FALSE
              ),
              logCPM.x = colDef(
                name = "logCPM", maxWidth = fs*6,
                aggregate = range_js,
                format = colFormat(digits = 2),
                style = function(value) bar_style(
                  width = value / max(.$logCPM.x), fontSize = fs
                ),
                filterMethod = js_greater
              ),
              logFC.x = colDef(
                "logFC", aggregate = "mean", 
                style = JS(
                  glue(
                    "function(rowInfo) {
                      var value = rowInfo.row['logFC.x']
                      if (value < 0) {
                        var color = '{{direction_colours[['Decreased']]}}'
                      } else if (value > 0) {
                        var color = '{{direction_colours[['Increased']]}}'
                      } else {
                        var color = '#000000'
                      }
                      return { color: color, fontSize: {{fs}} }
                    }",
                    .open = "{{", .close = "}}"
                  )
                ),
                format = colFormat(digits = 2), maxWidth = fs*6,
              ),
              FDR.x = colDef(
                name = "FDR", aggregate = "min", maxWidth = fs*6,
                format = colFormat(digits = 4),
                cell = \(value) sprint_pval(value),
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)", 
                  fontSize = fs
                ),
                filterMethod = js_less
              ),
              status.x = colDef("Status", aggregate = "unique", show = FALSE),
              range = colDef(
                name = "Signal Location", minWidth = 160,
                aggregate = granges_js
              ),
              region = colDef(
                name = "Genomic Region", aggregate = "unique"
              ),
              feature = colDef(
                name = "Feature", aggregate = "unique"
              ),
              logCPM.y = colDef(
                name = "logCPM", maxWidth = fs*6,
                aggregate = range_js,
                format = colFormat(digits = 2),
                style = function(value) bar_style(
                  width = value / max(results$logCPM), fontSize = fs
                ),
                filterMethod = js_greater
              ),
              logFC.y = colDef(
                name = "logFC", aggregate = range_js, maxWidth = fs*5,
                format = colFormat(digits = 2),
                cell = function(value) sprintf("%.2f", value),
                style =  JS(
                  glue(
                    "function(rowInfo) {
                      var value = rowInfo.row['logFC.y']
                      if (value < 0) {
                        var color = '{{direction_colours[['Decreased']]}}'
                      } else if (value > 0) {
                        var color = '{{direction_colours[['Increased']]}}'
                      } else {
                        var color = '#000000'
                      }
                      return { color: color, fontSize: {{fs}} }
                    }",
                    .open = "{{", .close = "}}"
                  )
                )
              ),
              FDR.y = colDef(
                name = "FDR",
                aggregate = "max",
                maxWidth = 10 + fs*5,
                format = colFormat(prefix = "< ", digits = 4),
                cell = \(value) sprint_pval(value),
                filterMethod = js_less,
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)",
                  fontSize = fs
                )
              ),
              status.y = colDef("Status", show = FALSE, aggregate = "unique"),
              d = colDef(
                "Distance To Peak", maxWidth = fs * 8, format = comma_col,
                aggregate = range_js
              )
            )[colnames(.)],
            columnGroups = list(
              colGroup("RNA", str_subset(names(.), "\\.x$")),
              colGroup(target, str_subset(names(.), "\\.y$"))
            ),
            theme = reactableTheme(style = list(fontSize = fs)),
            elementId = id
          )
        cp <- htmltools::em(
          glue(
            "Comparison of all DE genes with any associated DSA result also ",
            "considered as showing evidence of changed {target} signal. "
          )
        )
        tl <- tagList(
          tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue(
              "Reactable.downloadDataCSV('{id}', '{i}_deg_{target}_dsa.csv')"
            )
          ),
          tags$button(
            "Expand/Collapse All",
            onclick = glue("Reactable.toggleAllRowsExpanded('{id}')")
          ),
          tags$button(
            "Clear All Filters",
            onclick = glue("Reactable.setAllFilters('{id}', [])")
          ),
          tbl,
          cp
        )
        htmltools::div(
          class = "section level4", 
          htmltools::h4("DEGs with Diferential Signal"), 
          tl
        )
      }
    ) %>% 
    setNames(names(which(has_de)))
}
```


```{r print-tables-and-figures, echo=FALSE}
tagList(
  names(rna) %>% 
    lapply(
      \(x) {
        tagList(
          list(
            htmltools::div(
              id = x, class = "section level3 tabset",
              htmltools::h3(x, class = "tabset"),
              summary_tables[[x]],
              donut_plots[[x]],
              ma_plots[[x]],
              lfc_by_region[[x]],
              lfc_by_feature[[x]],
              volcano_plots[[x]],
              pval_hist[[x]],
              dsa_deg_tables[[x]]
            )
          )
        )
      }
    )
)
```


### `r target` Pattern GSEA 

#### GSEA By Direction of Regulation {.tabset}

GSEA analysis [@SubramanianGsea2005] was first performed by taking the genes mapped to various `r `target` signal regions, and checking for enrichment amongst the RNA-Seq results, ranking genes from up- to down-regulated, by significance of the logFC estimates.
The `R` package `fgsea` [@KorotkevichFgsea2019] was used all downstream GSEA analysis.

```{r run-gsea-dir}
ids_by_status <- results_tbl %>% 
  unite(group, any_of(c("status", "region", "feature"))) %>% 
  split(.$group) %>% 
  lapply(pull, "gene_id")
rna_dir_scores <- rna %>% 
  lapply(
    \(x) setNames(-sign(x$logFC) * log10(x$PValue), x$gene_id)
  )
rna_gsea_dir <- rna_dir_scores %>% 
  lapply(\(x) fgsea(ids_by_status, x, minSize = min(msigdb_params$size))) %>% 
  lapply(as_tibble) %>% 
  lapply(
    separate, pathway, 
    into = c("status", "region", "feature")[seq_len(2 + has_features)],
    sep = "_", extra = "merge", remove = FALSE
  ) %>%
  lapply(arrange, pval) %>% 
  lapply(
    mutate,
    padj = p.adjust(pval, enrich_params$adj),
    status = factor(status, levels = names(direction_colours)),
    region = factor(region, unname(regions)),
    leSize = map_int(leadingEdge, length),
    leadingEdge = map_chr(
      leadingEdge, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
    )
  ) %>% 
  lapply(
    dplyr::select, 
    pathway, any_of(c("status", "region", "feature")),
    padj, NES, ends_with("ize"), leadingEdge
  )
any_gsea_dir <- map_lgl(rna_gsea_dir, \(x) any(x$padj < enrich_params$alpha))
```

```{r tbl-gsea-dir, echo=FALSE}
if (any(any_gsea_dir)) {
  tagList(
    any_gsea_dir %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(x) {
          tbl <- rna_gsea_dir[[x]] %>% 
            dplyr::select(-pathway) %>% 
            dplyr::filter(padj < enrich_params$alpha) %>% 
            reactable(
              sortable = TRUE, filterable = TRUE, showPageSizeOptions = TRUE,
              columns = list(
                status = colDef(glue("{target} Status"), maxWidth = 100),
                region = colDef("Region Type", maxWidth = 120),
                feature = colDef("Feature Type", maxWidth = 100),
                padj = colDef(
                  "p<sub>adj</sub>", cell = \(value) sprint_pval(value),
                  filterMethod = js_less, html = TRUE, maxWidth = 80
                ),
                NES = colDef(
                  "NES", format = colFormat(digits = 2),
                  style = \(value) {
                    text_colour <- ifelse(
                      value > 0, 
                      direction_colours[["Increased"]], 
                      direction_colours[["Decreased"]]
                    )
                    list(color = text_colour)
                  }
                ),
                size = colDef("Group Size"),
                leSize = colDef("LE Size"),
                leadingEdge = colDef(
                  "Leading Edge", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150, maxWidth = 500
                )
              )[colnames(.)],
              defaultColDef = colDef(maxWidth = 80)
            )
          cp <- htmltools::em(
            glue(
              "GSEA results for {x}, testing an association between up- or down-", 
              "regulated genes for all {target} DSA groups incorporating all available ", 
              "annotations and features. A Normalized Enrichment Score (NES) > 0 ", 
              "indicates an association with up-regulation, with a negative score ",
              "being associated with down-regulation. The total number of genes ", 
              "mapped to each type of DSA group is provided along with those found ",
              "beyond the Leading Edge. Given that multiple peaks may map to a gene, ",
              "genes may be included in more than one grouping."
            )
          )
          htmltools::div(
            htmltools::div(
              id = paste0("tbl-gsea-dir", x) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "table", tbl, htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```


### Combined Enrichment Using Direction of Changed Expression

```{r load-gsea-dir}
gsea_dir <- file.path(all_paths$annotation, "gsea_dir.rds") %>% read_rds()
```


```{r integrate-gsea-dir}
changed_dir_enrich <- gsea_dir %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(changed_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  )
increased_dir_enrich <- gsea_dir %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(increased_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  )
decreased_dir_enrich <- gsea_dir %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(decreased_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  )
n_changed_dir_enrich <- map_int(changed_dir_enrich, nrow)
n_increased_dir_enrich <- map_int(increased_dir_enrich, nrow)
n_decreased_dir_enrich <- map_int(decreased_dir_enrich, nrow)
```

#### Result Tables {.tabset}

##### All Changed Sites {.tabset}

```{r tbl-all-changed-dir}
if (any(n_changed_dir_enrich > 0)) {
  tagList(
    n_changed_dir_enrich %>% 
      as.logical() %>% 
      setNames(names(gsea_dir)) %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(x) {
          tbl <- changed_dir_enrich[[x]] %>% 
            dplyr::select(-meta_p, cmn_ids) %>% 
            reactable(
              showPageSizeOptions = TRUE, sortable = TRUE, filterable = TRUE,
              columns = list(
                pathway = colDef(
                  "Gene Set", minWidth = 180,
                  cell = function(value) htmltools::tags$a(
                    href = gs_url[[value]],
                    target = "_blank",
                    str_replace_all(value, "_", " ")
                  ),
                  html = TRUE
                ),
                NES = colDef("NES", format = colFormat(digits = 2), maxWidth = 100),
                adj_p = colDef(
                  "P<sub>adj</sub>", cell = \(value) sprint_pval(value),
                  html = TRUE, maxWidth = 100, filterMethod = js_less
                ),
                n_cmn = colDef("Nbr Shared", maxWidth = 100, filterMethod = js_greater),
                genes = colDef(
                  "Shared Genes", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150
                )
              )
            )
          cp <- htmltools::em(
            glue(
              "Enrichment results for changed {target} sites were integrated ",
              "with GSEA results for {x}, which included the direction of ",
              "estimated logFC for gene expression. Genes within each gene set ",
              "which had significant DSA sites mapped to them, as well as being ",
              "included in the Leading Edge for GSEA were dentified as Shared ",
              "Genes. Whilst DSA sites were taken as those showing any evidence ", 
              "of changed signal, a positive Normalised Enrichment Score (NES) ",
              "indicates an association with up-regulated genes, whilst a ", 
              "negative score indicates an association with down-regulated genes."
            )
          )
          htmltools::div(
            htmltools::div(
              id = "tbl-changed-dir-enrich-" %>% 
                paste0(x) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level6",
              htmltools::h6(x),
              htmltools::div(
                class = "table", tbl, htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

##### Sites With Increased Signal {.tabset}

```{r tbl-increased-dir}
if (any(n_increased_dir_enrich > 0)) {
  tagList(
    n_increased_dir_enrich %>% 
      as.logical() %>% 
      setNames(names(gsea_dir)) %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(x) {
          tbl <- increased_dir_enrich[[x]] %>% 
            dplyr::select(-meta_p, cmn_ids) %>% 
            reactable(
              showPageSizeOptions = TRUE, sortable = TRUE, filterable = TRUE,
              columns = list(
                pathway = colDef(
                  "Gene Set", minWidth = 180,
                  cell = function(value) htmltools::tags$a(
                    href = gs_url[[value]],
                    target = "_blank",
                    str_replace_all(value, "_", " ")
                  ),
                  html = TRUE
                ),
                NES = colDef("NES", format = colFormat(digits = 2), maxWidth = 100),
                adj_p = colDef(
                  "P<sub>adj</sub>", cell = \(value) sprint_pval(value),
                  html = TRUE, maxWidth = 100, filterMethod = js_less
                ),
                n_cmn = colDef("Nbr Shared", maxWidth = 100, filterMethod = js_greater),
                genes = colDef(
                  "Shared Genes", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150
                )
              )
            )
          cp <- htmltools::em(
            glue(
              "Enrichment results for increased {target} sites were integrated ",
              "with GSEA results for {x}, which included the direction of ",
              "estimated logFC for gene expression. Genes within each gene set ",
              "which had significant DSA sites mapped to them, as well as being ",
              "included in the Leading Edge for GSEA were dentified as Shared ",
              "Genes. Whilst DSA sites were taken as those showing any evidence ", 
              "of increased signal, a positive Normalised Enrichment Score (NES) ",
              "indicates an association with up-regulated genes, whilst a ", 
              "negative score indicates an association with down-regulated genes."
            )
          )
          htmltools::div(
            htmltools::div(
              id = "tbl-increased-dir-enrich-" %>% 
                paste0(x) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level6",
              htmltools::h6(x),
              htmltools::div(
                class = "table", tbl, htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

##### Sites With Decreased Signal {.tabset}

```{r tbl-decreased-dir}
if (any(n_decreased_dir_enrich > 0)) {
  tagList(
    n_decreased_dir_enrich %>% 
      as.logical() %>% 
      setNames(names(gsea_dir)) %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(x) {
          tbl <- decreased_dir_enrich[[x]] %>% 
            dplyr::select(-meta_p, cmn_ids) %>% 
            reactable(
              showPageSizeOptions = TRUE, sortable = TRUE, filterable = TRUE,
              columns = list(
                pathway = colDef(
                  "Gene Set", minWidth = 180,
                  cell = function(value) htmltools::tags$a(
                    href = gs_url[[value]],
                    target = "_blank",
                    str_replace_all(value, "_", " ")
                  ),
                  html = TRUE
                ),
                NES = colDef("NES", format = colFormat(digits = 2), maxWidth = 100),
                adj_p = colDef(
                  "P<sub>adj</sub>", cell = \(value) sprint_pval(value),
                  html = TRUE, maxWidth = 100, filterMethod = js_less
                ),
                n_cmn = colDef("Nbr Shared", maxWidth = 100, filterMethod = js_greater),
                genes = colDef(
                  "Shared Genes", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150
                )
              )
            )
          cp <- htmltools::em(
            glue(
              "Enrichment results for decreased {target} sites were integrated ",
              "with GSEA results for {x}, which included the direction of ",
              "estimated logFC for gene expression. Genes within each gene set ",
              "which had significant DSA sites mapped to them, as well as being ",
              "included in the Leading Edge for GSEA were dentified as Shared ",
              "Genes. Whilst DSA sites were taken as those showing any evidence ", 
              "of decreased signal, a positive Normalised Enrichment Score (NES) ",
              "indicates an association with up-regulated genes, whilst a ", 
              "negative score indicates an association with down-regulated genes."
            )
          )
          htmltools::div(
            htmltools::div(
              id = "tbl-decreased-dir-enrich-" %>% 
                paste0(x) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level6",
              htmltools::h6(x),
              htmltools::div(
                class = "table", tbl, htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

### Network Plots {.tabset}

#### All Changed Sites {.tabset}

```{r network-changed-dir}
if (any(n_changed_dir_enrich >= min(network_params$size))) {
  changed_dir_network <- n_changed_dir_enrich %>% 
    is_weakly_greater_than(min(network_params$size)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        changed_dir_enrich[[x]] %>% 
          mutate(
            gene_set_size = msigdb %>% 
              dplyr::filter(gs_name %in% pathway) %>% 
              split(.$gs_name) %>% 
              .[pathway] %>% 
              map_int(nrow),
            prop = n_cmn / gene_set_size,
            cmn_ids = setNames(cmn_ids, pathway)
          ) %>% 
          make_tbl_graph(
            gs = .$cmn_ids,
            alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
            hit_col = "n_cmn", size_col = "gene_set_size",
            min_dist = network_params$max_distance,
            max_gs = max(network_params$size)
          ) %>% 
          ggraph(layout = net_layout, weights = oc) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(
              fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop
            ),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label, alpha = prop),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% 
                  str_trunc(60) %>% 
                  str_wrap(width = 18)
              ),
            repel = TRUE, bg.color = "white", bg.r = 0.1, 
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.2) +
          scale_alpha(range = c(0.5, 1), trans = "sqrt") +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
          guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
          labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
          theme_void() 
      }
    ) %>% 
    setNames(
      names(which(n_changed_dir_enrich >= min(network_params$size)))
    )
}
```

```{r print-network-changed-dir, echo=FALSE}
if (any(n_changed_dir_enrich >= min(network_params$size))) {
  tagList(
    names(changed_dir_network) %>% 
      lapply(
        \(x) {
          ## Export the image
          img_out <- file.path(fig_path, glue("changed_dir_network_{x}.{fig_type}"))
          fig_fun(filename = img_out, width = fw, height = fh)
          print(changed_dir_network[[x]])
          dev.off()
          ## Create html tags
          fig_link <- str_extract(img_out, "assets.+")
          cp <- htmltools::tags$em(
            glue(
              "Combined enrichment network for {x} and DSA for {target} using ",
              "all changed {target} sites."
            )
          )
          htmltools::div(
            htmltools::div(
              id = img_out %>% 
                basename() %>% 
                str_remove_all(glue(".{fig_type}$")) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "figure", style = "text-align: center",
                htmltools::img(src = fig_link, width = "100%"),
                htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

#### All Increased Sites {.tabset}

```{r network-increased-dir}
if (any(n_increased_dir_enrich >= min(network_params$size))) {
  increased_dir_network <- n_increased_dir_enrich %>% 
    is_weakly_greater_than(min(network_params$size)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        increased_dir_enrich[[x]] %>% 
          mutate(
            gene_set_size = msigdb %>% 
              dplyr::filter(gs_name %in% pathway) %>% 
              split(.$gs_name) %>% 
              .[pathway] %>% 
              map_int(nrow),
            prop = n_cmn / gene_set_size,
            cmn_ids = setNames(cmn_ids, pathway)
          ) %>% 
          make_tbl_graph(
            gs = .$cmn_ids,
            alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
            hit_col = "n_cmn", size_col = "gene_set_size",
            min_dist = network_params$max_distance,
            max_gs = max(network_params$size)
          ) %>% 
          ggraph(layout = net_layout, weights = oc) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(
              fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop
            ),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label, alpha = prop),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% 
                  str_trunc(60) %>% 
                  str_wrap(width = 18)
              ),
            repel = TRUE, bg.color = "white", bg.r = 0.1, 
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.2) +
          scale_alpha(range = c(0.5, 1), trans = "sqrt") +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
          guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
          labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
          theme_void() 
      }
    ) %>% 
    setNames(
      names(which(n_increased_dir_enrich >= min(network_params$size)))
    )
}
```

```{r print-network-increased-dir, echo=FALSE}
if (any(n_increased_dir_enrich >= min(network_params$size))) {
  tagList(
    names(increased_dir_network) %>% 
      lapply(
        \(x) {
          ## Export the image
          img_out <- file.path(fig_path, glue("increased_dir_network_{x}.{fig_type}"))
          fig_fun(filename = img_out, width = fw, height = fh)
          print(increased_dir_network[[x]])
          dev.off()
          ## Create html tags
          fig_link <- str_extract(img_out, "assets.+")
          cp <- htmltools::tags$em(
            glue(
              "Combined enrichment network for {x} and DSA for {target} using ",
              "all increased {target} sites."
            )
          )
          htmltools::div(
            htmltools::div(
              id = img_out %>% 
                basename() %>% 
                str_remove_all(glue(".{fig_type}$")) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "figure", style = "text-align: center",
                htmltools::img(src = fig_link, width = "100%"),
                htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

#### All Decreased Sites {.tabset}

```{r network-decreased-dir}
if (any(n_decreased_dir_enrich >= min(network_params$size))) {
  decreased_dir_network <- n_decreased_dir_enrich %>% 
    is_weakly_greater_than(min(network_params$size)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        decreased_dir_enrich[[x]] %>% 
          mutate(
            gene_set_size = msigdb %>% 
              dplyr::filter(gs_name %in% pathway) %>% 
              split(.$gs_name) %>% 
              .[pathway] %>% 
              map_int(nrow),
            prop = n_cmn / gene_set_size,
            cmn_ids = setNames(cmn_ids, pathway)
          ) %>% 
          make_tbl_graph(
            gs = .$cmn_ids,
            alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
            hit_col = "n_cmn", size_col = "gene_set_size",
            min_dist = network_params$max_distance,
            max_gs = max(network_params$size)
          ) %>% 
          ggraph(layout = net_layout, weights = oc) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(
              fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop
            ),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label, alpha = prop),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% 
                  str_trunc(60) %>% 
                  str_wrap(width = 18)
              ),
            repel = TRUE, bg.color = "white", bg.r = 0.1
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.2) +
          scale_alpha(range = c(0.5, 1), trans = "sqrt") +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
          guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
          labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
          theme_void() 
      }
    ) %>% 
    setNames(
      names(which(n_decreased_dir_enrich >= min(network_params$size)))
    )
}
```

```{r print-network-decreased-dir, echo=FALSE}
if (any(n_decreased_dir_enrich >= min(network_params$size))) {
  tagList(
    names(decreased_dir_network) %>% 
      lapply(
        \(x) {
          ## Export the image
          img_out <- file.path(fig_path, glue("decreased_dir_network_{x}.{fig_type}"))
          fig_fun(filename = img_out, width = fw, height = fh)
          print(decreased_dir_network[[x]])
          dev.off()
          ## Create html tags
          fig_link <- str_extract(img_out, "assets.+")
          cp <- htmltools::tags$em(
            glue(
              "Combined enrichment network for {x} and DSA for {target} using ",
              "all decreased {target} sites."
            )
          )
          htmltools::div(
            htmltools::div(
              id = img_out %>% 
                basename() %>% 
                str_remove_all(glue(".{fig_type}$")) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "figure", style = "text-align: center",
                htmltools::img(src = fig_link, width = "100%"),
                htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```


### Combined Enrichment Using Only Signficance of Changed Expression

```{r load-gsea-sig}
gsea_sig <- file.path(all_paths$annotation, "gsea_sig.rds") %>% read_rds()
```


```{r integrate-gsea-sig}
changed_sig_enrich <- gsea_sig %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(changed_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  )
increased_sig_enrich <- gsea_sig %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(increased_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  )
decreased_sig_enrich <- gsea_sig %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(decreased_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  )
n_changed_sig_enrich <- map_int(changed_sig_enrich, nrow)
n_increased_sig_enrich <- map_int(increased_sig_enrich, nrow)
n_decreased_sig_enrich <- map_int(decreased_sig_enrich, nrow)
```

### Result Tables {.tabset}

#### All Changed Sites {.tabset}

```{r tbl-all-changed-sig}
if (any(n_changed_sig_enrich > 0)) {
  tagList(
    n_changed_sig_enrich %>% 
      as.logical() %>% 
      setNames(names(gsea_sig)) %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(x) {
          tbl <- changed_sig_enrich[[x]] %>% 
            dplyr::select(-meta_p, cmn_ids) %>% 
            reactable(
              showPageSizeOptions = TRUE, sortable = TRUE, filterable = TRUE,
              columns = list(
                pathway = colDef(
                  "Gene Set", minWidth = 180,
                  cell = function(value) htmltools::tags$a(
                    href = gs_url[[value]],
                    target = "_blank",
                    str_replace_all(value, "_", " ")
                  ),
                  html = TRUE
                ),
                NES = colDef("NES", format = colFormat(digits = 2), maxWidth = 100),
                adj_p = colDef(
                  "P<sub>adj</sub>", cell = \(value) sprint_pval(value),
                  html = TRUE, maxWidth = 100, filterMethod = js_less
                ),
                n_cmn = colDef("Nbr Shared", maxWidth = 100, filterMethod = js_greater),
                genes = colDef(
                  "Shared Genes", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150
                )
              )
            )
          cp <- htmltools::em(
            glue(
              "Enrichment results for changed {target} sites were integrated ",
              "with GSEA results for {x}, which included the significance of ",
              "estimated logFC for gene expression. Genes within each gene set ",
              "which had significant DSA sites mapped to them, as well as being ",
              "included in the Leading Edge for GSEA were dentified as Shared ",
              "Genes. Whilst DSA sites were taken as those showing any evidence ", 
              "of changed signal, a positive Normalised Enrichment Score (NES) ",
              "indicates an association with up-regulated genes, whilst a ", 
              "negative score indicates an association with down-regulated genes."
            )
          )
          htmltools::div(
            htmltools::div(
              id = "tbl-changed-sig-enrich-" %>% 
                paste0(x) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "table", tbl, htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

#### Sites With Increased Signal {.tabset}

```{r tbl-increased-sig}
if (any(n_increased_sig_enrich > 0)) {
  n_increased_sig_enrich %>% 
    as.logical() %>% 
    setNames(names(gsea_sig)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        tbl <- increased_sig_enrich[[x]] %>% 
          dplyr::select(-meta_p, cmn_ids) %>% 
          reactable(
            showPageSizeOptions = TRUE, sortable = TRUE, filterable = TRUE,
            columns = list(
              pathway = colDef(
                "Gene Set", minWidth = 180,
                cell = function(value) htmltools::tags$a(
                  href = gs_url[[value]],
                  target = "_blank",
                  str_replace_all(value, "_", " ")
                ),
                html = TRUE
              ),
              NES = colDef("NES", format = colFormat(digits = 2), maxWidth = 100),
              adj_p = colDef(
                "P<sub>adj</sub>", cell = \(value) sprint_pval(value),
                html = TRUE, maxWidth = 100, filterMethod = js_less
              ),
              n_cmn = colDef("Nbr Shared", maxWidth = 100, filterMethod = js_greater),
              genes = colDef(
                "Shared Genes", cell = \(value) with_tooltip(value, width = 60),
                minWidth = 150
              )
            )
          )
        cp <- htmltools::em(
          glue(
            "Enrichment results for increased {target} sites were integrated ",
            "with GSEA results for {x}, which included only the significance of ",
            "estimated logFC for gene expression. Genes within each gene set ",
            "which had significant DSA sites mapped to them, as well as being ",
            "included in the Leading Edge for GSEA were dentified as Shared ",
            "Genes. Whilst DSA sites were taken as those showing any evidence ", 
            "of increased signal, a positive Normalised Enrichment Score (NES) ",
            "indicates an association with up-regulated genes, whilst a ", 
            "negative score indicates an association with down-regulated genes."
          )
        )
        htmltools::div(
          htmltools::div(
            id = "tbl-increased-sig-enrich-" %>% 
              paste0(x) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level5",
            htmltools::h5(x),
            htmltools::div(
              class = "table", tbl, htmltools::tags$caption(cp)
            )
          )
        )
      }
    )
}
```

#### Sites With Decreased Signal {.tabset}

```{r tbl-decreased-sig}
if (any(n_decreased_sig_enrich > 0)) {
  n_decreased_sig_enrich %>% 
    as.logical() %>% 
    setNames(names(gsea_sig)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        tbl <- decreased_sig_enrich[[x]] %>% 
          dplyr::select(-meta_p, cmn_ids) %>% 
          reactable(
            showPageSizeOptions = TRUE, sortable = TRUE, filterable = TRUE,
            columns = list(
              pathway = colDef(
                "Gene Set", minWidth = 180,
                cell = function(value) htmltools::tags$a(
                  href = gs_url[[value]],
                  target = "_blank",
                  str_replace_all(value, "_", " ")
                ),
                html = TRUE
              ),
              NES = colDef("NES", format = colFormat(digits = 2), maxWidth = 100),
              adj_p = colDef(
                "P<sub>adj</sub>", cell = \(value) sprint_pval(value),
                html = TRUE, maxWidth = 100, filterMethod = js_less
              ),
              n_cmn = colDef("Nbr Shared", maxWidth = 100, filterMethod = js_greater),
              genes = colDef(
                "Shared Genes", cell = \(value) with_tooltip(value, width = 60),
                minWidth = 150
              )
            )
          )
        cp <- htmltools::em(
          glue(
            "Enrichment results for decreased {target} sites were integrated ",
            "with GSEA results for {x}, which included only the significance of ",
            "estimated logFC for gene expression. Genes within each gene set ",
            "which had significant DSA sites mapped to them, as well as being ",
            "included in the Leading Edge for GSEA were dentified as Shared ",
            "Genes. Whilst DSA sites were taken as those showing any evidence ", 
            "of decreased signal, a positive Normalised Enrichment Score (NES) ",
            "indicates an association with up-regulated genes, whilst a ", 
            "negative score indicates an association with down-regulated genes."
          )
        )
        htmltools::div(
          htmltools::div(
            id = "tbl-decreased-sig-enrich-" %>% 
              paste0(x) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level5",
            htmltools::h5(x),
            htmltools::div(
              class = "table", tbl, htmltools::tags$caption(cp)
            )
          )
        )
      }
    )
}
```

### Network Plots {.tabset}

#### All Changed Sites {.tabset}

```{r network-changed-sig}
if (any(n_changed_sig_enrich >= min(network_params$size))) {
  changed_sig_network <- n_changed_sig_enrich %>% 
    is_weakly_greater_than(min(network_params$size)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        changed_sig_enrich[[x]] %>% 
          mutate(
            gene_set_size = msigdb %>% 
              dplyr::filter(gs_name %in% pathway) %>% 
              split(.$gs_name) %>% 
              .[pathway] %>% 
              map_int(nrow),
            prop = n_cmn / gene_set_size,
            cmn_ids = setNames(cmn_ids, pathway)
          ) %>% 
          make_tbl_graph(
            gs = .$cmn_ids,
            alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
            hit_col = "n_cmn", size_col = "gene_set_size",
            min_dist = network_params$max_distance,
            max_gs = max(network_params$size)
          ) %>% 
          ggraph(layout = net_layout, weights = oc) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(
              fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop
            ),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label, alpha = prop),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% 
                  str_trunc(60) %>% 
                  str_wrap(width = 18)
              ),
            repel = TRUE, bg.color = "white", bg.r = 0.1, 
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.2) +
          scale_alpha(range = c(0.5, 1), trans = "sqrt") +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
          guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
          labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
          theme_void() 
      }
    ) %>% 
    setNames(
      names(which(n_changed_sig_enrich >= min(network_params$size)))
    )
}
```

```{r print-network-changed-sig, echo=FALSE}
if (any(n_changed_sig_enrich >= min(network_params$size))) {
  tagList(
    names(changed_sig_network) %>% 
      lapply(
        \(x) {
          ## Export the image
          img_out <- file.path(fig_path, glue("changed_sig_network_{x}.{fig_type}"))
          fig_fun(filename = img_out, width = fw, height = fh)
          print(changed_sig_network[[x]])
          dev.off()
          ## Create html tags
          fig_link <- str_extract(img_out, "assets.+")
          cp <- htmltools::tags$em(
            glue(
              "Combined enrichment network for {x} and DSA for {target} using ",
              "all changed {target} sites."
            )
          )
          htmltools::div(
            htmltools::div(
              id = img_out %>% 
                basename() %>% 
                str_remove_all(glue(".{fig_type}$")) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "figure", style = "text-align: center",
                htmltools::img(src = fig_link, width = "100%"),
                htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

#### All Increased Sites {.tabset}

```{r network-increased-sig}
if (any(n_increased_sig_enrich >= min(network_params$size))) {
  increased_sig_network <- n_increased_sig_enrich %>% 
    is_weakly_greater_than(min(network_params$size)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        increased_sig_enrich[[x]] %>% 
          mutate(
            gene_set_size = msigdb %>% 
              dplyr::filter(gs_name %in% pathway) %>% 
              split(.$gs_name) %>% 
              .[pathway] %>% 
              map_int(nrow),
            prop = n_cmn / gene_set_size,
            cmn_ids = setNames(cmn_ids, pathway)
          ) %>% 
          make_tbl_graph(
            gs = .$cmn_ids,
            alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
            hit_col = "n_cmn", size_col = "gene_set_size",
            min_dist = network_params$max_distance,
            max_gs = max(network_params$size)
          ) %>% 
          ggraph(layout = net_layout, weights = oc) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(
              fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop
            ),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label, alpha = prop),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% 
                  str_trunc(60) %>% 
                  str_wrap(width = 18)
              ),
            repel = TRUE, bg.color = "white", bg.r = 0.1, 
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.2) +
          scale_alpha(range = c(0.5, 1), trans = "sqrt") +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
          guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
          labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
          theme_void() 
      }
    ) %>% 
    setNames(
      names(which(n_increased_sig_enrich >= min(network_params$size)))
    )
}
```

```{r print-network-increased-sig, echo=FALSE}
if (any(n_increased_sig_enrich >= min(network_params$size))) {
  tagList(
    names(increased_sig_network) %>% 
      lapply(
        \(x) {
          ## Export the image
          img_out <- file.path(fig_path, glue("increased_sig_network_{x}.{fig_type}"))
          fig_fun(filename = img_out, width = fw, height = fh)
          print(increased_sig_network[[x]])
          dev.off()
          ## Create html tags
          fig_link <- str_extract(img_out, "assets.+")
          cp <- htmltools::tags$em(
            glue(
              "Combined enrichment network for {x} and DSA for {target} using ",
              "all increased {target} sites."
            )
          )
          htmltools::div(
            htmltools::div(
              id = img_out %>% 
                basename() %>% 
                str_remove_all(glue(".{fig_type}$")) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "figure", style = "text-align: center",
                htmltools::img(src = fig_link, width = "100%"),
                htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

#### All Decreased Sites {.tabset}

```{r network-decreased-sig}
if (any(n_decreased_sig_enrich >= min(network_params$size))) {
  decreased_sig_network <- n_decreased_sig_enrich %>% 
    is_weakly_greater_than(min(network_params$size)) %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(x) {
        decreased_sig_enrich[[x]] %>% 
          mutate(
            gene_set_size = msigdb %>% 
              dplyr::filter(gs_name %in% pathway) %>% 
              split(.$gs_name) %>% 
              .[pathway] %>% 
              map_int(nrow),
            prop = n_cmn / gene_set_size,
            cmn_ids = setNames(cmn_ids, pathway)
          ) %>% 
          make_tbl_graph(
            gs = .$cmn_ids,
            alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
            hit_col = "n_cmn", size_col = "gene_set_size",
            min_dist = network_params$max_distance,
            max_gs = max(network_params$size)
          ) %>% 
          ggraph(layout = net_layout, weights = oc) +
          geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
          geom_node_point(
            aes(
              fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop
            ),
            shape = 21
          ) +
          geom_node_text(
            aes(label = label, alpha = prop),
            colour = "black", size = 3, 
            data = . %>%
              mutate(
                label = str_replace_all(label, "_", " ") %>% 
                  str_trunc(60) %>% 
                  str_wrap(width = 18)
              ),
            repel = TRUE, bg.color = "white", bg.r = 0.1
          ) +
          scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
          scale_fill_viridis_c(option = "inferno", begin = 0.2) +
          scale_alpha(range = c(0.5, 1), trans = "sqrt") +
          scale_size_continuous(range = c(1, 10)) +
          scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
          scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
          guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
          labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
          theme_void() 
      }
    ) %>% 
    setNames(
      names(which(n_decreased_sig_enrich >= min(network_params$size)))
    )
}
```

```{r print-network-decreased-sig, echo=FALSE}
if (any(n_decreased_sig_enrich >= min(network_params$size))) {
  tagList(
    names(decreased_sig_network) %>% 
      lapply(
        \(x) {
          ## Export the image
          img_out <- file.path(fig_path, glue("decreased_sig_network_{x}.{fig_type}"))
          fig_fun(filename = img_out, width = fw, height = fh)
          print(decreased_sig_network[[x]])
          dev.off()
          ## Create html tags
          fig_link <- str_extract(img_out, "assets.+")
          cp <- htmltools::tags$em(
            glue(
              "Combined enrichment network for {x} and DSA for {target} using ",
              "all decreased {target} sites."
            )
          )
          htmltools::div(
            htmltools::div(
              id = img_out %>% 
                basename() %>% 
                str_remove_all(glue(".{fig_type}$")) %>% 
                str_to_lower() %>% 
                str_replace_all("_", "-"),
              class = "section level5",
              htmltools::h5(x),
              htmltools::div(
                class = "figure", style = "text-align: center",
                htmltools::img(src = fig_link, width = "100%"),
                htmltools::tags$caption(cp)
              )
            )
          )
        }
      )
  )
}
```

### Highly Ranked Genes {.tabset}


```{r rna-grl-to-plot}
grl_to_plot <- rna %>% 
  lapply(inner_join, results_tbl, by = 'gene_id') %>% 
  bind_rows(.id = "dataset") %>% 
  dplyr::filter(grepl("(In|De)creased", status.y)) %>% 
  mutate(
    rank = rank(PValue.x) + rank(PValue.y)
  ) %>% 
  arrange(rank) %>% 
  dplyr::filter(
    gene_id %in% unique(gene_id)[1:5]
  ) %>% 
  split(.$gene_id) %>% 
  lapply(colToRanges, "range", seqinfo = seqinfo(results)) %>% 
  GRangesList()
y_lim <- bwfl[[target]] %>%
  lapply(import.bw, which = unlist(range(grl_to_plot))) %>%
  lapply(\(x) c(0, max(x$score))) %>% 
  unlist() %>% 
  range() %>% 
  list() %>% 
  setNames(target)
```

By choosing the 5 most highly ranked genes for differential expression in `r pander(names(rna))`, and which also had regions of significantly changed `r target` signal, a set of example plots are shown below.

```{r plot-all-top-rna-ranges}
.makeCaption <- function(.gr) {
  if (is.null(.gr)) return(NULL)
  id <- unique(.gr$gene_id)
  gn <- id2gene[id]
  cp <- glue(
    "The {width(range(.gr))}bp region with all changed peaks mapped to {gn}, ", 
    "as a highly-ranked DE gene in ",
    glue_collapse(.gr$dataset, sep = ", ", last = " & "),
    ", and with at least significantly changed DSA region for {target}. ",
    "Regions associated with differential signal are annotated by colour."
  )
}
htmltools::tagList(
  lapply(
    seq_along(grl_to_plot),
    \(i) {
      ## Export the image
      nm <- str_to_lower(names(grl_to_plot)[[i]])
      id <- unique(grl_to_plot[[i]]$gene_id)
      img_out <- file.path(fig_path, glue("{nm}.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      plotHFGC(
        grl_to_plot[[i]],
        hic = hic,
        features = feat_gr, featcol = feat_col_hfgc, 
        featsize = 1 + has_features, featstack = "dense",
        genes = hfgc_genes, genecol = gene_col,
        coverage = bwfl, linecol = line_col, ylim = y_lim,
        annotation = annot, annotcol = direction_colours,
        cytobands = bands_df, 
        collapseTranscripts = collapseTrans, maxTrans = 20,
        col.title = "black", background.title = "white", showAxis = FALSE,
        max = max(5e5, width(range(grl_to_plot[[i]])) * 1.1), 
        rotation.title = 90, zoom = 10, highlight = FALSE
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(.makeCaption(grl_to_plot[[i]]))
      htmltools::div(
        htmltools::div(
          id = nm,
          class = "section level3",
          htmltools::h3(id2gene[id]),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  )
)
```

