## Comparison With Transcriptomic Data

```{r define-rna-objects}
library(fgsea)
library(metap)
library(ggside)
results_tbl <- results %>% 
  as_tibble() %>% 
  dplyr::select(
    range, logCPM, logFC, PValue, FDR = !!sym(fdr_column), status,
    gene_id, any_of(c("region", "feature"))
  ) %>% 
  unnest(gene_id) %>% 
  mutate(gene_name = id2gene[gene_id])
rna <- lapply(rna, addDiffStatus)
has_de <- map_lgl(rna, \(x) any(x$FDR < 0.05))
dsa_status <- results_tbl %>% 
  dplyr::select(status, gene_id) %>% 
  unnest(gene_id) %>% 
  summarise(
    status = case_when(
      any(status == "Increased") & any(status == "Decreased") ~ "Ambiguous",
      any(status == "Increased") ~ "Increased",
      any(status == "Decreased") ~ "Decreased",
      any(status == "Ambiguous") ~ "Ambiguous",
      TRUE ~ "Unchanged"
    ) %>% 
      factor(levels = names(direction_colours)),
    .by = gene_id
  ) %>% 
  droplevels()
enrich_cols$pathway <- enrich_cols$gs_name
enrich_cols$NES <- colDef(
  format = colFormat(digits = 2), maxWidth = 100,
  style = \(value) {
    text_colour <- ifelse(
      value > 0, 
      direction_colours[["Increased"]], 
      direction_colours[["Decreased"]]
    )
    list(color = text_colour)
  }
)
enrich_cols$meta_p <- enrich_cols$p
enrich_cols$n_cmn <- enrich_cols$observed_gene_hits
enrich_cols$n_cmn$name <- "Nbr Common"
enrich_cols$genes <- enrich_cols$genes_with_hits
enrich_cols$genes$name <- "Common Genes"
enrich_cols$cmn_ids <- enrich_cols$gene_id
```

In order to compare DSA results for `r target`, gene-to-range mappings were summarised such that if a gene was mapped to any DSA results showing increased or decreased binding, they were given this status.
Genes with conflicting mappings were considered to be ambiguous, whilst those without any changed regions were assigned as being consistently unchanged.
These results were applied where summarised results were able to be presented, however, when comparing actual estimates for logFC etc, all DSA ranges were retained.


```{r summary-tables}
summary_tables <- names(rna) %>% 
  lapply(
    \(i) {
      x <- rna[[i]]
      id <- paste0("tbl-summary-", i) %>% 
        str_to_lower() %>% 
        str_replace_all("_", "-")
      tbl <- results_tbl %>% 
        full_join(x, by = "gene_id") %>% 
        mutate(
          status.x = fct_na_value_to_level(status.x, "Undetected"),
          status.y = fct_na_value_to_level(status.y, "Undetected")
        ) %>% 
        arrange(PValue.x) %>% 
        distinct(gene_id, .keep_all = TRUE) %>% 
        dplyr::select(gene_id, starts_with("status")) %>% 
        summarise(
          n = dplyr::n(), .by = starts_with("status")
        ) %>%
        pivot_wider(names_from = "status.y", values_from = "n", values_fill = 0) %>% 
        dplyr::select(
          starts_with("status"), any_of(c("Increased", "Decreased")), 
          everything()
        ) %>% 
        reactable(
          showPageSizeOptions = FALSE,
          columns = list(
            status.x = colDef(glue("{target} Status"), footer = "")
          ),
          defaultColDef = colDef(
            sortable = TRUE, format = comma_col,
            footer = function(values) comma(sum(values)),
            footerStyle = list(fontWeight = "bold")
          ),
          columnGroups = list(
            colGroup(
              "Gene Expression", str_subset(names(.), "ed$")
            )
          )
        )
      cp <- htmltools::em(
        glue(
          "Comparison between {target} status and gene expression in {i}."
        )
      )
      htmltools::div(
        htmltools::div(
          id = id,
          class = "section level5",
          htmltools::h5("Summary Table"),
          htmltools::div(
            class = "table", tbl, htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(rna))
```

```{r donut-plots}
donuts <- rna %>% 
  lapply(left_join, dsa_status, by = "gene_id") %>% 
  lapply(
    mutate, 
    status.x = fct_na_value_to_level(status.x, "Undetected"),
    status.y = fct_na_value_to_level(status.y, "Undetected"),
  ) %>% 
  lapply(
    dplyr::rename, RNA = status.x, "{target}" := status.y
  ) %>% 
  mclapply(
    plotSplitDonut,
    inner = !!sym(target), outer = "RNA",
    inner_palette = direction_colours, outer_palette = direction_colours,
    inner_glue = "{inner}\n{.data[[inner]]}\n{comma(n, 1)}\n{percent(p,0.1)}",
    inner_label_alpha = 0.8, inner_min_p = 0.04,
    outer_glue = "{.data[[outer]]} {comma(n, 1)}",
    outer_min_p = 0.001, outer_label_colour = "palette", 
    outer_rotate = TRUE,
    outer_label = "text",
    mc.cores = threads
  ) 
donut_plots <- names(donuts) %>% 
  lapply(
    \(x) {
      ## Export the image
      img_out <- file.path(fig_path, glue("donut_{x}.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      print(donuts[[x]])
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "All detected genes within the {x} dataset were compared to {target} ",
          "DSA results for {treat_levels[[2]]} Vs {treat_levels[[1]]}. Regions ",
          "with signal contained within the DSA results were previously mapped ",
          "to genes, and these mappings were used. Results for each gene were ",
          "summarised to account for multiple regions with different DSA results."
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level5",
          htmltools::h5("Donut Plot"),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(donuts))
```

```{r ma-plots}
ma_plots <- names(rna) %>% 
  lapply(
    \(i) {
      p <- rna[[i]] %>% 
        left_join(dsa_status, by = "gene_id") %>% 
        mutate(
          status.y = fct_na_value_to_level(status.y, "Undetected") %>% 
            fct_relabel(\(x) paste(target, x)),
          gene_name = id2gene[gene_id]
        ) %>% 
        droplevels() %>% 
        arrange(1 / PValue) %>% 
        ggplot(aes(logCPM, logFC)) +
        geom_point(aes(colour = status.x), alpha = 0.7) +
        geom_xsideboxplot(
          aes(logCPM, fct_rev(status.y), fill = status.y), orientation = "y",
          show.legend = FALSE
        ) +
        geom_ysideboxplot(
          aes(fct_rev(status.y), logFC, fill = status.y), orientation = "x"
        ) +
        geom_label_repel(
          aes(label = gene_name, colour = status.x),
          data = . %>% 
            dplyr::filter(grepl("(In|De)creased", status.x)) %>% 
            arrange(desc(abs(logFC))) %>% 
            dplyr::slice(1:5, .by = status.x),
          size = 3, show.legend = FALSE
        ) +
        annotate(
          "text", label = "logFC", angle = 90,
          x = min(rna[[i]]$logCPM) - 0.1 * diff(range(rna[[i]]$logCPM)),
          y = mean(range(rna[[i]]$logFC))
        ) +
        scale_xsidey_discrete(position = "right") +
        scale_y_continuous(position = "left") +
        scale_ysidex_discrete(labels = NULL) +
        scale_colour_manual(values = direction_colours) +
        scale_fill_manual(
          values = setNames(
            direction_colours, paste(target, names(direction_colours))
          )
        ) +
        coord_cartesian(xlim = range(rna[[i]]$logCPM), clip = 'off') +
        labs(colour = "DE Status", fill = paste(target, "Status"), y = NULL) +
        theme(
          axis.title.x = element_text(hjust = 0.4, vjust = 1),
          axis.title.y = element_text(hjust = 0.6, vjust = 1),
          ggside.panel.scale = 0.2, 
          ggside.axis.ticks.x = element_blank(),
          plot.margin = unit(c(1, 1, 1, 2), "lines")
        ) 
      
        ## Export the image
        img_out <- file.path(fig_path, glue("{i}_MA.{fig_type}"))
        fig_fun(filename = img_out, width = fw, height = fh)
        print(p)
        dev.off()
        
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "MA Plot for {i} with outer boxplots showing the distribution of ",
            "values for overall expression (logCPM) and logFC by {target} status"
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>% 
              basename() %>% 
              str_remove_all(glue(".{fig_type}$")) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level5",
            htmltools::h5("MA Plot"),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
    }
  ) %>% 
  setNames(names(rna))
```


```{r lfc-by-region}
pw_pl <- c(
  direction_colours[["Unchanged"]], RColorBrewer::brewer.pal(8, "Paired")
)
names(pw_pl) <- paste(
  target, 
  c(
    "Unchanged\nRNA Unchanged",
    "Unchanged\nRNA Decreased", "Decreased\nRNA Decreased", 
    "Decreased\nRNA Unchanged", "Increased\nRNA Unchanged", 
    "Unchanged\nRNA Increased", "Increased\nRNA Increased", 
    "Increased\nRNA Decreased", "Decreased\nRNA Increased"
  ) 
)
lfc_by_region <- names(rna) %>% 
  lapply(
    \(i) {
      x <- rna[[i]]
      p <- results_tbl %>% 
        inner_join(x, by = "gene_id") %>% 
        mutate(
          status.x = fct_relabel(status.x, \(x) paste(target, x)), 
          status.y = fct_relabel(status.y, \(x) paste("RNA", x)), 
          status = fct_cross(status.x, status.y, sep = "\n") %>% fct_rev()
        ) %>% 
        rename_with(
          \(x) str_replace_all(x, "\\.x$", paste0(" ", target))
        ) %>% 
        rename_with(
          \(x) str_replace_all(x, "\\.y$", " RNA")
        ) %>% 
        arrange(desc(`logFC RNA`), !!sym(glue("PValue {target}"))) %>% 
        distinct(gene_id, .keep_all = TRUE) %>%
        ggplot(aes(!!sym(glue("logFC {target}")), `logFC RNA`)) +
        geom_hline(yintercept = 0) +
        geom_vline(xintercept = 0) +
        geom_point(
          data = . %>% dplyr::filter(grepl("Unchanged.+Unchanged", status)),
          colour = direction_colours[["Unchanged"]], alpha = 0.5
        ) +
        geom_point(
          aes(colour = status),
          data = . %>% dplyr::filter(!grepl("Unchanged.+Unchanged", status)),
          alpha = 0.7
        ) +
        geom_smooth(se = FALSE, method = MASS::rlm, colour = "grey20") +
        geom_text(
          aes(label = lab),
          data = . %>% 
            dplyr::summarise(
              cor = cor(!!sym(glue("logFC {target}")), `logFC RNA`),
              "logFC {target}" := max(!!sym(glue("logFC {target}"))),
              `logFC RNA` = min(`logFC RNA`),
              lab = glue("rho == {round(cor, 3)}"),
              .by = region
            ) %>% 
            mutate(
              "logFC {target}" := max(!!sym(glue("logFC {target}"))),
              `logFC RNA` = min(`logFC RNA`)
            ),
          hjust = 1, vjust = 0, parse = TRUE
        ) +
        geom_label_repel(
          aes(label = gene_name, colour = status),
          data = . %>% 
            dplyr::filter(!grepl("Unchanged.+Unchanged", status)) %>% 
            dplyr::filter(
              (!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2 > 1
            ) %>% 
            dplyr::arrange(
              desc((!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2)
            ) %>% 
            dplyr::slice(1:2, .by = all_of(c("region", "status"))),
          show.legend = FALSE, size = 3, fontface = "italic"
        ) +
        facet_wrap(~region) +
        scale_colour_manual(values = pw_pl) +
        labs(
          x = glue("logFC ({target})", y = "logFC (RNA)"), 
          colour = "Combined\nStatus"
        ) +
        theme(legend.position = "bottom")
      
      ## Export the image
      img_out <- file.path(fig_path, glue("lfc_by_region_{i}.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "All detected genes within the {i} dataset were compared to {target}",
          "DSA results for {treat_levels[[2]]} Vs {treat_levels[[1]]}. ",
          "Where multiple signal regions wre mapped to a given gene, the most ",
          "highly-ranked by p-value was selected for visualisation. ",
          "Points are shown grouped by which genomic region the DSA result ", 
          "was mapped to. ",
          "Genes which were unchanged in {i} and with no changed DSA regions ",
          "are shown in grey. Robust linear regression lines are shown in ", 
          "black along with correlations in the bottom right corner of each panel."
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level5",
          htmltools::h5("Combined logFC"),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
      
    }
  ) %>% 
  setNames(names(rna))
```

```{r lfc-by-feature, echo = has_features}
lfc_by_feature <- list()
if (has_features) {
  lfc_by_feature <- names(rna) %>% 
    lapply(
      \(i) {
        x <- rna[[i]]
        p <- results_tbl %>% 
          unnest(feature, keep_empty = TRUE) %>% 
          inner_join(x, by = "gene_id") %>% 
          mutate(
            feature = feature %>% 
              str_replace_na("no_feature") %>% 
              fct_infreq(),
            status.x = fct_relabel(status.x, \(x) paste(target, x)), 
            status.y = fct_relabel(status.y, \(x) paste("RNA", x)), 
            status = fct_cross(status.x, status.y, sep = "\n") %>% fct_rev(),
            gene_name = id2gene[gene_id]
          ) %>% 
          distinct(range, feature, status, .keep_all = TRUE) %>% 
          rename_with(
            \(x) str_replace_all(x, "\\.x$", paste0(" ", target))
          ) %>% 
          rename_with(
            \(x) str_replace_all(x, "\\.y$", " RNA")
          ) %>% 
          arrange(
            desc(`logFC RNA`), !!sym(glue("PValue {target}")), desc(feature)
          ) %>% 
          distinct(gene_id, feature, .keep_all = TRUE) %>% 
          mutate(feature = fct_lump_n(feature, 8)) %>% 
          ggplot(aes(!!sym(glue("logFC {target}")), `logFC RNA`)) +
          geom_hline(yintercept = 0) +
          geom_vline(xintercept = 0) +
          geom_point(
            data = . %>% dplyr::filter(grepl("Unchanged.+Unchanged", status)),
            colour = direction_colours[["Unchanged"]], alpha = 0.5
          ) +
          geom_point(
            aes(colour = status),
            data = . %>% dplyr::filter(!grepl("Unchanged.+Unchanged", status)),
            alpha = 0.7
          ) +
          geom_smooth(se = FALSE, method = MASS::rlm, colour = "grey20") +
          geom_text(
            aes(label = lab),
            data = . %>% 
              dplyr::summarise(
                cor = cor(!!sym(glue("logFC {target}")), `logFC RNA`),
                "logFC {target}" := max(!!sym(glue("logFC {target}"))),
                `logFC RNA` = min(`logFC RNA`),
                lab = glue("rho == {round(cor, 3)}"),
                .by = feature
              ) %>% 
              mutate(
                "logFC {target}" := max(!!sym(glue("logFC {target}"))),
                `logFC RNA` = min(`logFC RNA`)
              ),
            hjust = 1, vjust = 0, parse = TRUE
          ) +
          geom_label_repel(
            aes(label = gene_name, colour = status),
            data = . %>% 
              dplyr::filter(!grepl("Unchanged.+Unchanged", status)) %>% 
              dplyr::filter(
                (!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2 > 1
              ) %>% 
              dplyr::arrange(
                desc((!!sym(glue("logFC {target}")))^2 + `logFC RNA`^2)
              ) %>% 
              dplyr::slice(1:2, .by = all_of(c("feature", "status"))),
            show.legend = FALSE, size = 3, fontface = "italic"
          ) +
          facet_wrap(~feature) +
          scale_colour_manual(values = pw_pl) +
          labs(
            x = glue("logFC ({target})", y = "logFC (RNA)"),
            colour = "Combined\nStatus"
          ) +
        theme(legend.position = "bottom")
        
        ## Export the image
        img_out <- file.path(fig_path, glue("lfc_by_feature_{i}.{fig_type}"))
        fig_fun(filename = img_out, width = fw, height = fh)
        print(p)
        dev.off()
        
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "All detected genes within the {i} dataset were compared to DSA ",
            "results for {target} when comparing {treat_levels[[2]]} Vs {treat_levels[[1]]}. ",
            "Where multiple signal regions wre mapped to a given gene, the most ",
            "highly-ranked by p-value was selected for visualisation within each feature. ",
            "Points are shown by which feature (if any) the DSA result was ", 
            "mapped to, merging features with low mappings into the 'Other' category. ",
            "Genes which were unchanged in {i} and with no changed DSA regions ",
            "are shown in grey. Robust linear regression lines are shown in ", 
            "black along with correlations in the bottom right corner of each panel."
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>% 
              basename() %>% 
              str_remove_all(glue(".{fig_type}$")) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level5",
            htmltools::h5("logFC By Feature"),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      }
    ) %>% 
    setNames(names(rna))
}
```

```{r volcano-plots, echo = any(has_de)}
volcano_plots <- list()
if (any(has_de)) {
  volcano_plots <- names(which(has_de)) %>% 
    lapply(
      \(i) {
        p <- rna[[i]] %>% 
          arrange(PValue) %>% 
          left_join(dsa_status, by = "gene_id") %>% 
          mutate(
            status.y = fct_na_value_to_level(status.y, "Undetected") %>% 
              fct_relabel(\(x) paste(target, x)),
            gene_name = id2gene[gene_id]
          ) %>% 
          droplevels() %>% 
          ggplot(aes(logFC, -log10(PValue), colour = status.x)) +
          geom_point(alpha = 0.7) +
          geom_label_repel(
            aes(label = gene_name),
            data = . %>% 
              dplyr::filter(grepl("creased", status.x)) %>% 
              distinct(gene_name, status.x, status.y, .keep_all = TRUE) %>% 
              dplyr::slice(1:2, .by = starts_with("status")),
            show.legend = FALSE, alpha = 0.7, size = 3, fontface = "italic"
          ) +
          geom_text(
            aes(label = label),
            data = . %>% 
              summarise(
                logFC = max(logFC), PValue = 1,
                n = dplyr::n(), .by = status.y
              ) %>% 
              mutate(
                logFC = max(logFC), label = paste("n =", comma(n))
              ),
            colour = "grey20", hjust = 1, vjust = 0, size = 3.5
          ) +
          facet_wrap(~status.y) +
          labs(
            x = "logFC (RNA)", y = expr(paste(-log[10], "PValue")), 
            colour = "DE Status"
          ) +
          scale_colour_manual(values = direction_colours)
        
        ## Export the image
        img_out <- file.path(fig_path, glue("volcano_{i}.{fig_type}"))
        fh <- ifelse(
          length(levels(p$data$status.y)) == 3, 2 * fh / 3, fh
        )
        fig_fun(filename = img_out, width = fw, height = fh)
        print(p)
        dev.off()
        
        ## Create html tags
        fig_link <- str_extract(img_out, "assets.+")
        cp <- htmltools::tags$em(
          glue(
            "Volcano plot for {i} showing genes separated by whether any DSA ",
            "results corresponded to changed {target} signal. The number of ",
            "genes in each panel is given in the bottom right corner. ",
            "Where possible, the two most highly ranked DE genes are labelled."
          )
        )
        htmltools::div(
          htmltools::div(
            id = img_out %>% 
              basename() %>% 
              str_remove_all(glue(".{fig_type}$")) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-"),
            class = "section level5",
            htmltools::h5("Volcano Plot"),
            htmltools::div(
              class = "figure", style = "text-align: center",
              htmltools::img(src = fig_link, width = "100%"),
              htmltools::tags$caption(cp)
            )
          )
        )
      }
    ) %>% 
    setNames(names(which(has_de)))
}
```
  
```{r pval-by-dsa}
pval_hist <- names(rna) %>% 
  lapply(
    \(i) {
      p <- rna[[i]] %>% 
        left_join(dsa_status, by = "gene_id") %>% 
        mutate(
          status.y = fct_na_value_to_level(status.y, "Undetected") %>% 
            fct_relabel(\(x) paste(target, x)),
        ) %>% 
        droplevels() %>% 
        ggplot(aes(PValue, after_stat(density))) +
        geom_histogram(
          fill = "grey70", colour = "black", binwidth = 0.01
        ) +
        facet_wrap(~status.y) +
        scale_y_continuous(expand = expansion(c(0, 0.05))) +
        scale_x_continuous(expand = expansion(0.025))
      y_max <- max(layer_scales(p)$y$range$range)
      p <- p +
        geom_text(
          aes(y = y, label = label),
          data = . %>% 
            summarise(
              n = dplyr::n(), .by = status.y
            ) %>% 
            mutate(
              PValue = 0.02, 
              y = y_max,
              label = paste("n =", comma(n))
            ),
          hjust = 0, vjust = 1, size = 3.5
        ) +
        labs(x = "PValue (RNA)", y = "Density")
      
      ## Export the image
      img_out <- file.path(fig_path, glue("pvals_{i}.{fig_type}"))
      fh <- ifelse(
        length(levels(p$data$status.y)) == 3, 2 * fh / 3, fh
      )
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "P-values for {i} separated by {target} status."
        )
      )
      htmltools::div(
        htmltools::div(
          id = img_out %>% 
            basename() %>% 
            str_remove_all(glue(".{fig_type}$")) %>% 
            str_to_lower() %>% 
            str_replace_all("_", "-"),
          class = "section level5",
          htmltools::h5("P-Value Histogram"),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  ) %>% 
  setNames(names(rna))
```

```{r dsa-deg-tables}
dsa_deg_tables <- list()
if (any(has_de) & has_ds) {
  dsa_deg_tables <- has_de %>% 
    which() %>% 
    names() %>% 
    lapply(
      \(i) {
        fs <- 12
        id <- paste0("tbl-", i, "-dsa") %>% 
          str_to_lower() %>% 
          str_replace_all("_", "-")
        tbl <- rna[[i]] %>% 
          dplyr::filter(FDR < 0.05) %>% 
          left_join(as_tibble(gtf_gene)[c("range", "gene_id")]) %>% 
          inner_join(
            dplyr::filter(results_tbl, grepl("(In|De)creased", status)),
            by = "gene_id"
          ) %>% 
          mutate(
            d = distance(GRanges(range.x), GRanges(range.y)),
            across(
              any_of("feature"), 
              \(x) map_chr(x, \(x) paste(sort(unique(x)), collapse = "; "))
            )
          ) %>% 
          dplyr::select(-range.x) %>% 
          dplyr::rename(range = range.y) %>% 
          dplyr::select(
            starts_with("gene"), ends_with(".x"), ends_with(".y"), 
            range, d, any_of(c("region", "feature"))
          ) %>% 
          dplyr::select(-starts_with("PValue")) %>% 
          reactable(
            showPageSizeOptions = nrow(.) > 10, groupBy = "gene_name",
            filterable = TRUE, sortable = TRUE,
            columns = list(
              gene_name = colDef(
                name = "Gene", maxWidth = fs*11,
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)",
                  fontSize = fs
                )
              ),
              gene_id = colDef(
                name = "ID", aggregate = "unique", cell = function(value) "",
                show = FALSE
              ),
              logCPM.x = colDef(
                name = "logCPM", maxWidth = fs*6,
                aggregate = range_js,
                format = colFormat(digits = 2),
                style = function(value) bar_style(
                  width = value / max(.$logCPM.x), fontSize = fs
                ),
                filterMethod = js_greater
              ),
              logFC.x = colDef(
                "logFC", aggregate = "mean", 
                style = JS(
                  glue(
                    "function(rowInfo) {
                      var value = rowInfo.row['logFC.x']
                      if (value < 0) {
                        var color = '{{direction_colours[['Decreased']]}}'
                      } else if (value > 0) {
                        var color = '{{direction_colours[['Increased']]}}'
                      } else {
                        var color = '#000000'
                      }
                      return { color: color, fontSize: {{fs}} }
                    }",
                    .open = "{{", .close = "}}"
                  )
                ),
                format = colFormat(digits = 2), maxWidth = fs*6,
              ),
              FDR.x = colDef(
                name = "FDR", aggregate = "min", maxWidth = fs*6,
                format = colFormat(digits = 4),
                cell = \(value) sprint_pval(value),
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)", 
                  fontSize = fs
                ),
                filterMethod = js_less
              ),
              status.x = colDef("Status", aggregate = "unique", show = FALSE),
              range = colDef(
                name = "Signal Location", minWidth = 160,
                aggregate = granges_js
              ),
              region = colDef(
                name = "Genomic Region", aggregate = "unique",
                filterInput = function(values, name) {
                  tags$select(
                    # Set to undefined to clear the filter
                    onchange = sprintf(
                      "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
                      id, name
                    ),
                    # "All" has an empty value to clear the filter, and is the default option
                    tags$option(value = "", "All"),
                    lapply(unique(values), tags$option),
                    "aria-label" = sprintf("Filter %s", name),
                    style = "width: 100%; height: 28px; font-size: 12px;"
                  )
                }
              ),
              feature = colDef(name = "Feature", aggregate = "unique"),
              logCPM.y = colDef(
                name = "logCPM", maxWidth = fs*6,
                aggregate = range_js,
                format = colFormat(digits = 2),
                style = function(value) bar_style(
                  width = value / max(results$logCPM), fontSize = fs
                ),
                filterMethod = js_greater
              ),
              logFC.y = colDef(
                name = "logFC", aggregate = range_js, maxWidth = fs*5,
                format = colFormat(digits = 2),
                cell = function(value) sprintf("%.2f", value),
                style =  JS(
                  glue(
                    "function(rowInfo) {
                      var value = rowInfo.row['logFC.y']
                      if (value < 0) {
                        var color = '{{direction_colours[['Decreased']]}}'
                      } else if (value > 0) {
                        var color = '{{direction_colours[['Increased']]}}'
                      } else {
                        var color = '#000000'
                      }
                      return { color: color, fontSize: {{fs}} }
                    }",
                    .open = "{{", .close = "}}"
                  )
                )
              ),
              FDR.y = colDef(
                name = "FDR",
                aggregate = "max",
                maxWidth = 10 + fs*5,
                format = colFormat(prefix = "< ", digits = 4),
                cell = \(value) sprint_pval(value),
                filterMethod = js_less,
                style = list(
                  borderRight = "1px solid rgba(0, 0, 0, 0.1)",
                  fontSize = fs
                )
              ),
              status.y = colDef("Status", show = FALSE, aggregate = "unique"),
              d = colDef(
                "Distance To Peak", maxWidth = fs * 8, format = comma_col,
                aggregate = range_js, filterMethod = js_less
              )
            )[colnames(.)],
            columnGroups = list(
              colGroup("RNA", str_subset(names(.), "\\.x$")),
              colGroup(target, str_subset(names(.), "\\.y$"))
            ),
            theme = reactableTheme(style = list(fontSize = fs)),
            elementId = id
          )
        cp <- htmltools::em(
          glue(
            "Comparison of all DE genes with any associated DSA result also ",
            "considered as showing evidence of changed {target} signal. "
          )
        )
        htmltools::div(
          id = id, class = "section level5", 
          htmltools::h5("DEGs with Differential Signal"), 
          tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue(
              "Reactable.downloadDataCSV('{id}', '{i}_deg_{target}_dsa.csv')"
            )
          ),
          tags$button(
            "Expand/Collapse All",
            onclick = glue("Reactable.toggleAllRowsExpanded('{id}')")
          ),
          tags$button(
            "Clear All Filters",
            onclick = glue("Reactable.setAllFilters('{id}', [])")
          ),
          tbl, tags$caption(cp)
        )
      }
    ) %>% 
    setNames(names(which(has_de)))
}
```

### Direct Comparisons


```{r print-tables-and-figures, echo=FALSE}
tagList(
  names(rna) %>% 
    lapply(
      \(x) {
        tagList(
          list(
            htmltools::div(
              id = x, class = "section level4 tabset",
              htmltools::h4(x, class = "tabset"),
              summary_tables[[x]],
              donut_plots[[x]],
              ma_plots[[x]],
              lfc_by_region[[x]],
              lfc_by_feature[[x]],
              volcano_plots[[x]],
              pval_hist[[x]],
              dsa_deg_tables[[x]]
            )
          )
        )
      }
    )
)
```


### `r target` Pattern GSEA 

GSEA analysis [@SubramanianGsea2005] was first performed by taking the genes mapped to various `r target` signal regions, and checking for enrichment amongst the RNA-Seq results.
Two analyses wre performed:

1. Ranking both `r target` signal and transcriptomic data by significance only, to test for any overall association with changed signal levels and DGE, then
2. Ranking both `r target` signal and transcriptomic data by both *direction of change* and *significance*, to test for any overall association between the direction of change in either -omics layer

The `R` package `fgsea` [@KorotkevichFgsea2019] was used for all downstream GSEA analysis.

### Overall Significance {.tabset}

```{r run-gsea-sig}
ids_by_status <- results_tbl %>%
  mutate(status = ifelse(status == "Unchanged", "Unchanged", "Changed")) %>% 
  unnest(any_of("feature")) %>% 
  unite(group, any_of(c("status", "region", "feature"))) %>% 
  split(.$group) %>% 
  lapply(pull, "gene_id")
rna_sig_scores <- rna %>% 
  lapply(\(x) setNames(-log10(x$PValue), x$gene_id))
rna_gsea_sig <- rna_sig_scores %>% 
  lapply(
    \(x) fgsea(
      ids_by_status, x, minSize = min(msigdb_params$size), scoreType = "pos"
    )
  ) %>% 
  lapply(as_tibble) %>% 
  lapply(
    separate, pathway, 
    into = c("status", "region", "feature")[seq_len(2 + has_features)],
    sep = "_", extra = "merge", remove = FALSE
  ) %>%
  lapply(arrange, pval) %>% 
  lapply(
    mutate,
    pathway = paste(target, pathway),
    padj = p.adjust(pval, enrich_params$adj),
    status = factor(status, levels = c("Unchanged", "Changed")),
    region = factor(region, unname(regions)),
    leSize = map_int(leadingEdge, length),
    leadingEdge = map_chr(
      leadingEdge, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
    )
  ) %>% 
  lapply(
    dplyr::select, 
    pathway, any_of(c("status", "region", "feature")),
    padj, NES, ends_with("ize"), leadingEdge
  ) %>% 
  lapply(dplyr::filter, leSize >= enrich_params$min_sig)
any_gsea_sig <- map_lgl(rna_gsea_sig, \(x) any(x$padj < enrich_params$alpha))
```


```{r tbl-gsea-sig, echo=FALSE}
if (any(any_gsea_sig)) {
  tagList(
    any_gsea_sig %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(x) {
          tbl_id <- paste0("tbl-gsea-sig", x) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-")
          tbl <- rna_gsea_sig[[x]] %>% 
            dplyr::select(-pathway) %>% 
            dplyr::filter(padj < enrich_params$alpha) %>% 
            reactable(
              sortable = TRUE, filterable = TRUE, showPageSizeOptions = TRUE,
              columns = list(
                status = colDef(
                  glue("{target} Status"), maxWidth = 100,
                  filterMethod = JS(
                    "function(rows, columnId, filterValue) {
                      return rows.filter(function(row) {
                        return row.values[columnId].indexOf(filterValue) !== -1
                      })
                    }"
                  ),
                  filterInput = function(values, name) {
                    tags$select(
                      # Set to undefined to clear the filter
                      onchange = sprintf(
                        "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
                        tbl_id, name
                      ),
                      # "All" has an empty value to clear the filter, and is the default option
                      tags$option(value = "", "All"),
                      lapply(unique(values), tags$option),
                      "aria-label" = sprintf("Filter %s", name),
                      style = "width: 100%; height: 28px; font-size: 12px;"
                    )
                  }
                ),
                region = colDef(
                  "Region Type", maxWidth = 120,
                  filterMethod = JS(
                    "function(rows, columnId, filterValue) {
                      return rows.filter(function(row) {
                        return row.values[columnId].indexOf(filterValue) !== -1
                      })
                    }"
                  ),
                  filterInput = function(values, name) {
                    tags$select(
                      # Set to undefined to clear the filter
                      onchange = sprintf(
                        "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
                        tbl_id, name
                      ),
                      # "All" has an empty value to clear the filter, and is the default option
                      tags$option(value = "", "All"),
                      lapply(unique(values), tags$option),
                      "aria-label" = sprintf("Filter %s", name),
                      style = "width: 100%; height: 28px; font-size: 12px;"
                    )
                  }
                ),
                feature = colDef(
                  "Feature Type", maxWidth = 100,
                  filterMethod = JS(
                    "function(rows, columnId, filterValue) {
                      return rows.filter(function(row) {
                        return row.values[columnId].indexOf(filterValue) !== -1
                      })
                    }"
                  ),
                  filterInput = function(values, name) {
                    tags$select(
                      # Set to undefined to clear the filter
                      onchange = sprintf(
                        "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
                        tbl_id, name
                      ),
                      # "All" has an empty value to clear the filter, and is the default option
                      tags$option(value = "", "All"),
                      lapply(unique(values), tags$option),
                      "aria-label" = sprintf("Filter %s", name),
                      style = "width: 100%; height: 28px; font-size: 12px;"
                    )
                  }
                ),
                padj = enrich_cols$adj_p,
                NES = enrich_cols$NES,
                size = colDef("Group Size"),
                leSize = colDef("LE Size"),
                leadingEdge = colDef(
                  "Leading Edge", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150, maxWidth = 500
                )
              )[colnames(.)],
              defaultColDef = colDef(maxWidth = 80),
              elementId = tbl_id
            )
          cp <- htmltools::em(
            glue(
              "GSEA results for {x}, testing an association between up- or down-", 
              "regulated genes for all {target} DSA groups incorporating all available ", 
              "annotations and features. A Normalized Enrichment Score (NES) > 0 ", 
              "indicates an association with differentially expressed genes. ", 
              "The total number of genes ", 
              "mapped to each type of DSA group is provided along with those found ",
              "beyond the Leading Edge. Given that multiple peaks may map to a gene, ",
              "genes may be included in more than one grouping."
            )
          )
          htmltools::div(
            id = tbl_id,
            class = "section level5",
            htmltools::h5(x),
            htmltools::div(
              class = "table", tbl, htmltools::tags$caption(cp)
            )
          )
        }
      )
  )
}
```


#### Directional Association {.tabset}

```{r run-gsea-dir}
ids_by_status <- results_tbl %>% 
  unnest(any_of("feature")) %>% 
  dplyr::filter(status != "Unchanged") %>% 
  unite(group, any_of(c("status", "region", "feature"))) %>% 
  split(.$group) %>% 
  lapply(pull, "gene_id")
rna_dir_scores <- rna %>% 
  lapply(
    \(x) setNames(-sign(x$logFC) * log10(x$PValue), x$gene_id)
  )
rna_gsea_dir <- rna_dir_scores %>% 
  lapply(\(x) fgsea(ids_by_status, x, minSize = min(msigdb_params$size))) %>% 
  lapply(as_tibble) %>% 
  lapply(
    separate, pathway, 
    into = c("status", "region", "feature")[seq_len(2 + has_features)],
    sep = "_", extra = "merge", remove = FALSE
  ) %>%
  lapply(arrange, pval) %>% 
  lapply(
    mutate,
    pathway = paste(target, pathway),
    padj = p.adjust(pval, enrich_params$adj),
    status = factor(status, levels = names(direction_colours)),
    region = factor(region, unname(regions)),
    leSize = map_int(leadingEdge, length),
    leadingEdge = map_chr(
      leadingEdge, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
    )
  ) %>% 
  lapply(
    dplyr::select, 
    pathway, any_of(c("status", "region", "feature")),
    padj, NES, ends_with("ize"), leadingEdge
  ) %>% 
  lapply(dplyr::filter, leSize >= enrich_params$min_sig)
any_gsea_dir <- map_lgl(rna_gsea_dir, \(x) any(x$padj < enrich_params$alpha))
```

```{r tbl-gsea-dir, echo=FALSE}
if (any(any_gsea_dir)) {
  tagList(
    any_gsea_dir %>% 
      which() %>% 
      names() %>% 
      lapply(
        \(x) {
          tbl_id <- paste0("tbl-gsea-dir", x) %>% 
              str_to_lower() %>% 
              str_replace_all("_", "-")
          tbl <- rna_gsea_dir[[x]] %>% 
            dplyr::select(-pathway) %>% 
            dplyr::filter(padj < enrich_params$alpha) %>% 
            reactable(
              sortable = TRUE, filterable = TRUE, showPageSizeOptions = TRUE,
              columns = list(
                status = colDef(
                  glue("{target} Status"), maxWidth = 100,
                  filterInput = function(values, name) {
                    tags$select(
                      # Set to undefined to clear the filter
                      onchange = sprintf(
                        "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
                        tbl_id, name
                      ),
                      # "All" has an empty value to clear the filter, and is the default option
                      tags$option(value = "", "All"),
                      lapply(unique(values), tags$option),
                      "aria-label" = sprintf("Filter %s", name),
                      style = "width: 100%; height: 28px; font-size: 12px;"
                    )
                  }
                ),
                region = colDef(
                  "Region Type", maxWidth = 120,
                  filterInput = function(values, name) {
                    tags$select(
                      # Set to undefined to clear the filter
                      onchange = sprintf(
                        "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
                        tbl_id, name
                      ),
                      # "All" has an empty value to clear the filter, and is the default option
                      tags$option(value = "", "All"),
                      lapply(unique(values), tags$option),
                      "aria-label" = sprintf("Filter %s", name),
                      style = "width: 100%; height: 28px; font-size: 12px;"
                    )
                  }
                ),
                feature = colDef(
                  "Feature Type", maxWidth = 100,
                  filterInput = function(values, name) {
                    tags$select(
                      # Set to undefined to clear the filter
                      onchange = sprintf(
                        "Reactable.setFilter('%s', '%s', event.target.value || undefined)",
                        tbl_id, name
                      ),
                      # "All" has an empty value to clear the filter, and is the default option
                      tags$option(value = "", "All"),
                      lapply(unique(values), tags$option),
                      "aria-label" = sprintf("Filter %s", name),
                      style = "width: 100%; height: 28px; font-size: 12px;"
                    )
                  }
                ),
                padj = enrich_cols$adj_p,
                NES = enrich_cols$NES,
                size = colDef("Group Size"),
                leSize = colDef("LE Size"),
                leadingEdge = colDef(
                  "Leading Edge", cell = \(value) with_tooltip(value, width = 60),
                  minWidth = 150, maxWidth = 500
                )
              )[colnames(.)],
              defaultColDef = colDef(maxWidth = 80),
              elementId = tbl_id
            )
          cp <- htmltools::em(
            glue(
              "GSEA results for {x}, testing an association between up- or down-", 
              "regulated genes for all {target} DSA groups incorporating all available ", 
              "annotations and features. A Normalized Enrichment Score (NES) > 0 ", 
              "indicates an association with up-regulation, with a negative score ",
              "being associated with down-regulation. The total number of genes ", 
              "mapped to each type of DSA group is provided along with those found ",
              "beyond the Leading Edge. Given that multiple peaks may map to a gene, ",
              "genes may be included in more than one grouping."
            )
          )
          htmltools::div(
            id = tbl_id,
            class = "section level4",
            htmltools::h4(x),
            htmltools::div(
              class = "table", tbl, htmltools::tags$caption(cp)
            )
          )
        }
      )
  )
}
```


### Combined Enrichment 

```{r load-gsea-rds}
gsea_dir <- file.path(all_paths$annotation, "gsea_dir.rds") %>% read_rds()
gsea_sig <- file.path(all_paths$annotation, "gsea_sig.rds") %>% read_rds()
```


```{r changed-rna-enrichment}
changed_rna_enrich <- gsea_sig %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(changed_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  ) %>% 
  .[map_int(., nrow) > 0]
tbl_changed_rna_enrich <- names(changed_rna_enrich) %>% 
  lapply(
    \(x) {
      id <- x %>% str_to_lower() %>% str_replace_all("_", "-")
      id <- paste0(id, "-changed-enrich-tbl")
      df <- changed_rna_enrich[[x]]
      tbl <- reactable(
        df, 
        searchable = TRUE, filterable = TRUE, showPageSizeOptions = TRUE,
        columns = enrich_cols[colnames(df)],
        elementId = id
      )
      cp <- tags$em(
        glue(
          "Results when combining directional GSEA results from {x} ",
          "with DSA enrichment results from {target}."
        )
      )
      htmltools::div(
        id = id, class = "section level5",
        h5("Combined Enrichment: DSA Changed"),
        tbl, tags$caption(cp)
      )
    }
  ) %>% 
  setNames(names(changed_rna_enrich))
network_changed_rna_enrich <- changed_rna_enrich %>% 
  map_int(nrow) %>% 
  is_weakly_greater_than(min(network_params$size)) %>% 
  which() %>% 
  names() %>%  
  sapply(
    \(x) {
      p <- changed_rna_enrich[[x]] %>% 
        mutate(
          gene_set_size = msigdb %>% 
            dplyr::filter(gs_name %in% pathway) %>% 
            split(.$gs_name) %>% 
            .[pathway] %>% 
            map_int(nrow),
          prop = n_cmn / gene_set_size,
          cmn_ids = setNames(cmn_ids, pathway)
        ) %>% 
        make_tbl_graph(
          gs = .$cmn_ids,
          alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
          hit_col = "n_cmn", size_col = "gene_set_size",
          min_dist = network_params$max_distance,
          max_gs = max(network_params$size)
        ) %>% 
        ggraph(layout = net_layout, weights = oc) +
        geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
        geom_node_point(
          aes(fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop),
          shape = 21
        ) +
        geom_node_text(
          aes(label = label, alpha = prop),
          colour = "black", size = 3, 
          data = . %>%
            mutate(
              label = str_replace_all(label, "_", " ") %>% 
                str_trunc(60) %>% 
                str_wrap(width = 18)
            ),
          repel = TRUE, bg.color = "white", bg.r = 0.1, 
        ) +
        scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
        scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
        scale_fill_viridis_c(option = "inferno", begin = 0.2) +
        scale_alpha(range = c(0.5, 1), trans = "sqrt") +
        scale_size_continuous(range = c(1, 10)) +
        scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
        scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
        guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
        labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
        theme_void() 
      
      ## Export the image
      img_out <- file.path(
        fig_path, glue("network_changed_rna_enrich_{x}.{fig_type}")
      )
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "The relationship between gene-sets considered enriched when ",
          "including GSEA results from {x} and changed sites from {target} ",
          "DSA analysis."
        )
      )
      htmltools::div(
        id = img_out %>% 
          basename() %>% 
          str_remove_all(glue(".{fig_type}$")) %>% 
          str_to_lower() %>% 
          str_replace_all("_", "-"),
        class = "section level5",
        htmltools::h5("Enrichment Network: DSA Changed"),
        htmltools::div(
          class = "figure", style = "text-align: center",
          htmltools::img(src = fig_link, width = "100%"),
          htmltools::tags$caption(cp)
        )
      )
      
    }, simplify = FALSE
  )  
```

```{r increased-rna-enrichment}
increased_rna_enrich <- gsea_dir %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(increased_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  ) %>% 
  .[map_int(., nrow) > 0]
tbl_increased_rna_enrich <- names(increased_rna_enrich) %>% 
  lapply(
    \(x) {
      id <- x %>% str_to_lower() %>% str_replace_all("_", "-")
      id <- paste0(id, "-increased-enrich-tbl")
      df <- increased_rna_enrich[[x]]
      tbl <- reactable(
        df, 
        searchable = TRUE, filterable = TRUE, showPageSizeOptions = TRUE,
        columns = enrich_cols[colnames(df)],
        elementId = id
      )
      cp <- tags$em(
        glue(
          "Results when combining directional GSEA results from {x} ",
          "with DSA enrichment results from {target}."
        )
      )
      htmltools::div(
        id = id, class = "section level5",
        h5("Combined Enrichment: DSA increased"),
        tbl, tags$caption(cp)
      )
    }
  ) %>% 
  setNames(names(increased_rna_enrich))
network_increased_rna_enrich <- increased_rna_enrich %>% 
  map_int(nrow) %>% 
  is_weakly_greater_than(min(network_params$size)) %>% 
  which() %>% 
  names() %>%  
  sapply(
    \(x) {
      p <- increased_rna_enrich[[x]] %>% 
        mutate(
          gene_set_size = msigdb %>% 
            dplyr::filter(gs_name %in% pathway) %>% 
            split(.$gs_name) %>% 
            .[pathway] %>% 
            map_int(nrow),
          prop = n_cmn / gene_set_size,
          cmn_ids = setNames(cmn_ids, pathway)
        ) %>% 
        make_tbl_graph(
          gs = .$cmn_ids,
          alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
          hit_col = "n_cmn", size_col = "gene_set_size",
          min_dist = network_params$max_distance,
          max_gs = max(network_params$size)
        ) %>% 
        ggraph(layout = net_layout, weights = oc) +
        geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
        geom_node_point(
          aes(fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop),
          shape = 21
        ) +
        geom_node_text(
          aes(label = label, alpha = prop),
          colour = "black", size = 3, 
          data = . %>%
            mutate(
              label = str_replace_all(label, "_", " ") %>% 
                str_trunc(60) %>% 
                str_wrap(width = 18)
            ),
          repel = TRUE, bg.color = "white", bg.r = 0.1, 
        ) +
        scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
        scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
        scale_fill_viridis_c(option = "inferno", begin = 0.2) +
        scale_alpha(range = c(0.5, 1), trans = "sqrt") +
        scale_size_continuous(range = c(1, 10)) +
        scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
        scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
        guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
        labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
        theme_void() 
      
      ## Export the image
      img_out <- file.path(
        fig_path, glue("network_increased_rna_enrich_{x}.{fig_type}")
      )
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "The relationship between gene-sets considered enriched when ",
          "including GSEA results from {x} and increased sites from {target} ",
          "DSA analysis."
        )
      )
      htmltools::div(
        id = img_out %>% 
          basename() %>% 
          str_remove_all(glue(".{fig_type}$")) %>% 
          str_to_lower() %>% 
          str_replace_all("_", "-"),
        class = "section level5",
        htmltools::h5("Enrichment Network: DSA Increased"),
        htmltools::div(
          class = "figure", style = "text-align: center",
          htmltools::img(src = fig_link, width = "100%"),
          htmltools::tags$caption(cp)
        )
      )
      
    }, simplify = FALSE
  )  
```

```{r decreased-rna-enrichment}
decreased_rna_enrich <- gsea_dir %>% 
  lapply(
    \(x) {
      x %>% 
        inner_join(decreased_enrich_df, by = c("pathway" = "gs_name")) %>% 
        dplyr::select(pathway, NES, pval, p, leadingEdge, gene_id) %>% 
        nest(p = all_of(c("pval", "p"))) %>% 
        mutate(
          meta_p = map_dbl(p, \(x) metap::maximump(unlist(x))$p),
          adj_p = p.adjust(meta_p, enrich_params$adj),
          cmn_ids = mapply(intersect, x = leadingEdge, y = gene_id),
          n_cmn = map_int(cmn_ids, length)
        ) %>% 
        arrange(meta_p) %>% 
        dplyr::filter(
          adj_p < enrich_params$alpha, n_cmn >= enrich_params$min_sig
        ) %>% 
        mutate(
          genes = map_chr(
            cmn_ids, \(x) paste(sort(unique(id2gene[x])), collapse = "; ")
          )
        ) %>% 
        dplyr::select(pathway, NES, meta_p, adj_p, n_cmn, genes, cmn_ids)
    }
  ) %>% 
  .[map_int(., nrow) > 0]
tbl_decreased_rna_enrich <- names(decreased_rna_enrich) %>% 
  lapply(
    \(x) {
      id <- x %>% str_to_lower() %>% str_replace_all("_", "-")
      id <- paste0(id, "-decreased-enrich-tbl")
      df <- decreased_rna_enrich[[x]]
      tbl <- reactable(
        df, 
        searchable = TRUE, filterable = TRUE, showPageSizeOptions = TRUE,
        columns = enrich_cols[colnames(df)],
        elementId = id
      )
      cp <- tags$em(
        glue(
          "Results when combining directional GSEA results from {x} ",
          "with DSA enrichment results from {target}."
        )
      )
      htmltools::div(
        id = id, class = "section level5",
        h5("Combined Enrichment: DSA Decreased"),
        tbl, tags$caption(cp)
      )
    }
  ) %>% 
  setNames(names(decreased_rna_enrich))
network_decreased_rna_enrich <- decreased_rna_enrich %>% 
  map_int(nrow) %>% 
  is_weakly_greater_than(min(network_params$size)) %>% 
  which() %>% 
  names() %>%  
  sapply(
    \(x) {
      p <- decreased_rna_enrich[[x]] %>% 
        mutate(
          gene_set_size = msigdb %>% 
            dplyr::filter(gs_name %in% pathway) %>% 
            split(.$gs_name) %>% 
            .[pathway] %>% 
            map_int(nrow),
          prop = n_cmn / gene_set_size,
          cmn_ids = setNames(cmn_ids, pathway)
        ) %>% 
        make_tbl_graph(
          gs = .$cmn_ids,
          alpha = enrich_params$alpha, p_col = "adj_p", gs_col = "pathway",
          hit_col = "n_cmn", size_col = "gene_set_size",
          min_dist = network_params$max_distance,
          max_gs = max(network_params$size)
        ) %>% 
        ggraph(layout = net_layout, weights = oc) +
        geom_edge_link(aes(width = oc^2, alpha = oc^2)) +
        geom_node_point(
          aes(fill = -log10(meta_p), size = n_cmn, stroke = 1.5 * prop),
          shape = 21
        ) +
        geom_node_text(
          aes(label = label, alpha = prop),
          colour = "black", size = 3, 
          data = . %>%
            mutate(
              label = str_replace_all(label, "_", " ") %>% 
                str_trunc(60) %>% 
                str_wrap(width = 18)
            ),
          repel = TRUE, bg.color = "white", bg.r = 0.1, 
        ) +
        scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
        scale_y_continuous(expand = expansion(c(0.1, 0.1))) +
        scale_fill_viridis_c(option = "inferno", begin = 0.2) +
        scale_alpha(range = c(0.5, 1), trans = "sqrt") +
        scale_size_continuous(range = c(1, 10)) +
        scale_edge_width(range = c(1, 6), limits = c(0, 1)) +
        scale_edge_alpha(range = c(0.1, 0.4), limits = c(0, 1))  +
        guides(edge_alpha = "none", edge_width = "none", alpha = "none") +
        labs(size = "Mapped\nGenes", fill = expr(paste("-", log[10], "p"))) +
        theme_void() 
      
      ## Export the image
      img_out <- file.path(
        fig_path, glue("network_decreased_rna_enrich_{x}.{fig_type}")
      )
      fig_fun(filename = img_out, width = fw, height = fh)
      print(p)
      dev.off()
      
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(
        glue(
          "The relationship between gene-sets considered enriched when ",
          "including GSEA results from {x} and decreased sites from {target} ",
          "DSA analysis."
        )
      )
      htmltools::div(
        id = img_out %>% 
          basename() %>% 
          str_remove_all(glue(".{fig_type}$")) %>% 
          str_to_lower() %>% 
          str_replace_all("_", "-"),
        class = "section level5",
        htmltools::h5("Enrichment Network: DSA Decreased"),
        htmltools::div(
          class = "figure", style = "text-align: center",
          htmltools::img(src = fig_link, width = "100%"),
          htmltools::tags$caption(cp)
        )
      )
      
    }, simplify = FALSE
  )  
```

Enrichment analysis was performed combined transcriptomic data with DSA results.
A simple strategy we used taking Wilkinson's maximum p-value method, which returns the probability of returning two p-values below the value of the largest p-value.
Transcriptomic enrichment was taken as the results from GSEA analysis, whilst enrichment results for `r target` were those obtained above.
Genes within the Leading Edge for GSEA and which were also mapped to the same pathway within DSA enrichment are returned as the common genes.

3 analysis were performed, with results only returned were significant combined enrichment was found for at least one pathway.

1. Enriched pathways within the DSA analysis using all changed sites, and transcriptomic GSEA taking only significance (i.e. non-directional results) to rank pathways
2. Enriched pathways within the DSA analysis using all increased sites, and transcriptomic GSEA taking directionality and significance to rank pathways
3. Enriched pathways within the DSA analysis using all decreased sites, and transcriptomic GSEA taking directionality and significance to rank pathways

For DSA enrichment, only the results obtained using the genomic background were used.


```{r print-dsa-rna-enrichment, echo=FALSE}
tagList(
  c(
    tbl_changed_rna_enrich, tbl_increased_rna_enrich, tbl_decreased_rna_enrich
  ) %>% 
    map_lgl(is, "shiny.tag") %>% 
    which() %>% 
    names() %>% 
    unique() %>% 
    lapply(
      \(x) {
        tagList(
          list(
            htmltools::div(
              id = x, class = "section level4 tabset",
              htmltools::h4(x, class = "tabset"),
              tbl_changed_rna_enrich[[x]],
              network_changed_rna_enrich[[x]],
              tbl_increased_rna_enrich[[x]],
              network_increased_rna_enrich[[x]],
              tbl_decreased_rna_enrich[[x]],
              network_decreased_rna_enrich[[x]]
            )
          )
        )
      }
    )
)
```




### Highly Ranked Genes {.tabset}


```{r rna-grl-to-plot}
grl_to_plot <- rna %>% 
  lapply(inner_join, results_tbl, by = 'gene_id') %>% 
  bind_rows(.id = "dataset") %>% 
  dplyr::filter(grepl("(In|De)creased", status.y)) %>% 
  mutate(
    rank = rank(PValue.x) + rank(PValue.y)
  ) %>% 
  arrange(rank) %>% 
  dplyr::filter(
    gene_id %in% unique(gene_id)[1:5]
  ) %>% 
  split(.$gene_id) %>% 
  lapply(colToRanges, "range", seqinfo = seqinfo(results)) %>% 
  GRangesList()
y_lim <- bwfl[[target]] %>%
  lapply(import.bw, which = unlist(range(grl_to_plot))) %>%
  lapply(\(x) c(0, max(x$score))) %>% 
  unlist() %>% 
  range() %>% 
  list() %>% 
  setNames(target)
```

By choosing the 5 genes highly tanked for both differential expression in any of `r pander(names(rna))`, and significantly changed `r target` signal, a set of example plots are shown below.

```{r plot-all-top-rna-ranges}
.makeCaption <- function(.gr) {
  if (is.null(.gr)) return(NULL)
  id <- unique(.gr$gene_id)
  gn <- id2gene[id]
  cp <- glue(
    "The {width(range(.gr))}bp region with all changed peaks mapped to {gn}, ", 
    "as a highly-ranked DE gene in ",
    glue_collapse(.gr$dataset, sep = ", ", last = " & "),
    ", and with at least significantly changed DSA region for {target}. ",
    "Regions associated with differential signal are annotated by colour."
  )
  cp
}
htmltools::tagList(
  lapply(
    seq_along(grl_to_plot),
    \(i) {
      ## Export the image
      nm <- str_to_lower(names(grl_to_plot)[[i]])
      id <- unique(grl_to_plot[[i]]$gene_id)
      img_out <- file.path(fig_path, glue("{nm}.{fig_type}"))
      fig_fun(filename = img_out, width = fw, height = fh)
      plotHFGC(
        grl_to_plot[[i]],
        hic = hic,
        features = feat_gr, featcol = hfgc_feat_col, featname = featnames,
        featsize = max(1, 0.8 * length(feat_gr)), featstack = "dense",
        genes = hfgc_genes, genecol = gene_col, 
        genesize = rep_len(c(2, 1), length(hfgc_genes)),
        coverage = bwfl, linecol = line_col, ylim = y_lim,
        annotation = annot, annotcol = direction_colours,
        cytobands = bands_df,
        collapseTranscripts = collapseTrans, maxTrans = 15,
        zoom = 20, max = max(1e4, width(range(grl_to_plot[[i]]))), 
        col.title = "black", background.title = "white", 
        rotation.title = 90, showAxis = FALSE, fontsize = 10,
        highlight = rgb(0, 0, 1, 0.3)
      )
      dev.off()
      ## Create html tags
      fig_link <- str_extract(img_out, "assets.+")
      cp <- htmltools::tags$em(.makeCaption(grl_to_plot[[i]]))
      htmltools::div(
        htmltools::div(
          id = nm,
          class = "section level4",
          htmltools::h4(id2gene[id]),
          htmltools::div(
            class = "figure", style = "text-align: center",
            htmltools::img(src = fig_link, width = "100%"),
            htmltools::tags$caption(cp)
          )
        )
      )
    }
  )
)
```

