## Normalisation

```{r load-extra-packages}
library(quantro)
library(qsmooth)
```


The `quantro` test was first applied [@HicksQuantro2015] to determine if treatment-specific binding distributions were found in the data.
Whilst this may not always be the case, the robustness of smooth-quantile normalisation (SQN) [@HicksSQN2017] will be applicable if data is drawn from different or near-identical distributions, and this method was applied grouping data by treatment.

```{r qtest, results='asis'}
qtest <- quantro(
  assay(filtered_counts, "logCPM"),
  groupFactor = filtered_counts$treat
)
pander(
  anova(qtest), 
  caption = paste(
    "*Results from qtest suggesting that the two treatment groups are drawn from",
    ifelse(
      qtest@anova$`Pr(>F)`[[1]] < 0.05,
      "different distributions.*",
      "the same distribution.*"
    )
  )
)
```


```{r add-qsmooth}
qs <-qsmooth(
  assay(filtered_counts, "logCPM"), group_factor = filtered_counts$treat
)
assay(filtered_counts, "qsmooth") <- qsmoothData(qs)
```

## Data Inspection {.tabset}

### QSmooth Weights

```{r plot-qs-weights, fig.height = 6, fig.cap = "*Quantile-specific weights used by the Smooth-Quantile normalisation. Low weights indicate signal quantiles which appear to be more specific within a group, whilst higher weights indicate similarity between groups.*"}
qsmoothPlotWeights(qs, xLab = "Quantiles", yLab = "Weights", mainLab = "QSmooth Weights")
```


### Density Plot: All Sliding Windows

```{r plot-density-windows, fig.height = 6, fig.width = 8, fig.cap = "*Density plot for __all windows prior to the selection of windows__ more likely to contain true signal. Retained windows will be those at the upper end, whilst discarded windows will be at the lower end.*"}
plotAssayDensities(
  window_counts, trans = "log1p", n_max = n_max, colour = "treat",
  linetype = "target"
) +
  scale_colour_manual(values = colours$treat) + 
  labs(
    x = "log(counts + 1)", y = "Density", colour = "Treat", linetype = "Target"
  )
```

### logCPM Densities

```{r plot-logcpm-densities, fig.height = 6, fig.cap = "*logCPM distributions for all retained windows A) prior to smooth quantile normalisation, and 2) after SQ normalisation.*"}
a <- filtered_counts %>% 
  plotAssayDensities("logCPM", colour = "treat", n_max = n_max) +
  scale_colour_manual(values = treat_colours) +
  labs(colour = "Treat")
b <- filtered_counts %>% 
  plotAssayDensities("qsmooth", colour = "treat", n_max = n_max) +
  scale_colour_manual(values = treat_colours) +
  labs(
    x = "Normalised logCPM", colour = "Treat"
  )
a + b + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
```

### RLE (Pre-Normalisation)

```{r plot-rle, dev = 'png', fig.height=6, fig.cap = glue("*RLE plot showing logCPM values. RLE values were calculated within each treatment group to account for the potentially different binding dynamics of {target}.*")}
filtered_counts %>% 
  plotAssayRle(
  "logCPM", fill = "treat", rle_group = "treat", n_max = n_max
) +
  geom_hline(yintercept = 0) + 
  facet_grid(.~treat, scales = "free_x", space = "free_x") +
  scale_x_discrete(labels = sample_labeller) +
  scale_fill_manual(values = treat_colours) +
  labs(x = "Sample", fill = "Treat")
```


### RLE (Post-Normalisation)

```{r plot-rle-normalised, dev = 'png', fig.height=6, fig.cap = glue("*RLE plot showing normalised logCPM. RLE values were calculated within each treatment group to account for the potentially different binding dynamics of {target}.*")}
filtered_counts %>% 
  plotAssayRle(
  "qsmooth", fill = "treat", rle_group = "treat", n_max = n_max
) +
  geom_hline(yintercept = 0) + 
  facet_grid(.~treat, scales = "free_x", space = "free_x") +
  scale_x_discrete(labels = sample_labeller) +
  scale_fill_manual(values = treat_colours) +
  labs(x = "Sample", fill = "Treat")
```


### PCA

```{r plot-pca, fig.height = 5, fig.cap = "*PCA plots for logCPM values A) before and B) after Smooth Quantile normalisation*"}
a <- plotAssayPCA(filtered_counts, "logCPM", colour = "treat", label = "label") +
  scale_colour_manual(values = treat_colours) +
  ggtitle("Pre-Normalisation")
b <- plotAssayPCA(filtered_counts, "qsmooth", colour = "treat", label = "label") +
  scale_colour_manual(values = treat_colours) +
  ggtitle("Post-Normalisation")
a + b + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
```


# Differential Binding Analysis

## Primary Analysis {.tabset}

```{r fit-model}
X <- model.matrix(~treat, data = colData(filtered_counts)) %>%
  set_colnames(str_remove(colnames(.), "treat"))
colData(filtered_counts)$design <- X
paired_cors <- block <- txt <- NULL
if (config$comparisons$paired) {
  block <- colData(filtered_counts)[[rep_col]]
  set.seed(1e6)
  ind <- sample.int(nrow(filtered_counts), n_max, replace = FALSE)
  paired_cors <- duplicateCorrelation(
    object = assay(filtered_counts, "qsmooth")[ind, ],
    design = X,
    block = block
  )$consensus.correlation
  txt <- glue(
    "Data were nested within {{rep_col}} as a potential source of correlation. ",
    "The estimated correlation within replicate samples was $\\hat{\\rho} = {{round(paired_cors, 3)}}$",
    .open = "{{", .close = "}}"
  )
}
fc <- ifelse(is.null(config$comparisons$fc), 1, config$comparisons$fc)
fit <- assay(filtered_counts, "qsmooth") %>%
  lmFit(design = X, block = block, correlation = paired_cors) %>% 
  treat(fc = fc, trend = TRUE, robust = FALSE)
fit_mu0 <- assay(filtered_counts, "qsmooth") %>%
  lmFit(design = X, block = block, correlation = paired_cors) %>% 
  treat(fc = 1, trend = TRUE, robust = FALSE)
res_cols <- c("logFC", "AveExpr", "t", "P.Value", "fdr")
rowData(filtered_counts) <- rowData(filtered_counts) %>% 
  .[!colnames(.) %in% c(res_cols, "p_mu0")] %>% 
  cbind(
    fit %>% 
      topTable(sort.by = "none", number = Inf) %>% 
      as.list %>% 
      setNames(res_cols) %>% 
      c(
        list(
          p_mu0 = topTable(fit_mu0, sort.by = "none", number = Inf)$P.Value
        )
      )
  ) 
```


After SQ-normalisation of logCPM values, the `limma-trend` method [@LawVoom2014] was applied to all retained windows.
A simple linear model was fitted taking `r treat_levels[[1]]` as the baseline and estimating the effects of `r treat_levels[[2]]` on `r target` binding within each sliding window, incorporating a trended prior-variance.
`r txt`
